### 2020-12-15 paper/authors.tex
Adds things needed to compile a paper

\collaboration{1000}{The LSST Dark Energy Science Collaboration (LSST DESC)}
\noaffiliation

\author[0000-0001-9789-9646]{Joe Zuntz}
\affiliation{Institute for Astronomy, University of Edinburgh, Edinburgh EH9 3HJ, United Kingdom}

%main partticipants and text contributors  (alphabetical)

\author[0000-0001-7956-0542]{Fran\c{c}ois Lanusse}
\affiliation{AIM, CEA, CNRS, Universit\'e Paris-Saclay, Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, F-91191 Gif-sur-Yvette, France}

\author{Alex Malz}
\affiliation{}

\author{Anze Slosar}
\affiliation{BNL}

% minor contributions and/or text writers (alphabetical)

### 2020-12-15 paper/paper.tex
Adds things needed to compile a paper

\PassOptionsToPackage{usenames,dvipsnames}{xcolor}
\documentclass[twocolumn,twocolappendix]{aastex63}

% Add your own macros here:
\pdfoutput=1 %for arXiv submission
%\usepackage{amsmath,amssymb,amstext}
\usepackage{amsmath,amstext}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage{ae,aecompl}
\usepackage[utf8]{inputenc}
\usepackage[figure,figure*]{hypcap}

\usepackage{url}
\urlstyle{same}

\usepackage{lineno}
\linenumbers
\modulolinenumbers[2]

% header settings
\shorttitle{Tomographic binning optimization}
\shortauthors{Author A et al.\ (LSST~DESC)}

% ======================================================================

\begin{document}
\title{The DESC Tomographic 3x2pt Optimization challenge}
\input{authors}

\begin{abstract}
Here is the abstract.
\end{abstract}

\keywords{methods: numerical -- methods: statistical -- dark energy  -- large-scale structure of the universe}

\accepted{June ??, 2020}
\submitjournal{the Astrophysical Journal Supplement}

%\tableofcontents%-----------------------------
%===========================
% BEGINNING OF THE MAIN TEXT
%===========================

\section{Introduction}
\label{sec:intro}

%=====================
% END OF THE MAIN TEXT
%=====================

\end{document}

### 2020-12-15 paper/paper.tex
adds minimal content

@@
 
+\newcommand{\placeholder}[1]{\textit{PLACEHOLDER: #1}}
+
 % header settings
@@
 \begin{abstract}
-Here is the abstract.
+The tomo challenge aims to compare methods for assigning photometrically observed galaxies to bins for a tomographic 3x2pt cosmological analysis, especially in cases where only restricted photometry can be used.
+We present this challenge, and the results of different methods when applied to the problem.
 \end{abstract}
@@
 
-\accepted{June ??, 2020}
-\submitjournal{the Astrophysical Journal Supplement}
+%\accepted{}
+%\submitjournal{the Astrophysical Journal Supplement}
 
@@
 
-\section{Introduction}
-\label{sec:intro}
+\section{Introduction \& Motivation}
+
+\begin{itemize}
+    \item Lensing/clustering/3x2pt 
+    \item Why bin galaxies for 3x2pt
+    \item Current binning standards
+    \item Constraints on potential binning methods
+    \item Goal of challenge: identify binning method(s) leading to best cosmology constraints 
+\end{itemize}
+
+\url{https://arxiv.org/abs/1901.06495}
+
+\url{https://arxiv.org/abs/astro-ph/0609338}
+
+\section{Challenge Design}
+
+\subsection{Philosophy}
+\begin{itemize}
+    \item Simplest challenge (optimistic data)
+    \item Openness?
+    \item Science-relevant metrics
+    \item Performance at scale
+\end{itemize}
+
+\subsection{Data}
+
+\begin{itemize}
+    \item CosmoDC2 - description
+    \item CosmoDC2 - discussion
+    \item Noise properties
+    \item Pre-selection 
+    \item Metacal as a limitation
+\end{itemize}
+
+\subsection{Metrics}
+\label{sec:metrics}
+
+\begin{itemize}
+    \item S/N of spectra, CCL
+    \item DETF Figure of merit, FireCrown(?)
+    \item mutual information
+\end{itemize}
+
+\placeholder{want to check performance at multiple stages of a cosmology analysis}
+
+\subsection{Infrastructure}
+\begin{itemize}
+    \item Pull request submissions
+    \item Data at NERSC
+\end{itemize}
+
+\subsection{Control Methods}
+\begin{itemize}
+    \item single bin
+    \item one galaxy per bin?
+    \item sanity check: random selection should be same as single bin
+\end{itemize}
+
+\section{Methods}
+\begin{itemize}
+    \item Gradient Boosting \@ jecampagne
+    \item Neural Network \@EiffL
+    \item Random Forest
+    \item \placeholder{cuts in photometry space without redshift estimation \@barber?}
+\end{itemize}
+
+\section{Results}
+\begin{itemize}
+    \item Big table of results
+    \item fisher matrix ellipses overplotted for top N methods
+    \item Plots of n(z)
+\end{itemize}
+
+\section{Discussion}
+\begin{itemize}
+    \item Overall effectiveness of method classes
+    \item Individual winner
+    \item Largest number of bins that proves workable
+\end{itemize}
+
+\subsection{Future directions}
+
+\begin{itemize}
+    \item spectroscopic incompleteness
+\end{itemize}
+
+\section{Conclusions}
 

### 2020-12-16 paper/paper.tex
Adds abstract

@@
 \linenumbers
-\modulolinenumbers[2]
+%\modulolinenumbers[2]
 
@@
 \begin{document}
-\title{The DESC Tomographic 3x2pt Optimization challenge}
+\title{The LSST DESC 3x2pt Tomography Optimization Challenge}
 \input{authors}
@@
 \begin{abstract}
-The tomo challenge aims to compare methods for assigning photometrically observed galaxies to bins for a tomographic 3x2pt cosmological analysis, especially in cases where only restricted photometry can be used.
-We present this challenge, and the results of different methods when applied to the problem.
+In this paper we present the results of the DESC 3x2pt tomography challenge, which aims to compare strategies for optimizing the photometric binning required for the main LSST 3x2pt analysis.
+This task is made particularly delicate in the context of a metacalibrated lensing survey, as only the photometry from the bands  included in the metacalibration process (riz and potentially g) can be used to define this tomographic binning.
+Using a set of realistic simulations including true galaxy redshifts and simulated photometry, the goal of the challenge is to propose a bin assignment strategy that maximizes the overall Figure of Merit of a 3x2pt analysis.
+This setting is idealized as it ignores spectroscopic completeness  issues for the training sample, but complex enough to address the main question of binning optimization.
+Based on the challenge dataset, we review and compare the different methods that have been proposed by participants.
+We find that even from this limited photometry information, various methods are able to yield binning schemes that achieve high FoM scores.
 \end{abstract}
 
-\keywords{methods: numerical -- methods: statistical -- dark energy  -- large-scale structure of the universe}
+\keywords{methods: statistical -- dark energy  -- large-scale structure of the universe}
 

### 2021-01-04 bin/batch_classify.py
add buzzard mode

@@ -62,15 +62,19 @@ def setup(queue):
                 queue.add_job(name, task, subset=gpu_or_cpu)
 
 
-def task(name, bands, classifier, nbin, config_index):
+def task(name, bands, classifier, nbin, config_index, buzzard=False):
     f = open(f"logs/{name}_log.txt", "w")
     print("Running",name, classifier, nbin)
     if classifier in off_by_one:
         nbin -= 1
     cmd = f"python bin/run_one_batch.py {name} {bands} {classifier} {nbin} {config_index}"
+    if buzzard:
+        cmd += " --buzzard"
     status = subprocess.call(cmd.split(), stderr=subprocess.STDOUT, stdout=f)
     return status
 
+def buzzard_task(name, bands, classifier, nbin, config_index):
+    return task(name, bands, classifier, nbin, config_index, buzzard=True)
 
 import argparse
 parser = argparse.ArgumentParser(description='Batch run jobs')
@@ -78,12 +82,18 @@ parser.add_argument('cpu_or_gpu', type=int, choices=[GPU, CPU], help='Run jobs f
 parser.add_argument('--setup', action='store_true', help='Set up the DB and exit')
 parser.add_argument('--query', action='store_true', help='Display next job')
 parser.add_argument('--list', action='store_true', help='List remaining jobs')
+parser.add_argument('--buzzard', action='store_true', help='Run on Buzzard')
 
 
 def main():
     args = parser.parse_args()
-    db = "classify.db"
-    queue = task_queue.SlurmTaskQueue(db, task, {"classifier": str, "bands":str, "nbin": int, "config_index":int})
+    if args.buzzard:
+        db = "classify_buzzard.db"
+        f = buzzard_task
+    else:
+        db = "classify.db"
+        f = task
+    queue = task_queue.SlurmTaskQueue(db, f, {"classifier": str, "bands":str, "nbin": int, "config_index":int})
     if args.setup:
         setup(queue)
     elif args.query:

### 2021-01-04 bin/run_one_batch.py

@@ -4,10 +4,17 @@ import sys
 sys.path.append(dir_path)
 import numpy as np
 
-task_name, bands, classifier, nbin, index = sys.argv[1:]
+task_name, bands, classifier, nbin, index = sys.argv[1:6]
 nbin = int(nbin)
 index = int(index)
 
+if len(sys.argv) > 6:
+    if sys.argv[6] == "--buzzard":
+        buzzard=True
+    else:
+        raise ValueError(f"Unknown cmd {sys.argv}")
+else:
+    buzzard=False
 
 scratch = os.environ['SCRATCH']
 
@@ -29,9 +36,14 @@ full_config = yaml.safe_load(open(fn))['run'][classifier]
 settings = list(full_config.values())[0]
 
 
-
-training_file = 'data-train/training-cut.hdf5'
-validation_file = 'secret/testing.hdf5'
+if buzzard:
+    training_file = 'buzzard/training-cut.hdf5'
+    validation_file = 'buzzard/testing.hdf5'
+    output_file = f'{scratch}/tomo_challenge_results_buzzard/{task_name}.npy'
+else:
+    training_file = 'data-train/training-cut.hdf5'
+    validation_file = 'secret/testing.hdf5'
+    output_file = f'{scratch}/tomo_challenge_results/{task_name}.npy'
 
 training_data = tc.load_data(
     training_file,
@@ -73,4 +85,4 @@ results = run_one(classifier, bands, settings, training_data, training_z, valida
              validation_z, metrics, metrics_fn)
 
 
-np.save(f'{scratch}/tomo_challenge_results/{task_name}.npy', results)
+np.save(output_file, results)

### 2021-01-04 paper/paper.tex
text for intro and motivationn

@@
 
-\section{Introduction \& Motivation}
+\section{Introduction}
+Weak gravitational lensing (WL) has emerged over the last decade as a powerful
+cosmological probe \citep{cfhtlens,des,kids,hsc}.  WL
+uses measurements of coherent shear distortion to the observed shapes of galaxies
+to track the evolution of large-scale gravitational fields.  It measures the integrated
+gravitational potential along lines of sight to source galaxies, and can thence constrain
+the laws of gravity, the expansion history of the Universe, and the history and growth
+of cosmic structure.
+
+WL has proven especially powerful in combination with galaxy clustering measurements,
+which can measure the density of matter up to an unknown bias function.  The high signal
+to noise of such measurements and the relative certainty of the redshift of these foreground
+samples breaks degeneracies in the systematic errors that affect WL.
+
+The \emph{3x2pt} method has become the standard method for performing this combination.
+In this method, two-point correlations are computed among and between two samples, the shapes of 
+background (source) galaxies and the locations of foreground galaxies, which trace foreground
+dark matter haloes.  The three combinations (source-source, source-lens, and lens-lens) are
+measured in either Fourier or configuration space, and can be predicted from a combination of 
+perturbation theory and simulation results.  The method has been used in the Dark Energy Survey, DES
+\citep{des-3x2pt}, and to combine Baryon Oscillation Spectroscopic Survey, BOSS, with the Kilo-Degree 
+Survey, KiDS \citep{kids-3x2pt}.
+
+Most lensing and 3x2pt analyses have been performed \emph{tomographically}, 
+binning galaxies by redshift.
+This approach captures almost all the available information in lensing data, since it measures
+an integrated effect and so galaxies nearby in redshift probe very similar fields.  For photometric
+foreground samples it is similarly near-lossless, since redshift estimates of such galaxies have
+large uncertainties\footnote{Spectroscopic foreground samples may be more likely to see significant 
+gains from moving beyond spectroscopic methods.}.  Binning galaxies by approximate redshift lets us 
+model galaxy bias, intrinsic alignments, and other systematic errors en masse in a more tractable way.
+While fully 3D methods have been proposed and prototyped \citep{kitching,more}, the tomographic 
+approach remains the standard within the field.
+
+Tomographic 3x2pt measurements will be a key science goal in the upcoming \emph{Stage IV} surveys,
+including the Rubin Observatory \citep{rubin} and the Euclid and Roman space telescopes \citep{euclid,roman}.
+
+In this paper we discuss and evaluate methods for assigning objects to tomographic bins, and 
+specifically the impact
+of using a limited set of color bands to make those assignments, motivated by requirements of using the
+metacalibration method for galaxy shape measurement to limit biases.  Tomographic methods have been 
+explored before in \citet{jain} and \citet{kitching2019}.  In this paper we describe the results of a 
+challenge using simulated Rubin-like data with a range of methods submitted by different teams.
+
+In \autoref{sec:motivation} we explain the need for methods that work on limited sets of bands in the context of upcoming lensing surveys.  \textbf{Complete structure...}
+
+\section{Motivation}
+\label{sec:motivation}
+
+The methodology used to assign galaxies to bins faces special challenges when we use a particularly
+effective approach for measuring galaxy shear, called \emph{metacalibration} (metacal),
+which was introduced in \citet{sheldonhuff} and developed further in \citet{sheldon}.
+In metacal, galaxy images are deconvolved from the point-spread function (PSF), sheared, and
+reconvolved before measurement, allowing one to directly compute the response of a metric
+to underlying shear and correct for it when computing two-point measurements.  This can
+almost completely remove errors due to model and estimator (noise) bias from WL data, at least
+when PSFs are well-measured and oversampled by the pixel grid.
+
+Furthermore, metacal provides a solution to the pernicious and general problem of
+\emph{selection biases} in WL.  Measurements of galaxy shear are known to have noise-induced errors
+that are
+highly covariant with those on other galaxy properties, including size and, most importantly, flux/
+magnitude.  It follows that a cut (or re-weighting) on galaxy magnitudes, or on any quantity derived
+from  them, will induce a shear bias since galaxies will preferentially fall either side of the cut
+depending on their shear.  Within metacal, we can solve this problem by performing the cut on both
+sheared variants of the catalog, and measuring the difference between the post-cut shear in the two
+variants.  In DES the corrected biases were found up to around 3\%, far larger than
+the requirements for the current generation of surveys \cite{des-y1-cat}.
+
+In summary: metacal allows us to correct for significant selection biases, but only if all selections
+are performed using only bands in which the PSF is measured well enough for deconvolution to be
+performed.  For Rubin, this means using only the $r$, $i$, $z$, and perhaps $g$ bands.  In this work we
+therefore study how well we can perform tomography using only this limited photometric information.  In
+particular, this limitation prevent us from using the most obvious method for tomography,
+and computnig a complete photometric probability density function (PDF) for each galaxy and assigning
+a bin using the mean or peak of that PDF.
+
+Simulation methods can also be used to correct for selection biases, provided that they match the
+real data to high accuracy.  If we determine that limiting the bands we can use for tomography results
+in a significant decrease in our overall signal-to-noise, outweighing the gain from the improved shape
+calibration then this might suggest moving to rely more on such methods\footnote{One can account for
+residual shear estimation uncertainty during parameter estimation by marginalizing over an unknown
+factor.  Widening the prior on this factor decreases the overall constraining power of the analysis.}.
 
 \begin{itemize}
-    \item Lensing/clustering/3x2pt 
-    \item Why bin galaxies for 3x2pt
-    \item Current binning standards
-    \item Constraints on potential binning methods
     \item Goal of challenge: identify binning method(s) leading to best cosmology constraints 

### 2021-01-04 paper/paper.tex
more content

@@
 challenge using simulated Rubin-like data with a range of methods submitted by different teams.
+We explore how many tomographic bins we can effectively generate using only a subset of bins, and compare methods submitted to the challenge.
 
@@
 
-\url{https://arxiv.org/abs/1901.06495}
-
-\url{https://arxiv.org/abs/astro-ph/0609338}
-
 \section{Challenge Design}
 
+The Dark Energy Science Collaboration (DESC) Tomographic Challenge was launched in May 2019 and 
+advertised among the LSST cosmology community, though it was open to outside entrants.
+
+In the challenge, participants assigned simulated galaxies to tomographic bins,
+using only their (g)riz bands and associated errors.  Their goal was to maximize the cosmological
+information in the final split data set; this is generally acheived by separating different bins by 
+redshift as much as possible.  Particpants were free to optimize or simply select nominal (target) 
+redshift bin edges from their training sample, but in either case had to classify galaxies into the 
+different bins.
+
 \subsection{Philosophy}
+
+Since this was a preliminary challenge designed to explore whether it is possible \emph{in theory}
+to use only (g)riz for tomography, we chose to simplify several aspects of the data.  In realistic 
+situations we will have access to limited sized training sets for WL photometric redshifts, since 
+spectra are time-consuming to obtain.  These training sets will also be highly incomplete compared
+to full catalogs, especially at faint magnitudes where spectroscopic coverage is sparse.  In this
+challenge we avoided both these issues -- the training samples we provided (see below) were comparable
+in size to the testing sample, and drawn from the same population.  Given this, the challenge
+represents a best-case scenario -- if no method succeeded on this easier case then more realistic
+cases would probably be impossible.
+
+Despite these simplifications, the other aspects of the challenge and process were designed to be
+as directly relevant as possible to the particular WL science case we focus on.  The data set
+was chosen to mimic the population, noise, and cuts we will use in real data (see \autoref{sec:data}, 
+and the metrics were designed to be as close to the science goals as possible, rather than a lower
+level representation (see \autoref{sec:metrics}). The data set was also large enough to pose a 
+reasonably realistic test of methods at the large data volume required for upcoming surveys, where
+$10^9$ -- $10^{10}$ galaxies will be measured.
+
+The challenge was open to any entrants, not just those already involved in DESC, but it was
+advertised primarily through Rubin, and so is perhaps best described as semi-public.
+Participants tested and ran their methods locally, but submitted code to be run as part of a central
+combined analysis.  No prize was offered apart from recognition.
+
 \begin{itemize}
-    \item Simplest challenge (optimistic data)
-    \item Openness?
-    \item Science-relevant metrics
     \item Performance at scale
@@
 \subsection{Data}
+\label{sec:data}
 
@@
 \label{sec:metrics}
+
+In this challenge we use only a source (background) population, since the metacal requirements
+do not apply to the foreground lens sample. We do, however, calculate
 
@@
 
+\appendix{Mistakes}
+The challenge organizers made a number of mistakes when building and running the challenge.
+These do not invalidate the process in any way, but we describe them here in the hope of
+being useful for future challenge designers.
+
+\begin{itemize}
+    \item Infrastructure for checking submissions
+    \item Guidance for users regarding caching and batch processing
+    \item Timing
+    \item Training set size
+    \item Advertisement
+\end{itemize}
+
 %=====================

### 2021-01-04 paper/paper.tex
metrics text

@@
 Participants tested and ran their methods locally, but submitted code to be run as part of a central
-combined analysis.  No prize was offered apart from recognition.
-
-\begin{itemize}
-    \item Performance at scale
-\end{itemize}
+combined analysis.  No prize was offered beyond recognition.
 
@@
 
-In this challenge we use only a source (background) population, since the metacal requirements
-do not apply to the foreground lens sample. We do, however, calculate
+In this challenge we use only a lensing source (background) population, since the metacal requirements
+do not apply to the foreground lens sample. We do, however, calculate 3x2pt metrics using the
+same source and lens populations, both because this scenario is one that will be used in some analyses,
+and because clustering-like statistics of lensing samples are important for studies of intrinsic
+alignments. We use two different sets of metrics, one based on the over signal to noise of the spectra 
+derived from the samples, and one based on a figure-of-merit for the constraints that would be obtained 
+using them.  \textbf{Discuss how correlated these turn out to be here}.
 
-\begin{itemize}
-    \item S/N of spectra, CCL
-    \item DETF Figure of merit, FireCrown(?)
-    \item mutual information
-\end{itemize}
+For each bin $i$ generated by a challenge entry, we make a histogram of the true redshifts 
+of each galaxy assigned to the bin.  This is then used as our number density $n_i(z)$ in the metrics 
+below. Both metrics reward bins that can
+cleanly separate galaxies by redshift, since this reduces the covariance between the samples. No
+method can perfectly separate them, so good methods reduce the tails of $n_i(z)$ that overlap with 
+neighbouring bins.
 
-\placeholder{want to check performance at multiple stages of a cosmology analysis}
+We developed two implementations of each of these metrics, one using the Core Cosmology Library 
+\citep{ccl} and one the JAX Cosmology Library \citep{jax-cosmo}, which provides differentiable theory 
+predictions using the JAX library \citep{jax}.  In production we use the latter since it provided a
+more stable calculation of metric 2.
+
+Metric 1 would be approximately measurable on real data without performing a full cosmology analysis,
+whereas metric 2 would be the final output of such an analysis.  We therefore simulate performance
+tests at multiple stages in the wider process.
+
+\subsubsection{Metric 1: Sigal-to-noise}
+Our first metric is the total signal-to-noise ratio of the power spectrum derived from the assigned
+$n_i(z)$ values:
+\begin{equation}
+    S = \mu^{T} C^{-1} \mu
+\end{equation}
+where $\mu$ is a stack of the theoretical predictions for the $C_\ell$ spectra for each tomographic 
+bin, in turn, and $C$ a Gaussian estimaate of the covariance between them, as described in
+\citet{wl-review} and summarized in \autoref{app:theory}.   We compute this metric for lensing alone, clustering alone, and the full 3x2pt combination.
+
+\subsubsection{Metric 2: Figure-of-merit}
+
+Metric 2 uses a figure of merit based on that presented by the Dark Energy Task Force \citep{detf}.  We 
+use the inverse area of a Fisher Matrix ellipse, which approximates the size in two axes 
+of the posterior PDF one would obtain on analysing the spectra:
+\begin{align}
+    F &= \left( \frac{\partial \mu}{\partial \theta} \right)^T C^{-1} \left( \frac{\partial \mu}{\partial \theta} \right), \\
+    S &= \frac{1}{2 \pi \sqrt{\det{([F^{-1}]_{\Omega_c,\sigma_8})}}}
+\end{align}
+where $[F^{-1}]_{\Omega_c,\sigma_8}$ extracts the 2x2 matrix for these two parameters.
 
@@
 
-\appendix{Mistakes}
+\appendix
+
+\section{Theory}\label{app:theory}
+In our metrics, the power spectrum between two tomographic bins $i$ and $j$ is computed using either the Core Cosmology Library \citep{ccl} or the JAX Cosmology Library \citep{jax-cosmo} as:
+\begin{equation}
+    C^{ij}_\ell = \int_0^{\infty} \frac{q_i(\chi) q_j(\chi)}{\chi^2} P\left(\frac{\ell +\frac{1}{2}}{\chi}, \chi \right) \mathrm{d}\chi
+\end{equation}
+where $\chi = \chi(z)$ is the comoving distance and $P$ the non-linear matter power spectrum at
+a fiducial cosmology.  The kernel functions $q_i(\chi)$ are different for the lensing and clustering samples:
+\begin{align}
+    q^{\mathrm{cluster}}_i(\chi) &= n_i(\chi)\\
+    q^{\mathrm{lensing}}_i(\chi) &= \frac{3}{2}\Omega_m \left(\frac{\mathrm{H}_0}{c}\right)^2 \frac{\chi}{a(\chi)} \int_\chi^{\infty} \frac{\chi' - \chi}{\chi'} n_i(\chi')\,\,\mathrm{d}\chi'
+\end{align}
+where $n_i(\chi) = n_i(z) \frac{\mathrm{d}z}{\mathrm{d}\chi}$.
+
+We evaluate these at 100 $\ell$ values from $\ell_\mathrm{min}=100$ to $\ell_\mathrm{max}=2000$.
+
+We use a Gaussian estimate for the covariance \citep{takada_jain}; this also incorporates the number of galaxies in the
+sample:
+\begin{equation}
+    \mathrm{Cov}(C^{ij}_\ell, C^{mn}_\ell) = \frac{1}{(2 \ell + 1)\Delta\ell f_\mathrm{sky}}(D^{im}_\ell + D^{jn}_\ell)(D^{in}_\ell + D^{jm}_\ell)
+\end{equation}
+where $D^{ij}_\ell = C^{ij}_\ell + N^{ij}_\ell$ and we assume an $f_\mathrm{sky}=0.25$.  The noise spectra are:
+\begin{align}
+N^{\mathrm{cluster},ij}_\ell = \delta_{ij} / n_i \\
+N^{\mathrm{lensing},ij}_\ell = \delta_{ij} \sigma_e^2 / n_i
+\end{align}
+where $n_i$ is the number density of galaxies in bin $i$ (we asssume equally weighted galaxies, and a 
+total number density over all bins of 20 galaxies per square arcminute) and $\sigma_e=0.26$.
+
+\section{Mistakes}
 The challenge organizers made a number of mistakes when building and running the challenge.

### 2021-01-18 paper/paper.tex
Paper text

@@
 \usepackage[figure,figure*]{hypcap}
-
+\usepackage{natbib}
 \usepackage{url}
@@
 
-\begin{itemize}
-    \item Goal of challenge: identify binning method(s) leading to best cosmology constraints 
-\end{itemize}
-
 \section{Challenge Design}
@@
     F &= \left( \frac{\partial \mu}{\partial \theta} \right)^T C^{-1} \left( \frac{\partial \mu}{\partial \theta} \right), \\
-    S &= \frac{1}{2 \pi \sqrt{\det{([F^{-1}]_{\Omega_c,\sigma_8})}}}
+    S &= \frac{1}{2 \pi \sqrt{\det{([F^{-1}]_{p_1, p_2})}}}
 \end{align}
-where $[F^{-1}]_{\Omega_c,\sigma_8}$ extracts the 2x2 matrix for these two parameters.
+where $[F^{-1}]_{p_1, p_2}$ extracts the 2x2 matrix for two parameters.  We use both the original
+DETF variant in which $p_1 = w_0, p_2 = w_a$, representing the constraining power on dark energy 
+evolution, and a version in which $p_1 = \Omega_c$ and $p_2 = \sigma_8$, representing constraining
+power on overall cosmic structure amplitude and its evolution.
 
@@
 \subsection{Control Methods}
-\begin{itemize}
-    \item single bin
-    \item one galaxy per bin?
-    \item sanity check: random selection should be same as single bin
-\end{itemize}
+
+Two control methods were used to test the challenge machinery and metrics.
+The first method randomly assigned galaxies to each tomographic bin, resulting
+in bins which each had the same $n(z)$, on average.  Using this method, we expect
+that increasing the number of bins should not increase any metric, since bins are
+maximally correlated.  We find this to be true for all our metrics.
+
+Second, we use a method which uses only the $i$-band to select tomographic bins,
+splitting galaxies in equal-sized bins in this magnitude. We expect only a small
+increase in metrics with the number of bins, and also that the addition of the $g$-band
+should make no difference to metrics.  Again, we find this to be true.
+
+One additional method was submitted to the challenge with the caveat that it could
+never be applied to real data and should therefore considered a form of control method.
+This method was {\sc Utopia}, which found, for each object in the testing data, the
+nearest neighbour object in the training sample.  The high density, completeness, and
+representativeness of our training set made this possible; realistics spectroscopic training
+sets are too sparse.  The method performed well, as expected, 
+but was not the highest scoring overall.
 
 \section{Methods}
-\begin{itemize}
-    \item Gradient Boosting \@ jecampagne
-    \item Neural Network \@EiffL
-    \item Random Forest
-    \item \placeholder{cuts in photometry space without redshift estimation \@barber?}
-\end{itemize}
+
+\subsection{ {\sc RandomForest} }
+This method was submitted by the challenge organizers and used a random
+forest algorithm to assign galaxies.
+
+Random forests \cite{breiman2001} train a collection of individual \emph{decision trees},
+each of which consists of a set of bifurcating comparisons in which different features in
+the data (in this case, magnitudes and errors) are compared to criterion values to choose
+which branch to follow.  Each ``leaf'' (end comparison) selects a classification bin for
+an input galaxy, and the tree is trained to choose comparison features and comparison values
+which maximally split the discrimination at each step, and in final leaf purity.
+
+Since decision trees tend to over-fit data, random forests generate a suite of decision
+trees, each randomly choosing a subset of features on which to train at each bifurcation.
+An average over the trained trees is taken as the overall classification.
+
+In this implementation we chose nominal bin edges by splittng ensuring equal numbers of training set 
+objects were in each bin, and then trained the forest to map magnitudes, magnitude errors, and
+galaxy sizes to bin indices.  We use the {\sc scikit-learn} implementation of the random forest
+\citep{scikit-learn}.
+
+% Please describe your methods with some very general / high level background theory
+% and a focus on the spsecific implementation.  Add any citations you like with bibtex in
+% paper.bib
+% Feel free to rename your method, but please leave a comment explaining so that I can do the
+% same in the tables.
+
+\subsection{ {\sc Autokeras LSTM} }
+
+\subsection{ {\sc CNN} }
+
+\subsection{ {\sc ComplexSOM} }
+
+\subsection{ {\sc ENSEMBLE} }
+
+\subsection{ {\sc GPzBinning} }
+
+\subsection{ {\sc JaxCNN} }
+
+\subsection{ {\sc JaxResNet} }
+
+\subsection{ {\sc LGBM} }
+
+\subsection{ {\sc LSTM} }
+
+\subsection{ {\sc NeuralNetwork 1} }
+
+\subsection{ {\sc NeuralNetwork 2} }
+
+\subsection{ {\sc PCACluster} }
+
+\subsection{ {\sc SimpleSOM} }
+
+\subsection{ {\sc TCN} }
+
+\subsection{ {\sc UTOPIA} }
+
+\subsection{ {\sc ZotBin} }
+
+\subsection{ {\sc ZotNet} }
+
+\subsection{ {\sc funbins} }
+
+\subsection{ {\sc mlpqna} }
+
+\subsection{ {\sc myCombinedClassifiers} }
 
 \section{Results}
+
+\begin{table*}[]
+\begin{tabular}{l|llll|llll|}
+                & \multicolumn{4}{c|}{\textbf{riz}}      & \multicolumn{4}{c|}{\textbf{griz}}                                \\ \hline
+\textbf{Method} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} \\ \hline
+{\sc Autokeras\_LSTM } & 31.1 & 74.9    & 103.4    & nan    & 44.1             & 67.5             & 122.2             & 98.4\\
+{\sc CNN } & 27.1 & 50.8    & 76.7    & 103.3    & 30.5             & 57.7             & 95.4             & 122.4\\
+{\sc ComplexSOM } & 38.0 & 52.1    & 94.4    & 101.6    & 34.9             & 45.0             & 91.6             & 100.3\\
+{\sc ENSEMBLE1 } & 65.6 & 65.6    & 94.8    & 0.0    & 0.0             & 0.0             & 0.0             & 0.0\\
+{\sc funbins } & 36.8 & 82.5    & 122.1    & 141.8    & 42.1             & 100.4             & 142.5             & 167.2\\
+{\sc GPzBinning } & 26.2 & 49.7    & 80.8    & 111.7    & 27.8             & 55.3             & 87.2             & 126.9\\
+{\sc IBandOnly } & 38.0 & 50.0    & 54.4    & 57.3    & 38.0             & 50.0             & 54.4             & 57.3\\
+{\sc JaxCNN } & 59.9 & 101.9    & 105.2    & nan    & 79.7             & 125.2             & 150.0             & nan\\
+{\sc JaxResNet } & 73.7 & 111.1    & 131.6    & nan    & 82.2             & 126.0             & nan             & 161.5\\
+{\sc LGBM } & 27.0 & 50.8    & 78.8    & 108.2    & 30.3             & 57.4             & 92.9             & 125.2\\
+{\sc LSTM } & 26.7 & 51.1    & 81.4    & 103.6    & 30.2             & 57.1             & 95.9             & 126.9\\
+{\sc mlpqna } & 39.3 & 64.7    & 93.2    & 121.8    & 42.8             & 72.2             & 109.0             & 133.7\\
+{\sc myCombinedClassifiers } & 35.3 & 60.5    & 83.5    & 120.4    & 39.4             & 63.9             & 93.1             & 148.9\\
+{\sc NeuralNetwork\_1 } & 0.0 & 0.0    & 0.0    & 0.0    & 0.0             & 0.0             & 0.0             & 0.0\\
+{\sc NeuralNetwork\_2 } & 0.0 & 0.0    & 0.0    & 0.0    & 0.0             & 0.0             & 0.0             & 0.0\\
+{\sc PCACluster } & 29.5 & 68.1    & 75.8    & 0.0    & 50.1             & 72.7             & 90.8             & 0.0\\
+{\sc Random } & 1.2 & 1.2    & 1.2    & 1.2    & 1.2             & 1.2             & 1.2             & 1.2\\
+{\sc RandomForest } & 39.5 & 65.2    & 93.2    & 118.5    & 43.0             & 72.8             & 110.1             & 136.5\\
+{\sc SimpleSOM } & 39.5 & 59.9    & 77.9    & 98.5    & 42.2             & 73.2             & 108.0             & 134.8\\
+{\sc TCN } & 40.0 & 65.3    & 90.9    & 110.5    & 44.1             & 72.8             & 108.0             & 134.0\\
+{\sc UTOPIA } & 39.0 & 63.7    & 90.8    & 112.2    & 43.1             & 74.0             & 112.8             & 139.3\\
+{\sc ZotBin } & 64.8 & 106.4    & 121.8    & 135.3    & 77.6             & 120.0             & 141.8             & 154.7\\
+{\sc ZotNet } & 73.6 & 111.2    & 131.8    & 145.9    & 83.7             & 128.5             & 150.1             & 167.2\\
+\end{tabular}
+\caption{INCOMPLETE, AND AUTO-GENERATED SO NO POINT EDITING! Values of the 3x2pt DETF figure-of-merit achieved by entrant methods on the 
+CosmoDC2 part of the challenge.}
+\end{table*}
+
 \begin{itemize}
@@
     \item Individual winner
-    \item Largest number of bins that proves workable
 \end{itemize}
+
+\subsection{Individual winners}
+
+RUN WINNERS WITH UGRIZY TO GET AN IDEA OF THE LOSS FROM DROPPING U AND Y.
+
+\subsection{Large bin numbers}
+
+It is clear from values above that fears about restricted photometry catastrophically limiting
+the information needed for tomography are unfounded.  In both the CosmoDC2 and Buzzard analyses,
+even the $riz$ bands alone are sufficient to split objects into nine reasonably separated
+tomographic bins.
+Once bins become this narrow the primary analysis concern will become the calibration of overall
+bin $n(z)$ values, rather than initial tomography, so we consider this to be reassuring.
+
+The inclusion of the $g$ band typically increases metrics by around 10--20\%, including for the
+highest scoring methods.  This value should be weighed against the challenge of high-accuracy
+PSF measurement for this band, but until the end of the survey this is unlikely to be
+the limiting factor in overall precision.
 
@@
 \section{Conclusions}
+
+\bibliography{paper}
 
@@
 \begin{itemize}
+    \item Splitting classification from nominal bin choice
     \item Infrastructure for checking submissions

### 2021-01-20 paper/paper.tex
add stacking method

@@
 \usepackage{url}
+\usepackage{listings}
+
 \urlstyle{same}
@@
 
-\subsection{ {\sc myCombinedClassifiers} }
+\subsection{ {\sc Stacked Generalization} }
+
+\emph{Stacked generalization} consists of stacking the output of several individual estimators and 
+using a classifier to compute the final prediction. Stacking methods use the strength of each 
+individual estimator by using their outputs as input of a final estimator.  This entry stacked
+a \emph{Gradient Tree Boosting} (GTB) classifier and then used a Logistic Regression as the final step.
+
+GTB is a machine learning method which combines decision trees (like the Random Forest, RF)
+by performing a weighted mean average of individual trees \citep{Friedman:2002we,RefWorks:1634}. 
+The prediction $F(X)$ of the ensemble of trees for a new sample $X$ is given by
+\begin{equation}
+F(X) = \sum_k w_k F_k(X)
+\end{equation}
+where $F_k(X)$ is the individual tree answer and $w_k$ a weight per tree. Unlike the Random Forest, the 
+tree depth is rather small and the training  of the $k$-th tree is based on the errors of the
+$(k-1)$-th tree.  Like the RF, a randomly chosen fraction of features ($\approx 50\%$) is 
+omitted at each split, and this step is changed at each tree (leading to a Stochastic Gradient Descent 
+algorithm).  GTB learning is essentially non-parallelizable and so cannot use a map-reduce paradigm, 
+unlike RF.It suffers from the opposite problem to RF's overfitting, and is subject
+to high bias, so one generally combines many weak individual learners.
+
+The two methods have their own systematics; they notably differ in the feature importances. Stacking 
+the two is therefore especially effective.  The submitted method used {\sc SKLearn}, and used 50 models 
+of each type combined using its {\sc StackingClassifier}.
+
+% \lstset{language=Python}
+% \lstset{frame=lines}
+% \lstset{caption={The code}}
+% \lstset{label={lst:code_direct}}
+% \lstset{basicstyle=\footnotesize}
+% \begin{lstlisting}
+% estimators = [ 
+% ('gd', make_pipeline(StandardScaler(),
+%     GradientBoostingClassifier(n_estimators=n_estimators, verbose=1))),
+
+% ('rf', make_pipeline(StandardScaler(),
+%     RandomForestClassifier(n_estimators=n_estimators,verbose=1)))  ]
+
+% classifier = StackingClassifier(estimators=estimators,
+%                                final_estimator=LogisticRegression(max_iter=5000))
+% \end{lstlisting}
 

### 2021-01-26 paper/paper.tex
extra sentence

@@
 
+An iterative precedure was used to progressively change the target bin edges to optimize one of the metrics (FOMs or SNRs) between the trainings of the classifiers.
+In general, we recover different choices for each statistic, but the differences are relatively marginal.
+This procedure was done by hand for this challenge, but could be automated and applied to any kind of classifier.
+
 % \lstset{language=Python}

### 2021-01-26 bin/batch_metrics.py
change to secret testing results

@@ -55,7 +55,6 @@ def setup(queue):
 
         for nbin in bins:
             name = f"{classifier}_{nbin}_{config_index}"
-            plot_distributionsnt(name)
             queue.add_job(name, {})
 
 
@@ -63,7 +62,7 @@ def setup(queue):
 def task(name):
     print(name)
     sys.stdout.flush()
-    results_dir = "/global/cscratch1/sd/zuntz/tomo_challenge_results"
+    results_dir = "/global/cscratch1/sd/zuntz/tomo_challenge_secret_testing_results"
     result_file = f'{results_dir}/{name}.npy'
     img_file = f'{results_dir}/plots/{name}.png'
     metric_file = f'{results_dir}/metrics/{name}.yml'

### 2021-01-26 bin/run_metrics.py

@@ -13,7 +13,7 @@ with h5py.File("secret/testing.hdf5", "r") as f:#change later
     truth_z = f["redshift_true"][:]
 print("Loaded truth data")
 
-results_dir = "/global/cscratch1/sd/zuntz/tomo_challenge_results/"
+results_dir = "/global/cscratch1/sd/zuntz/tomo_challenge_secret_testing_results/"
 completed = sorted(glob.glob(results_dir + "*.npy"))
 
 for result_file in completed:

### 2021-01-26 tomo_challenge/classifiers/tensorflow.py
make tensorfor code use v1

@@ -8,7 +8,10 @@ import matplotlib.pyplot as plt
 import warnings
 import h5py
 
-import tensorflow as tf
+#import tensorflow as tf
+import tensorflow.compat.v1 as tf
+tf.disable_v2_behavior()
+
 from tensorflow import keras
 # from dbn.tensorflow import SupervisedDBNClassification # version conflict with tf_v2 (had to put it inside the DBN class)
 

### 2021-01-27 paper/paper.tex
move file names to avoid clash and update tabl

@@
 {\sc myCombinedClassifiers } & 35.3 & 60.5    & 83.5    & 120.4    & 39.4             & 63.9             & 93.1             & 148.9\\
-{\sc NeuralNetwork\_1 } & 0.0 & 0.0    & 0.0    & 0.0    & 0.0             & 0.0             & 0.0             & 0.0\\
-{\sc NeuralNetwork\_2 } & 0.0 & 0.0    & 0.0    & 0.0    & 0.0             & 0.0             & 0.0             & 0.0\\
+{\sc NeuralNetwork1 } & 76.3 & 117.6    & 135.8    & 0.0    & 81.9             & 132.9             & 158.0             & 0.0\\
+{\sc NeuralNetwork2 } & 30.5 & 48.0    & 57.8    & 136.1    & 49.4             & 102.7             & 122.2             & 140.0\\
 {\sc PCACluster } & 29.5 & 68.1    & 75.8    & 0.0    & 50.1             & 72.7             & 90.8             & 0.0\\

### 2021-01-27 paper/paper.tex
add GPZ description

@@
 \subsection{ {\sc GPzBinning} }
+GPzBin is based on GPz \citep{Almosallam2016a,Almosallam2016b}, a machine learning
+regression algorithm originally developed for calculating the photometric
+redshifts of galaxies \citep{Gomes2017,Duncan2018,Hatfield2020} (but now also
+applied to other problems in physics \citep{Peng2019,Hatfield2020}). The
+algorithm is based on a Gaussian Process (GP; essentially an un-parametrised continuous
+function defined everywhere with Gaussian uncertainties); GPz specifically uses 
+a sparse GP framework, a fast and a scalable approximation
+of a full process \citep{Rasmussen2006}.
+
+The mean $\mu(x)$ and variance $\sigma(x)^2$ for predictions as a function of
+the input parameters $x$ (typically the photometry) are both linear combinations
+of a finite number of `basis functions': multi-variate Gaussian kernels with
+associated weights, locations and covariances. The algorithm seeks to find the
+optimal parameters of the basis functions such that the mean and variance
+functions are the most likely functions to have generated the data. The key
+innovations introduced by GPz include a) input-dependent variance estimations
+(heteroscedastic noise), b) a `cost-sensitive learning' framework where the
+algorithm can be tailored for the precise science goal, and c) properly
+accounting for uncertainty contributions from both variance in the data
+($\beta^{-1}_{\star}$) as well as uncertainty from lack of data ($\nu$) in a
+given part of parameter space (by marginalising over the functions supported by
+the GP that could have produced the data).
+
+%During the training stage of the algorithm GPz is inferring the locations and
+%spreads that describe the basis functions; during the validation stage it is
+%inferring the appropriate complexity of the model, essentially how many basis
+%functions to use.
+
+GPzBin is a simple extension to GPz for the problem of tomographic binning. For
+a given number of tomographic bins it selects redshift bin edges $z_{i}$ such
+that there are an equal number of training galaxies in each bin. Testing data
+galaxies then are assigned to the bin corresponding to their GPz predicted
+photometric redshift. The code allows for two selection criteria for removing
+galaxies for which it was not possible to make a good photometric redshift
+prediction: 1) cutting galaxies close to the boundaries of bins based on an
+`edge strictness' parameter $U$ which removes galaxies where $\mu\pm U \times
+\sigma \lessgtr z_i$ and 2) cutting galaxies based on the degree $E$ to which
+GPz is having to extrapolate to make the redshift prediction, removing galaxies
+with $\nu>E\sigma^2$ (see Figure 12 of \cite{Hatfield2020}). Cutting galaxies
+removes some signal at the possible benefit of improving binning purity. As the
+training and test data were sampled from the same distribution in this data
+challenge we might not expect removing galaxies to give huge improvements in
+binning quality, but cuts based on $E$ and $U$ might become more useful in
+future studies where the training and test data have substantially different
+colour-magnitude distributions.
+
+When used in GPzBin for this data challenge the GPz settings
+in Table \ref{table-settings} were used (see
+\cite{Almosallam2016a,Almosallam2016b} for precise definitions and
+interpretations).
+
+%GPz requires very little fine-tuning. The most important parameter is $m$, the number of basis functions. A higher $m$ corresponds to higher model complexity and longer training times.
+
+\begin{table*}[]
+% \tiny
+\begin{center}
+    \begin{tabular}{| l | l | l |}
+    \hline
+    Parameter   &   Value   &   Description\\  
+    \hline 
+    $m$         &   100     &   Number of basis functions; complexity of GP\\
+
+    maxIter     &   500     &   Maximum number of iterations (comparisons with 
+                                the validation data) permitted\\
+
+    maxAttempts &   50      &   Maximum number of iterations to attempt if there 
+                                is no progress on the validation set\\
+
+    method      &   GPVC    &   Type of bespoke covariances used on each basis 
+                                function (see \cite{Almosallam2016b} for 
+                                the different options)\\
+
+    normalize   &   True    &   Pre-process the input by subtracting the means
+                                and dividing by the standard deviations\\
+
+    joint       &   True    &   Jointly learn a prior linear mean-function 
+                                (learn the function means and variances 
+                                jointly)\\
+
+    heteroscedastic  & True & Model noise as well as point estimates \\
+    \hline
+    \end{tabular}
+\caption{Parameter settings in the GPz entry.}
+\end{center}
+\label{table-settings}
+\end{table*}
 
@@
 \begin{table*}[]
-\begin{tabular}{l|llll|llll|}
+\begin{tabular}{|l|llll|llll|}
                 & \multicolumn{4}{c|}{\textbf{riz}}      & \multicolumn{4}{c|}{\textbf{griz}}                                \\ \hline

### 2021-01-28 paper/authors.tex
Adds abstract
text for intro and motivationn
Update authors.tex
add text and authors from clecio

@@
 \noaffiliation
+
+% paper leads
 
@@
 %main partticipants and text contributors  (alphabetical)
+\author{Bela Abolfathi}
+\noaffiliation
+
+\author{David Alonso}
+\noaffiliation
+
+\author{Abby Bault}
+\noaffiliation
+
+\author{Cl\'{e}cio R. Bom}
+\affiliation{Centro Brasileiro de Pesquisas F\'isicas, Rua Dr. Xavier Sigaud 150, 22290-180 Rio de Janeiro, RJ, Brazil}
+\affiliation{Centro Federal de Educa\c{c}\~{a}o Tecnol\'{o}gica Celso Suckow da Fonseca, Rodovia M\'{a}rcio Covas, lote J2, quadra J - Itagua\'{i} (Brazil)}
+
+\author{Massimo Brescia}
+\noaffiliation
+%\affiliation{INAF Osservatorio Astronomico di Capodimonte, via Moiariello 16, I-80131, Napoli Italy}
+
+\author{Jean-Eric Campagne}
+\noaffiliation
+
+\author[0000-0002-3787-4196]{Stefano Cavuoti}
+\noaffiliation
+%\affiliation{INAF Osservatorio Astronomico di Capodimonte, via Moiariello 16, I-80131, Napoli Italy}
+
+\author{Eduardo S. Cypriano} $^3$
+\affiliation{Universidade de S\~ao Paulo, IAG, Rua do Mat\~ao 1225, S\~ao Paulo, SP, Brazil}
+
+\author{Bernardo M. O. Fraga}
+\affiliation{Centro Brasileiro de Pesquisas F\'isicas, Rua Dr. Xavier Sigaud 150, 22290-180 Rio de Janeiro, RJ, Brazil}
+
+\author{Elizabeth J. Gonzalez} $^{4,5}$.
+\affiliation{Instituto de Astronom\'{\i}a Te\'orica y Experimental (IATE-CONICET), Laprida 854, X5000BGR, C\'ordoba, Argentina.}
+
+\affiliation{Observatorio Astron\'omico de C\'ordoba, Universidad Nacional de C\'ordoba, Laprida 854, X5000BGR, C\'ordoba, Argentina.}
+
+\author{Dylan Green}
+\noaffiliation
+
+\author{Peter Hatfield}
+\noaffiliation
+
+\author{David Kirkby}
+\noaffiliation
 
 \author[0000-0001-7956-0542]{Fran\c{c}ois Lanusse}
-\affiliation{AIM, CEA, CNRS, Universit\'e Paris-Saclay, Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, F-91191 Gif-sur-Yvette, France}
+\noaffiliation
+%\affiliation{AIM, CEA, CNRS, Universit\'e Paris-Saclay, Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, F-91191 Gif-sur-Yvette, France}
 
 \author{Alex Malz}
-\affiliation{}
+\noaffiliation
+
+\author{Andrina Nicola}
+\noaffiliation
+
+\author{Erfan Nourbakhsh}
+\noaffiliation
+
+\author{Andy Park}
+\noaffiliation
 
 \author{Anze Slosar}
-\affiliation{BNL}
+\noaffiliation
+
+\author{Gabriel Teixeira}
+\affiliation{Centro Brasileiro de Pesquisas F\'isicas, Rua Dr. Xavier Sigaud 150, 22290-180 Rio de Janeiro, RJ, Brazil}
+
+\author{Angus Wright}
+\noaffiliation
 

### 2021-01-28 paper/paper.tex
add text and authors from clecio

@@
 
-\subsection{ {\sc Autokeras LSTM} }
+\subsection{ {\sc LSTM} } 
+\label{ClecioLSTM} 
+Methods \ref{ClecioLSTM} - \ref{ClecioEnsemble} were
+submitted by a team of authors CRB, BMOF, GT, ESC, and EJG.  All were built using TensorFlow, and
+chose bin edges such that an equal number of galaxies is assigned to each bin, and trained the
+network using the galaxies' magnitudes and colours.
+
+The LSTM method uses a Bidirectional Long Short Term Memory network to assign galaxies to redshift
+bins.
+
+Recurrent Neural Networks (RNNs)\citep{schuster, medsker, pascanu} are a type of NN capable of
+analyzing sequential data, where data points have a strong correlation with their neighbours. Long
+Short Term Memory (LSTM) units are a particular type of RNN capable of retaining relevant
+information from previous points while at the same time removing unnecessary data. This is achieved
+through a series of gates with learnable weights connected to different activation functions to
+balance the amount of data retained or removed.
+
+In some cases, relevant information can come both from data points coming before or after each point
+being analyzed. In such cases, two LSTMs can be combined, each going in a different direction, to
+create effectively a bidirectional LSTM cell \citep{schuster}.
+
+In this implementation, we chose bin edges such that an equal number of galaxies is assigned to each
+bin, and trained the network using the galaxies' magnitudes and colours. The net was implemented
+using Tensorflow and Keras. The Deep model is composed of a series of convolutional blocks (1d
+convolutional layer with a tanh activation followed by a MaxPooling layer) followed by a
+bidirectional LSTM layer and a series of fully connected layers.
+
+\subsection{ {\sc Autokeras LSTM}}
+Autokeras \citep{autokeras} is an AutoML system based on Keras. Given a general neural network
+architecture and a dataset, it searches for the best possible detailed configuration for the problem
+at hand.
+
+We started from the basic architecture and bin assignment scheme mentioned in Section
+\ref{ClecioLSTM} and let Autokeras search for the configuration that results in the lowest
+validation loss. We kept the order of the layer blocks fixed, i.e., convolutional blocks,
+bidirectional LSTM, and Dense blocks. We let the other hyperparameters free, such as the number of
+neurons,  Dropout rates, maxpooling layers, the number of convolutional filters, and strides.
+
+\subsection{ {\sc LGBM} }
+This method uses a Light Gradient Boosted Machine (LightGBM) \citep{lgbm} to assign galaxies to
+redshift bins.
+
+LightGBM uses a histogram-based algorithm that assigns continuous feature values to discrete bins
+instead of other pre-sort-based algorithms for decision tree learning. LightGBM also grows the trees
+leaf-wise, which tends to achieve lower losses than level-wise tree growth. This method tends to
+overfit for small data so that a \textit{max\_depth} parameter can be specified to limit tree depth.
+
+In this implementation, we chose bin edges such that an equal number of galaxies is assigned to each
+bin, and trained the network using the galaxies' magnitudes and colours.
+
+\subsection{ {\sc TCN}}
+This method uses a Temporal-Convolutional (TCN) \citep{baitcn} network to assign galaxies to redshift
+bins.
+
+Recurrent Neural Networks (such as LSTMs) are considered to be the state-of-the-art model for
+sequence modelling. However, research indicates that certain convolutional architectures can achieve
+human-level performance in these tasks \citep{dauphin}.
+
+TCNs are Fully Convolutional Networks with causal convolutions, in which the output at time
+\textit{t} is convolved only with elements of time \textit{t} and lower in the previous layer. This
+new approach prevents leakage of information from the future to the past. This simple model has one
+disadvantage, in that it needs a very deep architecture or large filters to achieve a long history
+size.
+
+By using dilated convolutions, where a fixed step is introduced between adjacent filter taps, the
+receptive field of the TCN can be increased without increasing the number of convolutional filters.
+Furthermore, residual blocks \citep{he} are also used to stabilize deeper and larger networks. These
+modifications cause TCNs to have a longer memory than traditional RNNs with the same capacity,  and
+also require low memory for training.
+
+We used the Keras implementation of TCN \citep{kerastcn} and chose bin edges such that an equal
+number of galaxies is assigned to each bin, and trained the network using the galaxies' magnitudes
+and colours.
 
 \subsection{ {\sc CNN} } 
+This method uses an optimized Convolutional Neural Network (CNN) to assign
+galaxies to redshift bins.
+
+CNNs \citep{lecun2015deep} are layers inspired in pattern recognition types developed in mammals' brains.
+They consist of kernels that are convolved with the data as it flows through the net. CNN-based
+models have emerged as state-of-the-art in several computer vision tasks, such as image
+classification and pose estimations, as well as having applicability in a range of different fields.
+
+We combine the Convolutional layers are with dimensionality reduction (pooling layers) and different
+activation functions. Here, we used Autokeras to optimize a general convolutional architecture,
+using the galaxies' magnitudes and colours to assign them to redshift bins, which were chosen such
+that an equal number of galaxies were assigned to each.
+
+\subsection{ {\sc Ensemble} } 
+\label{ClecioEnsemble} 
+This method combines different neural network
+architectures to assign galaxies to redshift bins.
+
+Deep Ensemble models combine different neural network architectures to obtain better performance
+than any of the nets alone. In this method, we used the Bidirectional LSTM optimized by Autokeras, a
+ResNet \citep{resnet} and a Fully Convolutional Network (FCN) \citep{fcn}. The predictions of these
+models were averaged to get the final predicted bin.
 
@@
 \subsection{ {\sc GPzBinning} }
+This method was submitted by author PH.
+
 GPzBin is based on GPz \citep{Almosallam2016a,Almosallam2016b}, a machine learning
@@
 
-\subsection{ {\sc LGBM} }
-
-\subsection{ {\sc LSTM} }
-
 \subsection{ {\sc NeuralNetwork 1} }
@@
 \subsection{ {\sc SimpleSOM} }
-
-\subsection{ {\sc TCN} }
 
@@
 \subsection{ {\sc Stacked Generalization} }
+This method was submitted by author JEC.
 

### 2021-01-29 paper/paper.tex
add MLPQNA

@@
 \subsection{ {\sc mlpqna} }
+This entry was submitted by authors SC and MB.
+
+The Multi Layer Perceptron trained by Quasi Newton Algorithm 
+({\sc MLPQNA}; \citealp{Brescia12}), is a feed-forward neural network 
+for multi-class classification and single/multi regression use cases.
+
+The architecture is a classical Multi Layer Perceptron (MLP; \citealp{Rosenblatt1961})
+with one or more hidden layers, on which the supervised learning paradigm is
+based on the Quasi Newton Algorithm (QNA), implemented through the L-BFGS rule
+\citep{Nocedal80}. L-BFGS is a solver that approximates the Hessian matrix,
+representing the second-order partial derivative of a function. Further, it
+approximates the inverse of the Hessian matrix to perform parameter updates.
+MLPQNA uses the least square loss function with the Tikhonov regularization
+\citep{Tikhonov77} for regression. It uses the cross-entropy \citep{deBoer05}
+and softmax \citep{Sutton98} as output error evaluation criteria for
+classification.
+
+Besides several scientific contexts and projects in which MLPQNA has succesfully
+been used, it was recently the kernel embedded into the method {\sc
+METAPHOR} \citep{cavuoti20} for photo-z estimation, which participated in the
+LSST Photometric Redshift PDF Challenge \citep{schmidt20}.
+
+The {\sc MLPQNA} python implementation used here
+uses a customized Scipy version of the L-BFGS built-in rule.
+It used two hidden layers with $2N-1$ and $N+1$ nodes respectively,
+where $N$ is the number of input features (depending on the experiment).
+The decay was set to 0.1.
 

### 2021-01-29 paper/paper.tex
rename and update FFNN table

@@
 {\sc SimpleSOM } & 39.5 & 59.9    & 77.9    & 98.5    & 42.2             & 73.2             & 108.0             & 134.8\\
+{\sc TensorFlow\_FFNN } & 39.0 & 64.3    & 90.3    & 114.7    & 43.0             & 72.0             & 112.3             & 137.2\\
 {\sc TCN } & 40.0 & 65.3    & 90.9    & 110.5    & 44.1             & 72.8             & 108.0             & 134.0\\

### 2021-01-29 paper/paper.tex
Add PCACluster

@@
 \subsection{ {\sc PCACluster} }
+This entry was submitted by author DG.
+
+Principal Component Analysis (PCA) reduces the dimensionality of a dataset by
+calculating eigenvectors that align with the principal axes of variation in the
+data \citep{doi:10.1098}. {\sc PCACluster} uses PCA to reduce the flux \& color
+data set to three dimensions.
+
+To generate the PCA dimensions used by this method the PCA algorithm takes in
+the $r$-band flux and $ri$ and $iz$ colors and outputs three principal
+components.  If the data set provides the $g$-band, the algorithm also uses the
+$gr$ color when determining eigenvectors. Tomographic bins are then determined
+in this three dimensional space using a unique clustering algorithm where each
+observation is assigned to a bin by determining which cluster centroid is
+closest to that observation.
+
+In order to determine optimal centroid positions, clustering is framed as a
+gradient descent problem that seeks to maximize the requested metric: either the
+SNR or the FoM. This approach of identifying clusters using gradient descent is
+inspired by the similar approach to K-means clustering in
+\cite{NIPS1994_a1140a3d}.
+
+This entry uses {\sc jax\_cosmo} to take the derivative of the requested metric
+and classification function combination with respect to the centroid locations
+to calculate gradients and weight updates \citep{jax-cosmo}.  The speed of
+training PCACluster is directly proportional to the number of requested
+centroids (the number of requested tomographic bins) and the amount of training
+data used.
 

### 2021-01-29 bin/run_one_batch.py
update sub script

@@ -43,7 +43,7 @@ if buzzard:
 else:
     training_file = 'data-train/training-cut.hdf5'
     validation_file = 'secret/testing.hdf5'
-    output_file = f'{scratch}/tomo_challenge_results/{task_name}.npy'
+    output_file = f'{scratch}/tomo_challenge_secret_testing_results/{task_name}.npy'
 
 training_data = tc.load_data(
     training_file,

### 2021-01-29 tomo_challenge/classifiers/neural_network.py
disable NN model saving

@@ -104,10 +104,10 @@ class NeuralNetwork(Tomographer):
         labels = training_z
 
         # If model is already trained, we just load the weights
-        if os.path.exists(self.export_name):
-            with open(self.export_name, 'rb') as file:
-                self.model = serialization.from_bytes(self.model, pickle.load(file))
-            return
+        # if os.path.exists(self.export_name):
+        #     with open(self.export_name, 'rb') as file:
+        #         self.model = serialization.from_bytes(self.model, pickle.load(file))
+        #     return
 
         lr = 0.001
         optimizer = optim.Adam(learning_rate=lr).create(self.model)
@@ -149,8 +149,8 @@ class NeuralNetwork(Tomographer):
                 print('iter: %d; Loss : %f'%(i,loss))
 
         # Export model to disk
-        with open(self.export_name, 'wb') as file:
-            pickle.dump(serialization.to_bytes(optimizer.target), file)
+        # with open(self.export_name, 'wb') as file:
+        #     pickle.dump(serialization.to_bytes(optimizer.target), file)
 
         self.model = optimizer.target
 

### 2021-01-29 paper/paper.tex
narrow table
more discussion

@@
     method      &   GPVC    &   Type of bespoke covariances used on each basis 
-                                function (see \cite{Almosallam2016b} for 
-                                the different options)\\
+                                function (see \citealp{Almosallam2016b})\\
 
@@
 CosmoDC2 part of the challenge.}
+\label{tab:cosmodc2}
 \end{table*}
@@
 \begin{itemize}
-    \item Big table of results
     \item fisher matrix ellipses overplotted for top N methods
@@
 
-\subsection{Individual winners}
+-
 
-RUN WINNERS WITH UGRIZY TO GET AN IDEA OF THE LOSS FROM DROPPING U AND Y.
-
-\subsection{Large bin numbers}
-
+\subsection{General outcomes}
 It is clear from values above that fears about restricted photometry catastrophically limiting
@@
 even the $riz$ bands alone are sufficient to split objects into nine reasonably separated
-tomographic bins.
-Once bins become this narrow the primary analysis concern will become the calibration of overall
+tomographic bins.  Once bins become this narrow the primary analysis concern will become the calibration of overall
 bin $n(z)$ values, rather than initial tomography, so we consider this to be reassuring.
@@
 the limiting factor in overall precision.
+
+Methods using fixed bins with equal numbers by nominal redshift tend to plateau at scores of
+120-140 in Table \ref{tab:cosmodc2}, suggesting that a range of machine learning models can
+effectively classify by bin, at least in this case of representative, extensive training
+information.  Notably, this is the case even for the UTOPIA entry, which uses the nearest 
+training set neighbour to assign bins and thus was explicitly designed to work only
+for this idealised scenario; it should give an upper limit when using the same fixed bins.
+That other methods with the same nominal bin choice acheive scores close to it once again
+confirms that many entries are close to the best possible score for this approach.
+
+Several methods that instead try to optimize nominal bin edges break this
+barrier and acheive very high scores.  This included ZotNet, JaxResNet, and JaxCNN.  The
+Stacked Generalization method similarly included manually tuned bin edges to maximize
+score for this scenario. The typical 15--20\% improvement in FoM scores when optimizing
+nominal bin edges makes clear that this easy win should is compelling.
+
+Notably, though, the method that received the highest overall score in Table \ref{tab:cosmodc2}, 
+{\sc funbins} did so with a simpler approach - it used a random forest to assign galaxies to bins,
+but chose nominal bin edges using an equal width bins in \emph{comoving distance} $\chi$, converting
+this to redshift using a fiducial cosmology. The success of this simple solution suggests this
+as an easy heuristic for well-optimized nominal edges.
 

### 2021-02-01 paper/paper.tex
add funbins

@@
 
-\subsection{ {\sc RandomForest} }
+\subsection{ {\sc RandomForest} } \label{sec:randomforest}
 This method was submitted by the challenge organizers and used a random
@@
 \subsection{ {\sc funbins} }
+This method uses the random forest algorithm described in section \ref{randomforest} to assign galaxies to tomographic bins.
+There are 3 options for determining the bins used to assign labels to the training data - \textit{log}, \textit{random}, and \textit{chi}.
+
+The first option, \textit{log}, calculates bin edges such that the number of galaxies in each bin grows logarithmically instead of being constant as in section 4.1.
+This option uses increasingly large bins for more distant galaxies and tests the relative importance of nearby galaxies.
+
+The second option, \textit{random}, draws the intermediate bin edges from a uniform distribution while fixing the outer edges to the limits of the data.
+This option is not theoretically motivated and just to study the sensitivity to the binning.
+
+The last option, \textit{chi}, calculates redshift bin edges whose corresponding comoving distances are equally spaced, using Planck15 \citep{Planck15} cosmology for the distance-redshift relationship.
+This should lead to roughly equal shot noise for the angular power in each bin.
+This method was used in the scores shown below.
+
+Once the bin edges are determined by either of the methods above, the random forest is trained as described in section \ref{randomforest}.
 
@@
 
+\section{Acknowledgements}
+ESC acknowledges support from the FAPESP (\#2019/19687-2) and CNPq  (\#308539/2018-4).
+
 \bibliography{paper}

### 2021-02-02 paper/paper.tex
add more tables

@@
 \textbf{Method} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} \\ \hline
-{\sc Autokeras\_LSTM } & 31.1 & 74.9    & 103.4    & nan    & 44.1             & 67.5             & 122.2             & 98.4\\
+{\sc Autokeras\_LSTM } & 31.1 & 74.9    & 103.4    & --    & 44.1             & 67.5             & 122.2             & 98.4\\
 {\sc CNN } & 27.1 & 50.8    & 76.7    & 103.3    & 30.5             & 57.7             & 95.4             & 122.4\\
 {\sc ComplexSOM } & 38.0 & 52.1    & 94.4    & 101.6    & 34.9             & 45.0             & 91.6             & 100.3\\
-{\sc ENSEMBLE1 } & 65.6 & 65.6    & 94.8    & 0.0    & 0.0             & 0.0             & 0.0             & 0.0\\
+{\sc ENSEMBLE1 } & 65.6 & 65.6    & 94.8    & --    & --             & --             & --             & --\\
 {\sc funbins } & 36.8 & 82.5    & 122.1    & 141.8    & 42.1             & 100.4             & 142.5             & 167.2\\
@@
 {\sc IBandOnly } & 38.0 & 50.0    & 54.4    & 57.3    & 38.0             & 50.0             & 54.4             & 57.3\\
-{\sc JaxCNN } & 59.9 & 101.9    & 105.2    & nan    & 79.7             & 125.2             & 150.0             & nan\\
-{\sc JaxResNet } & 73.7 & 111.1    & 131.6    & nan    & 82.2             & 126.0             & nan             & 161.5\\
+{\sc JaxCNN } & 59.9 & 101.9    & 105.2    & --    & 79.7             & 125.2             & 150.0             & --\\
+{\sc JaxResNet } & 73.7 & 111.1    & 131.6    & --    & 82.2             & 126.0             & --             & 161.5\\
 {\sc LGBM } & 27.0 & 50.8    & 78.8    & 108.2    & 30.3             & 57.4             & 92.9             & 125.2\\
@@
 {\sc myCombinedClassifiers } & 35.3 & 60.5    & 83.5    & 120.4    & 39.4             & 63.9             & 93.1             & 148.9\\
-{\sc NeuralNetwork1 } & 76.3 & 117.6    & 135.8    & 0.0    & 81.9             & 132.9             & 158.0             & 0.0\\
+{\sc NeuralNetwork1 } & 76.3 & 117.6    & 135.8    & --    & 81.9             & 132.9             & 158.0             & --\\
 {\sc NeuralNetwork2 } & 30.5 & 48.0    & 57.8    & 136.1    & 49.4             & 102.7             & 122.2             & 140.0\\
-{\sc PCACluster } & 29.5 & 68.1    & 75.8    & 0.0    & 50.1             & 72.7             & 90.8             & 0.0\\
+{\sc PCACluster } & 29.5 & 68.1    & 75.8    & --    & 50.1             & 72.7             & 90.8             & --\\
 {\sc Random } & 1.2 & 1.2    & 1.2    & 1.2    & 1.2             & 1.2             & 1.2             & 1.2\\
@@
 
+\section{Complete results}
+
+\begin{table*}[]
+\begin{tabular}{|l|llll|llll|}
+                & \multicolumn{4}{c|}{\textbf{riz}}      & \multicolumn{4}{c|}{\textbf{griz}}                                \\ \hline
+\textbf{Method} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} \\ \hline
+{\sc Autokeras\_LSTM } & 22.3 & 32.7    & 61.8    & --    & 26.3             & 43.4             & --             & 72.2\\
+{\sc CNN } & 18.0 & 36.9    & 59.8    & 70.3    & 17.7             & 40.7             & 62.6             & 74.5\\
+{\sc ComplexSOM } & 30.4 & 47.8    & 57.2    & 74.9    & 25.0             & 48.5             & 63.4             & 75.0\\
+{\sc ENSEMBLE1 } & 53.8 & 53.8    & 74.6    & 74.6    & 52.2             & 52.2             & 82.9             & 82.9\\
+{\sc funbins } & 23.2 & 50.5    & 65.0    & 72.2    & 26.0             & 57.4             & 81.3             & 91.7\\
+{\sc GPzBinning } & 16.8 & 46.4    & 61.4    & 70.7    & 17.6             & 39.7             & 67.0             & 82.5\\
+{\sc IBandOnly } & 11.4 & 15.9    & 20.8    & 22.0    & 11.4             & 15.9             & 20.8             & 22.0\\
+{\sc JaxCNN } & 38.6 & 59.0    & 69.7    & --    & 43.7             & 71.5             & 83.1             & 95.4\\
+{\sc JaxResNet } & 39.6 & 61.7    & --    & --    & 45.9             & 63.3             & --             & --\\
+{\sc LGBM } & 18.2 & 37.1    & 60.9    & 70.7    & 17.8             & 39.5             & 61.6             & 75.4\\
+{\sc LSTM } & 18.0 & 35.3    & 61.1    & 69.9    & 17.9             & 38.8             & 60.9             & 74.7\\
+{\sc mlpqna } & 24.1 & 54.2    & 65.7    & 73.2    & 27.2             & 51.5             & 70.4             & 83.4\\
+{\sc myCombinedClassifiers } & 18.4 & 33.0    & 56.1    & 85.3    & 22.1             & 38.8             & 68.3             & 97.2\\
+{\sc NeuralNetwork1 } & 42.8 & 66.8    & 82.0    & 93.7    & 46.6             & 74.1             & 90.9             & 105.4\\
+{\sc NeuralNetwork2 } & 21.9 & 52.1    & 60.8    & 76.5    & 22.2             & 47.0             & 59.9             & 73.9\\
+{\sc PCACluster } & 22.3 & 49.2    & 56.4    & --    & 21.0             & 44.8             & 57.6             & --\\
+{\sc Random } & 0.8 & 0.8    & 0.8    & 0.8    & 0.8             & 0.8             & 0.8             & 0.8\\
+{\sc RandomForest } & 23.9 & 53.4    & 65.0    & 72.5    & 27.3             & 52.1             & 70.4             & 83.5\\
+{\sc SimpleSOM } & 26.1 & 46.3    & 64.3    & 70.7    & 28.3             & 51.9             & 70.8             & 84.7\\
+{\sc TensorFlow\_FFNN } & -- & --    & --    & --    & --             & --             & --             & --\\
+{\sc TCN } & 26.6 & 53.7    & 65.7    & 74.6    & 27.0             & 49.2             & 69.4             & 82.4\\
+{\sc UTOPIA } & 24.9 & 53.8    & 65.3    & 73.6    & 27.9             & 52.9             & 70.1             & 83.1\\
+{\sc ZotBin } & 30.8 & 57.4    & 68.2    & 76.5    & 44.6             & 66.2             & --             & 89.4\\
+{\sc ZotNet } & 40.3 & 62.4    & 75.7    & 86.0    & 48.0             & 71.0             & 86.6             & 98.4\\
+\end{tabular}
+\caption{INCOMPLETE, AND AUTO-GENERATED SO NO POINT EDITING! Values of the 3x2pt DETF figure-of-merit achieved by entrant methods on the 
+Buzzard part of the challenge.}
+\label{tab:buzzard_3x2}
+\end{table*}
+
+\begin{table*}[]
+\begin{tabular}{|l|llll|llll|}
+                & \multicolumn{4}{c|}{\textbf{riz}}      & \multicolumn{4}{c|}{\textbf{griz}}                                \\ \hline
+\textbf{Method} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} \\ \hline
+{\sc Autokeras\_LSTM } & 13.8 & 30.2    & 31.1    & --    & 26.4             & 7.7             & 43.0             & 44.6\\
+{\sc CNN } & 7.0 & 29.8    & 37.5    & 38.1    & 6.2             & 35.8             & 45.0             & 48.7\\
+{\sc ComplexSOM } & 17.3 & 19.7    & 25.0    & 25.4    & 6.5             & 12.2             & 19.3             & 21.5\\
+{\sc ENSEMBLE1 } & 36.0 & 36.0    & 38.6    & --    & --             & --             & --             & --\\
+{\sc funbins } & 23.3 & 31.6    & 36.5    & 39.8    & 31.2             & 43.0             & 50.4             & 54.0\\
+{\sc GPzBinning } & 6.9 & 23.9    & 33.1    & 35.1    & 6.5             & 36.0             & 46.0             & 47.8\\
+{\sc IBandOnly } & 0.8 & 2.6    & 3.2    & 3.6    & 0.8             & 2.6             & 3.2             & 3.6\\
+{\sc JaxCNN } & 9.3 & 27.9    & 29.0    & --    & 3.3             & 38.0             & 42.7             & --\\
+{\sc JaxResNet } & 2.1 & 30.8    & 37.9    & --    & 2.2             & 31.1             & --             & 44.7\\
+{\sc LGBM } & 7.0 & 30.1    & 37.5    & 39.6    & 6.3             & 36.3             & 45.4             & 48.6\\
+{\sc LSTM } & 6.9 & 29.3    & 36.8    & 38.2    & 6.4             & 35.5             & 44.6             & 48.9\\
+{\sc mlpqna } & 23.1 & 35.0    & 37.3    & 38.0    & 30.0             & 43.9             & 48.9             & 51.4\\
+{\sc myCombinedClassifiers } & 1.2 & 8.4    & 28.8    & 39.1    & 1.7             & 10.9             & 37.2             & 49.3\\
+{\sc NeuralNetwork1 } & 1.6 & 34.0    & 36.7    & --    & 2.1             & 37.7             & 46.3             & --\\
+{\sc NeuralNetwork2 } & 19.2 & 21.8    & 22.4    & 35.8    & 8.5             & 18.6             & 22.5             & 36.7\\
+{\sc PCACluster } & 7.8 & 11.3    & 14.2    & --    & 12.6             & 12.0             & 21.6             & --\\
+{\sc Random } & 0.0 & 0.0    & 0.0    & 0.0    & 0.0             & 0.0             & 0.0             & 0.0\\
+{\sc RandomForest } & 23.5 & 34.9    & 38.4    & 39.6    & 30.0             & 44.5             & 50.1             & 53.2\\
+{\sc SimpleSOM } & 20.5 & 30.7    & 35.1    & 36.7    & 28.8             & 45.4             & 50.7             & 53.8\\
+{\sc TensorFlow\_FFNN } & 22.6 & 34.4    & 36.7    & 37.7    & 29.7             & 43.4             & 48.3             & 50.9\\
+{\sc TCN } & 23.6 & 36.0    & 38.0    & 38.7    & 27.6             & 41.2             & 46.3             & 49.4\\
+{\sc UTOPIA } & 22.0 & 31.1    & 35.1    & 36.3    & 29.8             & 43.2             & 49.6             & 52.3\\
+{\sc ZotBin } & 4.8 & 24.8    & 30.9    & 29.1    & 4.0             & 27.3             & 35.9             & 38.1\\
+{\sc ZotNet } & 2.2 & 29.9    & 33.1    & 34.6    & 2.3             & 33.2             & 42.3             & 43.0\\
+\end{tabular}
+\caption{INCOMPLETE, AND AUTO-GENERATED SO NO POINT EDITING! Values of the weak lensing $\Omega_c - \sigma_8$
+figure-of-merit achieved by entrant methods on the CosmoDC2 part of the challenge.}
+\label{tab:dc2_wl}
+\end{table*}
+
+\begin{table*}[]
+\begin{tabular}{|l|llll|llll|}
+                & \multicolumn{4}{c|}{\textbf{riz}}      & \multicolumn{4}{c|}{\textbf{griz}}                                \\ \hline
+\textbf{Method} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} \\ \hline
+{\sc Autokeras\_LSTM } & 2.9 & 10.5    & 8.2    & --    & 5.7             & 5.0             & --             & 11.5\\
+{\sc CNN } & 1.8 & 9.5    & 12.3    & 13.2    & 1.8             & 9.4             & 14.7             & 16.2\\
+{\sc ComplexSOM } & 1.9 & 6.2    & 7.5    & 8.4    & 1.0             & 7.9             & 13.2             & 15.9\\
+{\sc ENSEMBLE1 } & 11.4 & 11.4    & 13.7    & 13.7    & 13.9             & 13.9             & 17.9             & 17.9\\
+{\sc funbins } & 4.4 & 8.8    & 10.8    & 11.8    & 7.2             & 13.1             & 16.3             & 17.4\\
+{\sc GPzBinning } & 1.7 & 9.3    & 11.3    & 12.5    & 2.1             & 10.6             & 15.0             & 16.7\\
+{\sc IBandOnly } & 0.2 & 0.3    & 0.4    & 0.4    & 0.2             & 0.3             & 0.4             & 0.4\\
+{\sc JaxCNN } & 4.0 & 10.0    & 11.6    & --    & 4.4             & 13.1             & 14.9             & 15.5\\
+{\sc JaxResNet } & 4.2 & 10.2    & --    & --    & 4.8             & 11.3             & --             & --\\
+{\sc LGBM } & 1.8 & 9.5    & 12.5    & 13.4    & 1.8             & 9.6             & 14.5             & 16.5\\
+{\sc LSTM } & 1.8 & 9.6    & 12.6    & 13.3    & 1.9             & 9.6             & 14.5             & 16.5\\
+{\sc mlpqna } & 4.8 & 11.2    & 13.0    & 13.7    & 6.9             & 13.8             & 17.3             & 18.3\\
+{\sc myCombinedClassifiers } & 1.5 & 6.6    & 11.0    & 13.1    & 2.3             & 9.5             & 14.9             & 17.6\\
+{\sc NeuralNetwork1 } & 3.9 & 10.2    & 12.0    & 13.3    & 5.0             & 14.5             & 16.4             & 17.9\\
+{\sc NeuralNetwork2 } & 2.3 & 9.8    & 11.1    & 12.6    & 2.8             & 7.0             & 9.8             & 14.1\\
+{\sc PCACluster } & 3.8 & 6.0    & 4.2    & --    & 1.2             & 2.5             & 5.0             & --\\
+{\sc Random } & 0.0 & 0.0    & 0.0    & 0.0    & 0.0             & 0.0             & 0.0             & 0.0\\
+{\sc RandomForest } & 4.6 & 10.8    & 12.6    & 13.1    & 6.8             & 13.8             & 17.2             & 18.2\\
+{\sc SimpleSOM } & 4.7 & 10.4    & 12.0    & 12.7    & 7.3             & 15.0             & 17.6             & 18.5\\
+{\sc TensorFlow\_FFNN } & -- & --    & --    & --    & --             & --             & --             & --\\
+{\sc TCN } & 5.0 & 10.9    & 12.9    & 13.4    & 6.2             & 12.2             & 15.8             & 16.4\\
+{\sc UTOPIA } & 4.2 & 9.9    & 11.6    & 12.1    & 6.6             & 13.0             & 16.2             & 17.2\\
+{\sc ZotBin } & 5.7 & 8.8    & 10.8    & 10.5    & 3.6             & 10.0             & --             & 12.4\\
+{\sc ZotNet } & 3.8 & 10.1    & 11.5    & 11.7    & 4.3             & 11.3             & 13.1             & 13.4\\
+\end{tabular}
+\caption{INCOMPLETE, AND AUTO-GENERATED SO NO POINT EDITING! Values of the weak lensing $\Omega_c - \sigma_8$
+figure-of-merit achieved by entrant methods on the Buzzard part of the challenge.}
+\label{tab:buzzard_wl}
+\end{table*}
+
 \section{Mistakes}
 The challenge organizers made a number of mistakes when building and running the challenge.
-These do not invalidate the process in any way, but we describe them here in the hope of
+These do not invalidate our process or results, but we describe them here in the hope of
 being useful for future challenge designers.

### 2021-02-02 paper/paper.tex
another method

@@
 \begin{abstract}
-In this paper we present the results of the DESC 3x2pt tomography challenge, which aims to compare strategies for optimizing the photometric binning required for the main LSST 3x2pt analysis.
-This task is made particularly delicate in the context of a metacalibrated lensing survey, as only the photometry from the bands  included in the metacalibration process (riz and potentially g) can be used to define this tomographic binning.
-Using a set of realistic simulations including true galaxy redshifts and simulated photometry, the goal of the challenge is to propose a bin assignment strategy that maximizes the overall Figure of Merit of a 3x2pt analysis.
-This setting is idealized as it ignores spectroscopic completeness  issues for the training sample, but complex enough to address the main question of binning optimization.
-Based on the challenge dataset, we review and compare the different methods that have been proposed by participants.
-We find that even from this limited photometry information, various methods are able to yield binning schemes that achieve high FoM scores.
+In this paper we present the results of the DESC 3x2pt tomography challenge,
+which aims to compare strategies for optimizing the photometric binning required
+for the main LSST 3x2pt analysis. This task is made particularly delicate in the
+context of a metacalibrated lensing survey, as only the photometry from the
+bands  included in the metacalibration process (riz and potentially g) can be
+used to define this tomographic binning. Using a set of realistic simulations
+including true galaxy redshifts and simulated photometry, the goal of the
+challenge is to propose a bin assignment strategy that maximizes the overall
+Figure of Merit of a 3x2pt analysis. This setting is idealized as it ignores
+spectroscopic completeness  issues for the training sample, but complex enough
+to address the main question of binning optimization. Based on the challenge
+dataset, we review and compare the different methods that have been proposed by
+participants. We find that even from this limited photometry information,
+various methods are able to yield binning schemes that achieve high FoM scores.
 \end{abstract}
@@
 variants.  In DES the corrected biases were found up to around 3\%, far larger than
-the requirements for the current generation of surveys \cite{des-y1-cat}.
+the requirements for the current generation of surveys \citep{des-y1-cat}.
 
@@
 
-Random forests \cite{breiman2001} train a collection of individual \emph{decision trees},
+Random forests \citep{breiman2001} train a collection of individual \emph{decision trees},
 each of which consists of a set of bifurcating comparisons in which different features in
@@
 \subsection{ {\sc ComplexSOM} }
+  The {\tt ComplexSOM} method implements an additional optimization layer on the methodology used by {\tt SimpleSOM}.
+
+  Let ${\sf C}^{\rm group}_\ell$ be the matrix containing all auto- and cross-power spectra between members of a large set of galaxy samples (called ``groups'' here).
+We can compress this large set into a smaller set of samples (called ``bins'' here), by combining different groups.
+Let ${\sf A}$ be the assignment matrix determining these bins, such that $A_{\alpha,i}=1$ if group $i$ is assigned to bin $\alpha$, and 0 otherwise.
+In that case, the redshift distribution of bin $\alpha$ is given by $N_\alpha(z)=\sum_i A_{\alpha,i}n_i(z)$, where $n_i(z)$ is the redshift distribution of group $i$.
+Defining the normalized assignment matrix ${\sf B}$ as
+  \begin{equation}
+    B_{\alpha,i}=A_{\alpha,i}\frac{\int dz\,n_i(z)}{\sum_j A_{\alpha,j}\int dz\,n_j(z)},
+  \end{equation}
+  ${\sf C}^{\rm group}_\ell$ is related to the matrix of cross-power spectra between pairs of bins via:
+  \begin{equation}
+    {\sf C}^{\rm bin}_\ell={\sf B}\,{\sf C}^{\rm group}_\ell\,{\sf B}^T.
+  \end{equation}
+
+  The rationale behind the {\tt ComplexSOM} method is based on the observation that, while ${\sf C}^{\rm group}_\ell$ is a slow quantity to calculate for a large number of groups, ${\sf B}$ and therefore ${\sf C}^{\rm bin}_\ell$ are very fast.
+Thus, given an initial set of groups characterizing the distribution of the full sample in colour space, ${\sf C}^{\rm group}_\ell$ can be precomputed once, and used to find the optimal assignment matrix ${\sf B}$ that maximizes the figure of merit in Eq.
+{\bf reference to FoM equation}, which can be written in terms of ${\sf C}^{\rm bin}_\ell$ alone as
+  \begin{equation}\label{eq:fom_tr}
+    F_{\theta\theta'}=\sum_\ell f_{\rm sky}\frac{2\ell+1}{2}{\rm Tr}\left[\partial_\theta{\sf C}^{\rm bin}_\ell({\sf C}^{\rm bin}_\ell)^{-1}\partial_{\theta'}{\sf C}^{\rm bin}_\ell({\sf C}_\ell^{\rm bin})^{-1}\right].
+  \end{equation}
+
+  Note that the general problem of finding the matrix ${\sf B}$ that optimally compresses the information on a given parameter has an analytical solution in terms of Karnhunen-L\^oeve eigenvectors \citep{astro-ph/9603021}.
+However, the assignment matrix used in this case has the additional constraints that it must be positive, discrete and binary (i.e.
+groups are either in a bin or they are not).
+Thus, finding the absolute maximum figure of merit would involve a brute-force evaluation of all matrix elements $A_{\alpha,j}$, which is an intractable $N_{\rm bin}^{N_{\rm group}}$ problem.
+We thus parametrize ${\sf A}$ in terms of a small set of continuous parameters, for which standard minimization routines can be used.
+In particular we use $N_{\rm bin}-1$ parameters, given by the edges of the redshift bins, and a group is assigned to a bin if its mean redshift falls within its edges.
+We also explored the possibility of ``trashing'' groups if less than a fraction $p_{\rm in}$ of its redshift distribution fell within its preassigned bin, treating $p_{\rm in}$ as an additional free parameter.
+We found that adding this extra freedom did not translate into a significant improvement in the final level of data compression.
+
+  Once the SOM has been generated, the {\tt ComplexSOM} method proceeds as follows.
+First, the full SOM is distilled into a smaller set of ${\cal O}(100)$ groups based by combining SOM cells with similar moments of their redshift distributions, thus ensuring that this process does not induce any catastrophic $N(z)$ broadening {\bf you may wanna refer to the {\tt SimpleSOM} method here}.
+The large ${\sf C}^{\rm group}_\ell$ is then precomputed for the initial set of groups.
+Finally, we find the optimal assignment matrix, defined in terms of the free redshift-edge parameters, by maximizing the figure of merit in Eq.
+\ref{eq:fom_tr} using Powell's minimization method \citep{10.1093/comjnl/7.2.155}.
 
@@
 GPz is having to extrapolate to make the redshift prediction, removing galaxies
-with $\nu>E\sigma^2$ (see Figure 12 of \cite{Hatfield2020}). Cutting galaxies
+with $\nu>E\sigma^2$ (see Figure 12 of \citealp{Hatfield2020}). Cutting galaxies
 removes some signal at the possible benefit of improving binning purity. As the
@@
 in Table \ref{table-settings} were used (see
-\cite{Almosallam2016a,Almosallam2016b} for precise definitions and
+\citealp{Almosallam2016a,Almosallam2016b} for precise definitions and
 interpretations).
@@
 inspired by the similar approach to K-means clustering in
-\cite{NIPS1994_a1140a3d}.
+\citet{NIPS1994_a1140a3d}.
 
@@
 
+\subsection{Splitting nominal bin choice from assignment}
+There were two ways to obtain good scores in the challenge: either one could choose the best
+nominal bin edges, or find the best method to put galaxies into those bins once they are chosen.
+These two are not completely disconnected: some theoretically powerful choices of nominal edges
+could prove impossible to assign in practice.  But it would still have been useful to separate
+the challenge into two separate parts, fixing the edges while optimizing classifiers, and vice
+versa.   We can retrospectively determine this since we understand the classifiers, but the picture
+would have been clearer doing so in advance.
+
+\subsection{Infrastructure}
+
 \begin{itemize}
-    \item Splitting classification from nominal bin choice
+    \item 
     \item Infrastructure for checking submissions

### 2021-02-02 paper/paper.tex
more appendix

@@
 The challenge organizers made a number of mistakes when building and running the challenge.
-These do not invalidate our process or results, but we describe them here in the hope of
+These do not invalidate our process or results, but we describe some of them here in the hope of
 being useful for future challenge designers.
@@
 \subsection{Infrastructure}
+Since submissions could use a wide variety of libraries and dependencies, it was extremely
+difficult after the challenge was complete to build appropriate environments for all the challenges
+in order to run the methods on additional data sets. In retrospect, a robust and careful
+continuous integration tool with a mechanism for entrants to specify an environment (for example, a 
+container or a {\sc conda} requirements file) could have been used to simplify this and ensure
+all methods could be easily run by the organizers.
 
-\begin{itemize}
-    \item 
-    \item Infrastructure for checking submissions
-    \item Guidance for users regarding caching and batch processing
-    \item Timing
-    \item Training set size
-    \item Advertisement
-\end{itemize}
+Once additional issue, though, was that many methods required GPUs to train in a reasonable time,
+and specifying an environment on these systems can be more complicated.
+
+\subsection{Guidance on caching \& batch processing}
+Since the training process for the challenge could be slow for some algorithms, many entries
+very reasonably cached trained models or other values for later use.  In some cases these
+cached values were then incorrectly picked up by subsequent runs on new data sets, leading
+to (at best) crashes and at worst silently poor scores.  We should have provided an explicit
+caching mechanism for entrants to avoid such issues.
+
+Additionally, the challenge supplied the data described in Section \ref{sec:data} to entrants,
+and some assumed that it was safe to hard-code specific paths to them or
+otherwise assume fixed data inputs.  This led to problems when switching to new test data sets. 
+Again, a requirement enforced by continuous integration could have checked this.
 

### 2021-02-02 tomo_challenge/classifiers/PQNLD.py
very quick and dirty enable of parallel PQNLD

@@ -52,10 +52,15 @@ gr = None
 dev = None
 kohonen = None
 
-training_cat_file = "training_bpz.cat"
-training_col_file = "training_bpz.columns"
+task_id = os.environ.get("SLURM_ARRAY_TASK_ID", "666")
+
+run_file = f"bpz_run_{task_id}.py"
+training_cat_file = f"training_bpz_{task_id}.cat"
+training_col_file = f"training_bpz_{task_id}.columns"
+bpz_file = f"training_bpz_{task_id}.bpz"
+ascii_file = f"training_bpz_{task_id}.asc"
 #validation_cat_file = "validato" # not used
-validation_col_file = "validation_bpz.columns"
+validation_col_file = f"validation_bpz_{task_id}.columns"
 
 def init_r_packages():
     global base, stats, gr, dev, kohonen
@@ -121,8 +126,8 @@ class PQNLD(Tomographer):
           true redshift for the training sample
 
         """
-        remove("training_bpz.asc")
-        remove("training_bpz.bpz")
+        remove(ascii_file)
+        remove(bpz_file)
         remove(training_col_file)
         remove(training_cat_file)
         remove(validation_col_file)
@@ -272,16 +277,16 @@ class PQNLD(Tomographer):
         if not os.path.exists("bpz-1.99.3-py3/"): 
             os.system("bash INSTALL.sh")
         
-        if not os.path.exists("training_bpz.bpz"): 
+        if not os.path.exists(bpz_file): 
             print("Running BPZ on the training data")
             curdir=os.getcwd()
             os.chdir('bpz-1.99.3-py3')
-            os.system("echo '#NEW' > bpz_run.py")
-            os.system("echo import sys >> bpz_run.py")
-            os.system("echo import os >> bpz_run.py")
-            os.system("echo 'os.environ[\"HOME\"]=\""+curdir+"\"' >>bpz_run.py")
-            os.system("echo 'os.environ[\"BPZPATH\"]=\""+curdir+"/bpz-1.99.3-py3/\"' >> bpz_run.py")
-            os.system("echo 'os.environ[\"NUMERIX\"]=\"numpy\"' >> bpz_run.py")
+            os.system("echo '#NEW' > " + run_file)
+            os.system("echo import sys >> " + run_file)
+            os.system("echo import os >> " + run_file)
+            os.system("echo 'os.environ[\"HOME\"]=\""+curdir + "\"' >> " + run_file)
[…]

### 2021-02-02 tomo_challenge/classifiers/PQNLD.py
one more missing obj

@@ -287,7 +287,7 @@ class PQNLD(Tomographer):
             os.system("echo 'os.environ[\"HOME\"]=\""+curdir + "\"' >> " + run_file)
             os.system("echo 'os.environ[\"BPZPATH\"]=\""+curdir+"/bpz-1.99.3-py3/\"' >> " + run_file)
             os.system("echo 'os.environ[\"NUMERIX\"]=\"numpy\"' >> " + run_file)
-            argv = ["\'bpz.py\'", "\'../training_bpz.cat\'",
+            argv = ["\'bpz.py\'", f"\'../{training_cat_file}\'",
                 "-PRIOR", "NGVS", "-SPECTRA", "CWWSB_capak.list",
                 "-ZMIN", "0.001", "-ZMAX", "7.000",
                 "-INTERP", "10", "-NEW_AB", "yes",

### 2021-02-02 bpz-1.99.3-py3/bpz_tools.py
tell BPZ not to recreate files

@@ -710,8 +710,11 @@ def ABflux(sed,filter,madau='yes'):
     #fmax=max(f)
     #f=where(less(f,fmax*ab_clip),0.,f)
 
-    print('Writing AB file ',ABoutput)
-    put_data(ABoutput,(z_ab,f))
+    if not os.path.exists(ABoutput):
+        print('Writing AB file ',ABoutput)
+        put_data(ABoutput,(z_ab,f))
+    else:
+        print("Skipping existing ", ABoutput)
 
 
 def VegatoAB(m_vega,filter,Vega=Vega):

### 2021-02-04 paper/paper.tex
add angus text

@@
 \subsection{ {\sc SimpleSOM} }
+The SimpleSOM algorithm submitted by AHW utilises a self-organising map
+\citep[SOM,][]{Kohonen:1982} to perform  a discretisation of the
+high-dimensional colour-magnitude-space of the challenge training and reference
+datasets.
+
+We utilise a branched version of the {\tt
+kohonen}\footnote{\url{https://github.com/ AngusWright/kohonen.git}} package
+\citep{Wright/etal:2020b, Wehrens/Kruisselbrink:2018, Wehrens/Lutgarde:2007}
+within R \citep{R}, which we integrate with  the challenge's {\tt python}
+classes using the {\tt rpy2} interface\footnote{\url{https://rpy2.github.io}}. 
+We train a $101 \times 101$ hexagonal-cell SOM with toroidal topology on the
+maximal combination of (non-repeated) colours, $z$-band magnitude, and so-called
+`band-triplets'. For example, for the `griz' setup, our SOM is trained on the
+combination of  $z$, $g-r$, $g-i$, $g-z$, $r-i$, $r-z$, $i-z$, $g-r-(r-i)$,
+$r-i-(i-z)$. This combination proved optimal during testing on the DC2 dataset,
+as determined by the 3x2pt SNR metric. The training and reference data then
+propagated into this trained SOM, producing like-for-like groupings between the
+two datasets. We then compute the mean redshift of the training sample sources
+within each cell and the number of reference sources per-cell. By rank-ordering
+the cells in mean training-sample redshift, we construct the cumulative
+distribution function of reference sample source counts, and split cells into
+${n_{\rm tomo}}$ equal-N tomographic bins using this function.
+
+In addition to this base functionality, the SimpleSOM method includes a
+mechanism for distilling $N_{\rm cell}$ SOM cells into $N_{\rm group}<N_{\rm
+cell}$ groups of cells, while maintaining optimal redshift resolution. This is
+done by invoking a hierarchical clustering of SOM cells \citep[see Appendix B of
+][]{Wright/etal:2020a}, where in this case we combine cells based on the
+similarity of the cells Nz moments (mean, median, and normalised median absolute
+deviation from median) using complete-linkage clustering. The result of this
+procedure is the construction of fewer discrete groups of training and reference
+data, while not introducing pathological redshift-distribution broadening (as
+happens when, e.g., training using a lower resolution SOM and/or clustering
+cells based on distances in colour-magnitude space).
 
 \subsection{ {\sc UTOPIA} }
+
+The `Useless Technique that OutPerforms Intelligent Alternatives' is a method
+submitted by AHW that is designed,  primarily, as a demonstration of the care
+that must be taken in the interpretation of the tomographic challenge results. 
+The method performs  the simplest-possible direct mapping between the training
+and reference data, by performing a nearest-neighbour  assignment of training
+sample galaxies to each reference sample source in $n_{\rm band}$ magnitude-only
+space. Each reference  sample galaxy is then assigned the redshift of its
+matching training-sample source, and these redshifts are  used to bin the
+reference galaxies into $n_{\rm tomo}$ equal-N tomographic bins. Importantly,
+the method was specifically  implemented so as to use only the minimum
+information contained within the available bands (by using magnitudes alone,
+without  any colours, etc.).
+
+In this way UTOPIA is the simplest possible algorithm (utilising all available
+photometric bands) that one can  implement for tomography. It is also a method
+whose performance is optimal in the limit of an infinitely large,  perfectly
+representative training sample; i.e. the limit in which this challenge operates.
+As the requirements of perfect  training-sample representivity and large size
+are violated, UTOPIA ought to assign redshifts to reference  galaxies from
+increasingly distant regions of the multi-dimensional magnitude-space, resulting
+in poorer tomographic binning.  Therefore, UTOPIA acts as a warning against
+over-interpretation of the tomographic challenge results, as success in the
+challenge  need not translate to optimal-performance/usefulness under realistic
+survey conditions. Instead, the challenge results should be interpreted
+holistically, by comparing the performance between different classes of methods,
+rather than individuals.
 

### 2021-02-08 tomo_challenge/classifiers/PQNLD.py
more attempts to make PQNLD work for griz

@@ -503,8 +503,9 @@ class PQNLD(Tomographer):
                                    ["0.000001","0.000001","0.000001","0.000001"])
             #Output the columns file 
             columns.to_csv(validation_col_file,sep=' ',index=None,header=None,mode='w')
-            columns = pd.DataFrame(["M_0"],
-                                   ["7"])
+            columns = pd.DataFrame(np.transpose(
+                                  np.array([["M_0"],
+                                            ["7"]])))
             #Output the columns file 
             columns.to_csv(validation_col_file,sep=' ',index=None,header=None,mode='a')
         elif self.bands == 'grizy':
@@ -521,8 +522,9 @@ class PQNLD(Tomographer):
                                    ["0.000001","0.000001","0.000001","0.000001","0.000001"])
             #Output the columns file 
             columns.to_csv(validation_col_file,sep=' ',index=None,header=None,mode='w')
-            columns = pd.DataFrame(["M_0"],
-                                   ["7"])
+            columns = pd.DataFrame(np.transpose(
+                                  np.array([["M_0"],
+                                            ["7"]])))
             #Output the columns file 
             columns.to_csv(validation_col_file,sep=' ',index=None,header=None,mode='a')
         elif self.bands == 'ugriz':
@@ -539,8 +541,9 @@ class PQNLD(Tomographer):
                                    ["0.000001","0.000001","0.000001","0.000001","0.000001"])
             #Output the columns file 
             columns.to_csv(validation_col_file,sep=' ',index=None,header=None,mode='w')
-            columns = pd.DataFrame(["M_0"],
-                                   ["9"])
+            columns = pd.DataFrame(np.transpose(
+                                  np.array([["M_0"],
+                                            ["9"]])))
             #Output the columns file 
             columns.to_csv(validation_col_file,sep=' ',index=None,header=None,mode='a')
         elif self.bands == 'ugrizy':
@@ -557,8 +560,9 @@ class PQNLD(Tomographer):
                                    ["0.000001","0.000001","0.000001","0.000001","0.000001","0.000001"])
             #Output the columns file 
             columns.to_csv(validation_col_file,sep=' ',index=None,header=None,mode='w')
-            columns = pd.DataFrame(["M_0"],
-                                   ["9"])
+            columns = pd.DataFrame(np.transpose(
+                                  np.array([["M_0"],
+                                            ["9"]])))
             #Output the columns file 
             columns.to_csv(validation_col_file,sep=' ',index=None,header=None,mode='a')
 

### 2021-02-08 tomo_challenge/classifiers/PQNLD.py
PQNLD fix in training as well as apply.

@@ -195,8 +195,9 @@ class PQNLD(Tomographer):
                                    ["0.000001","0.000001","0.000001","0.000001"]])))
             #Output the columns file 
             columns.to_csv(training_col_file,sep=' ',index=None,header=None,mode='w')
-            columns = pd.DataFrame(["M_0","Z_S"],
-                                   ["7","9"])
+            columns = pd.DataFrame(np.transpose(
+                                  np.array([["M_0","Z_S"],
+                                            ["7","9"]])))
             #Output the columns file 
             columns.to_csv(training_col_file,sep=' ',index=None,header=None,mode='a')
         elif self.bands == 'grizy':
@@ -213,8 +214,9 @@ class PQNLD(Tomographer):
                                    ["0.000001","0.000001","0.000001","0.000001","0.000001"]])))
             #Output the columns file 
             columns.to_csv(training_col_file,sep=' ',index=None,header=None,mode='w')
-            columns = pd.DataFrame(["M_0","Z_S"],
-                                   ["7","11"])
+            columns = pd.DataFrame(np.transpose(
+                                  np.array([["M_0","Z_S"],
+                                            ["7","11"]])))
             #Output the columns file 
             columns.to_csv(training_col_file,sep=' ',index=None,header=None,mode='a')
         elif self.bands == 'ugriz':
@@ -231,8 +233,9 @@ class PQNLD(Tomographer):
                                    ["0.000001","0.000001","0.000001","0.000001","0.000001"]])))
             #Output the columns file 
             columns.to_csv(training_col_file,sep=' ',index=None,header=None,mode='w')
-            columns = pd.DataFrame(["M_0","Z_S"],
-                                   ["9","11"])
+            columns = pd.DataFrame(np.transpose(
+                                  np.array([["M_0","Z_S"],
+                                            ["9","11"]])))
             #Output the columns file 
             columns.to_csv(training_col_file,sep=' ',index=None,header=None,mode='a')
         elif self.bands == 'ugrizy':
@@ -249,8 +252,9 @@ class PQNLD(Tomographer):
                                    ["0.000001","0.000001","0.000001","0.000001","0.000001","0.000001"]])))
             #Output the columns file 
             columns.to_csv(training_col_file,sep=' ',index=None,header=None,mode='w')
-            columns = pd.DataFrame(["M_0","Z_S"],
-                                   ["9","13"])
+            columns = pd.DataFrame(np.transpose(
+                                  np.array([["M_0","Z_S"],
[…]

### 2021-02-08 paper/results/make_figures.py
add g-band loss fig

@@ -1,0 +1,154 @@
+import yaml
+import tabulate
+import numpy as np
+from astropy.table import Table
+import matplotlib
+matplotlib.use('agg')
+import matplotlib.pyplot as plt
+
+# set default styles
+plt.style.use('StyleSheet.mplstyle')
+
+
+
+method_names = [
+    "Autokeras_LSTM",
+    "CNN",
+    "ComplexSOM",
+    "ENSEMBLE1",
+    "funbins",
+    "GPzBinning",
+    "IBandOnly",
+    "JaxCNN",
+    "JaxResNet",
+    "LGBM",
+    "LSTM",
+#     "MineCraft",
+    "mlpqna",
+    "myCombinedClassifiers",
+    "NeuralNetwork1",
+    "NeuralNetwork2",
+    "PCACluster",
+    "Random",
+    "RandomForest",
+    "SimpleSOM",
+    "TensorFlow_FFNN",
+    "TCN",
+    "UTOPIA",
+    "ZotBin",
+    "ZotNet",
+]
+
+metrics = ["SNR_ww", "SNR_gg", "SNR_3x2", "FOM_ww", "FOM_gg", "FOM_3x2", "FOM_DETF_ww", "FOM_DETF_gg", "FOM_DETF_3x2"]
+
+
+class EverythingIsZero:
+    def __getitem__(self, x):
+        return 0
+    
+class EverythingIsNan:
+    def __getitem__(self, x):
+        return np.nan
+
+
+def make_table(results, metric):
+    N = np.array([3, 5, 7, 9])
+    row_format = r"{0} & {1} & {2}    & {3}    & {4}    & {5}             & {6}             & {7}             & {8}\\"
+    data = {}
+    for row in results:
+        name  = row['method']
+        bands = row['bands']
+        bins = row['bins']
+        data[name, bands, bins] = row[metric]
+
+
+    for name in method_names:
+        disp_name = rf"{{\sc {name} }}".replace("_", r"\_")
+        row = [disp_name]
+        for i, bands in enumerate(['riz', 'griz']):
+            for n in N:
+                val  = data.get((name, bands, n), np.nan)
+                if np.isnan(val) or val == 0:
+                    row.append(f"--")
+                else:
+                    row.append(f"{val:.1f}")
+
+        print(row_format.format(*row))
+
+
+def load_table(base):
[…]

### 2021-02-08 paper/paper.tex
more text and plots

@@
 participants. We find that even from this limited photometry information,
-various methods are able to yield binning schemes that achieve high FoM scores.
+various methods are able to yield binning schemes that achieve high figure-of-merit scores.
 \end{abstract}
@@
     S = \mu^{T} C^{-1} \mu
+\label{eq:snr}
 \end{equation}
@@
 
-Metric 2 uses a figure of merit based on that presented by the Dark Energy Task Force \citep{detf}.  We 
+Metric 2 uses a figure of merit (FOM) based on that presented by the Dark Energy Task Force \citep{detf}.  We 
 use the inverse area of a Fisher Matrix ellipse, which approximates the size in two axes 
@@
     S &= \frac{1}{2 \pi \sqrt{\det{([F^{-1}]_{p_1, p_2})}}}
+\label{eq:fom}
 \end{align}
@@
 \subsection{ {\sc ComplexSOM} }
-  The {\tt ComplexSOM} method implements an additional optimization layer on the methodology used by {\tt SimpleSOM}.
-
-  Let ${\sf C}^{\rm group}_\ell$ be the matrix containing all auto- and cross-power spectra between members of a large set of galaxy samples (called ``groups'' here).
-We can compress this large set into a smaller set of samples (called ``bins'' here), by combining different groups.
-Let ${\sf A}$ be the assignment matrix determining these bins, such that $A_{\alpha,i}=1$ if group $i$ is assigned to bin $\alpha$, and 0 otherwise.
-In that case, the redshift distribution of bin $\alpha$ is given by $N_\alpha(z)=\sum_i A_{\alpha,i}n_i(z)$, where $n_i(z)$ is the redshift distribution of group $i$.
+The {\tt ComplexSOM} method implements an additional optimization layer on the
+methodology used by {\tt SimpleSOM}.
+
+Let ${\sf C}^{\rm group}_\ell$ be the matrix containing all auto- and
+cross-power spectra between members of a large set of galaxy samples (called
+``groups'' here). We can compress this large set into a smaller set of samples
+(called ``bins'' here), by combining different groups. Let ${\sf A}$ be the
+assignment matrix determining these bins, such that $A_{\alpha,i}=1$ if group
+$i$ is assigned to bin $\alpha$, and 0 otherwise. In that case, the redshift
+distribution of bin $\alpha$ is given by $N_\alpha(z)=\sum_i
+A_{\alpha,i}n_i(z)$, where $n_i(z)$ is the redshift distribution of group $i$.
+
 Defining the normalized assignment matrix ${\sf B}$ as
@@
 \end{equation}
-  ${\sf C}^{\rm group}_\ell$ is related to the matrix of cross-power spectra between pairs of bins via:
+
+${\sf C}^{\rm group}_\ell$ is related to the matrix of cross-power spectra
+between pairs of bins via:
 \begin{equation}
@@
 
-  The rationale behind the {\tt ComplexSOM} method is based on the observation that, while ${\sf C}^{\rm group}_\ell$ is a slow quantity to calculate for a large number of groups, ${\sf B}$ and therefore ${\sf C}^{\rm bin}_\ell$ are very fast.
-Thus, given an initial set of groups characterizing the distribution of the full sample in colour space, ${\sf C}^{\rm group}_\ell$ can be precomputed once, and used to find the optimal assignment matrix ${\sf B}$ that maximizes the figure of merit in Eq.
-{\bf reference to FoM equation}, which can be written in terms of ${\sf C}^{\rm bin}_\ell$ alone as
+The rationale behind the {\tt ComplexSOM} method is based on the observation
+that, while ${\sf C}^{\rm group}_\ell$ is a slow quantity to calculate for a
+large number of groups, ${\sf B}$ and therefore ${\sf C}^{\rm bin}_\ell$ are
+very fast. Thus, given an initial set of groups characterizing the
+distribution of the full sample in colour space, ${\sf C}^{\rm group}_\ell$
+can be precomputed once, and used to find the optimal assignment matrix ${\sf
+B}$ that maximizes the figure of merit in Equation \ref{eq:fom},
+which can be written in terms of ${\sf C}^{\rm bin}_\ell$ alone as
+
 \begin{equation}\label{eq:fom_tr}
@@
 
-  Note that the general problem of finding the matrix ${\sf B}$ that optimally compresses the information on a given parameter has an analytical solution in terms of Karnhunen-L\^oeve eigenvectors \citep{astro-ph/9603021}.
-However, the assignment matrix used in this case has the additional constraints that it must be positive, discrete and binary (i.e.
-groups are either in a bin or they are not).
-Thus, finding the absolute maximum figure of merit would involve a brute-force evaluation of all matrix elements $A_{\alpha,j}$, which is an intractable $N_{\rm bin}^{N_{\rm group}}$ problem.
-We thus parametrize ${\sf A}$ in terms of a small set of continuous parameters, for which standard minimization routines can be used.
-In particular we use $N_{\rm bin}-1$ parameters, given by the edges of the redshift bins, and a group is assigned to a bin if its mean redshift falls within its edges.
-We also explored the possibility of ``trashing'' groups if less than a fraction $p_{\rm in}$ of its redshift distribution fell within its preassigned bin, treating $p_{\rm in}$ as an additional free parameter.
-We found that adding this extra freedom did not translate into a significant improvement in the final level of data compression.
-
-  Once the SOM has been generated, the {\tt ComplexSOM} method proceeds as follows.
-First, the full SOM is distilled into a smaller set of ${\cal O}(100)$ groups based by combining SOM cells with similar moments of their redshift distributions, thus ensuring that this process does not induce any catastrophic $N(z)$ broadening {\bf you may wanna refer to the {\tt SimpleSOM} method here}.
-The large ${\sf C}^{\rm group}_\ell$ is then precomputed for the initial set of groups.
-Finally, we find the optimal assignment matrix, defined in terms of the free redshift-edge parameters, by maximizing the figure of merit in Eq.
-\ref{eq:fom_tr} using Powell's minimization method \citep{10.1093/comjnl/7.2.155}.
+Note that the general problem of finding the matrix ${\sf B}$ that optimally
+compresses the information on a given parameter has an analytical solution in
+terms of Karnhunen-L\^oeve eigenvectors \citep{astro-ph/9603021}. However, the
+assignment matrix used in this case has the additional constraints that it
+must be positive, discrete and binary (i.e. groups are either in a bin or they
+are not). Thus, finding the absolute maximum figure of merit would involve a
+brute-force evaluation of all matrix elements $A_{\alpha,j}$, which is an
+intractable $N_{\rm bin}^{N_{\rm group}}$ problem. We thus parametrize ${\sf
+A}$ in terms of a small set of continuous parameters, for which standard
+minimization routines can be used. In particular we use $N_{\rm bin}-1$
+parameters, given by the edges of the redshift bins, and a group is assigned
+to a bin if its mean redshift falls within its edges. We also explored the
+possibility of ``trashing'' groups if less than a fraction $p_{\rm in}$ of its
+redshift distribution fell within its preassigned bin, treating $p_{\rm in}$
+as an additional free parameter. We found that adding this extra freedom did
+not translate into a significant improvement in the final level of data
+compression.
+
+Once the SOM has been generated, the {\tt ComplexSOM} method proceeds as
+follows. First, the full SOM is distilled into a smaller set of ${\cal O}(100)$
+groups based by combining SOM cells with similar moments of their redshift
+distributions, thus ensuring that this process does not induce any catastrophic
+$N(z)$ broadening {\bf you may wanna refer to the {\tt SimpleSOM} method here}.
+The large ${\sf C}^{\rm group}_\ell$ is then precomputed for the initial set of
+groups. Finally, we find the optimal assignment matrix, defined in terms of the
+free redshift-edge parameters, by maximizing the figure of merit in Eq.
+\ref{eq:fom_tr} using Powell's minimization method
+\citep{10.1093/comjnl/7.2.155}.
 
@@
 
+\section{Overall results}
+The methods described above were run on a data set, previously unseen by entrants but drawn
+from the same underlying population, of size 8.6M objects for CosmoDC2 and 5.4M objects for Buzzard.
+A complete suite of the metrics described in Section \ref{sec:metrics} was run on each of the two
+for each method.  In total, each method was run on 16 scenarios, each combination of: griz vs riz,
+Buzzard vs CosmoDC2, and 3,5,7, and 9 bins requested.  Nine metrics were calculated for each
+scenario: clustering, lensing, and 3x2pt for each of the SNR metric (Equation \ref{eq:snr}, the $w_0-w_a$ 
+DETF FOM, and the $\Omega_c - \sigma_8$ FOM (Equation \ref{eq:fom}).
+
+The DETF results for each method are shown in Table \ref{tab:cosmodc2}.  Further results for other
+metrics are shown in Appendix \ref{sec:complete}.  Missing entries occured in two ways.
+The first, marked with an asterisk, is when a method did not run for the given scenario,
+either due to memory overflow or taking an extremely long time.
+The second, marked with a dash, is when a method ran, but generated at least one bin with an extremely small number
+count, such that the spectrum could not be calculated.
+
 \begin{table*}[]
@@
 {\sc ComplexSOM } & 38.0 & 52.1    & 94.4    & 101.6    & 34.9             & 45.0             & 91.6             & 100.3\\
-{\sc ENSEMBLE1 } & 65.6 & 65.6    & 94.8    & --    & --             & --             & --             & --\\
+{\sc ENSEMBLE1 } & 65.6 & 65.6    & 94.8    & *    & *             & *             & *             & *\\
 {\sc funbins } & 36.8 & 82.5    & 122.1    & 141.8    & 42.1             & 100.4             & 142.5             & 167.2\\
@@
 {\sc myCombinedClassifiers } & 35.3 & 60.5    & 83.5    & 120.4    & 39.4             & 63.9             & 93.1             & 148.9\\
-{\sc NeuralNetwork1 } & 76.3 & 117.6    & 135.8    & --    & 81.9             & 132.9             & 158.0             & --\\
+{\sc NeuralNetwork1 } & 76.3 & 117.6    & 135.8    & 109.9    & 81.9             & 132.9             & 158.0             & 96.9\\
 {\sc NeuralNetwork2 } & 30.5 & 48.0    & 57.8    & 136.1    & 49.4             & 102.7             & 122.2             & 140.0\\
-{\sc PCACluster } & 29.5 & 68.1    & 75.8    & --    & 50.1             & 72.7             & 90.8             & --\\
+{\sc PCACluster } & 29.5 & 68.1    & 75.8    & *    & 50.1             & 72.7             & 90.8             & *\\
 {\sc Random } & 1.2 & 1.2    & 1.2    & 1.2    & 1.2             & 1.2             & 1.2             & 1.2\\
@@
 
+\subsection{Metric comparisons}
+Figure \ref{fig:metrics} shows a comparison of some of the different metrics used in the challenge.
+
+\begin{figure}[htbp]
+% \centering
+\includegraphics[width=0.9\linewidth]{results/metric_comparisons.pdf}
+\caption{A comparison of the metrics used in our challenge. The x-axis is shared, and shows
+the 3x2pt $w_0 - w_a$ (DETF) figure of merit on the CosmoDC2 data set for all methods, 
+with colours labelling the number
+of bins used.  Each of the panels shows a metric which is the same except for a single change, 
+which is labelled on its y-axis.  For example, the top panel still shows the 3x2pt $w_0 - w_a$ (DETF) figure of merit,
+but for the Buzzard data set.}
+\label{fig:metrics}
+\end{figure}
+
+\subsection{Impact of losing g-band}
+Figure \ref{fig:loss} shows the ratio of the FOM between methods using the griz and the same method
+using the riz.  Each point represents one method with a specific number of bins, indicated by colour.
+For lower scoring methods and for smaller numbers of bins there is more scatter in the ratio,
+but for the highest scoring algorithms and configurations the loss when losing g-band is a little
+more stable, at around 10--15\%.
+
+\begin{figure*}
+\includegraphics[width=0.9\linewidth]{results/g_band_loss.pdf}
+\caption{The degradation in constraining power over all methods
+ when comparing them on methods including and excluding
+the g-band, in terms of the ratio of the $w_0-w-a$ figure of merit.  The left-hand panel shows
+results for the CosmoDC2 scenarios, and the right-hand panel for the Buzzard scenarios.  Colours
+indicate the number of bins used in the analysis.}
+\label{fig:loss}
+\end{figure*}
+
+\subsection{Trained edges vs fixed edges}
+
+Some of the methods in the challenge accepted target bin edges as an input to their algorithms,
+and then used classifiers to assign objects to those bins.  Other methods optimized the values
+of bin edges themselves.
+
 \begin{itemize}
@@
 \section{Discussion}
-\begin{itemize}
-    \item Overall effectiveness of method classes
-    \item Individual winner
-\end{itemize}
-
--
-
-\subsection{General outcomes}
 It is clear from values above that fears about restricted photometry catastrophically limiting
@@
 \begin{itemize}
-    \item spectroscopic incompleteness
+    \item spectroscopic incompleteness / unrepresentativeness
+    \item data set volume
+    \item comparison of optimal choices in different science cases
 \end{itemize}
-
-\section{Conclusions}
 
@@
 \section{Theory}\label{app:theory}
-In our metrics, the power spectrum between two tomographic bins $i$ and $j$ is computed using either the Core Cosmology Library \citep{ccl} or the JAX Cosmology Library \citep{jax-cosmo} as:
+
+In our metrics, the power spectrum between two tomographic bins $i$ and $j$ is
+computed using either the Core Cosmology Library \citep{ccl} or the JAX
+Cosmology Library \citep{jax-cosmo} as:
+
 \begin{equation}
@@
 \section{Complete results}
+\label{sec:complete}
 
@@
 Since the training process for the challenge could be slow for some algorithms, many entries
-very reasonably cached trained models or other values for later use.  In some cases these
+(very reasonably) cached trained models or other values for later use.  In some cases these
 cached values were then incorrectly picked up by subsequent runs on new data sets, leading

### 2021-02-08 paper/results/make_figures.py

@@ -68,8 +68,10 @@ def make_table(results, metric):
         for i, bands in enumerate(['riz', 'griz']):
             for n in N:
                 val  = data.get((name, bands, n), np.nan)
-                if np.isnan(val) or val == 0:
+                if np.isnan(val):
                     row.append(f"--")
+                elif val == 0:
+                    row.append(f"*")
                 else:
                     row.append(f"{val:.1f}")
 
@@ -102,8 +104,68 @@ def pair_riz_griz(data):
     return data[riz], data[griz]
 
 
+def color_plot(data, x, y, ax):
+    c = data['bins']
+    cmap = matplotlib.cm.Dark2
+    for i, nbin in enumerate([3,5,7,9]):
+        s = data['bins']==nbin
+        ax.plot(x[s], y[s], '.', markersize=8, color=cmap.colors[i], label=f'$n_b={nbin}$')
+
+
+def plot_metric_comparisons(dc2, buzzard, filename):
+    assert (dc2['method'] == buzzard['method']).all()
+    assert (dc2['bins'] == buzzard['bins']).all()
+
+
+    fig, ax = plt.subplots(4, 1, figsize=(4,9), sharex=True)
+    x = dc2['FOM_DETF_3x2'].copy()
+    x[x==0] = np.nan
+    y = buzzard['FOM_DETF_3x2'].copy()
+    y[y==0] = np.nan
+    color_plot(dc2, x, y, ax[0])
+    ax[0].set_ylabel("Buzzard FOM")
+    ax[0].set_ylim(0)
+    ax[0].legend(ncol=2, frameon=True)
+
+
+    x = dc2['FOM_DETF_3x2'].copy()
+    x[x==0] = np.nan
+    y = dc2['FOM_3x2'].copy() / 1000
+    y[y==0] = np.nan
+    color_plot(dc2, x, y, ax[1])
+    ax[1].set_ylabel(r"$\Omega_c - \sigma_8$ FOM / 1000")
+    ax[1].set_ylim(0)
+
+
+    x = dc2['FOM_DETF_3x2'].copy()
+    x[x==0] = np.nan
+    y = dc2['FOM_DETF_ww'].copy()
+    y[y==0] = np.nan
+    color_plot(dc2, x, y, ax[2])
+    ax[2].set_ylabel(r"Lensing FOM")
[…]

### 2021-02-08 paper/paper.tex
add tables and plots

@@
 \textbf{Method} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} \\ \hline
-{\sc Autokeras\_LSTM } & 31.1 & 74.9    & 103.4    & --    & 44.1             & 67.5             & 122.2             & 98.4\\
-{\sc CNN } & 27.1 & 50.8    & 76.7    & 103.3    & 30.5             & 57.7             & 95.4             & 122.4\\
-{\sc ComplexSOM } & 38.0 & 52.1    & 94.4    & 101.6    & 34.9             & 45.0             & 91.6             & 100.3\\
-{\sc ENSEMBLE1 } & 65.6 & 65.6    & 94.8    & *    & *             & *             & *             & *\\
-{\sc funbins } & 36.8 & 82.5    & 122.1    & 141.8    & 42.1             & 100.4             & 142.5             & 167.2\\
-{\sc GPzBinning } & 26.2 & 49.7    & 80.8    & 111.7    & 27.8             & 55.3             & 87.2             & 126.9\\
-{\sc IBandOnly } & 38.0 & 50.0    & 54.4    & 57.3    & 38.0             & 50.0             & 54.4             & 57.3\\
-{\sc JaxCNN } & 59.9 & 101.9    & 105.2    & --    & 79.7             & 125.2             & 150.0             & --\\
-{\sc JaxResNet } & 73.7 & 111.1    & 131.6    & --    & 82.2             & 126.0             & --             & 161.5\\
-{\sc LGBM } & 27.0 & 50.8    & 78.8    & 108.2    & 30.3             & 57.4             & 92.9             & 125.2\\
-{\sc LSTM } & 26.7 & 51.1    & 81.4    & 103.6    & 30.2             & 57.1             & 95.9             & 126.9\\
-{\sc mlpqna } & 39.3 & 64.7    & 93.2    & 121.8    & 42.8             & 72.2             & 109.0             & 133.7\\
-{\sc myCombinedClassifiers } & 35.3 & 60.5    & 83.5    & 120.4    & 39.4             & 63.9             & 93.1             & 148.9\\
-{\sc NeuralNetwork1 } & 76.3 & 117.6    & 135.8    & 109.9    & 81.9             & 132.9             & 158.0             & 96.9\\
-{\sc NeuralNetwork2 } & 30.5 & 48.0    & 57.8    & 136.1    & 49.4             & 102.7             & 122.2             & 140.0\\
-{\sc PCACluster } & 29.5 & 68.1    & 75.8    & *    & 50.1             & 72.7             & 90.8             & *\\
-{\sc Random } & 1.2 & 1.2    & 1.2    & 1.2    & 1.2             & 1.2             & 1.2             & 1.2\\
-{\sc RandomForest } & 39.5 & 65.2    & 93.2    & 118.5    & 43.0             & 72.8             & 110.1             & 136.5\\
-{\sc SimpleSOM } & 39.5 & 59.9    & 77.9    & 98.5    & 42.2             & 73.2             & 108.0             & 134.8\\
-{\sc TensorFlow\_FFNN } & 39.0 & 64.3    & 90.3    & 114.7    & 43.0             & 72.0             & 112.3             & 137.2\\
-{\sc TCN } & 40.0 & 65.3    & 90.9    & 110.5    & 44.1             & 72.8             & 108.0             & 134.0\\
-{\sc UTOPIA } & 39.0 & 63.7    & 90.8    & 112.2    & 43.1             & 74.0             & 112.8             & 139.3\\
-{\sc ZotBin } & 64.8 & 106.4    & 121.8    & 135.3    & 77.6             & 120.0             & 141.8             & 154.7\\
-{\sc ZotNet } & 73.6 & 111.2    & 131.8    & 145.9    & 83.7             & 128.5             & 150.1             & 167.2\\
+\input{results/tables/table_FOM_DETF_3x2_dc2.tex}
 \end{tabular}
@@
 \textbf{Method} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} \\ \hline
-{\sc Autokeras\_LSTM } & 13.8 & 30.2    & 31.1    & --    & 26.4             & 7.7             & 43.0             & 44.6\\
-{\sc CNN } & 7.0 & 29.8    & 37.5    & 38.1    & 6.2             & 35.8             & 45.0             & 48.7\\
-{\sc ComplexSOM } & 17.3 & 19.7    & 25.0    & 25.4    & 6.5             & 12.2             & 19.3             & 21.5\\
-{\sc ENSEMBLE1 } & 36.0 & 36.0    & 38.6    & --    & --             & --             & --             & --\\
-{\sc funbins } & 23.3 & 31.6    & 36.5    & 39.8    & 31.2             & 43.0             & 50.4             & 54.0\\
-{\sc GPzBinning } & 6.9 & 23.9    & 33.1    & 35.1    & 6.5             & 36.0             & 46.0             & 47.8\\
-{\sc IBandOnly } & 0.8 & 2.6    & 3.2    & 3.6    & 0.8             & 2.6             & 3.2             & 3.6\\
-{\sc JaxCNN } & 9.3 & 27.9    & 29.0    & --    & 3.3             & 38.0             & 42.7             & --\\
-{\sc JaxResNet } & 2.1 & 30.8    & 37.9    & --    & 2.2             & 31.1             & --             & 44.7\\
-{\sc LGBM } & 7.0 & 30.1    & 37.5    & 39.6    & 6.3             & 36.3             & 45.4             & 48.6\\
-{\sc LSTM } & 6.9 & 29.3    & 36.8    & 38.2    & 6.4             & 35.5             & 44.6             & 48.9\\
-{\sc mlpqna } & 23.1 & 35.0    & 37.3    & 38.0    & 30.0             & 43.9             & 48.9             & 51.4\\
-{\sc myCombinedClassifiers } & 1.2 & 8.4    & 28.8    & 39.1    & 1.7             & 10.9             & 37.2             & 49.3\\
-{\sc NeuralNetwork1 } & 1.6 & 34.0    & 36.7    & --    & 2.1             & 37.7             & 46.3             & --\\
-{\sc NeuralNetwork2 } & 19.2 & 21.8    & 22.4    & 35.8    & 8.5             & 18.6             & 22.5             & 36.7\\
-{\sc PCACluster } & 7.8 & 11.3    & 14.2    & --    & 12.6             & 12.0             & 21.6             & --\\
-{\sc Random } & 0.0 & 0.0    & 0.0    & 0.0    & 0.0             & 0.0             & 0.0             & 0.0\\
-{\sc RandomForest } & 23.5 & 34.9    & 38.4    & 39.6    & 30.0             & 44.5             & 50.1             & 53.2\\
-{\sc SimpleSOM } & 20.5 & 30.7    & 35.1    & 36.7    & 28.8             & 45.4             & 50.7             & 53.8\\
-{\sc TensorFlow\_FFNN } & 22.6 & 34.4    & 36.7    & 37.7    & 29.7             & 43.4             & 48.3             & 50.9\\
-{\sc TCN } & 23.6 & 36.0    & 38.0    & 38.7    & 27.6             & 41.2             & 46.3             & 49.4\\
-{\sc UTOPIA } & 22.0 & 31.1    & 35.1    & 36.3    & 29.8             & 43.2             & 49.6             & 52.3\\
-{\sc ZotBin } & 4.8 & 24.8    & 30.9    & 29.1    & 4.0             & 27.3             & 35.9             & 38.1\\
-{\sc ZotNet } & 2.2 & 29.9    & 33.1    & 34.6    & 2.3             & 33.2             & 42.3             & 43.0\\
+\input{results/tables/table_FOM_ww_dc2.tex}
 \end{tabular}
@@
 \textbf{Method} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} \\ \hline
-{\sc Autokeras\_LSTM } & 2.9 & 10.5    & 8.2    & --    & 5.7             & 5.0             & --             & 11.5\\
-{\sc CNN } & 1.8 & 9.5    & 12.3    & 13.2    & 1.8             & 9.4             & 14.7             & 16.2\\
-{\sc ComplexSOM } & 1.9 & 6.2    & 7.5    & 8.4    & 1.0             & 7.9             & 13.2             & 15.9\\
-{\sc ENSEMBLE1 } & 11.4 & 11.4    & 13.7    & 13.7    & 13.9             & 13.9             & 17.9             & 17.9\\
-{\sc funbins } & 4.4 & 8.8    & 10.8    & 11.8    & 7.2             & 13.1             & 16.3             & 17.4\\
-{\sc GPzBinning } & 1.7 & 9.3    & 11.3    & 12.5    & 2.1             & 10.6             & 15.0             & 16.7\\
-{\sc IBandOnly } & 0.2 & 0.3    & 0.4    & 0.4    & 0.2             & 0.3             & 0.4             & 0.4\\
-{\sc JaxCNN } & 4.0 & 10.0    & 11.6    & --    & 4.4             & 13.1             & 14.9             & 15.5\\
-{\sc JaxResNet } & 4.2 & 10.2    & --    & --    & 4.8             & 11.3             & --             & --\\
-{\sc LGBM } & 1.8 & 9.5    & 12.5    & 13.4    & 1.8             & 9.6             & 14.5             & 16.5\\
-{\sc LSTM } & 1.8 & 9.6    & 12.6    & 13.3    & 1.9             & 9.6             & 14.5             & 16.5\\
-{\sc mlpqna } & 4.8 & 11.2    & 13.0    & 13.7    & 6.9             & 13.8             & 17.3             & 18.3\\
-{\sc myCombinedClassifiers } & 1.5 & 6.6    & 11.0    & 13.1    & 2.3             & 9.5             & 14.9             & 17.6\\
-{\sc NeuralNetwork1 } & 3.9 & 10.2    & 12.0    & 13.3    & 5.0             & 14.5             & 16.4             & 17.9\\
-{\sc NeuralNetwork2 } & 2.3 & 9.8    & 11.1    & 12.6    & 2.8             & 7.0             & 9.8             & 14.1\\
-{\sc PCACluster } & 3.8 & 6.0    & 4.2    & --    & 1.2             & 2.5             & 5.0             & --\\
-{\sc Random } & 0.0 & 0.0    & 0.0    & 0.0    & 0.0             & 0.0             & 0.0             & 0.0\\
-{\sc RandomForest } & 4.6 & 10.8    & 12.6    & 13.1    & 6.8             & 13.8             & 17.2             & 18.2\\
-{\sc SimpleSOM } & 4.7 & 10.4    & 12.0    & 12.7    & 7.3             & 15.0             & 17.6             & 18.5\\
-{\sc TensorFlow\_FFNN } & -- & --    & --    & --    & --             & --             & --             & --\\
-{\sc TCN } & 5.0 & 10.9    & 12.9    & 13.4    & 6.2             & 12.2             & 15.8             & 16.4\\
-{\sc UTOPIA } & 4.2 & 9.9    & 11.6    & 12.1    & 6.6             & 13.0             & 16.2             & 17.2\\
-{\sc ZotBin } & 5.7 & 8.8    & 10.8    & 10.5    & 3.6             & 10.0             & --             & 12.4\\
-{\sc ZotNet } & 3.8 & 10.1    & 11.5    & 11.7    & 4.3             & 11.3             & 13.1             & 13.4\\
+\input{results/tables/table_FOM_ww_buzzard.tex}
+
 \end{tabular}

### 2021-02-08 paper/results/make_figures.py

@@ -29,6 +29,7 @@ method_names = [
     "NeuralNetwork1",
     "NeuralNetwork2",
     "PCACluster",
+    "PQNLD",
     "Random",
     "RandomForest",
     "SimpleSOM",
@@ -51,7 +52,7 @@ class EverythingIsNan:
         return np.nan
 
 
-def make_table(results, metric):
+def make_table(results, metric, filename):
     N = np.array([3, 5, 7, 9])
     row_format = r"{0} & {1} & {2}    & {3}    & {4}    & {5}             & {6}             & {7}             & {8}\\"
     data = {}
@@ -61,7 +62,7 @@ def make_table(results, metric):
         bins = row['bins']
         data[name, bands, bins] = row[metric]
 
-
+    f = open(filename, 'w')
     for name in method_names:
         disp_name = rf"{{\sc {name} }}".replace("_", r"\_")
         row = [disp_name]
@@ -75,7 +76,9 @@ def make_table(results, metric):
                 else:
                     row.append(f"{val:.1f}")
 
-        print(row_format.format(*row))
+        f.write(row_format.format(*row))
+        f.write("\n")
+    f.close()
 
 
 def load_table(base):
@@ -200,21 +203,18 @@ def plot_g_band_loss(dc2, buzzard, filename):
     fig.savefig(filename)
     plt.close(fig)
  
+def make_tex_tables(dc2, buzzard, dirname):
+    for name, data in [("dc2", dc2), ("buzzard", buzzard)]:
+        for metric in metrics:
+            fn = f'{dirname}/table_{metric}_{name}.tex'
+            print(fn)
+            make_table(data, metric, fn)
+
 
 if __name__ == '__main__':
     dc2 = load_table('cosmodc2')
     buzzard = load_table('buzzard')
+
     plot_metric_comparisons(dc2, buzzard, "metric_comparisons.pdf")
-    # plot_g_band_loss(dc2, buzzard, "g_band_loss.pdf")
-    # plot_g_band_loss(dc2, buzzard, "g_band_loss.pdf")
-
-    # make_table(dc2, 'FOM_DETF_3x2')
-
-    # pylab.scatter(dc2['bins'], dc2['FOM_DETF_3x2'])
-    # pylab.show()
-    #dc2.sort("FOM_ww")
-    # pylab.show()
-    # buzzard = load_table('buzzard')
-    # print(buzzard)
[…]

### 2021-02-09 tomo_challenge/classifiers/PQNLD.py
FFS

@@ -500,11 +500,12 @@ class PQNLD(Tomographer):
                            "z","g-r-(r-i)",
                            "r-i-(i-z)")
             # Filter columns AB/Vega zp_error zp_offset
-            columns = pd.DataFrame(["g_SDSS","r_SDSS","i_SDSS","z_SDSS"],
+            columns = pd.DataFrame(np.transpose(np.array([
+                                   ["g_SDSS","r_SDSS","i_SDSS","z_SDSS"],
                                    ["1,2","3,4","5,6","7,8"],
                                    ["AB","AB","AB","AB"],
                                    ["0.01","0.01","0.01","0.01"],
-                                   ["0.000001","0.000001","0.000001","0.000001"])
+                                   ["0.000001","0.000001","0.000001","0.000001"]])))
             #Output the columns file 
             columns.to_csv(validation_col_file,sep=' ',index=None,header=None,mode='w')
             columns = pd.DataFrame(np.transpose(
@@ -519,11 +520,12 @@ class PQNLD(Tomographer):
                            "z-y","z","g-r-(r-i)",
                            "r-i-(i-z)","i-z-(z-y)")
             # Filter columns AB/Vega zp_error zp_offset
-            columns = pd.DataFrame(["g_SDSS","r_SDSS","i_SDSS","z_SDSS","y_SDSS"],
+            columns = pd.DataFrame(np.transpose(np.array([
+                                   ["g_SDSS","r_SDSS","i_SDSS","z_SDSS","y_SDSS"],
                                    ["1,2","3,4","5,6","7,8","9,10"],
                                    ["AB","AB","AB","AB","AB"],
                                    ["0.01","0.01","0.01","0.01","0.01"],
-                                   ["0.000001","0.000001","0.000001","0.000001","0.000001"])
+                                   ["0.000001","0.000001","0.000001","0.000001","0.000001"]])))
             #Output the columns file 
             columns.to_csv(validation_col_file,sep=' ',index=None,header=None,mode='w')
             columns = pd.DataFrame(np.transpose(
@@ -538,11 +540,12 @@ class PQNLD(Tomographer):
                            "z","u-g-(g-r)","g-r-(r-i)",
                            "r-i-(i-z)")
             # Filter columns AB/Vega zp_error zp_offset
-            columns = pd.DataFrame(["u_SDSS","g_SDSS","r_SDSS","i_SDSS","z_SDSS"],
+            columns = pd.DataFrame(np.transpose(np.array([
[…]

### 2021-02-10 paper/results/make_figures.py
adding edge compare

@@ -128,7 +128,7 @@ def plot_metric_comparisons(dc2, buzzard, filename):
     color_plot(dc2, x, y, ax[0])
     ax[0].set_ylabel("Buzzard FOM")
     ax[0].set_ylim(0)
-    ax[0].legend(ncol=2, frameon=True)
+    ax[0].legend(frameon=True)
 
 
     x = dc2['FOM_DETF_3x2'].copy()

### 2021-02-11 bpz-1.99.3-py3/useful.py
tweak bpz old thing

@@ -248,7 +248,7 @@ def params_file(file):
         #     raise 'Keyword not string!'
         if len(pieces)<2:
             mensaje='No value(s) for parameter  '+key
-            raise mensaje
+            raise ValueError(mensaje)
         dict[key]=tuple(pieces[1:]) 
         if len(dict[key])==1: dict[key]=dict[key][0]
     return dict

### 2021-02-11 tomo_challenge/classifiers/lstm_flax.py
fixes to lstm

@@ -80,7 +80,7 @@ class Flax_LSTM(Tomographer):
         self.opt = options
         self.n_bins = options['bins']
         self.n_features = options['n_feats']
-        self.model = get_classifier(self.n_features, self.n_bins)
+        self.model = get_classifier(self.n_bins, self.n_features)#, self.n_bins)
         self.scaler = MinMaxScaler()
     
     def load_data(self, fname, take_colors=True, cutoff=0.0):
@@ -125,7 +125,7 @@ class Flax_LSTM(Tomographer):
          
     def train(self, training_data, training_z, batch_size=512, epochs=20):
         x_train = self.scaler.fit_transform(training_data)
-        x_train = np.expand_dims(x_train, axis=1)
+        x_train = np.expand_dims(x_train, axis=-1)
         lr = 0.001
         optimizer = optim.Adam(learning_rate=lr).create(self.model)
 
@@ -149,8 +149,9 @@ class Flax_LSTM(Tomographer):
 
         losses = []
         for e in range(epochs):
-            for i, (x_train, labels) in enumerate(get_batches().as_numpy_iterator()):
-                optimizer, loss = train_step(optimizer, x_train, labels)
+            print("Running training epoch ", e)
+            for i, (x_train1, labels) in enumerate(get_batches().as_numpy_iterator()):
+                optimizer, loss = train_step(optimizer, x_train1, labels)
                 losses.append(loss)
 
             print('Epoch {}\nLoss = {}'.format(e, loss))

### 2021-02-12 paper/paper.tex
adding edge compare
more paper text and content

@@
 
-\section{Results}
+\section{Results \& Discussion}
 
@@
 count, such that the spectrum could not be calculated.
+
+It is clear from values here that fears about restricted photometry catastrophically limiting
+the information needed for tomography are unfounded.  In both the CosmoDC2 and Buzzard analyses,
+even the $riz$ bands alone are sufficient to split objects into nine reasonably separated
+tomographic bins.  Once bins become this narrow the primary analysis concern will become the calibration of overall
+bin $n(z)$ values, rather than initial tomography, so we consider this to be reassuring.
 
@@
 \end{table*}
+
+\begin{figure}[htbp]
+% \centering
+% \includegraphics[width=0.9\linewidth]{results/metric_comparisons.pdf}
+\caption{Add $n(z)$ plots here for funbins.}
+\label{fig:metrics}
+\end{figure}
 
@@
 
+This is true in particular for the highest scoring methods, so we may assume the behaviour to be
+generic. The value of the extra FoM that could be gained should be weighed against 
+the challenge of high-accuracy
+PSF measurement for this band, and the cost in time and effort of determining it, 
+but until the end of the survey this is unlikely to be the limiting factor in overall precision.
+
 \begin{figure*}
@@
 and then used classifiers to assign objects to those bins.  Other methods optimized the values
-of bin edges themselves.
-
-\begin{itemize}
-    \item fisher matrix ellipses overplotted for top N methods
-    \item Plots of n(z)
-\end{itemize}
-
-\section{Discussion}
-It is clear from values above that fears about restricted photometry catastrophically limiting
-the information needed for tomography are unfounded.  In both the CosmoDC2 and Buzzard analyses,
-even the $riz$ bands alone are sufficient to split objects into nine reasonably separated
-tomographic bins.  Once bins become this narrow the primary analysis concern will become the calibration of overall
-bin $n(z)$ values, rather than initial tomography, so we consider this to be reassuring.
-
-The inclusion of the $g$ band typically increases metrics by around 10--20\%, including for the
-highest scoring methods.  This value should be weighed against the challenge of high-accuracy
-PSF measurement for this band, but until the end of the survey this is unlikely to be
-the limiting factor in overall precision.
-
-Methods using fixed bins with equal numbers by nominal redshift tend to plateau at scores of
+of bin edges themselves, by optimizing the metric.  For the runs shown here, these methods all
+targeted one of the 3x2pt metrics. \textbf{Check this is correct}.
+
+A comparison of trained-edge and fixed-edge methods on the CosmoDC2 griz scenarios
+is shown in Figure \ref{fig:edges}.  One important point to note is that while trained-edge
+methods dominate the highest scores for the 3x2pt DETF metric, on which they trained, they
+fare much worse on the lensing-only metrics (the methods were not re-trained on the lensing metrics;
+we show the metrics for the same assigned bins).  This suggests that the difference between
+optimal bins for different science cases is a significant one, and thus that analysis pipelines
+should make it straightforward to calibrate multiple configurations and pass them to
+subsequent science anlysis.
+
+\begin{figure*}
+\includegraphics[width=0.9\columnwidth]{results/edge_comparison.pdf}
+\caption{A comparison of results from methods which used fixed target edges (dashed blue)
+with those which optimized target bins (solid orange).  Methods were trained on the 3x2pt metric
+shown in the upper panel; the lower performance for these classifications 
+for the lensing-only metric shown in the lower panel illustrates the need for science case-specific
+binning choices. In each case, classification is done using griz bands on the CosmoDC2 data set.}
+\label{fig:edges}
+\end{figure*}
+
+Among methods with fixed bins, those using equal numbers in each tend to plateau at scores of
 120-140 in Table \ref{tab:cosmodc2}, suggesting that a range of machine learning models can
@@
 
-Several methods that instead try to optimize nominal bin edges break this
+Several of the methods with optimized target bin edges break this
 barrier and acheive very high scores.  This included ZotNet, JaxResNet, and JaxCNN.  The

### 2021-02-12 paper/results/make_figures.py

@@ -40,6 +40,18 @@ method_names = [
     "ZotNet",
 ]
 
+
+methods_with_trained_edges = [
+    "ComplexSOM",
+    "JaxCNN",
+    "JaxResNet",
+    "NeuralNetwork1",
+    "NeuralNetwork2",
+    "PCACluster",
+    "ZotBin",
+    "ZotNet",
+]
+
 metrics = ["SNR_ww", "SNR_gg", "SNR_3x2", "FOM_ww", "FOM_gg", "FOM_3x2", "FOM_DETF_ww", "FOM_DETF_gg", "FOM_DETF_3x2"]
 
 
@@ -203,6 +215,67 @@ def plot_g_band_loss(dc2, buzzard, filename):
     fig.savefig(filename)
     plt.close(fig)
  
+
+def plot_edge_type_comparison(data, filename):
+    _, data = pair_riz_griz(data)
+    fig, ax = plt.subplots(2, 1, figsize=(5,8), sharex=True)
+
+    untrained = [m for m in method_names if m not in methods_with_trained_edges]
+
+    def get(method):
+        s = data['method'] == method
+        m = data[s]
+        x = m['bins']
+        y = m['FOM_DETF_3x2']
+        z = m['FOM_DETF_ww']
+        y[y==0] = np.nan
+        z[z==0] = np.nan
+        return x, y, z
+
+
+    color =  '#1f77b4'
+    alpha = 0.5
+    lw = 1
+    style =  "--"
+    need_legend = True
+    for method in untrained:
+        x, y, z = get(method)
+        label = "Fixed edges" if need_legend else None
+        need_legend = False
+        ax[0].plot(x, y, color, alpha=alpha, lw=lw, ls=style, label=label)
+        ax[1].plot(x, z, color, alpha=alpha, lw=lw, ls=style)
+
+    color =  '#ff7f0e'
+    alpha = 1.0
+    lw = 3
+    style =  "-"
+    need_legend = True
+
+    for method in methods_with_trained_edges:
+        x, y, z = get(method)
+        label = "Trained edges" if need_legend else None
+        need_legend = False
+        ax[0].plot(x, y, color, alpha=alpha, lw=lw, ls=style, label=label)
+        ax[1].plot(x, z, color, alpha=alpha, lw=lw, ls=style)
+
+    ax[0].tick_params(which='minor', length=0, axis='x')
[…]

### 2021-02-12 paper/paper.tex
more in data section

@@
 
-\begin{itemize}
-    \item CosmoDC2 - description
-    \item CosmoDC2 - discussion
-    \item Noise properties
-    \item Pre-selection 
-    \item Metacal as a limitation
-\end{itemize}
+We run challenge entries on two data sets, CosmoDC2 and Buzzard, each of which is split into training, 
+validation, and testing components.  In each case participants were supplied with the training and 
+validation data, and the testing data was kept secret for final evaluation. In each case the three
+data sets were random sub-selections of the full data sets, with the training and testing sets 25\% of 
+the full data each and the validation 50\%. In practice many entrants trained on a subset of
+this, so to make the final evaluation phase tractable we trained on 1M objects for both data sets.
+
+Each of these data set come from cosmological simulations which provide truth values of galaxy magnitudes
+and sizes.  We simulate and add noise to the objects as described below.
+
+\subsubsection{CosmoDC2}
+
+The CosmoDC2 simulation was designed to support the Dark Energy Science
+Collaboration and is described in detail in \citet{cosmodc2}.  It covered 440
+deg${}^2$ of sky, and used the dark matter particles from the Outer Rim
+simulations \citep{outer_rim}.  The UniverseMachine \citep{universe_machine}
+simulations were then used, in a combination of empirical and semi-analytic
+methods, to assign galaxies with a limited set of specified properties to halos.
+
+These objects were then matched to outputs of the Galacticus model
+\citep{galacticus}, which generated complete galaxy properties and which was run
+on an additional DESC simulation, AlphaQ, and included all the LSST bands.  The
+simulation is complete to $r=28$, and contains objects up to redshift 3.  We
+include the ultra-faint galaxies in the simulation, not assigned to a specific
+halo.
+
+After the post-processing and selections described below, and a cut to a specific geographic region, 
+we used around 36M objects from the CosmoDC2 simulations.
+
+One limitation of the CosmoDC2 catalogs we used, found during the challenge,
+was that the matching to Galacticus led to only a limited number of SEDs being
+used at high redshifts, and thus to many galaxies in the simulations sharing
+similar characteristics at these redshifts, such as colour-colour relations.
+It was unknown whether this limitation would have any practical impact on the challenge,
+and was one of the reasons for adpoting the additional Buzzard catalog.
+
+\subsubsection{Buzzard}
+
+The Buzzard catalog was developed to support Dark Energy Survey Year 1 (DES-Y1)
+data analysis, and is described in detail in \citet{buzzard}.  It used dark
+matter simulations from L-GADGET2 \cite{lgadget} and then added galaxies using
+the {\sc AddGals} algorithm \citet{addgals}, which matches a set of galaxy
+property probability densities to data using sub-halo abundance matching. 
+Buzzard galaxies extend to around redshift 2.3, and were shown to be a good
+match (after appropriate selections) to DES-Y1 observable catalogs.  Magnitudes
+using the LSST band passes were provided in the catalog.
+
+After the post-processing and selections described below, and a cut to a
+specific geographic region,  we used around 20M objects from the Buzzard
+simulations.
+
+\subsubsection{Post-processing}
+
+We add noise to the truth values in the extra-galactic catalogs to simulate real observations.
+In each case we simulate observations corresponding to the first year Rubin observations
+using the DESC {\sc TXPipe} framework, following the methodology of \citet{ivezic_jones_lupton}
+and assuming the numerical values for telescope and sky characteristics therein.
+
+In both simulations we add two cuts to approximately simulate the selection used in a real
+lensing catalog.  In both cases we apply a cut on the combined $riz$ signal to noise of
+$S/N > 10$, and a size cut $T / T_\mathrm{psf} > 0.5$, where
+$T$ is the trace $I_{xx} + I_{yy}$ of the moments matrix and measures the squared radius of the
+galaxy compared to that of the PSF.
 
@@
 
-\subsection{ {\sc ENSEMBLE} }
-
 \subsection{ {\sc GPzBinning} }
@@
 
-\section{Overall results}
+\subsection{Overall results}
 The methods described above were run on a data set, previously unseen by entrants but drawn

### 2021-02-12 paper/results/make_figures.py

@@ -76,7 +76,7 @@ def make_table(results, metric, filename):
 
     f = open(filename, 'w')
     for name in method_names:
-        disp_name = rf"{{\sc {name} }}".replace("_", r"\_")
+        disp_name = rf"{{\sc {name} }}".replace("_", r"\_").replace("myCombinedClassifiers", "Stacked Generalization")
         row = [disp_name]
         for i, bands in enumerate(['riz', 'griz']):
             for n in N:

### 2021-02-12 paper/paper.tex
new plots and things

@@
 Participants tested and ran their methods locally, but submitted code to be run as part of a central
-combined analysis.  No prize was offered beyond recognition.
+combined analysis on the NERSC supercomputing facility.  No prize was offered apart from recognition.
 
@@
 
+We used the second Buzzard catalog for two reasons.  First, to check the significance of
+an issue with galaxy colour assignments in CosmoDC2, described below.  And second, to check
+for over-fitting in the design of models themselves, and thus to determine whether two
+reasonable but different galaxy populations can be optimized by the same methods; this will
+tell us whether our conclusions might be reasonable for real data. We discuss correlations
+between metrics in Section \ref{sec:metric-results}.
+
 Each of these data set come from cosmological simulations which provide truth values of galaxy magnitudes
@@
 simulations \citep{outer_rim}.  The UniverseMachine \citep{universe_machine}
-simulations were then used, in a combination of empirical and semi-analytic
+simulations were then used, with the GalSampler technique \citep{galsampler},
+in a combination of empirical and semi-analytic
 methods, to assign galaxies with a limited set of specified properties to halos.
@@
 was that the matching to Galacticus led to only a limited number of SEDs being
-used at high redshifts, and thus to many galaxies in the simulations sharing
+used at high redshifts, and thus too many galaxies in the simulations sharing
 similar characteristics at these redshifts, such as colour-colour relations.
@@
 
-Metric 2 uses a figure of merit (FOM) based on that presented by the Dark Energy Task Force \citep{detf}.  We 
+Metric 2 uses a figure of merit (FOM) based on that presented by the Dark Energy Task Force (DETF) \citep{detf}.  We 
 use the inverse area of a Fisher Matrix ellipse, which approximates the size in two axes 
@@
 either due to memory overflow or taking an extremely long time.
-The second, marked with a dash, is when a method ran, but generated at least one bin with an extremely small number
-count, such that the spectrum could not be calculated.
-
-It is clear from values here that fears about restricted photometry catastrophically limiting
-the information needed for tomography are unfounded.  In both the CosmoDC2 and Buzzard analyses,
-even the $riz$ bands alone are sufficient to split objects into nine reasonably separated
-tomographic bins.  Once bins become this narrow the primary analysis concern will become the calibration of overall
-bin $n(z)$ values, rather than initial tomography, so we consider this to be reassuring.
+The second, marked with a dash, is when a method ran, but generated at least one bin with an extremely 
+small number count, such that the spectrum could not be calculated.
 
@@
 
+Figure \ref{fig:funbin_nz} shows the number density obtained by the highest scoring method in
+the fiducial metric, {\sc funbins}, for both the riz and griz scenarios.
+It is clear from this that fears about restricted photometry catastrophically limiting
+the information needed for tomography are unfounded, at least in our idealised scenario.
+In both the CosmoDC2 and Buzzard analyses,
+even the $riz$ bands alone are sufficient to split objects into nine reasonably separated
+tomographic bins.  Once bins become this narrow the primary analysis concern will become the calibration 
+of overall bin $n(z)$ values, rather than initial tomography, so we consider this to be reassuring.
+
+There is a clear widening of the $n(z)$ when losing the g-band.  This widening leads to a loss in 
+constraining power, which is discussed more in Section \ref{sec:gband}.
+
 \begin{figure}[htbp]
 % \centering
-% \includegraphics[width=0.9\linewidth]{results/metric_comparisons.pdf}
-\caption{Add $n(z)$ plots here for funbins.}
-\label{fig:metrics}
+\includegraphics[width=1\linewidth]{results/funbins_nz.pdf}
+\caption{$n(z)$ values from funbins}
+\label{fig:funbin_nz}
 \end{figure}
 
-\subsection{Metric comparisons}
+\subsection{Metric comparisons} \label{sec:metric-results}
 Figure \ref{fig:metrics} shows a comparison of some of the different metrics used in the challenge.
+We plot, using assignments for methods and for all bin counts, the relationship between our fiducial
+metric, the DETF 3x2pt on CosmoDC2, and other metrics we measured.
+
+\begin{enumrate}
+\item Generally good relationship, especially at high end
+\item Domination of metric by clustering
+\item Bimodality in SNR
+\end{enumerate}
 
@@
 
-\subsection{Impact of losing g-band}
+\subsection{Impact of losing g-band} \label{sec:gband}
 Figure \ref{fig:loss} shows the ratio of the FOM between methods using the griz and the same method

### 2021-02-12 paper/results/make_figures.py

@@ -1,3 +1,4 @@
+import os
 import yaml
 import tabulate
 import numpy as np
@@ -5,10 +6,14 @@ from astropy.table import Table
 import matplotlib
 matplotlib.use('agg')
 import matplotlib.pyplot as plt
+import subprocess
 
 # set default styles
 plt.style.use('StyleSheet.mplstyle')
 
+
+blue_color =  '#1f77b4'
+orange_color =  '#ff7f0e'
 
 
 method_names = [
@@ -233,7 +238,7 @@ def plot_edge_type_comparison(data, filename):
         return x, y, z
 
 
-    color =  '#1f77b4'
+    color =  blue_color
     alpha = 0.5
     lw = 1
     style =  "--"
@@ -245,7 +250,7 @@ def plot_edge_type_comparison(data, filename):
         ax[0].plot(x, y, color, alpha=alpha, lw=lw, ls=style, label=label)
         ax[1].plot(x, z, color, alpha=alpha, lw=lw, ls=style)
 
-    color =  '#ff7f0e'
+    color =  orange_color
     alpha = 1.0
     lw = 3
     style =  "-"
@@ -276,6 +281,47 @@ def plot_edge_type_comparison(data, filename):
     plt.close(fig)
 
 
+def plot_funbins_nz(filename):
+    riz_file = "./cosmodc2/bins/funbins_9_riz_0.npy"
+    griz_file = "./cosmodc2/bins/funbins_9_griz_0.npy"
+    z_file = "./cosmodc2/bins/z.npz"
+
+
+    if not os.path.exists(riz_file):
+        print("Data not downloaded for n(z) plot - not overwriting")
+        return
+
+    fig, ax = plt.subplots(2, 1, figsize=(6,6), sharex=True, sharey=True)
+    colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
+
+    riz = np.load(riz_file)
+    griz = np.load(griz_file)
+    z = np.load(z_file)['arr_0']
+
+
+    nbin = 9
+
+    for i in range(nbin):
+        w = np.where(griz == i)
+        weight = np.repeat(1e-5, w[0].size)
+        ax[0].hist(z[w], bins=50, histtype='step', ls='-', color=colors[i], weights=weight, lw=3)
+
+    for i in range(nbin):
+        w = np.where(riz == i)
+        weight = np.repeat(1e-5, w[0].size)
[…]

### 2021-02-17 paper/paper.tex
decribe infrastructure

@@
 \subsection{Infrastructure}
-\begin{itemize}
-    \item Pull request submissions
-    \item Data at NERSC
-\end{itemize}
+
+The challenge was structured as a python library, in which each classification
+method was a class.  The abstract base parent class, {\sc Tomographer}, 
+and other challenge machinery, performed initialization,
+configuration option handling, data loading, and computing derived quantities
+such as colours from magnitudes.
+
+Participants were expected to subclass
+{\sc Tomographer} and implement two methods, {\sc train} and {\sc apply}.  Each
+was required to accept as an argument an array or dictionary of galaxy properties
+(magnitudes, and, optionally, sizes, errors, and colours).  The {\sc train} method
+was also passed an array of true redshift values, which could be used however they wished.
+Methods were submitted to the challenge in the form of a pull request to the challenge
+repostitory\footnote{\url{http://github.com/LSSTDESC/tomo_challenge}}.
+
+The training and validation parts of the data set were made available for training
+and hyper-parameter tuning. Once the challenge period was complete, the algorithms
+were collected from the pull requests and run at the National Energy Research
+Supercomputing Center (NERSC). If a method failed due to a difference in the
+runtime environment or the hardware at NERSC, the participants and organizers tried
+to amend it after the formal challenge period, though this was not always possible.
 
@@
 never be applied to real data and should therefore considered a form of control method.
-This method was {\sc Utopia}, which found, for each object in the testing data, the
+This method was {\sc Utopia}, which, as described below, found, for each object in the testing data, the
 nearest neighbour object in the training sample.  The high density, completeness, and

### 2021-02-17 paper/paper.tex
fixes
new text and placeholder plots

@@
 metacalibration method for galaxy shape measurement to limit biases.  Tomographic methods have been 
-explored before in \citet{jain} and \citet{kitching2019}.  In this paper we describe the results of a 
+explored before in \citet{jain} and \citet{kitching2019} \textbf{more}.  In this paper we describe the results of a 
 challenge using simulated Rubin-like data with a range of methods submitted by different teams.
@@
 galaxy compared to that of the PSF.
+
+The overall number density of the two data sets is shown in Figure \ref{fig:initial_nz},
+and a set of colour-colour diagrams  in Figure \ref{fig:colour_colour}.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=1\linewidth]{results/initial_data.pdf}
+\caption{The underlying number density $n(z)$ for the two catalogs used in the challenge,
+after the initial cuts described in the text.}
+\label{fig:initial_nz}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=1\linewidth]{results/colour_colour.pdf}
+\caption{A colour-colour diagram for the two catalogs used in the challenge, showing the bands
+available to participants.}
+\label{fig:colour_colour}
+\end{figure}
 
@@
 neurons,  Dropout rates, maxpooling layers, the number of convolutional filters, and strides.
+
+A technical problem, involving conflicting versions of libraries and incompatible hardware,
+meant that method could not be run in the final challenge.
 
@@
 \includegraphics[width=1\linewidth]{results/funbins_nz.pdf}
-\caption{$n(z)$ values from funbins}
+\caption{The tomographic $n(z)$ values generated by the winning method in the challenge, {\sc funbins},
+generating nine bins. 
+The upper panel shows results for the method applied to riz bands, and the lower on griz. The
+increased spread and overlap of the bins in the lower panel decreases the overall constraining
+power and signal-to-noise.}
 \label{fig:funbin_nz}
@@
 
-\begin{enumrate}
-\item Generally good relationship, especially at high end
-\item Domination of metric by clustering
-\item Bimodality in SNR
-\end{enumerate}
+The different metrics are generally well-correlated (this also holds true for the metrics not 
+shown here), especially at the highest values of the metrics representing the most successful
+methods. This is particularly so when comparing the Buzzard and CosmoDC2 metrics, which is encouraging 
+as it implies these methods are broadly stable to reasonable changes in the underlying galaxy
+distribution, and hence on the survey properties.  The largest outlier comes from the 
+{\sc Neural Network} entry.
+
+The relatively broader spread in the lensing-only metric illustrates how our metric is dominated
+by the stronger constraining power in the clustering part of the data set.  This arises because
+we do not include a galaxy bias model in our FOM, and so the constraining power of the clustering
+is artifically high.
+However, since the clustering metric then largely becomes a proxy for how well separated
+galaxy bins are, this is not a major limitation of the metrics.
+
+Finally, there is a bimodal structure in the relationship between the FOM and SNR metrics.
+This largely traces the split between models which train the target bin edges vs those
+which fix them, as described below in Section \ref{sec:train-vs-fix}.
 
@@
 
-\subsection{Trained edges vs fixed edges}
+\subsection{Trained edges vs fixed edges}\label{sec:train-vs-fix}
 
@@
 
-\begin{figure*}
+\begin{figure}
 \includegraphics[width=0.9\columnwidth]{results/edge_comparison.pdf}
@@
 \label{fig:edges}
-\end{figure*}
+\end{figure}
 
@@
 
-\subsection{Future directions}
-
-\begin{itemize}
-    \item spectroscopic incompleteness / unrepresentativeness
-    \item data set volume
-    \item comparison of optimal choices in different science cases
-\end{itemize}
+\section{Conclusions}
+The Tomography Challenge has proved a useful mechanism for encouraging the development
+of methods for this science question.  We encourage this challenge structure for other
+well-specified science cases, and DESC has subsequently run other challenges along the same lines
+such is its N5K beyond-limber challege\footnote{\url{https://github.com/LSSTDESC/N5K}}.
+We describe in Appendix \ref{sec:mistakes} some lessons we have learned from this challenge
+that may be useful for future ones.
+
+The results of the challenge have shown first that three or four-band photometry
+is sufficient for specifying tomography even up to a large number of bins, if
+the training data is complete enough: degneracies between colour do not make this impossible.
+Excluding the g-band from the data reduces the constraining power by a noticeable but
+not catastrophic amount.
+
+Finally, we have shown that in general there is a signficant advantage to optimizing
+target bins afresh when considering a new science case; this can be a significant boost
+to the final constraining power.
+Despite this, evenly spaced bins in fiducial co-moving distance have proven
+a good general choice in the winning method in the challenge, {\sc funbins}.
+
+The challenge was heavily simplified, with the intention of testing whether
+effective bin-assignments with limited photometry is possible even in theory,
+and of generating as many potential methods for the task as possible.  With
+both these acheived, future directions for simulating the problem are clear.
+The most obvious is limited spectroscopic completeness and size.  Realistic
+training data sets will be far smaller than the millions of objects we used here,
+and the difficulty of measuring spectra for faint galaxies will make the them
+incomplete and highly unrepresentative. Machine learning methods will be particularly
+affected by the latter.  Future challenges must explore these important limitations
+using simulated incompleteness.
 
@@
 
-\section{Mistakes}
+\section{Mistakes} \label{sec:mistakes}
 The challenge organizers made a number of mistakes when building and running the challenge.

### 2021-02-17 paper/results/make_figures.py

@@ -4,12 +4,9 @@ import tabulate
 import numpy as np
 from astropy.table import Table
 import matplotlib
-matplotlib.use('agg')
-import matplotlib.pyplot as plt
 import subprocess
 
 # set default styles
-plt.style.use('StyleSheet.mplstyle')
 
 
 blue_color =  '#1f77b4'
@@ -320,6 +317,37 @@ def plot_funbins_nz(filename):
     fig.savefig(filename)
     plt.close(fig)
 
+def make_initial_nz(filename):
+    dc2_file = "dc2_validation.hdf5"
+    buzzard_file = "buzzard_validation.hdf5"
+    if not os.path.exists(buzzard_file):
+        print("Data not downloaded for colour-colour plot - not overwriting")
+        print("Download from https://portal.nersc.gov/cfs/lsst/txpipe/tomo_challenge_data/ugrizy_buzzard/validation.hdf5")
+        print("and https://portal.nersc.gov/cfs/lsst/txpipe/tomo_challenge_data/griz/validation.hdf5")
+        return
+
+    #buzz = h5py.File(buzzard_file)
+    #dc2 = h5py.File(dc2_file)
+    buzz_z = buzz['redshift_true'][:]
+    fig, ax = plt.subplots(figsize=(6,6))
+    ax.hist(buzz_z, bins=100, histtype='step', label='Buzzard')
+    del buzz_z
+    dc2_z = dc2['redshift_true'][:]
+    ax.hist(dx2_z, bins=100, histtype='step', label='CosmoDC2')
+    fig.tight_layout()
+    fig.savefig(filename)
+    plt.close(fig)
+
+def make_colour_colour(filename):
+    dc2_file = "dc2_validation.hdf5"
+    buzzard_file = "buzzard_validation.hdf5"
+    if not os.path.exists(buzzard_file):
+        print("Data not downloaded for colour-colour plot - not overwriting")
+        print("Download from https://portal.nersc.gov/cfs/lsst/txpipe/tomo_challenge_data/ugrizy_buzzard/validation.hdf5")
+        print("and https://portal.nersc.gov/cfs/lsst/txpipe/tomo_challenge_data/griz/validation.hdf5")
+        return
+    
+
 
 
 def make_tex_tables(dc2, buzzard, dirname):
@@ -331,6 +359,10 @@ def make_tex_tables(dc2, buzzard, dirname):
 
 
 if __name__ == '__main__':
[…]

### 2021-02-17 paper/results/make_figures.py
plotting script

@@ -328,15 +328,39 @@ def make_initial_nz(filename):
 
     #buzz = h5py.File(buzzard_file)
     #dc2 = h5py.File(dc2_file)
+    buzz = h5py.File(buzzard_file)
+    dc2 = h5py.File(dc2_file)
     buzz_z = buzz['redshift_true'][:]
     fig, ax = plt.subplots(figsize=(6,6))
     ax.hist(buzz_z, bins=100, histtype='step', label='Buzzard')
     del buzz_z
     dc2_z = dc2['redshift_true'][:]
     ax.hist(dx2_z, bins=100, histtype='step', label='CosmoDC2')
+    ax.legend(frameon=True)
     fig.tight_layout()
     fig.savefig(filename)
     plt.close(fig)
+
+
+def colour_colour_cat(ax, cat, colour, thin, label):
+    g = cat['g_mag'][:][::thin]
+    r = cat['r_mag'][:][::thin]
+    i = cat['i_mag'][:][::thin]
+    z = cat['z_mag'][:][::thin]
+
+    gr = g - r
+    ri = r - i
+    gz = g - z
+    iz = i - z
+
+    ax[0].scatter(gr, ri, c=colour, s=1, label=label)
+    ax[1].scatter(gz, iz, c=colour, s=1, label=label)
+
+    ax[0].set_xlabel("g - r")
+    ax[0].set_ylabel("r - i")
+    ax[1].set_xlabel("g - z")
+    ax[1].set_ylabel("i - z")
+
 
 def make_colour_colour(filename):
     dc2_file = "dc2_validation.hdf5"
@@ -346,8 +370,21 @@ def make_colour_colour(filename):
         print("Download from https://portal.nersc.gov/cfs/lsst/txpipe/tomo_challenge_data/ugrizy_buzzard/validation.hdf5")
         print("and https://portal.nersc.gov/cfs/lsst/txpipe/tomo_challenge_data/griz/validation.hdf5")
         return
-    
 
+
+    fig, ax = plt.subplots(2, 1, figsize=(4, 8))
+
+    buzz = h5py.File(buzzard_file)
+    dc2 = h5py.File(dc2_file)
+
+    colour_colour_cat(ax, buzz, blue_color, 1000, "Buzzard")
+    colour_colour_cat(ax, buzz, orange_color, 1000, "CosmoDC2")
+
+    ax[0].legend(frameon=True)
+    
+    fig.tight_layout()
[…]

### 2021-02-17 paper/results/make_figures.py
add colour-colour and nz plots

@@ -5,6 +5,7 @@ import numpy as np
 from astropy.table import Table
 import matplotlib
 import subprocess
+import h5py
 
 # set default styles
 
@@ -328,14 +329,19 @@ def make_initial_nz(filename):
 
     #buzz = h5py.File(buzzard_file)
     #dc2 = h5py.File(dc2_file)
-    buzz = h5py.File(buzzard_file)
-    dc2 = h5py.File(dc2_file)
+    buzz = h5py.File(buzzard_file, 'r')
+    dc2 = h5py.File(dc2_file, 'r')
     buzz_z = buzz['redshift_true'][:]
-    fig, ax = plt.subplots(figsize=(6,6))
-    ax.hist(buzz_z, bins=100, histtype='step', label='Buzzard')
+    dc2_z = dc2['redshift_true'][:]
+    fig, ax = plt.subplots(figsize=(5,5))
+    w1 = np.repeat(1e-5, buzz_z.size)
+    w2 = np.repeat(1e-5, dc2_z.size)
+    ax.hist(buzz_z, bins=50, histtype='step', label='Buzzard', linewidth=3, weights=w1)
     del buzz_z
     dc2_z = dc2['redshift_true'][:]
-    ax.hist(dx2_z, bins=100, histtype='step', label='CosmoDC2')
+    ax.hist(dc2_z, bins=50, histtype='step', label='CosmoDC2', linewidth=3, weights=w2)
+    ax.set_xlabel("z")
+    ax.set_ylabel("Counts / $10^5$")
     ax.legend(frameon=True)
     fig.tight_layout()
     fig.savefig(filename)
@@ -343,18 +349,22 @@ def make_initial_nz(filename):
 
 
 def colour_colour_cat(ax, cat, colour, thin, label):
-    g = cat['g_mag'][:][::thin]
-    r = cat['r_mag'][:][::thin]
-    i = cat['i_mag'][:][::thin]
-    z = cat['z_mag'][:][::thin]
+    if 'g_mag' in cat.keys():
+        fmt = '{}_mag'
+    else:
+        fmt = 'mcal_mag_{}'
+    g = cat[fmt.format('g')][:][::thin]
+    r = cat[fmt.format('r')][:][::thin]
+    i = cat[fmt.format('i')][:][::thin]
+    z = cat[fmt.format('z')][:][::thin]
 
     gr = g - r
     ri = r - i
     gz = g - z
     iz = i - z
 
[…]

### 2021-02-17 paper/paper.tex
new plots

@@
 property probability densities to data using sub-halo abundance matching. 
-Buzzard galaxies extend to around redshift 2.3, and were shown to be a good
+Buzzard galaxies extend to around redshift 2.3 (significantly shallower than CosmoDC2)
+and were shown to be a good
 match (after appropriate selections) to DES-Y1 observable catalogs.  Magnitudes
@@
 \begin{figure}[htbp]
-\centering
-\includegraphics[width=1\linewidth]{results/initial_data.pdf}
+% \centering
+\includegraphics[width=0.9\linewidth]{results/initial_data.pdf}
 \caption{The underlying number density $n(z)$ for the two catalogs used in the challenge,
@@
 \begin{figure}[htbp]
-\centering
-\includegraphics[width=1\linewidth]{results/colour_colour.pdf}
+% \centering
+\includegraphics[width=0.9\linewidth]{results/colour_colour.pdf}
 \caption{A colour-colour diagram for the two catalogs used in the challenge, showing the bands

### 2021-02-17 paper/results/make_figures.py
squah figs

@@ -333,7 +333,7 @@ def make_initial_nz(filename):
     dc2 = h5py.File(dc2_file, 'r')
     buzz_z = buzz['redshift_true'][:]
     dc2_z = dc2['redshift_true'][:]
-    fig, ax = plt.subplots(figsize=(5,5))
+    fig, ax = plt.subplots(figsize=(5,4))
     w1 = np.repeat(1e-5, buzz_z.size)
     w2 = np.repeat(1e-5, dc2_z.size)
     ax.hist(buzz_z, bins=50, histtype='step', label='Buzzard', linewidth=3, weights=w1)
@@ -382,7 +382,7 @@ def make_colour_colour(filename):
         return
 
 
-    fig, ax = plt.subplots(2, 1, figsize=(4, 8))
+    fig, ax = plt.subplots(2, 1, figsize=(4, 6))
 
     buzz = h5py.File(buzzard_file, 'r')
     dc2 = h5py.File(dc2_file, 'r')

### 2021-02-22 paper/paper.tex
clarify caption
add text from DK

@@
 
-\subsection{ {\sc ZotBin} }
-
-\subsection{ {\sc ZotNet} }
+\subsection{ {\sc ZotBin and ZotNet} }
+
+This pair of related methods was submitted by author DPK and its associated code is available at
+\url{https://github.com/dkirkby/zotbin}.
+
+The ZotBin method consists of three stages: preprocessing, grouping, and bin optimization.  The ZotNet method
+uses the same preprocessing, then skips directly to bin optimization.  Both methods optimize the final bins to
+maximize an arbitrary linear combination of the three metrics defined in \ref{sec:metrics}, using the JAX
+library~\cite{jax} for efficient computation. The goal of ZotNet is to find a nearly optimal set of bins (for
+a specified linear combination of metrics) using a neural network, at the expense of interpretability of the
+results.  The ZotBin method aims to perform almost as well as ZotNet, but with a much more interpretable mapping
+from input features to final bins.
+
+Both methods transform the $n$ input fluxes into $n-1$ colors and one flux. The data are very sparse in this
+$n$-dimensional feature space and exhibit complex correlations. The preprocessing step transforms to a new
+$n$-dimensional space where the data is nearly uncorrelated and dense on the unit hypercube $[0,1]^n$. This is accomplished
+by learning a normalizing flow~\cite{2019arXiv191202762P} that maps the complex input probability density
+into a multivariate unit Gaussian, then applying the error function to yield a uniform distribution. Although
+the resulting transformation is non-linear, it is invertible and differentiable, and thus provides an interpretable
+smooth mapping.
+
+The second stage of ZotBin starts by dividing the unit hypercube $[0,1]^n$ from the preprocessing step
+into a regular lattice of ${\cal O}$(10K) cells which, by construction, each contain a similar number
+of training samples.  Next, adjacent cells are iteratively merged into $M \sim 100$ groups of cells by
+combining at each step the pair of groups that are most ``similar''.  We define similarity as the product
+of similarities in feature and redshift space.  Redshift similarity is based on the histogram of redshifts
+associated with each group, interpreted as a vector.
+
+The final stage of ZotBin selects linear combinations of the $M$ groups to form the $N$ output bins. This is
+accomplished by optimizing the specified metric combination with respect to the $M\times N$ matrix of linear
+weights.
+
+The ZotBin method defines a fully invertible sequence of steps that transform the input feature space into
+the final bins, so that each bin is associated with a well-defined region in color / flux space. The ZotNet method,
+on the other hand, trains a neural network to directly map from the preprocessed unit hypercube $[0,1]^n$ into
+$N$ output bins by optimizing the specified metric combination. The resulting map is not invertible, so less
+interpretable, but also less constrained than ZotBin, so expected to achieve somewhat better performance.
 
@@
 \caption{The tomographic $n(z)$ values generated by the winning method in the challenge, {\sc funbins},
-generating nine bins. 
+generating nine bins with the CosmoDC2 data. 
 The upper panel shows results for the method applied to riz bands, and the lower on griz. The

### 2021-03-03 paper/paper.tex
Added firt pass as NN method

@@
 
-\subsection{ {\sc NeuralNetwork 1} }
-
-\subsection{ {\sc NeuralNetwork 2} }
+\subsection{ {\sc NeuralNetwork 1 \& 2} }
+
+The {\sc NeuralNetwork} approach combines the two following ideas:
+\begin{itemize}
+	\item Use a simple neural network to parameterize a bin assignment function $f_\theta$ taking available photometry as the input and returning a bin number. 
+	\item Optimize the parameters of $f_\theta$ as to maximize a target score, either the total SNR \autoref{eq:snr}, or the DETF FoM \autoref{eq:fom}.
+\end{itemize}
+The interest of this approach is that it actually bypasses the need for a photometric redshift estimate in allocating a galaxy to a bin, and directly solves the problem of optimal bin assignment given available photometry, as to maximize the cosmological information.
+
+Details of the particular neural network are mostly irrelevant, the main difficulty in this approach is to be able to compute the back-propagation of gradients through the cosmological loss function as to update the parameters $\theta$ of the bin assignment function.
+This is was made possible for this challenge by the \texttt{jax-cosmo} library \citep{jax-cosmo}, which allows the computation of both metrics using the JAX automatic differentiation framework \citep{jax2018github}.
+
+In practice, to parameterize $f_\theta$ we use a simple dense neural network, composed of 3 linear layers of size 500 with leaky\_relu activation function and output batch normalization.
+The last layer of the model is a linear layer with an output size matching the number of bins, and softmax activation.
+We implement this model using the JAX-based Flax neural network library \citep{flax2020github}.
+
+Training is performed using batches of 5000 galaxies, over 2000 iterations under the ADAM optimizer \citep{adam} with a base learning rate of 0.001.
+The loss function $\mathcal{L}$ used for training directly derive from the challenge metrics, and constitutes the only difference between the NeuralNetwork 1 \& NeuralNetwork 2 entries:
+
+\begin{itemize}
+	\item NeuralNetwork 1: $\mathcal{L}_1 = 1 / S_{FoM}$
+	\item NeuralNetwork 2: $\mathcal{L}_2 = - S_{SNR} $
+\end{itemize}
+
+%\subsection{ {\sc NeuralNetwork 2} }
 

### 2021-03-04 paper/paper.tex
add references

@@
 Weak gravitational lensing (WL) has emerged over the last decade as a powerful
-cosmological probe \citep{cfhtlens,des,kids,hsc}.  WL
+cosmological probe \citep{cfhtlens,rcslens,des,kids,hsc}.  WL
 uses measurements of coherent shear distortion to the observed shapes of galaxies
@@
 model galaxy bias, intrinsic alignments, and other systematic errors en masse in a more tractable way.
-While fully 3D methods have been proposed and prototyped \citep{kitching,more}, the tomographic 
-approach remains the standard within the field.
+While fully 3D methods have been proposed, prototyped, and shown to have significant promise, 
+\citep{heavens,kitching}, the tomographic approach remains the standard within the field.
 
 Tomographic 3x2pt measurements will be a key science goal in the upcoming \emph{Stage IV} surveys,
-including the Rubin Observatory \citep{rubin} and the Euclid and Roman space telescopes \citep{euclid,roman}.
-
-In this paper we discuss and evaluate methods for assigning objects to tomographic bins, and 
-specifically the impact
+including the Rubin Observatory \citep{rubin} and the Euclid and Roman space telescopes
+\citep{euclid,roman}. The general question of optimizing tomography was recently discussed in detail
+in \citet{rainbow} using a self-organizing map approach to target up to five tomographic bins.  For this
+configuration they find that equally spaced redshift bins are a good proxy
+for an optimized method, but note, importantly, that we should not be constrained to trying to directly
+match our bin selections to specific tomographic bin edges.
+
+In this paper we discuss and evaluate the related challenge
 of using a limited set of color bands to make those assignments, motivated by requirements of using the
-metacalibration method for galaxy shape measurement to limit biases.  Tomographic methods have been 
-explored before in \citet{jain} and \citet{kitching2019} \textbf{more}.  In this paper we describe the results of a 
+metacalibration method for galaxy shape measurement to limit biases.  
+A similar question, with different motivations, was explored in \citet{jain}, who find that 3-band 
+gri tomography could be effective for tomographic selection, a result we will echo here in a new context: in this paper we describe the results of a 
 challenge using simulated Rubin-like data with a range of methods submitted by different teams.
@@
 sheared variants of the catalog, and measuring the difference between the post-cut shear in the two
-variants.  In DES the corrected biases were found up to around 3\%, far larger than
+variants.  In DES the corrected biases were found up to around 4\%, far larger than
 the requirements for the current generation of surveys \citep{des-y1-cat}.
@@
 data analysis, and is described in detail in \citet{buzzard}.  It used dark
-matter simulations from L-GADGET2 \cite{lgadget} and then added galaxies using
-the {\sc AddGals} algorithm \citet{addgals}, which matches a set of galaxy
+matter simulations from L-GADGET2 \cite{gadget2} and then added galaxies using
+an algorithm that matches a set of galaxy
 property probability densities to data using sub-halo abundance matching. 
@@
 bin, in turn, and $C$ a Gaussian estimaate of the covariance between them, as described in
-\citet{wl-review} and summarized in \autoref{app:theory}.   We compute this metric for lensing alone, clustering alone, and the full 3x2pt combination.
+\citet{takada_jain} and summarized in \autoref{app:theory}.   We compute this metric for lensing alone, clustering alone, and the full 3x2pt combination.
 
@@
 receptive field of the TCN can be increased without increasing the number of convolutional filters.
-Furthermore, residual blocks \citep{he} are also used to stabilize deeper and larger networks. These
+Furthermore, residual blocks \citep{resnet} are also used to stabilize deeper and larger networks. These
 modifications cause TCNs to have a longer memory than traditional RNNs with the same capacity,  and
@@
 Details of the particular neural network are mostly irrelevant, the main difficulty in this approach is to be able to compute the back-propagation of gradients through the cosmological loss function as to update the parameters $\theta$ of the bin assignment function.
-This is was made possible for this challenge by the \texttt{jax-cosmo} library \citep{jax-cosmo}, which allows the computation of both metrics using the JAX automatic differentiation framework \citep{jax2018github}.
+This is was made possible for this challenge by the \texttt{jax-cosmo} library \citep{jax-cosmo}, which allows the computation of both metrics using the JAX automatic differentiation framework \citep{jax}.
 

### 2021-03-10 tomo_challenge/classifiers/PQNLD.py
more tweaks

@@ -126,11 +126,12 @@ class PQNLD(Tomographer):
           true redshift for the training sample
 
         """
-        remove(ascii_file)
-        remove(bpz_file)
-        remove(training_col_file)
-        remove(training_cat_file)
-        remove(validation_col_file)
+        #remove(ascii_file)
+        print("NOT REMOVING FILES")
+        #remove(bpz_file)
+        #remove(training_col_file)
+        #remove(training_cat_file)
+        #remove(validation_col_file)
         print("Initialising")
         #Number of tomographic bins 
         n_bin = self.opt['bins']
@@ -281,7 +282,9 @@ class PQNLD(Tomographer):
         if not os.path.exists("bpz-1.99.3-py3/"): 
             os.system("bash INSTALL.sh")
         
-        if not os.path.exists(bpz_file): 
+        if os.path.exists(bpz_file):
+            print("SKPPING BPZ RUN")
+        else:
             print("Running BPZ on the training data")
             curdir=os.getcwd()
             os.chdir('bpz-1.99.3-py3')

### 2021-03-10 tomo_challenge/classifiers/__init__.py

@@ -41,8 +41,8 @@ else:
             gpus = tf.config.experimental.list_physical_devices('GPU')
             for gpu in gpus:
                 if tf_mem:
-                    tf.config.experimental.set_memory_growth(gpu, True)
-                    print("Setting memory_growth=True")
+#                    tf.config.experimental.set_memory_growth(gpu, True)
+                    print("NOT Setting memory_growth=True")
         except AttributeError:
             gpus = []
 

### 2021-03-10 tomo_challenge/classifiers/lstm_flax.py

@@ -26,6 +26,9 @@ import jax.random as rand
 import jax.numpy as jnp
 import jax
 from .. import jax_metrics as j_metrics
+import pickle, os
+array_id = os.environ.get("SLURM_ARRAY_TASK_ID", "666")
+model_file =  f"lstm_{array_id}.pickle"
 
 def get_classifier(n_bin, n_features):
     class BinningNN(nn.Module):
@@ -123,7 +126,10 @@ class Flax_LSTM(Tomographer):
 
 
          
-    def train(self, training_data, training_z, batch_size=512, epochs=20):
+    def train(self, training_data, training_z, batch_size=512, epochs=1):
+        print(training_data.shape)
+        training_data = training_data[::50]
+        training_z = training_z[::50]
         x_train = self.scaler.fit_transform(training_data)
         x_train = np.expand_dims(x_train, axis=-1)
         lr = 0.001
@@ -159,6 +165,8 @@ class Flax_LSTM(Tomographer):
    
         self.model = optimizer.target
 
+#        with open(model_file, "wb") as outfile:
+#            picklae.dump(self.model, outfile)
 
     def apply (self, data):
         """Applies training to the data.

### 2021-03-10 paper/paper.tex
add JaxCNN and JaxResNet

@@
 
-\subsection{ {\sc CNN} } 
+\subsection{ {\sc CNN} } \label{sec:cnn}
 This method uses an optimized Convolutional Neural Network (CNN) to assign
@@
 
-\subsection{ {\sc JaxCNN} }
-
-\subsection{ {\sc JaxResNet} }
+\subsection{{\sc JaxCNN \& JaxResNet} }
+
+This methi submitted by the author AP.
+
+JaxCNN and JaxResNet are a convolutional neural network based algorithm that
+maps the relationship between the color-magnitude data of galaxies to their bin
+indices. Convolutional neural network (CNN) models, as described in section
+\ref{sec:cnn}, consist of several layers, each consisting of a linear
+convolution operator followed by polling layers and a nonlinear activation
+function. Residual Network (ResNet) models take one step further and learn the
+\textit{identity} mapping by skipping, or adding \textit{shortcut} connections
+to, one or more layers \citep{resnet}. ResNet helps solve the vanishing gradient
+problem, hence we can have a deeper neural network.
+
+JaxCNN uses 2-layer convolutional neural networks and JaxResNet uses a stack of
+3 layers consisting of $1 \times 1, 3 \times 3$, and $1 \times 1$ convolutions
+with shortcut connections added to each stack. Both methods use the Rectified
+Linear Unit (ReLU) activation function under the ADAM optimizer \cite{adam} with
+a learning rate of 0.001.
+
+We chose the learning algorithm for both methods to minimize the reciprocal of
+the figure of merit $F$, where $F$ is defined in eqref{E:FOM}. We implement
+these models using the JAX-based Flax neural network library \cite{jax}.
 
@@
 \subsection{ {\sc funbins} }
-This method uses the random forest algorithm described in section \ref{randomforest} to assign galaxies to tomographic bins.
+This method uses the random forest algorithm described in section \ref{sec:randomforest} to assign galaxies to tomographic bins.
 There are 3 options for determining the bins used to assign labels to the training data - \textit{log}, \textit{random}, and \textit{chi}.
@@
 
-Once the bin edges are determined by either of the methods above, the random forest is trained as described in section \ref{randomforest}.
+Once the bin edges are determined by either of the methods above, the random forest is trained as described in section \ref{sec:randomforest}.
 

### 2021-03-10 paper/paper.tex
completed run through

@@
-\PassOptionsToPackage{usenames,dvipsnames}{xcolor}
+\PassOptionsToPackage{usenames,dvipsnames}{xcolour}
 \documentclass[twocolumn,twocolappendix]{aastex63}
@@
 \usepackage{url}
+\usepackage{mdwlist}
 \usepackage{listings}
@@
 \shorttitle{Tomographic binning optimization}
-\shortauthors{Author A et al.\ (LSST~DESC)}
+\shortauthors{Zuntz et al.\ (LSST~DESC)}
 
@@
 In this paper we present the results of the DESC 3x2pt tomography challenge,
-which aims to compare strategies for optimizing the photometric binning required
+which aimed to compare strategies for optimizing the photometric binning required
 for the main LSST 3x2pt analysis. This task is made particularly delicate in the
-context of a metacalibrated lensing survey, as only the photometry from the
+context of a metacalibrated lensing catalog, as only the photometry from the
 bands  included in the metacalibration process (riz and potentially g) can be
 used to define this tomographic binning. Using a set of realistic simulations
-including true galaxy redshifts and simulated photometry, the goal of the
-challenge is to propose a bin assignment strategy that maximizes the overall
-Figure of Merit of a 3x2pt analysis. This setting is idealized as it ignores
-spectroscopic completeness  issues for the training sample, but complex enough
+generating galaxy redshifts and simulated photometry, the goal of the
+challenge was to propose bin assignment strategies that maximized various
+figures of merit for a 3x2pt analysis. This challenge was highly idealized: it ignored
+spectroscopic completeness and limited size for the training sample, but was complex enough
 to address the main question of binning optimization. Based on the challenge
@@
 various methods are able to yield binning schemes that achieve high figure-of-merit scores.
+We further find that the g band adds around 15\% to typical figures of merit, and
+that there is significant value in optimizing different bin assignments for specific
+science cases.
 \end{abstract}
@@
 
-The \emph{3x2pt} method has become the standard method for performing this combination.
+The \emph{3x2pt} method has become a standard method for performing this combination.
 In this method, two-point correlations are computed among and between two samples, the shapes of 
@@
 large uncertainties\footnote{Spectroscopic foreground samples may be more likely to see significant 
-gains from moving beyond spectroscopic methods.}.  Binning galaxies by approximate redshift lets us 
+gains from moving beyond tomographic methods.}.  Binning galaxies by approximate redshift also lets us 
 model galaxy bias, intrinsic alignments, and other systematic errors en masse in a more tractable way.
@@
 \citep{heavens,kitching}, the tomographic approach remains the standard within the field.
+
+We are free to assign galaxies to different tomographic bins in any way we wish; changing the choice
+can potentially the ease with which we can calibrate
+the bins, but any choice can lead to correct results. The bins need not even be purely tomographic: we can happily correlate bins with peaks at different redshifts if that is useful.
+We should choose, then, an assignment algorithm that maximises the constraining power for a science
+case of interest.  This challenge explores such algorithms.
 
@@
 In this paper we discuss and evaluate the related challenge
-of using a limited set of color bands to make those assignments, motivated by requirements of using the
+of using a limited set of colour bands to make those assignments, motivated by requirements of using the
 metacalibration method for galaxy shape measurement to limit biases.  
@@
 
-In \autoref{sec:motivation} we explain the need for methods that work on limited sets of bands in the context of upcoming lensing surveys.  \textbf{Complete structure...}
+In \autoref{sec:motivation} we explain the need for methods that work on limited sets of bands in the context of upcoming lensing surveys.
+In \autoref{sec:data} we describe the simulated data used in the challenge.
+\autoref{sec:methods} briefly summarizes the submitted methods; they are described in much more detail in \autoref{app:methods}.  In \autoref{sec:results} we discuss the
+results of the challenge, and then conclude in \autoref{sec:conclusion}.
 
@@
 almost completely remove errors due to model and estimator (noise) bias from WL data, at least
-when PSFs are well-measured and oversampled by the pixel grid.
+when PSFs are well-measured and oversampled by the pixel grid.  The method has been successfully
+applied in the Dark Energy Survey Year 1 and Year 3 analyses \citep{des-y1-cat, des-y3-cat},
+and application to early Rubin data is planned.
 
@@
 that are
-highly covariant with those on other galaxy properties, including size and, most importantly, flux/
+highly covariant with those on other galaxy properties, including size and, most importantly, flux and
 magnitude.  It follows that a cut (or re-weighting) on galaxy magnitudes, or on any quantity derived
 from  them, will induce a shear bias since galaxies will preferentially fall either side of the cut
-depending on their shear.  Within metacal, we can solve this problem by performing the cut on both
+depending on their shear.  Within metacal, we can compute and thus account for this bias
+very accurately, by performing the cut on both
 sheared variants of the catalog, and measuring the difference between the post-cut shear in the two
@@
 particular, this limitation prevent us from using the most obvious method for tomography,
-and computnig a complete photometric probability density function (PDF) for each galaxy and assigning
+and computing a complete photometric probability density function (PDF) for each galaxy and assigning
 a bin using the mean or peak of that PDF.
@@
 In the challenge, participants assigned simulated galaxies to tomographic bins,
-using only their (g)riz bands and associated errors.  Their goal was to maximize the cosmological
+using only their (g)riz bands and associated errors, and the galay size.  Their goal was to maximize the cosmological
 information in the final split data set; this is generally acheived by separating different bins by 
@@
 Since this was a preliminary challenge designed to explore whether it is possible \emph{in theory}
-to use only (g)riz for tomography, we chose to simplify several aspects of the data.  In realistic 
+to use only the (g)riz bands for tomography, we chose to simplify several aspects of the data.  In realistic 
 situations we will have access to limited sized training sets for WL photometric redshifts, since 
@@
 to full catalogs, especially at faint magnitudes where spectroscopic coverage is sparse.  In this
-challenge we avoided both these issues -- the training samples we provided (see below) were comparable
+challenge we avoided both these issues -- the training samples we provided (see \autoref{sec:data}) were comparable
 in size to the testing sample, and drawn from the same population.  Given this, the challenge
@@
 as directly relevant as possible to the particular WL science case we focus on.  The data set
-was chosen to mimic the population, noise, and cuts we will use in real data (see \autoref{sec:data}, 
+was chosen to mimic the population, noise, and cuts we will use in real data (see \autoref{sec:data-proc}) 
 and the metrics were designed to be as close to the science goals as possible, rather than a lower
@@
 We used the second Buzzard catalog for two reasons.  First, to check the significance of
-an issue with galaxy colour assignments in CosmoDC2, described below.  And second, to check
+a limitation of galaxy colour assignments in CosmoDC2, described below.  And second, to check
 for over-fitting in the design of models themselves, and thus to determine whether two
@@
 tell us whether our conclusions might be reasonable for real data. We discuss correlations
-between metrics in Section \ref{sec:metric-results}.
+between metric scores on the two data sets in Section \ref{sec:metric-results}.
 
@@
 The CosmoDC2 simulation was designed to support the Dark Energy Science
-Collaboration and is described in detail in \citet{cosmodc2}.  It covered 440
+Collaboration and is described in detail in \citet{cosmodc2}.  It covers 440
 deg${}^2$ of sky, and used the dark matter particles from the Outer Rim
@@
 simulation is complete to $r=28$, and contains objects up to redshift 3.  We
-include the ultra-faint galaxies in the simulation, not assigned to a specific
-halo.
+include the ultra-faint galaxies, not assigned to a specific halo, in our sample.
 
@@
 
-One limitation of the CosmoDC2 catalogs we used, found during the challenge,
+One limitation of the CosmoDC2 catalogs, found as part of the challenge,
 was that the matching to Galacticus led to only a limited number of SEDs being
@@
 
-\subsubsection{Post-processing}
+\subsubsection{Post-processing} \label{sec:data-proc}
 
 We add noise to the truth values in the extra-galactic catalogs to simulate real observations.
-In each case we simulate observations corresponding to the first year Rubin observations
-using the DESC {\sc TXPipe} framework, following the methodology of \citet{ivezic_jones_lupton}
+In each case we simulate first year Rubin observations
+using the DESC {\sc TXPipe} framework\footnote{\url{https://github.com/LSSTDESC/TXPipe}}, following the methodology of \citet{ivezic_jones_lupton}
 and assuming the numerical values for telescope and sky characteristics therein.
@@
 $T$ is the trace $I_{xx} + I_{yy}$ of the moments matrix and measures the squared radius of the
-galaxy compared to that of the PSF.
+galaxy compared to that of a fixed Rubin PSF of size $0.75$ arc-seconds.
 
@@
 derived from the samples, and one based on a figure-of-merit for the constraints that would be obtained 
-using them.  \textbf{Discuss how correlated these turn out to be here}.
+using them. The relationships between these metrics in our results is discussed in \autoref{sec:metric-results}.
 
@@
 Metric 1 would be approximately measurable on real data without performing a full cosmology analysis,
-whereas metric 2 would be the final output of such an analysis.  We therefore simulate performance
-tests at multiple stages in the wider process.
+whereas metric 2 would be the final output of such an analysis.  They therefore test performance at what
+would be multiple stages in the wider pipeline.
 
@@
 where $\mu$ is a stack of the theoretical predictions for the $C_\ell$ spectra for each tomographic 
-bin, in turn, and $C$ a Gaussian estimaate of the covariance between them, as described in
+bin, in turn, and $C$ a Gaussian estimate of the covariance between them, as described in
 \citet{takada_jain} and summarized in \autoref{app:theory}.   We compute this metric for lensing alone, clustering alone, and the full 3x2pt combination.
@@
 where $[F^{-1}]_{p_1, p_2}$ extracts the 2x2 matrix for two parameters.  We use both the original
-DETF variant in which $p_1 = w_0, p_2 = w_a$, representing the constraining power on dark energy 
+DETF variant in which $p_1 = w_0$ and $p_2 = w_a$, representing the constraining power on dark energy 
 evolution, and a version in which $p_1 = \Omega_c$ and $p_2 = \sigma_8$, representing constraining
@@
 runtime environment or the hardware at NERSC, the participants and organizers tried
-to amend it after the formal challenge period, though this was not always possible.
+to amend it after the formal challenge period ended, though this was not always possible.
 
@@
 The first method randomly assigned galaxies to each tomographic bin, resulting
-in bins which each had the same $n(z)$, on average.  Using this method, we expect
-that increasing the number of bins should not increase any metric, since bins are
+in bins which each had the same $n(z)$, on average (the random noise fluctuations
+are very small when using this number of galaxies).  Using this method, we expect
+that increasing the number of bins should not increase any metric score, since bins are
 maximally correlated.  We find this to be true for all our metrics.
 
-Second, we use a method which uses only the $i$-band to select tomographic bins,
+Second, we used a method which used only the $i$-band to select tomographic bins,
 splitting galaxies in equal-sized bins in this magnitude. We expect only a small
-increase in metrics with the number of bins, and also that the addition of the $g$-band
-should make no difference to metrics.  Again, we find this to be true.
+increase in metric score with the number of bins, and also that the addition of the $g$-band
+should make no difference to scores.  Again, we find this to be true.
 
@@
 never be applied to real data and should therefore considered a form of control method.
-This method was {\sc Utopia}, which, as described below, found, for each object in the testing data, the
+This method was {\sc Utopia}, which, as described in \ref{sec:utopia}, found, for each object in the testing data, the
 nearest neighbour object in the training sample.  The high density, completeness, and
@@
 
-\section{Methods}
+\section{Submitted Methods} \label{sec:methods}
+
+Twenty five methods were submitted to the challenge, including the control methods described above.
+Most used one machine learning method of one type or another to perform the primary classification,
+rather than trying to perform a full photometric redshift analysis.
+
+The methods were (with the submitting authors listed in brackets):
+% \setlength{\itemsep}{0pt}%
+\begin{description*}
+	\item[\textsc{Random}] the random assignment control method (organizers)
+	\item[\textsc{IBandOnly}] the i-band only control method (organizers)
+	\item[\textsc{RandomForest}] a collection of individual decision trees (organizers)
+	\item[\textsc{LSTM}] a Long Short-Term Memory neural network (CRB, BMOF, GT, ESC, EJG)
+	\item[\textsc{AutokerasLSTM}] an automaticaly configured LSTM variant\footnote{A technical problem, involving conflicting versions of libraries and incompatible hardware, meant that AutoKerasLSTM could not be run in the final challenge.} (CRB, BMOF, GT, ESC, EJG)
+	\item[\textsc{LGBM}] a Light Gradient Boosted Machine, a histogram-based tree algorthm (CRB, BMOF, GT, ESC, EJG)
+	\item[\textsc{TCN}] a Temporal Convolution Network (CRB, BMOF, GT, ESC, EJG)
+	\item[\textsc{CNN}] a Convolutional Neural Network (CRB, BMOF, GT, ESC, EJG)
+	\item[\textsc{Ensemble}] a combination of three other network methods (CRB, BMOF, GT, ESC, EJG)
+	\item[\textsc{SimpleSOM}] a Self-Organizing Map (AHW)
+	\item[\textsc{PQNLD}] an extension of {\sc SimpleSom} combining with template-based redshift estimation (AHW)
+	\item[\textsc{ComplexSOM}] an extension of {\sc SimpleSom} with an additional assignment optimization (DA, AHW)
+	\item[\textsc{GPzBinning}] a Gaussian Process redshift estimator and binner (PH)
+	\item[\textsc{JaxCNN/JaxResnet}] two related CNN-based bin edge optimizers (AP)
+	\item[\textsc{NeuralNetwork1/2}] a dense network optimizing bin assignment for two metrics (AP)
+	\item[\textsc{PCACluster}] a Principal Component Analysis of fluxes then optimized clustering (DG)
+	\item[\textsc{Utopia}] a nearest-neighbour method, not applicable to real data (AHW)
+	\item[\textsc{ZotBin/ZotNet}] two related neural network methods with a preprocesing normalizing flow (DK)
+	\item[\textsc{FFNN}] a feed-forward neural network (EN)
+	\item[\textsc{FunBins}] a random forest with various nominal bin edge selection (AB)
+	\item[\textsc{MLPQNA}] a multi-layer perceptron (SC and MB)
+	\item[\textsc{Stacked Generalization}] an combination of gradient tree boosting and logistic regression (JEC)
+\end{description*}
+
+These methods are described in full in \autoref{app:methods}
+
+\section{Results \& Discussion} \label{sec:results}
+
+\subsection{Result types}
+
+The methods described above were run on a data set, previously unseen by entrants but drawn
+from the same underlying population, of size 8.6M objects for CosmoDC2 and 5.4M objects for Buzzard.
+A complete suite of the metrics described in Section \ref{sec:metrics} was run on each of the two
+for each method.  In total, each method was run on 16 scenarios, each combination of: griz vs riz,
+Buzzard vs CosmoDC2, and 3,5,7, and 9 bins requested.  Nine metrics were calculated for each
+scenario: clustering, lensing, and 3x2pt for each of the SNR metric (Equation \ref{eq:snr}), the $w_0-w_a$ 
+DETF FOM, and the $\Omega_c - \sigma_8$ FOM (Equation \ref{eq:fom}).
+
+\subsection{Overview}
+
+The DETF results for each method are shown in Table \ref{tab:cosmodc2}.  Complete results for all metrics
+are availanble online.  Missing entries occured in two ways.
+The first, marked with an asterisk, is when a method did not run for the given scenario,
+either due to memory overflow or taking an extremely long time.
+The second, marked with a dash, is when a method ran, but generated at least one bin with an extremely 
+small number count, such that the spectrum could not be calculated.
+
+The challenge did not require that entrants clasified all galaxies; ambiguous
+galaxies could be discarded at the cost of number density.  Several methods
+took advantage of this.  The ComplexSom method (see \autoref{sec:csom}) found
+it did not improve scores, but as noted in \autoref{sec:pqnld} this may be an artifact
+of the high density of the training data, and so could be explored in future challenges.
+
+\begin{table*}[]
+\begin{tabular}{|l|llll|llll|}
+\hline
+                & \multicolumn{4}{c|}{\textbf{riz}}      & \multicolumn{4}{c|}{\textbf{griz}}                                \\ \hline
+\textbf{Method} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} \\ \hline
+\input{results/tables/table_FOM_DETF_3x2_dc2.tex}
+\hline
+\end{tabular}
+\caption{Values of the 3x2pt DETF ($w_0,w_a$) figure-of-merit achieved by entrant methods on the 
+CosmoDC2 part of the challenge.}
+\label{tab:cosmodc2}
+\end{table*}
+
+Figure \ref{fig:funbin_nz} shows the number density obtained by the highest scoring method in
+the fiducial metric, {\sc funbins}, for both the riz and griz scenarios.
+It is clear from this that fears about restricted photometry catastrophically limiting
+the information needed for tomography are unfounded, at least in our idealised scenario.
+In both the CosmoDC2 and Buzzard analyses,
+even the $riz$ bands alone are sufficient to split objects into nine reasonably separated
+tomographic bins.  Once bins become this narrow the primary analysis concern will become the calibration 
+of overall bin $n(z)$ values, rather than initial tomography, so we consider this to be reassuring.
+
+There is a clear widening of the $n(z)$ when losing the g-band.  This widening leads to a loss in 
+constraining power, which is discussed more in Section \ref{sec:gband}.
+
+For comparison, we also show results using all six LSST bands, ugrizy; these could not be used
+with the metacalibration method. The improvement in the sharpness of the bins is noticeable but not large,
+and does not translate into large increases in the figures of merit: the DETF score for CosmoDC2
+is 169.7, barely higher than the score for griz.  While this is only a general indication and not
+a thorough exploration, it is encouraging.
+
+\begin{figure}[htbp]
+% \centering
+\includegraphics[width=1\linewidth]{results/funbins_nz.pdf}
+\caption{The tomographic $n(z)$ values generated by the winning method in the challenge, {\sc funbins},
+generating nine bins with the CosmoDC2 data.
+The upper panel shows results for the method applied to riz bands, and the middle on griz. The
+increased spread and overlap of the bins in the middle panel decreases the overall constraining
+power and signal-to-noise.  The lower panel is for comparison and is not part of the challenge,
+and illutrates results using all the ugrizy bands.}
+\label{fig:funbin_nz}
+\end{figure}
+
+\subsection{Metric comparisons} \label{sec:metric-results}
+Figure \ref{fig:metrics} shows a comparison of some of the different metrics used in the challenge.
+We plot, using assignments for methods and for all bin counts, the relationship between our fiducial
+metric, the DETF 3x2pt on CosmoDC2, and other metrics we measured.
+
+The different metrics are generally well-correlated (this also holds true for the metrics not 
+shown here), especially at the highest values of the metrics representing the most successful
+methods. This is particularly so when comparing the Buzzard and CosmoDC2 metrics, which is encouraging 
+as it implies these methods are broadly stable to reasonable changes in the underlying galaxy
+distribution, and hence on the survey properties.  The largest outlier comes from the 
+{\sc Neural Network} entry.
+
+The relatively broader spread in the lensing-only metric illustrates how our metric is dominated
+by the stronger constraining power in the clustering part of the data set.  This arises because
+we do not include a galaxy bias model in our FOM, and so the constraining power of the clustering
+is artifically high.
+However, since the clustering metric then largely becomes a proxy for how well separated
+galaxy bins are, this is not a major limitation of the metrics.
+
+Finally, there is a bimodal structure in the relationship between the FOM and SNR metrics.
+This largely traces the split between models which train the target bin edges vs those
+which fix them, as described below in Section \ref{sec:train-vs-fix}.
+
+\begin{figure}[htbp]
+% \centering
+\includegraphics[width=0.9\linewidth]{results/metric_comparisons.pdf}
+\caption{A comparison of the metrics used in our challenge. The x-axis is shared, and shows
+the 3x2pt $w_0 - w_a$ (DETF) figure of merit on the CosmoDC2 data set for all methods, 
+with colours labelling the number
+of bins used.  Each of the panels shows a metric which is the same except for a single change, 
+which is labelled on its y-axis.  For example, the top panel still shows the 3x2pt $w_0 - w_a$ (DETF) figure of merit,
+but for the Buzzard data set.}
+\label{fig:metrics}
+\end{figure}
+
+\subsection{Impact of losing g-band} \label{sec:gband}
+Figure \ref{fig:loss} shows the ratio of the FOM between methods using the griz and the same method
+using the riz.  Each point represents one method with a specific number of bins, indicated by colour.
+For lower scoring methods and for smaller numbers of bins there is more scatter in the ratio,
+but for the highest scoring algorithms and configurations the loss when losing g-band is a little
+more stable, at around 10--15\%.
+
+The value of the extra FoM that could be gained should be weighed against 
+the challenge of high-accuracy
+PSF measurement for this band, and the cost in time and effort of determining it, 
+but until the end of the survey this is unlikely to be the limiting factor in overall precision.
+
+\begin{figure*}
+\includegraphics[width=0.9\linewidth]{results/g_band_loss.pdf}
+\caption{The degradation in constraining power over all methods
+ when comparing them on methods including and excluding
+the g-band, in terms of the ratio of the $w_0-w-a$ figure of merit.  The left-hand panel shows
+results for the CosmoDC2 scenarios, and the right-hand panel for the Buzzard scenarios.  Colours
+indicate the number of bins used in the analysis.}
+\label{fig:loss}
+\end{figure*}
+
+\subsection{Trained edges vs fixed edges}\label{sec:train-vs-fix}
+
+Some of the methods in the challenge accepted target bin edges as an input to their algorithms,
+and then used classifiers to assign objects to those bins.  Other methods optimized the values
+of bin edges themselves, by optimizing the metric.  For the runs shown here, these methods mostly
+targeted one of the 3x2pt metrics.
+
+A comparison of trained-edge and fixed-edge methods on the CosmoDC2 griz scenarios
+is shown in Figure \ref{fig:edges}.  One important point to note is that while trained-edge
+methods dominate the highest scores for the 3x2pt DETF metric, on which they trained, they
+fare much worse on the lensing-only metrics (the methods were not re-trained on the lensing metrics;
+we show the metrics for the same assigned bins).  This suggests that the difference between
+optimal bins for different science cases is a significant one, and thus that analysis pipelines
+should make it straightforward to calibrate multiple configurations and pass them to
+subsequent science anlysis.
+
+\begin{figure}
+\includegraphics[width=0.9\columnwidth]{results/edge_comparison.pdf}
+\caption{A comparison of results from methods which used fixed target edges (dashed blue)
+with those which optimized target bins (solid orange).  Methods were trained on the 3x2pt metric
+shown in the upper panel; the lower performance for these classifications 
+for the lensing-only metric shown in the lower panel illustrates the need for science case-specific
+binning choices. In each case, classification is done using griz bands on the CosmoDC2 data set.}
+\label{fig:edges}
+\end{figure}
+
+Among methods with fixed bins, those using equal numbers in each tend to plateau at scores of
+120-140 in Table \ref{tab:cosmodc2}, suggesting that a range of machine learning models can
+effectively classify by bin, at least in this case of representative, extensive training
+information.  Notably, this is the case even for the UTOPIA entry, which uses the nearest 
+training set neighbour to assign bins and thus was explicitly designed to work only
+for this idealised scenario; it should give an upper limit when using the same fixed bins.
+That other methods with the same nominal bin choice acheive scores close to it once again
+confirms that many entries are close to the best possible score for this approach.
+
+Several of the methods with optimized target bin edges break this
+barrier and acheive very high scores.  This included ZotNet, JaxResNet, and JaxCNN.  The
+Stacked Generalization method similarly included manually tuned bin edges to maximize
+score for this scenario. The typical 15--20\% improvement in FoM scores when optimizing
+nominal bin edges makes clear that this easy win should be compelling.
+
+Notably, though, the method that received the highest overall score in Table \ref{tab:cosmodc2}, 
+{\sc funbins} did so with a simpler approach - it used a random forest to assign galaxies to bins,
+but chose nominal bin edges using an equal width bins in \emph{comoving distance} $\chi$, converting
+this to redshift using a fiducial cosmology. The success of this simple solution suggests this
+as an easy heuristic for well-optimized nominal edges.
+
+\section{Conclusions} \label{sec:conclusion}
+The Tomography Challenge has proved a useful mechanism for encouraging the development
+of methods for this science question.  We encourage this challenge structure for other
+well-specified science cases, and DESC has subsequently run other challenges along the same lines
+such is its N5K beyond-limber challege\footnote{\url{https://github.com/LSSTDESC/N5K}}.
+We describe in Appendix \ref{sec:mistakes} some lessons we have learned from this challenge
+that may be useful for future ones.
+
+The results of the challenge have shown first that three or four-band photometry
+is sufficient for specifying tomography even up to a large number of bins, if
+the training data is complete enough: degneracies between colour do not make this impossible.
+Excluding the g-band from the data reduces the constraining power by a noticeable but
+not catastrophic amount.
+
+Finally, we have shown that in general there is a signficant advantage to optimizing
+target bins afresh when considering a new science case; this can be a significant boost
+to the final constraining power.
+Despite this, evenly spaced bins in fiducial co-moving distance have proven
+a good general choice in the winning method in the challenge, {\sc funbins}.
+
+The challenge was heavily simplified, with the intention of testing whether
+effective bin-assignments with limited photometry is possible even in theory,
+and of generating as many potential methods for the task as possible.  With
+both these acheived, future directions for simulating the problem are clear.
+The most obvious is limited spectroscopic completeness and size.  Realistic
+training data sets will be far smaller than the millions of objects we used here,
+and the difficulty of measuring spectra for faint galaxies will make the them
+incomplete and highly unrepresentative. Machine learning methods will be particularly
+affected by the latter problem.  Future challenges must explore these important limitations
+using simulated incompleteness.
+
+\section{Acknowledgements}
+ESC acknowledges support from the FAPESP (\#2019/19687-2) and CNPq  (\#308539/2018-4).
+
+\bibliography{paper}
+
+\appendix
+
+\section{Theory}\label{app:theory}
+
+In our metrics, the power spectrum between two tomographic bins $i$ and $j$ is
+computed using either the Core Cosmology Library \citep{ccl} or the JAX
+Cosmology Library \citep{jax-cosmo} as:
+
+\begin{equation}
+    C^{ij}_\ell = \int_0^{\infty} \frac{q_i(\chi) q_j(\chi)}{\chi^2} P\left(\frac{\ell +\frac{1}{2}}{\chi}, \chi \right) \mathrm{d}\chi
+\end{equation}
+where $\chi = \chi(z)$ is the comoving distance and $P$ the non-linear matter power spectrum at
+a fiducial cosmology.  The kernel functions $q_i(\chi)$ are different for the lensing and clustering samples:
+\begin{align}
+    q^{\mathrm{cluster}}_i(\chi) &= n_i(\chi)\\
+    q^{\mathrm{lensing}}_i(\chi) &= \frac{3}{2}\Omega_m \left(\frac{\mathrm{H}_0}{c}\right)^2 \frac{\chi}{a(\chi)} \int_\chi^{\infty} \frac{\chi' - \chi}{\chi'} n_i(\chi')\,\,\mathrm{d}\chi'
+\end{align}
+where $n_i(\chi) = n_i(z) \frac{\mathrm{d}z}{\mathrm{d}\chi}$.
+
+We evaluate these at 100 $\ell$ values from $\ell_\mathrm{min}=100$ to $\ell_\mathrm{max}=2000$.
+
+We use a Gaussian estimate for the covariance \citep{takada_jain}; this also incorporates the number of galaxies in the
+sample:
+\begin{equation}
+    \mathrm{Cov}(C^{ij}_\ell, C^{mn}_\ell) = \frac{1}{(2 \ell + 1)\Delta\ell f_\mathrm{sky}}(D^{im}_\ell + D^{jn}_\ell)(D^{in}_\ell + D^{jm}_\ell)
+\end{equation}
+where $D^{ij}_\ell = C^{ij}_\ell + N^{ij}_\ell$ and we assume an $f_\mathrm{sky}=0.25$.  The noise spectra are:
+\begin{align}
+N^{\mathrm{cluster},ij}_\ell = \delta_{ij} / n_i \\
+N^{\mathrm{lensing},ij}_\ell = \delta_{ij} \sigma_e^2 / n_i
+\end{align}
+where $n_i$ is the number density of galaxies in bin $i$ (we asssume equally weighted galaxies, and a 
+total number density over all bins of 20 galaxies per square arcminute) and $\sigma_e=0.26$.
+
+\section{Method Details} \label{app:methods}
+This appendix details the different methods submitted to the challenge.
 
@@
 an input galaxy, and the tree is trained to choose comparison features and comparison values
-which maximally split the discrimination at each step, and in final leaf purity.
-
-Since decision trees tend to over-fit data, random forests generate a suite of decision
+which maximize discrimination at each step and final leaf purity.
+
+Since decision trees tend to over-fit, random forests generate a suite of 
 trees, each randomly choosing a subset of features on which to train at each bifurcation.
@@
 
-% Please describe your methods with some very general / high level background theory
-% and a focus on the spsecific implementation.  Add any citations you like with bibtex in
-% paper.bib
-% Feel free to rename your method, but please leave a comment explaining so that I can do the
-% same in the tables.
-
 \subsection{ {\sc LSTM} } 
@@
 Methods \ref{ClecioLSTM} - \ref{ClecioEnsemble} were
-submitted by a team of authors CRB, BMOF, GT, ESC, and EJG.  All were built using TensorFlow, and
+submitted by a team of authors CRB, BMOF, GT, ESC, and EJG.  All were built using TensorFlow and/or Keras, and
 chose bin edges such that an equal number of galaxies is assigned to each bin, and trained the
@@
 
-Recurrent Neural Networks (RNNs)\citep{schuster, medsker, pascanu} are a type of NN capable of
+Recurrent Neural Networks (RNNs) \citep{schuster, medsker, pascanu} are a type of NN capable of
 analyzing sequential data, where data points have a strong correlation with their neighbours. Long
@@
 
-In this implementation, we chose bin edges such that an equal number of galaxies is assigned to each
-bin, and trained the network using the galaxies' magnitudes and colours. The net was implemented
-using Tensorflow and Keras. The Deep model is composed of a series of convolutional blocks (1d
+The Deep model is composed of a series of convolutional blocks (1d
 convolutional layer with a tanh activation followed by a MaxPooling layer) followed by a
@@
 validation loss. We kept the order of the layer blocks fixed, i.e., convolutional blocks,
-bidirectional LSTM, and Dense blocks. We let the other hyperparameters free, such as the number of
-neurons,  Dropout rates, maxpooling layers, the number of convolutional filters, and strides.
+bidirectional LSTM, and dense blocks. We left the other hyperparameters free, such as the number of
+neurons,  dropout rates, maxpooling layers, the number of convolutional filters, and strides.
 
 A technical problem, involving conflicting versions of libraries and incompatible hardware,
-meant that method could not be run in the final challenge.
+meant that this method could not be run in the final challenge.
 
@@
 
-In this implementation, we chose bin edges such that an equal number of galaxies is assigned to each
-bin, and trained the network using the galaxies' magnitudes and colours.
-
 \subsection{ {\sc TCN}}
-This method uses a Temporal-Convolutional (TCN) \citep{baitcn} network to assign galaxies to redshift
+This method uses a Temporal-Convolutional network (TCN) \citep{baitcn} to assign galaxies to redshift
 bins.
@@
 receptive field of the TCN can be increased without increasing the number of convolutional filters.
-Furthermore, residual blocks \citep{resnet} are also used to stabilize deeper and larger networks. These
+Residual blocks \citep{resnet} are also used to stabilize deeper and larger networks. These
 modifications cause TCNs to have a longer memory than traditional RNNs with the same capacity,  and
@@
 
-\subsection{ {\sc ComplexSOM} }
-The {\tt ComplexSOM} method implements an additional optimization layer on the
-methodology used by {\tt SimpleSOM}.
-
-Let ${\sf C}^{\rm group}_\ell$ be the matrix containing all auto- and
-cross-power spectra between members of a large set of galaxy samples (called
-``groups'' here). We can compress this large set into a smaller set of samples
-(called ``bins'' here), by combining different groups. Let ${\sf A}$ be the
-assignment matrix determining these bins, such that $A_{\alpha,i}=1$ if group
-$i$ is assigned to bin $\alpha$, and 0 otherwise. In that case, the redshift
-distribution of bin $\alpha$ is given by $N_\alpha(z)=\sum_i
-A_{\alpha,i}n_i(z)$, where $n_i(z)$ is the redshift distribution of group $i$.
-
-Defining the normalized assignment matrix ${\sf B}$ as
-\begin{equation}
-B_{\alpha,i}=A_{\alpha,i}\frac{\int dz\,n_i(z)}{\sum_j A_{\alpha,j}\int dz\,n_j(z)},
-\end{equation}
-
-${\sf C}^{\rm group}_\ell$ is related to the matrix of cross-power spectra
-between pairs of bins via:
-\begin{equation}
-{\sf C}^{\rm bin}_\ell={\sf B}\,{\sf C}^{\rm group}_\ell\,{\sf B}^T.
-\end{equation}
-
-The rationale behind the {\tt ComplexSOM} method is based on the observation
-that, while ${\sf C}^{\rm group}_\ell$ is a slow quantity to calculate for a
-large number of groups, ${\sf B}$ and therefore ${\sf C}^{\rm bin}_\ell$ are
-very fast. Thus, given an initial set of groups characterizing the
-distribution of the full sample in colour space, ${\sf C}^{\rm group}_\ell$
-can be precomputed once, and used to find the optimal assignment matrix ${\sf
-B}$ that maximizes the figure of merit in Equation \ref{eq:fom},
-which can be written in terms of ${\sf C}^{\rm bin}_\ell$ alone as
-
-\begin{equation}\label{eq:fom_tr}
-F_{\theta\theta'}=\sum_\ell f_{\rm sky}\frac{2\ell+1}{2}{\rm Tr}\left[\partial_\theta{\sf C}^{\rm bin}_\ell({\sf C}^{\rm bin}_\ell)^{-1}\partial_{\theta'}{\sf C}^{\rm bin}_\ell({\sf C}_\ell^{\rm bin})^{-1}\right].
-\end{equation}
-
-Note that the general problem of finding the matrix ${\sf B}$ that optimally
-compresses the information on a given parameter has an analytical solution in
-terms of Karnhunen-L\^oeve eigenvectors \citep{astro-ph/9603021}. However, the
-assignment matrix used in this case has the additional constraints that it
-must be positive, discrete and binary (i.e. groups are either in a bin or they
-are not). Thus, finding the absolute maximum figure of merit would involve a
-brute-force evaluation of all matrix elements $A_{\alpha,j}$, which is an
-intractable $N_{\rm bin}^{N_{\rm group}}$ problem. We thus parametrize ${\sf
-A}$ in terms of a small set of continuous parameters, for which standard
-minimization routines can be used. In particular we use $N_{\rm bin}-1$
-parameters, given by the edges of the redshift bins, and a group is assigned
-to a bin if its mean redshift falls within its edges. We also explored the
-possibility of ``trashing'' groups if less than a fraction $p_{\rm in}$ of its
-redshift distribution fell within its preassigned bin, treating $p_{\rm in}$
-as an additional free parameter. We found that adding this extra freedom did
-not translate into a significant improvement in the final level of data
-compression.
-
-Once the SOM has been generated, the {\tt ComplexSOM} method proceeds as
-follows. First, the full SOM is distilled into a smaller set of ${\cal O}(100)$
-groups based by combining SOM cells with similar moments of their redshift
-distributions, thus ensuring that this process does not induce any catastrophic
-$N(z)$ broadening {\bf you may wanna refer to the {\tt SimpleSOM} method here}.
-The large ${\sf C}^{\rm group}_\ell$ is then precomputed for the initial set of
-groups. Finally, we find the optimal assignment matrix, defined in terms of the
-free redshift-edge parameters, by maximizing the figure of merit in Eq.
-\ref{eq:fom_tr} using Powell's minimization method
-\citep{10.1093/comjnl/7.2.155}.
-
-\subsection{ {\sc GPzBinning} }
-This method was submitted by author PH.
-
-GPzBin is based on GPz \citep{Almosallam2016a,Almosallam2016b}, a machine learning
-regression algorithm originally developed for calculating the photometric
-redshifts of galaxies \citep{Gomes2017,Duncan2018,Hatfield2020} (but now also
-applied to other problems in physics \citep{Peng2019,Hatfield2020}). The
-algorithm is based on a Gaussian Process (GP; essentially an un-parametrised continuous
-function defined everywhere with Gaussian uncertainties); GPz specifically uses 
-a sparse GP framework, a fast and a scalable approximation
-of a full process \citep{Rasmussen2006}.
-
-The mean $\mu(x)$ and variance $\sigma(x)^2$ for predictions as a function of
-the input parameters $x$ (typically the photometry) are both linear combinations
-of a finite number of `basis functions': multi-variate Gaussian kernels with
-associated weights, locations and covariances. The algorithm seeks to find the
-optimal parameters of the basis functions such that the mean and variance
-functions are the most likely functions to have generated the data. The key
-innovations introduced by GPz include a) input-dependent variance estimations
-(heteroscedastic noise), b) a `cost-sensitive learning' framework where the
-algorithm can be tailored for the precise science goal, and c) properly
-accounting for uncertainty contributions from both variance in the data
-($\beta^{-1}_{\star}$) as well as uncertainty from lack of data ($\nu$) in a
-given part of parameter space (by marginalising over the functions supported by
-the GP that could have produced the data).
-
-%During the training stage of the algorithm GPz is inferring the locations and
-%spreads that describe the basis functions; during the validation stage it is
-%inferring the appropriate complexity of the model, essentially how many basis
-%functions to use.
-
-GPzBin is a simple extension to GPz for the problem of tomographic binning. For
-a given number of tomographic bins it selects redshift bin edges $z_{i}$ such
-that there are an equal number of training galaxies in each bin. Testing data
-galaxies then are assigned to the bin corresponding to their GPz predicted
-photometric redshift. The code allows for two selection criteria for removing
-galaxies for which it was not possible to make a good photometric redshift
-prediction: 1) cutting galaxies close to the boundaries of bins based on an
-`edge strictness' parameter $U$ which removes galaxies where $\mu\pm U \times
-\sigma \lessgtr z_i$ and 2) cutting galaxies based on the degree $E$ to which
-GPz is having to extrapolate to make the redshift prediction, removing galaxies
-with $\nu>E\sigma^2$ (see Figure 12 of \citealp{Hatfield2020}). Cutting galaxies
-removes some signal at the possible benefit of improving binning purity. As the
-training and test data were sampled from the same distribution in this data
-challenge we might not expect removing galaxies to give huge improvements in
-binning quality, but cuts based on $E$ and $U$ might become more useful in
-future studies where the training and test data have substantially different
-colour-magnitude distributions.
-
-When used in GPzBin for this data challenge the GPz settings
-in Table \ref{table-settings} were used (see
-\citealp{Almosallam2016a,Almosallam2016b} for precise definitions and
-interpretations).
-
-%GPz requires very little fine-tuning. The most important parameter is $m$, the number of basis functions. A higher $m$ corresponds to higher model complexity and longer training times.
-
-\begin{table*}[]
-% \tiny
-\begin{center}
-    \begin{tabular}{| l | l | l |}
-    \hline
-    Parameter   &   Value   &   Description\\  
-    \hline 
-    $m$         &   100     &   Number of basis functions; complexity of GP\\
-
-    maxIter     &   500     &   Maximum number of iterations (comparisons with 
-                                the validation data) permitted\\
-
-    maxAttempts &   50      &   Maximum number of iterations to attempt if there 
-                                is no progress on the validation set\\
-
-    method      &   GPVC    &   Type of bespoke covariances used on each basis 
-                                function (see \citealp{Almosallam2016b})\\
-
-    normalize   &   True    &   Pre-process the input by subtracting the means
-                                and dividing by the standard deviations\\
-
-    joint       &   True    &   Jointly learn a prior linear mean-function 
-                                (learn the function means and variances 
-                                jointly)\\
-
-    heteroscedastic  & True & Model noise as well as point estimates \\
-    \hline
-    \end{tabular}
-\caption{Parameter settings in the GPz entry.}
-\end{center}
-\label{table-settings}
-\end{table*}
-
-\subsection{{\sc JaxCNN \& JaxResNet} }
-
-This methi submitted by the author AP.
-
-JaxCNN and JaxResNet are a convolutional neural network based algorithm that
-maps the relationship between the color-magnitude data of galaxies to their bin
-indices. Convolutional neural network (CNN) models, as described in section
-\ref{sec:cnn}, consist of several layers, each consisting of a linear
-convolution operator followed by polling layers and a nonlinear activation
-function. Residual Network (ResNet) models take one step further and learn the
-\textit{identity} mapping by skipping, or adding \textit{shortcut} connections
-to, one or more layers \citep{resnet}. ResNet helps solve the vanishing gradient
-problem, hence we can have a deeper neural network.
-
-JaxCNN uses 2-layer convolutional neural networks and JaxResNet uses a stack of
-3 layers consisting of $1 \times 1, 3 \times 3$, and $1 \times 1$ convolutions
-with shortcut connections added to each stack. Both methods use the Rectified
-Linear Unit (ReLU) activation function under the ADAM optimizer \cite{adam} with
-a learning rate of 0.001.
-
-We chose the learning algorithm for both methods to minimize the reciprocal of
-the figure of merit $F$, where $F$ is defined in eqref{E:FOM}. We implement
-these models using the JAX-based Flax neural network library \cite{jax}.
-
-\subsection{ {\sc NeuralNetwork 1 \& 2} }
-
-The {\sc NeuralNetwork} approach combines the two following ideas:
-\begin{itemize}
-	\item Use a simple neural network to parameterize a bin assignment function $f_\theta$ taking available photometry as the input and returning a bin number. 
-	\item Optimize the parameters of $f_\theta$ as to maximize a target score, either the total SNR \autoref{eq:snr}, or the DETF FoM \autoref{eq:fom}.
-\end{itemize}
-The interest of this approach is that it actually bypasses the need for a photometric redshift estimate in allocating a galaxy to a bin, and directly solves the problem of optimal bin assignment given available photometry, as to maximize the cosmological information.
-
-Details of the particular neural network are mostly irrelevant, the main difficulty in this approach is to be able to compute the back-propagation of gradients through the cosmological loss function as to update the parameters $\theta$ of the bin assignment function.
-This is was made possible for this challenge by the \texttt{jax-cosmo} library \citep{jax-cosmo}, which allows the computation of both metrics using the JAX automatic differentiation framework \citep{jax}.
-
-In practice, to parameterize $f_\theta$ we use a simple dense neural network, composed of 3 linear layers of size 500 with leaky\_relu activation function and output batch normalization.
-The last layer of the model is a linear layer with an output size matching the number of bins, and softmax activation.
-We implement this model using the JAX-based Flax neural network library \citep{flax2020github}.
-
-Training is performed using batches of 5000 galaxies, over 2000 iterations under the ADAM optimizer \citep{adam} with a base learning rate of 0.001.
-The loss function $\mathcal{L}$ used for training directly derive from the challenge metrics, and constitutes the only difference between the NeuralNetwork 1 \& NeuralNetwork 2 entries:
-
-\begin{itemize}
-	\item NeuralNetwork 1: $\mathcal{L}_1 = 1 / S_{FoM}$
-	\item NeuralNetwork 2: $\mathcal{L}_2 = - S_{SNR} $
-\end{itemize}
-
-%\subsection{ {\sc NeuralNetwork 2} }
-
-\subsection{ {\sc PCACluster} }
-This entry was submitted by author DG.
-
-Principal Component Analysis (PCA) reduces the dimensionality of a dataset by
-calculating eigenvectors that align with the principal axes of variation in the
-data \citep{doi:10.1098}. {\sc PCACluster} uses PCA to reduce the flux \& color
-data set to three dimensions.
-
-To generate the PCA dimensions used by this method the PCA algorithm takes in
-the $r$-band flux and $ri$ and $iz$ colors and outputs three principal
-components.  If the data set provides the $g$-band, the algorithm also uses the
-$gr$ color when determining eigenvectors. Tomographic bins are then determined
-in this three dimensional space using a unique clustering algorithm where each
-observation is assigned to a bin by determining which cluster centroid is
-closest to that observation.
-
-In order to determine optimal centroid positions, clustering is framed as a
-gradient descent problem that seeks to maximize the requested metric: either the
-SNR or the FoM. This approach of identifying clusters using gradient descent is
-inspired by the similar approach to K-means clustering in
-\citet{NIPS1994_a1140a3d}.
-
-This entry uses {\sc jax\_cosmo} to take the derivative of the requested metric
-and classification function combination with respect to the centroid locations
-to calculate gradients and weight updates \citep{jax-cosmo}.  The speed of
-training PCACluster is directly proportional to the number of requested
-centroids (the number of requested tomographic bins) and the amount of training
-data used.
-
 \subsection{ {\sc SimpleSOM} }
-The SimpleSOM algorithm submitted by AHW utilises a self-organising map
+The SimpleSOM algorithm submitted by author AHW utilises a self-organising map
 \citep[SOM,][]{Kohonen:1982} to perform  a discretisation of the
@@
 ][]{Wright/etal:2020a}, where in this case we combine cells based on the
-similarity of the cells Nz moments (mean, median, and normalised median absolute
+similarity of the cells' $n(z) $moments (mean, median, and normalised median absolute
 deviation from median) using complete-linkage clustering. The result of this
@@
 
-\subsection{ {\sc UTOPIA} }
+\subsection{PQNLD} \label{sec:pqnld}
+
+The `por que no los dos' algorithm, submitted by author AHW, is a development of the SimpleSOM algorithm, adding the 
+additional complexity of also computing template-based photometric redshift point-estimates using BPZ 
+\citep{Benitez:2000}. Photo-$z$ point-estimates are derived using the re-calibrated 
+template set of \cite{Capak:2004} in combination with 
+the Bayesian redshift prior from \cite{Raichoor/etal:2014}.
+
+The algorithm seeks to leverage information from both template-based 
+photometric redshift estimates and machine-learning to improve the allocation of discretised cells in 
+colour-magnitude-space to tomographic bins, by flagging and discarding sources which reside in cells that have 
+significant colour-redshift degeneracies. In practice this is achieved by performing a quality-control step prior to 
+the assignment of SOM cells to tomographic bins, whereby each cell $i$ is flagged as having a catastrophic 
+discrepancy between its mean training $z$ and mean photo-$z$ if: 
+\begin{equation}
+\frac{|\langle z_{\rm train,i}\rangle - \langle z_{\rm phot,i}\rangle  -\langle z_{\rm train} - z_{\rm phot}\rangle|}
+{\sigma[z_{\rm train} - z_{\rm phot}]} \geq 5.
+\end{equation}
+
+This flagging is inspired by similar quality control that is implemented in the redshift calibration procedure of
+\citet{Wright/etal:2020a}, which removes pathological cells in the SOM prior to the construction of calibrated 
+redshift distributions for KiDS. We note, though, that this procedure is unlikely to have a significant influence in this 
+challenge, as it is primarily designed to identify and remove cells with considerable covariate shift between the 
+training and reference samples; a problem which does not exist (by construction) in this challenge.
+
+\subsection{ {\sc ComplexSOM} } \label{sec:csom}
+This method was submitted by author DA.
+
+The {\tt ComplexSOM} method implements an additional optimization layer on the
+methodology used by {\tt SimpleSOM}.
+
+Let ${\sf C}^{\rm group}_\ell$ be the matrix containing all auto- and
+cross-power spectra between members of a large set of galaxy samples (called
+``groups'' here). We can compress this large set into a smaller set of samples
+(called ``bins'' here), by combining different groups. Let ${\sf A}$ be the
+assignment matrix determining these bins, such that $A_{\alpha,i}=1$ if group
+$i$ is assigned to bin $\alpha$, and 0 otherwise. In that case, the redshift
+distribution of bin $\alpha$ is given by $N_\alpha(z)=\sum_i
+A_{\alpha,i}n_i(z)$, where $n_i(z)$ is the redshift distribution of group $i$.
+
+Defining the normalized assignment matrix ${\sf B}$ as
+\begin{equation}
+B_{\alpha,i}=A_{\alpha,i}\frac{\int dz\,n_i(z)}{\sum_j A_{\alpha,j}\int dz\,n_j(z)},
+\end{equation}
+
+${\sf C}^{\rm group}_\ell$ is related to the matrix of cross-power spectra
+between pairs of bins via:
+\begin{equation}
+{\sf C}^{\rm bin}_\ell={\sf B}\,{\sf C}^{\rm group}_\ell\,{\sf B}^T.
+\end{equation}
+
+The rationale behind the {\tt ComplexSOM} method is based on the observation
+that, while ${\sf C}^{\rm group}_\ell$ is a slow quantity to calculate for a
+large number of groups, ${\sf B}$ and therefore ${\sf C}^{\rm bin}_\ell$ are
+very fast. Thus, given an initial set of groups characterizing the
+distribution of the full sample in colour space, ${\sf C}^{\rm group}_\ell$
+can be precomputed once, and used to find the optimal assignment matrix ${\sf
+B}$ that maximizes the figure of merit in Equation \ref{eq:fom},
+which can be written in terms of ${\sf C}^{\rm bin}_\ell$ alone as
+
+\begin{equation}\label{eq:fom_tr}
+F_{\theta\theta'}=\sum_\ell f_{\rm sky}\frac{2\ell+1}{2}{\rm Tr}\left[\partial_\theta{\sf C}^{\rm bin}_\ell({\sf C}^{\rm bin}_\ell)^{-1}\partial_{\theta'}{\sf C}^{\rm bin}_\ell({\sf C}_\ell^{\rm bin})^{-1}\right].
+\end{equation}
+
+Note that the general problem of finding the matrix ${\sf B}$ that optimally
+compresses the information on a given parameter has an analytical solution in
+terms of Karhunen-L\`oeve eigenvectors \citep{astro-ph/9603021}. However, the
+assignment matrix used in this case has the additional constraints that it
+must be positive, discrete and binary (i.e. groups are either in a bin or they
+are not). Thus, finding the absolute maximum figure of merit would involve a
+brute-force evaluation of all matrix elements $A_{\alpha,j}$, which is an
+intractable $N_{\rm bin}^{N_{\rm group}}$ problem. We thus parametrize ${\sf
+A}$ in terms of a small set of continuous parameters, for which standard
+minimization routines can be used. In particular we use $N_{\rm bin}-1$
+parameters, given by the edges of the redshift bins, and a group is assigned
+to a bin if its mean redshift falls within its edges. We also explored the
+possibility of ``trashing'' groups if less than a fraction $p_{\rm in}$ of its
+redshift distribution fell within its preassigned bin, treating $p_{\rm in}$
+as an additional free parameter. We found that adding this extra freedom did
+not translate into a significant improvement in the final level of data
+compression.
+
+Once the SOM has been generated, the {\tt ComplexSOM} method proceeds as
+follows. First, the full SOM is distilled into a smaller set of ${\cal O}(100)$
+groups based by combining SOM cells with similar moments of their redshift
+distributions, thus ensuring that this process does not induce any catastrophic
+$N(z)$ broadening.
+The large ${\sf C}^{\rm group}_\ell$ is then precomputed for the initial set of
+groups. Finally, we find the optimal assignment matrix, defined in terms of the
+free redshift-edge parameters, by maximizing the figure of merit in Eq.
+\ref{eq:fom_tr} using Powell's minimization method
+\citep{10.1093/comjnl/7.2.155}.
+
+\subsection{ {\sc UTOPIA} }\label{sec:utopia}
 
@@
 holistically, by comparing the performance between different classes of methods,
-rather than individuals.
+rather than individuals. The authors have endeavoured to do so in the conclusions presented here.
+
+\subsection{ {\sc GPzBinning} }
+This method was submitted by author PH.
+
+GPzBin is based on GPz \citep{Almosallam2016a,Almosallam2016b}, a machine learning
+regression algorithm originally developed for calculating the photometric
+redshifts of galaxies \citep{Gomes2017,Duncan2018,Hatfield2020} but now also
+applied to other problems in physics \citep{Peng2019,Hatfield2020}. The
+algorithm is based on a Gaussian Process (GP; essentially an un-parametrised continuous
+function defined everywhere with Gaussian uncertainties); GPz specifically uses 
+a sparse GP framework, a fast and a scalable approximation
+of a full process \citep{Rasmussen2006}.
+
+The mean $\mu(x)$ and variance $\sigma(x)^2$ for predictions as a function of
+the input parameters $x$ (typically the photometry) are both linear combinations
+of a finite number of `basis functions': multi-variate Gaussian kernels with
+associated weights, locations and covariances. The algorithm seeks to find the
+optimal parameters of the basis functions such that the mean and variance
+functions are the most likely functions to have generated the data. The key
+innovations introduced by GPz include a) input-dependent variance estimations
+(heteroscedastic noise), b) a `cost-sensitive learning' framework where the
+algorithm can be tailored for the precise science goal, and c) properly
+accounting for uncertainty contributions from both variance in the data
+($\beta^{-1}_{\star}$) as well as uncertainty from lack of data ($\nu$) in a
+given part of parameter space (by marginalising over the functions supported by
+the GP that could have produced the data).
+
+%During the training stage of the algorithm GPz is inferring the locations and
+%spreads that describe the basis functions; during the validation stage it is
+%inferring the appropriate complexity of the model, essentially how many basis
+%functions to use.
+
+GPzBin is a simple extension to GPz for the problem of tomographic binning. For
+a given number of tomographic bins it selects redshift bin edges $z_{i}$ such
+that there are an equal number of training galaxies in each bin. Testing sample
+galaxies are then assigned to the bin corresponding to their GPz predicted
+photometric redshift. The code allows for two selection criteria for removing
+galaxies for which it was not possible to make a good photometric redshift
+prediction: 1) cutting galaxies close to the boundaries of bins based on an
+`edge strictness' parameter $U$ which removes galaxies where $\mu\pm U \times
+\sigma \lessgtr z_i$ and 2) cutting galaxies based on the degree $E$ to which
+GPz is having to extrapolate to make the redshift prediction, removing galaxies
+with $\nu>E\sigma^2$ (see Figure 12 of \citealp{Hatfield2020}). Cutting galaxies
+removes some signal at the possible benefit of improving binning purity. As the
+training and test data were sampled from the same distribution in this data
+challenge we might not expect removing galaxies to give huge improvements in
+binning quality, but cuts based on $E$ and $U$ might become more useful in
+future studies where the training and test data have substantially different
+colour-magnitude distributions.
+
+We used these settings in GPz for this challenge (see
+\citealp{Almosallam2016a,Almosallam2016b} for definitions): x=100, maxIter=500, 
+maxAttempts$=50$, method=GPVC, normalize=True, and joint=True.
+
+%GPz requires very little fine-tuning. The most important parameter is $m$, the number of basis functions. A higher $m$ corresponds to higher model complexity and longer training times.
+
+\subsection{{\sc JaxCNN \& JaxResNet} }
+
+This method was submitted by the author AP.
+
+JaxCNN and JaxResNet are convolutional neural network based algorithms that
+map the relationship between the colour-magnitude data of galaxies to their bin
+indices. Convolutional neural network (CNN) models, as described in section
+\ref{sec:cnn}, consist of several layers, each consisting of a linear
+convolution operator followed by polling layers and a nonlinear activation
+function. Residual Network (ResNet) models take one step further and learn the
+\textit{identity} mapping by skipping, or adding \textit{shortcut} connections
+to, one or more layers \citep{resnet}. ResNet helps solve the vanishing gradient
+problem, allowing a deeper neural network.
+
+JaxCNN uses 2-layer convolutional neural networks and JaxResNet uses a stack of
+3 layers consisting of $1 \times 1, 3 \times 3$, and $1 \times 1$ convolutions
+with shortcut connections added to each stack. Both methods use the Rectified
+Linear Unit (ReLU) activation function under the ADAM optimizer \cite{adam} with
+a learning rate of 0.001.
+
+We chose the learning algorithm for both methods to minimize the reciprocal of
+the figure of merit $F$, where $F$ is defined in eqref{E:FOM}. We implement
+these models using the JAX-based Flax neural network library \cite{jax}.
+
+\subsection{ {\sc NeuralNetwork 1 \& 2} }
+This method was submitted by the author FL.
+
+The {\sc NeuralNetwork} approach combines the two following ideas:
+\begin{itemize}
+	\item Using a simple neural network to parameterize a bin assignment function $f_\theta$ taking available photometry as the input and returning a bin number. 
+	\item Optimizing the parameters of $f_\theta$ as to maximize a target score, either the total SNR \autoref{eq:snr}, or the DETF FoM \autoref{eq:fom}.
+\end{itemize}
+The method directly solves the problem of optimal bin assignment given available photometry, to maximize the cosmological information, without estimating a redshift.
+
+Details of the particular neural network are mostly irrelevant; the main difficulty in this approach is to be able to compute the back-propagation of gradients through the cosmological loss function to update the parameters $\theta$ of the bin assignment function.
+This is was made possible for this challenge by the \textsc{jax-cosmo} library \citep{jax-cosmo}, which allows the computation of both metrics using the JAX automatic differentiation framework \citep{jax}.
+
+In practice, to parameterize $f_\theta$ we use a simple dense neural network, composed of 3 linear layers of size 500 with leaky\_relu activation function and output batch normalization.
+The last layer of the model is a linear layer with an output size matching the number of bins, and softmax activation.
+We implement this model using the JAX-based \textsc{Flax} neural network library \citep{flax2020github}.
+
+Training is performed using batches of 5000 galaxies, over 2000 iterations under the ADAM optimizer \citep{adam} with a base learning rate of 0.001.
+The loss functions $\mathcal{L}$ used for training directly derive from the challenge metrics, and constitute the only difference between the NeuralNetwork 1 \& NeuralNetwork 2 entries:
+
+\begin{itemize}
+	\item NeuralNetwork 1: $\mathcal{L}_1 = 1 / S_{FoM}$
+	\item NeuralNetwork 2: $\mathcal{L}_2 = - S_{SNR} $
+\end{itemize}
+
+%\subsection{ {\sc NeuralNetwork 2} }
+
+\subsection{ {\sc PCACluster} }
+This entry was submitted by author DG.
+
+Principal Component Analysis (PCA) reduces the dimensionality of a dataset by
+calculating eigenvectors that align with the principal axes of variation in the
+data \citep{doi:10.1098}. {\sc PCACluster} uses PCA to reduce the flux \& colour
+data set to three dimensions.
+
+To generate the PCA dimensions used by this method the PCA algorithm takes in
+the $r$-band flux and $ri$ and $iz$ colours and outputs three principal
+components.  If the data set provides the $g$-band, the algorithm also uses the
+$gr$ colour when determining eigenvectors. Tomographic bins are then determined
+in this three dimensional space using a clustering algorithm where each
+observation is assigned to a bin by determining which cluster centroid is
+closest to that observation.
+
+In order to determine optimal centroid positions, clustering is framed as a
+gradient descent problem that seeks to maximize the requested metric: either the
+SNR or the FoM. This approach of identifying clusters using gradient descent is
+inspired by the similar approach to K-means clustering in
+\citet{NIPS1994_a1140a3d}.
+
+This entry uses {\sc jax\_cosmo} to take the derivative of the requested metric
+and classification function combination with respect to the centroid locations
+to calculate gradients and weight updates \citep{jax-cosmo}.  The speed of
+training PCACluster is directly proportional to the number of requested
+centroids (the number of requested tomographic bins) and the amount of training
+data used.
 
@@
 
-Both methods transform the $n$ input fluxes into $n-1$ colors and one flux. The data are very sparse in this
+Both methods transform the $n$ input fluxes into $n-1$ colours and one flux. The data are very sparse in this
 $n$-dimensional feature space and exhibit complex correlations. The preprocessing step transforms to a new
@@
 The ZotBin method defines a fully invertible sequence of steps that transform the input feature space into
-the final bins, so that each bin is associated with a well-defined region in color / flux space. The ZotNet method,
+the final bins, so that each bin is associated with a well-defined region in colour / flux space. The ZotNet method,
 on the other hand, trains a neural network to directly map from the preprocessed unit hypercube $[0,1]^n$ into
@@
 \subsection{ {\sc funbins} }
-This method uses the random forest algorithm described in section \ref{sec:randomforest} to assign galaxies to tomographic bins.
-There are 3 options for determining the bins used to assign labels to the training data - \textit{log}, \textit{random}, and \textit{chi}.
+This method uses the random forest algorithm described in section \ref{sec:randomforest} to assign galaxies to tomographic bins, but modifies the target bin edge selection method.
+There are 3 options for determining the edges used to assign labels to the training data: \textit{log}, \textit{random}, and \textit{chi}.
 
@@
 The second option, \textit{random}, draws the intermediate bin edges from a uniform distribution while fixing the outer edges to the limits of the data.
-This option is not theoretically motivated and just to study the sensitivity to the binning.
+This option is not theoretically motivated and designed only to study the sensitivity to the binning.
 
@@
 This should lead to roughly equal shot noise for the angular power in each bin.
-This method was used in the scores shown below.
+This method was used in the scores shown in this paper.
 
 Once the bin edges are determined by either of the methods above, the random forest is trained as described in section \ref{sec:randomforest}.
+
+\subsection{FFNN}
+This entry was submitted by author EN.
+
+The method is a Feed Forward Neural Network (FFNN), a classic type of neural network.
+This implementation used the Keras API in the TensorFlow library. Three components, a flattener, and two 
+dense layers of ReLu neurons, were optimized by ADAM on the sparse categorical cross-entropy of the classifications.
+A MinMaxScaler was used to pre-process the data, and the training stopped when there was no improvement in the  
+validation loss for three consecutive epochs, to avoid overfitting.
 
@@
 The architecture is a classical Multi Layer Perceptron (MLP; \citealp{Rosenblatt1961})
-with one or more hidden layers, on which the supervised learning paradigm is
-based on the Quasi Newton Algorithm (QNA), implemented through the L-BFGS rule
+with one or more hidden layers, on which the supervised learning paradigm is run
+using the Quasi Newton Algorithm (QNA), implemented through the L-BFGS rule
 \citep{Nocedal80}. L-BFGS is a solver that approximates the Hessian matrix,
@@
 Besides several scientific contexts and projects in which MLPQNA has succesfully
-been used, it was recently the kernel embedded into the method {\sc
+been used, it was recently used as the kernel embedded into the method {\sc
 METAPHOR} \citep{cavuoti20} for photo-z estimation, which participated in the
@@
 \end{equation}
-where $F_k(X)$ is the individual tree answer and $w_k$ a weight per tree. Unlike the Random Forest, the 
+where $F_k(X)$ is the individual tree answer and $w_k$ a weight per tree. Unlike the RF, the 
 tree depth is rather small and the training  of the $k$-th tree is based on the errors of the
@@
 algorithm).  GTB learning is essentially non-parallelizable and so cannot use a map-reduce paradigm, 
-unlike RF.It suffers from the opposite problem to RF's overfitting, and is subject
+unlike RF. It suffers from the opposite problem to RF's overfitting, and is subject
 to high bias, so one generally combines many weak individual learners.
@@
 This procedure was done by hand for this challenge, but could be automated and applied to any kind of classifier.
-
-% \lstset{language=Python}
-% \lstset{frame=lines}
-% \lstset{caption={The code}}
-% \lstset{label={lst:code_direct}}
-% \lstset{basicstyle=\footnotesize}
-% \begin{lstlisting}
-% estimators = [ 
-% ('gd', make_pipeline(StandardScaler(),
-%     GradientBoostingClassifier(n_estimators=n_estimators, verbose=1))),
-
-% ('rf', make_pipeline(StandardScaler(),
-%     RandomForestClassifier(n_estimators=n_estimators,verbose=1)))  ]
-
-% classifier = StackingClassifier(estimators=estimators,
-%                                final_estimator=LogisticRegression(max_iter=5000))
-% \end{lstlisting}
-
-\section{Results \& Discussion}
-
-\subsection{Overall results}
-The methods described above were run on a data set, previously unseen by entrants but drawn
-from the same underlying population, of size 8.6M objects for CosmoDC2 and 5.4M objects for Buzzard.
-A complete suite of the metrics described in Section \ref{sec:metrics} was run on each of the two
-for each method.  In total, each method was run on 16 scenarios, each combination of: griz vs riz,
-Buzzard vs CosmoDC2, and 3,5,7, and 9 bins requested.  Nine metrics were calculated for each
-scenario: clustering, lensing, and 3x2pt for each of the SNR metric (Equation \ref{eq:snr}, the $w_0-w_a$ 
-DETF FOM, and the $\Omega_c - \sigma_8$ FOM (Equation \ref{eq:fom}).
-
-The DETF results for each method are shown in Table \ref{tab:cosmodc2}.  Further results for other
-metrics are shown in Appendix \ref{sec:complete}.  Missing entries occured in two ways.
-The first, marked with an asterisk, is when a method did not run for the given scenario,
-either due to memory overflow or taking an extremely long time.
-The second, marked with a dash, is when a method ran, but generated at least one bin with an extremely 
-small number count, such that the spectrum could not be calculated.
-
-\begin{table*}[]
-\begin{tabular}{|l|llll|llll|}
-                & \multicolumn{4}{c|}{\textbf{riz}}      & \multicolumn{4}{c|}{\textbf{griz}}                                \\ \hline
-\textbf{Method} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} \\ \hline
-\input{results/tables/table_FOM_DETF_3x2_dc2.tex}
-\end{tabular}
-\caption{INCOMPLETE, AND AUTO-GENERATED SO NO POINT EDITING! Values of the 3x2pt DETF figure-of-merit achieved by entrant methods on the 
-CosmoDC2 part of the challenge.}
-\label{tab:cosmodc2}
-\end{table*}
-
-Figure \ref{fig:funbin_nz} shows the number density obtained by the highest scoring method in
-the fiducial metric, {\sc funbins}, for both the riz and griz scenarios.
-It is clear from this that fears about restricted photometry catastrophically limiting
-the information needed for tomography are unfounded, at least in our idealised scenario.
-In both the CosmoDC2 and Buzzard analyses,
-even the $riz$ bands alone are sufficient to split objects into nine reasonably separated
-tomographic bins.  Once bins become this narrow the primary analysis concern will become the calibration 
-of overall bin $n(z)$ values, rather than initial tomography, so we consider this to be reassuring.
-
-There is a clear widening of the $n(z)$ when losing the g-band.  This widening leads to a loss in 
-constraining power, which is discussed more in Section \ref{sec:gband}.
-
-\begin{figure}[htbp]
-% \centering
-\includegraphics[width=1\linewidth]{results/funbins_nz.pdf}
-\caption{The tomographic $n(z)$ values generated by the winning method in the challenge, {\sc funbins},
-generating nine bins with the CosmoDC2 data. 
-The upper panel shows results for the method applied to riz bands, and the lower on griz. The
-increased spread and overlap of the bins in the lower panel decreases the overall constraining
-power and signal-to-noise.}
-\label{fig:funbin_nz}
-\end{figure}
-
-\subsection{Metric comparisons} \label{sec:metric-results}
-Figure \ref{fig:metrics} shows a comparison of some of the different metrics used in the challenge.
-We plot, using assignments for methods and for all bin counts, the relationship between our fiducial
-metric, the DETF 3x2pt on CosmoDC2, and other metrics we measured.
-
-The different metrics are generally well-correlated (this also holds true for the metrics not 
-shown here), especially at the highest values of the metrics representing the most successful
-methods. This is particularly so when comparing the Buzzard and CosmoDC2 metrics, which is encouraging 
-as it implies these methods are broadly stable to reasonable changes in the underlying galaxy
-distribution, and hence on the survey properties.  The largest outlier comes from the 
-{\sc Neural Network} entry.
-
-The relatively broader spread in the lensing-only metric illustrates how our metric is dominated
-by the stronger constraining power in the clustering part of the data set.  This arises because
-we do not include a galaxy bias model in our FOM, and so the constraining power of the clustering
-is artifically high.
-However, since the clustering metric then largely becomes a proxy for how well separated
-galaxy bins are, this is not a major limitation of the metrics.
-
-Finally, there is a bimodal structure in the relationship between the FOM and SNR metrics.
-This largely traces the split between models which train the target bin edges vs those
-which fix them, as described below in Section \ref{sec:train-vs-fix}.
-
-\begin{figure}[htbp]
-% \centering
-\includegraphics[width=0.9\linewidth]{results/metric_comparisons.pdf}
-\caption{A comparison of the metrics used in our challenge. The x-axis is shared, and shows
-the 3x2pt $w_0 - w_a$ (DETF) figure of merit on the CosmoDC2 data set for all methods, 
-with colours labelling the number
-of bins used.  Each of the panels shows a metric which is the same except for a single change, 
-which is labelled on its y-axis.  For example, the top panel still shows the 3x2pt $w_0 - w_a$ (DETF) figure of merit,
-but for the Buzzard data set.}
-\label{fig:metrics}
-\end{figure}
-
-\subsection{Impact of losing g-band} \label{sec:gband}
-Figure \ref{fig:loss} shows the ratio of the FOM between methods using the griz and the same method
-using the riz.  Each point represents one method with a specific number of bins, indicated by colour.
-For lower scoring methods and for smaller numbers of bins there is more scatter in the ratio,
-but for the highest scoring algorithms and configurations the loss when losing g-band is a little
-more stable, at around 10--15\%.
-
-This is true in particular for the highest scoring methods, so we may assume the behaviour to be
-generic. The value of the extra FoM that could be gained should be weighed against 
-the challenge of high-accuracy
-PSF measurement for this band, and the cost in time and effort of determining it, 
-but until the end of the survey this is unlikely to be the limiting factor in overall precision.
-
-\begin{figure*}
-\includegraphics[width=0.9\linewidth]{results/g_band_loss.pdf}
-\caption{The degradation in constraining power over all methods
- when comparing them on methods including and excluding
-the g-band, in terms of the ratio of the $w_0-w-a$ figure of merit.  The left-hand panel shows
-results for the CosmoDC2 scenarios, and the right-hand panel for the Buzzard scenarios.  Colours
-indicate the number of bins used in the analysis.}
-\label{fig:loss}
-\end{figure*}
-
-\subsection{Trained edges vs fixed edges}\label{sec:train-vs-fix}
-
-Some of the methods in the challenge accepted target bin edges as an input to their algorithms,
-and then used classifiers to assign objects to those bins.  Other methods optimized the values
-of bin edges themselves, by optimizing the metric.  For the runs shown here, these methods all
-targeted one of the 3x2pt metrics. \textbf{Check this is correct}.
-
-A comparison of trained-edge and fixed-edge methods on the CosmoDC2 griz scenarios
-is shown in Figure \ref{fig:edges}.  One important point to note is that while trained-edge
-methods dominate the highest scores for the 3x2pt DETF metric, on which they trained, they
-fare much worse on the lensing-only metrics (the methods were not re-trained on the lensing metrics;
-we show the metrics for the same assigned bins).  This suggests that the difference between
-optimal bins for different science cases is a significant one, and thus that analysis pipelines
-should make it straightforward to calibrate multiple configurations and pass them to
-subsequent science anlysis.
-
-\begin{figure}
-\includegraphics[width=0.9\columnwidth]{results/edge_comparison.pdf}
-\caption{A comparison of results from methods which used fixed target edges (dashed blue)
-with those which optimized target bins (solid orange).  Methods were trained on the 3x2pt metric
-shown in the upper panel; the lower performance for these classifications 
-for the lensing-only metric shown in the lower panel illustrates the need for science case-specific
-binning choices. In each case, classification is done using griz bands on the CosmoDC2 data set.}
-\label{fig:edges}
-\end{figure}
-
-Among methods with fixed bins, those using equal numbers in each tend to plateau at scores of
-120-140 in Table \ref{tab:cosmodc2}, suggesting that a range of machine learning models can
-effectively classify by bin, at least in this case of representative, extensive training
-information.  Notably, this is the case even for the UTOPIA entry, which uses the nearest 
-training set neighbour to assign bins and thus was explicitly designed to work only
-for this idealised scenario; it should give an upper limit when using the same fixed bins.
-That other methods with the same nominal bin choice acheive scores close to it once again
-confirms that many entries are close to the best possible score for this approach.
-
-Several of the methods with optimized target bin edges break this
-barrier and acheive very high scores.  This included ZotNet, JaxResNet, and JaxCNN.  The
-Stacked Generalization method similarly included manually tuned bin edges to maximize
-score for this scenario. The typical 15--20\% improvement in FoM scores when optimizing
-nominal bin edges makes clear that this easy win should is compelling.
-
-Notably, though, the method that received the highest overall score in Table \ref{tab:cosmodc2}, 
-{\sc funbins} did so with a simpler approach - it used a random forest to assign galaxies to bins,
-but chose nominal bin edges using an equal width bins in \emph{comoving distance} $\chi$, converting
-this to redshift using a fiducial cosmology. The success of this simple solution suggests this
-as an easy heuristic for well-optimized nominal edges.
-
-\section{Conclusions}
-The Tomography Challenge has proved a useful mechanism for encouraging the development
-of methods for this science question.  We encourage this challenge structure for other
-well-specified science cases, and DESC has subsequently run other challenges along the same lines
-such is its N5K beyond-limber challege\footnote{\url{https://github.com/LSSTDESC/N5K}}.
-We describe in Appendix \ref{sec:mistakes} some lessons we have learned from this challenge
-that may be useful for future ones.
-
-The results of the challenge have shown first that three or four-band photometry
-is sufficient for specifying tomography even up to a large number of bins, if
-the training data is complete enough: degneracies between colour do not make this impossible.
-Excluding the g-band from the data reduces the constraining power by a noticeable but
-not catastrophic amount.
-
-Finally, we have shown that in general there is a signficant advantage to optimizing
-target bins afresh when considering a new science case; this can be a significant boost
-to the final constraining power.
-Despite this, evenly spaced bins in fiducial co-moving distance have proven
-a good general choice in the winning method in the challenge, {\sc funbins}.
-
-The challenge was heavily simplified, with the intention of testing whether
-effective bin-assignments with limited photometry is possible even in theory,
-and of generating as many potential methods for the task as possible.  With
-both these acheived, future directions for simulating the problem are clear.
-The most obvious is limited spectroscopic completeness and size.  Realistic
-training data sets will be far smaller than the millions of objects we used here,
-and the difficulty of measuring spectra for faint galaxies will make the them
-incomplete and highly unrepresentative. Machine learning methods will be particularly
-affected by the latter.  Future challenges must explore these important limitations
-using simulated incompleteness.
-
-\section{Acknowledgements}
-ESC acknowledges support from the FAPESP (\#2019/19687-2) and CNPq  (\#308539/2018-4).
-
-\bibliography{paper}
-
-\appendix
-
-\section{Theory}\label{app:theory}
-
-In our metrics, the power spectrum between two tomographic bins $i$ and $j$ is
-computed using either the Core Cosmology Library \citep{ccl} or the JAX
-Cosmology Library \citep{jax-cosmo} as:
-
-\begin{equation}
-    C^{ij}_\ell = \int_0^{\infty} \frac{q_i(\chi) q_j(\chi)}{\chi^2} P\left(\frac{\ell +\frac{1}{2}}{\chi}, \chi \right) \mathrm{d}\chi
-\end{equation}
-where $\chi = \chi(z)$ is the comoving distance and $P$ the non-linear matter power spectrum at
-a fiducial cosmology.  The kernel functions $q_i(\chi)$ are different for the lensing and clustering samples:
-\begin{align}
-    q^{\mathrm{cluster}}_i(\chi) &= n_i(\chi)\\
-    q^{\mathrm{lensing}}_i(\chi) &= \frac{3}{2}\Omega_m \left(\frac{\mathrm{H}_0}{c}\right)^2 \frac{\chi}{a(\chi)} \int_\chi^{\infty} \frac{\chi' - \chi}{\chi'} n_i(\chi')\,\,\mathrm{d}\chi'
-\end{align}
-where $n_i(\chi) = n_i(z) \frac{\mathrm{d}z}{\mathrm{d}\chi}$.
-
-We evaluate these at 100 $\ell$ values from $\ell_\mathrm{min}=100$ to $\ell_\mathrm{max}=2000$.
-
-We use a Gaussian estimate for the covariance \citep{takada_jain}; this also incorporates the number of galaxies in the
-sample:
-\begin{equation}
-    \mathrm{Cov}(C^{ij}_\ell, C^{mn}_\ell) = \frac{1}{(2 \ell + 1)\Delta\ell f_\mathrm{sky}}(D^{im}_\ell + D^{jn}_\ell)(D^{in}_\ell + D^{jm}_\ell)
-\end{equation}
-where $D^{ij}_\ell = C^{ij}_\ell + N^{ij}_\ell$ and we assume an $f_\mathrm{sky}=0.25$.  The noise spectra are:
-\begin{align}
-N^{\mathrm{cluster},ij}_\ell = \delta_{ij} / n_i \\
-N^{\mathrm{lensing},ij}_\ell = \delta_{ij} \sigma_e^2 / n_i
-\end{align}
-where $n_i$ is the number density of galaxies in bin $i$ (we asssume equally weighted galaxies, and a 
-total number density over all bins of 20 galaxies per square arcminute) and $\sigma_e=0.26$.
-
-\section{Complete results}
-\label{sec:complete}
-
-\begin{table*}[]
-\begin{tabular}{|l|llll|llll|}
-                & \multicolumn{4}{c|}{\textbf{riz}}      & \multicolumn{4}{c|}{\textbf{griz}}                                \\ \hline
-\textbf{Method} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} \\ \hline
-{\sc Autokeras\_LSTM } & 22.3 & 32.7    & 61.8    & --    & 26.3             & 43.4             & --             & 72.2\\
-{\sc CNN } & 18.0 & 36.9    & 59.8    & 70.3    & 17.7             & 40.7             & 62.6             & 74.5\\
-{\sc ComplexSOM } & 30.4 & 47.8    & 57.2    & 74.9    & 25.0             & 48.5             & 63.4             & 75.0\\
-{\sc ENSEMBLE1 } & 53.8 & 53.8    & 74.6    & 74.6    & 52.2             & 52.2             & 82.9             & 82.9\\
-{\sc funbins } & 23.2 & 50.5    & 65.0    & 72.2    & 26.0             & 57.4             & 81.3             & 91.7\\
-{\sc GPzBinning } & 16.8 & 46.4    & 61.4    & 70.7    & 17.6             & 39.7             & 67.0             & 82.5\\
-{\sc IBandOnly } & 11.4 & 15.9    & 20.8    & 22.0    & 11.4             & 15.9             & 20.8             & 22.0\\
-{\sc JaxCNN } & 38.6 & 59.0    & 69.7    & --    & 43.7             & 71.5             & 83.1             & 95.4\\
-{\sc JaxResNet } & 39.6 & 61.7    & --    & --    & 45.9             & 63.3             & --             & --\\
-{\sc LGBM } & 18.2 & 37.1    & 60.9    & 70.7    & 17.8             & 39.5             & 61.6             & 75.4\\
-{\sc LSTM } & 18.0 & 35.3    & 61.1    & 69.9    & 17.9             & 38.8             & 60.9             & 74.7\\
-{\sc mlpqna } & 24.1 & 54.2    & 65.7    & 73.2    & 27.2             & 51.5             & 70.4             & 83.4\\
-{\sc myCombinedClassifiers } & 18.4 & 33.0    & 56.1    & 85.3    & 22.1             & 38.8             & 68.3             & 97.2\\
-{\sc NeuralNetwork1 } & 42.8 & 66.8    & 82.0    & 93.7    & 46.6             & 74.1             & 90.9             & 105.4\\
-{\sc NeuralNetwork2 } & 21.9 & 52.1    & 60.8    & 76.5    & 22.2             & 47.0             & 59.9             & 73.9\\
-{\sc PCACluster } & 22.3 & 49.2    & 56.4    & --    & 21.0             & 44.8             & 57.6             & --\\
-{\sc Random } & 0.8 & 0.8    & 0.8    & 0.8    & 0.8             & 0.8             & 0.8             & 0.8\\
-{\sc RandomForest } & 23.9 & 53.4    & 65.0    & 72.5    & 27.3             & 52.1             & 70.4             & 83.5\\
-{\sc SimpleSOM } & 26.1 & 46.3    & 64.3    & 70.7    & 28.3             & 51.9             & 70.8             & 84.7\\
-{\sc TensorFlow\_FFNN } & -- & --    & --    & --    & --             & --             & --             & --\\
-{\sc TCN } & 26.6 & 53.7    & 65.7    & 74.6    & 27.0             & 49.2             & 69.4             & 82.4\\
-{\sc UTOPIA } & 24.9 & 53.8    & 65.3    & 73.6    & 27.9             & 52.9             & 70.1             & 83.1\\
-{\sc ZotBin } & 30.8 & 57.4    & 68.2    & 76.5    & 44.6             & 66.2             & --             & 89.4\\
-{\sc ZotNet } & 40.3 & 62.4    & 75.7    & 86.0    & 48.0             & 71.0             & 86.6             & 98.4\\
-\end{tabular}
-\caption{INCOMPLETE, AND AUTO-GENERATED SO NO POINT EDITING! Values of the 3x2pt DETF figure-of-merit achieved by entrant methods on the 
-Buzzard part of the challenge.}
-\label{tab:buzzard_3x2}
-\end{table*}
-
-\begin{table*}[]
-\begin{tabular}{|l|llll|llll|}
-                & \multicolumn{4}{c|}{\textbf{riz}}      & \multicolumn{4}{c|}{\textbf{griz}}                                \\ \hline
-\textbf{Method} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} \\ \hline
-\input{results/tables/table_FOM_ww_dc2.tex}
-\end{tabular}
-\caption{INCOMPLETE, AND AUTO-GENERATED SO NO POINT EDITING! Values of the weak lensing $\Omega_c - \sigma_8$
-figure-of-merit achieved by entrant methods on the CosmoDC2 part of the challenge.}
-\label{tab:dc2_wl}
-\end{table*}
-
-\begin{table*}[]
-\begin{tabular}{|l|llll|llll|}
-                & \multicolumn{4}{c|}{\textbf{riz}}      & \multicolumn{4}{c|}{\textbf{griz}}                                \\ \hline
-\textbf{Method} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} \\ \hline
-\input{results/tables/table_FOM_ww_buzzard.tex}
-
-\end{tabular}
-\caption{INCOMPLETE, AND AUTO-GENERATED SO NO POINT EDITING! Values of the weak lensing $\Omega_c - \sigma_8$
-figure-of-merit achieved by entrant methods on the Buzzard part of the challenge.}
-\label{tab:buzzard_wl}
-\end{table*}
 

### 2021-03-10 paper/results/make_figures.py

@@ -282,6 +282,7 @@ def plot_edge_type_comparison(data, filename):
 def plot_funbins_nz(filename):
     riz_file = "./cosmodc2/bins/funbins_9_riz_0.npy"
     griz_file = "./cosmodc2/bins/funbins_9_griz_0.npy"
+    ugrizy_file = "./cosmodc2/bins/funbins_ugrizy.npy"
     z_file = "./cosmodc2/bins/z.npz"
 
 
@@ -289,29 +290,37 @@ def plot_funbins_nz(filename):
         print("Data not downloaded for n(z) plot - not overwriting")
         return
 
-    fig, ax = plt.subplots(2, 1, figsize=(6,6), sharex=True, sharey=True)
+    fig, ax = plt.subplots(3, 1, figsize=(6,9), sharex=True, sharey=True)
     colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
 
     riz = np.load(riz_file)
     griz = np.load(griz_file)
+    ugrizy = np.load(ugrizy_file)
     z = np.load(z_file)['arr_0']
 
 
     nbin = 9
 
     for i in range(nbin):
-        w = np.where(griz == i)
+        w = np.where(riz == i)
         weight = np.repeat(1e-5, w[0].size)
         ax[0].hist(z[w], bins=50, histtype='step', ls='-', color=colors[i], weights=weight, lw=3)
 
     for i in range(nbin):
-        w = np.where(riz == i)
+        w = np.where(griz == i)
         weight = np.repeat(1e-5, w[0].size)
         ax[1].hist(z[w], bins=50, histtype='step', ls='-', color=colors[i], weights=weight, lw=3)
 
+
+    for i in range(nbin):
+        w = np.where(ugrizy == i)
+        weight = np.repeat(1e-5, w[0].size)
+        ax[2].hist(z[w], bins=50, histtype='step', ls='-', color=colors[i], weights=weight, lw=3)
+
     ax[0].set_ylabel('riz counts / $10^5$')
     ax[1].set_ylabel('griz counts / $10^5$')
-    ax[1].set_xlabel('Redshift')
+    ax[2].set_ylabel('ugrizy counts / $10^5$')
+    ax[2].set_xlabel('Redshift')
     ax[0].set_xlim(0, 3)
     fig.tight_layout()
     fig.subplots_adjust(hspace=0.05)

### 2021-03-11 paper/authors.tex
Update authors.tex
added affiliations
add some ORCIDS

@@
 \affiliation{Institute for Astronomy, University of Edinburgh, Edinburgh EH9 3HJ, United Kingdom}
+
+\author[0000-0001-7956-0542]{Fran\c{c}ois Lanusse}
+\affiliation{AIM, CEA, CNRS, Universit\'e Paris-Saclay, Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, F-91191 Gif-sur-Yvette, France}
 
@@
 
-\author{Cl\'{e}cio R. Bom}
+\author[0000-0003-4383-2969]{Cl\'{e}cio R. Bom}
 \affiliation{Centro Brasileiro de Pesquisas F\'isicas, Rua Dr. Xavier Sigaud 150, 22290-180 Rio de Janeiro, RJ, Brazil}
@@
 \author{Massimo Brescia}
-\noaffiliation
-%\affiliation{INAF Osservatorio Astronomico di Capodimonte, via Moiariello 16, I-80131, Napoli Italy}
+\affiliation{INAF - Osservatorio Astronomico di Capodimonte, via Moiariello 16, I-80131, Napoli Italy}
 
-\author{Jean-Eric Campagne}
-\noaffiliation
+\author[0000-0002-1590-6927]{Jean-Eric Campagne}
+\affiliation{Universit\'e Paris-Saclay, CNRS/IN2P3, IJCLab, 91405 Orsay, France}
 
 \author[0000-0002-3787-4196]{Stefano Cavuoti}
-\noaffiliation
-%\affiliation{INAF Osservatorio Astronomico di Capodimonte, via Moiariello 16, I-80131, Napoli Italy}
+\affiliation{INAF - Osservatorio Astronomico di Capodimonte, via Moiariello 16, I-80131, Napoli Italy}
+\affiliation{INFN - Sezione di Napoli, via Cinthia 21, I-80126 Napoli, Italy}
+\affiliation{Department of Physics ``E. Pancini'', University of Naples Federico II, via Cintia, 21, I-80126 Napoli, Italy}
 
@@
 
-\author{Elizabeth J. Gonzalez} $^{4,5}$.
+\author[0000-0002-0226-9893]{Elizabeth J. Gonzalez} $^{4,5}$.
 \affiliation{Instituto de Astronom\'{\i}a Te\'orica y Experimental (IATE-CONICET), Laprida 854, X5000BGR, C\'ordoba, Argentina.}
@@
 \noaffiliation
-
-\author[0000-0001-7956-0542]{Fran\c{c}ois Lanusse}
-\noaffiliation
-%\affiliation{AIM, CEA, CNRS, Universit\'e Paris-Saclay, Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, F-91191 Gif-sur-Yvette, France}
 

### 2021-03-11 paper/paper.tex
add complexsom authors

@@
 	\item[\textsc{PQNLD}] an extension of {\sc SimpleSom} combining with template-based redshift estimation (AHW)
-	\item[\textsc{ComplexSOM}] an extension of {\sc SimpleSom} with an additional assignment optimization (DA, AHW)
+	\item[\textsc{ComplexSOM}] an extension of {\sc SimpleSom} with an additional assignment optimization (DA, AHW, AN, EG, AB)
 	\item[\textsc{GPzBinning}] a Gaussian Process redshift estimator and binner (PH)
@@
 \subsection{ {\sc ComplexSOM} } \label{sec:csom}
-This method was submitted by author DA.
+This method was submitted by authors DA, AHW, AN, EG, and AB.
 

### 2021-03-12 paper/authors.tex
add complexsom authors
move sentence
more orcids
add massive tables

@@
 %main partticipants and text contributors  (alphabetical)
-\author{Bela Abolfathi}
-\noaffiliation
+\author[0000-0003-1820-8486]{Bela Abolfathi}
+\affiliation{Department of Physics and Astronomy, University of California, Irvine, CA 92697, USA}
 
-\author{David Alonso}
-\noaffiliation
+\author[0000-0002-4598-9719]{David Alonso}
+\affiliation{Astrophysics, University of Oxford, DWB, Keble Road, Oxford OX1 3RH, UK}
 
-\author{Abby Bault}
-\noaffiliation
+\author[0000-0002-9964-1005]{Abby Bault}
+\affiliation{Department of Physics and Astronomy, University of California, Irvine, CA 92697, USA}
 
@@
 
-\author{Massimo Brescia}
+\author[0000-0001-9506-5680]{Massimo Brescia}
 \affiliation{INAF - Osservatorio Astronomico di Capodimonte, via Moiariello 16, I-80131, Napoli Italy}
+
+\author{Adam Broussard}
+\noaffiliation
 
@@
 
+\author[0000-0003-1530-8713]{Eric Gawiser}
+\affiliation{Department of Physics and Astronomy, Rutgers, the State University, Piscataway, NJ 08854, USA}
+
 \author[0000-0002-0226-9893]{Elizabeth J. Gonzalez} $^{4,5}$.
@@
 
-\author{Dylan Green}
-\noaffiliation
+\author[0000-0002-0676-3661]{Dylan Green}
+\affiliation{Department of Physics and Astronomy, University of California, Irvine, CA 92697, USA}
 
@@
 
-\author{David Kirkby}
+\author{Kartheik Iyer}
 \noaffiliation
+
+\author[0000-0002-8828-5463]{David Kirkby}
+\affiliation{Department of Physics and Astronomy, University of California, Irvine, CA 92697, USA}
 
@@
 \author{Andy Park}
-\noaffiliation
+\affiliation{Department of Physics and Astronomy, University of California, Irvine, CA 92697, USA}
 
 \author{Anze Slosar}
-\noaffiliation
+\affiliation{Brookhaven National Laboratory, Upton NY 11973, USA}
 
@@
 
-% minor contributions and/or text writers (alphabetical)
+% DESC Builder authors
+\author{Yao-Yuan Mao}
+\noaffiliation

### 2021-03-12 paper/paper.tex
move sentence
add massive tables

@@
 \usepackage{listings}
-
+\usepackage{rotating}
+% \usepackage{pdflscape}
 \urlstyle{same}
@@
 
-One additional method was submitted to the challenge with the caveat that it could
-never be applied to real data and should therefore considered a form of control method.
-This method was {\sc Utopia}, which, as described in \ref{sec:utopia}, found, for each object in the testing data, the
-nearest neighbour object in the training sample.  The high density, completeness, and
-representativeness of our training set made this possible; realistics spectroscopic training
-sets are too sparse.  The method performed well, as expected, 
-but was not the highest scoring overall.
-
 \section{Submitted Methods} \label{sec:methods}
@@
 
-The methods were (with the submitting authors listed in brackets):
+These methods are described in full in \autoref{app:methods}.  Briefly, they were (with the submitting authors listed in brackets):
 % \setlength{\itemsep}{0pt}%
@@
 	\item[\textsc{PCACluster}] a Principal Component Analysis of fluxes then optimized clustering (DG)
-	\item[\textsc{Utopia}] a nearest-neighbour method, not applicable to real data (AHW)
+	\item[\textsc{Utopia}] a nearest-neighbour method, optimised in the limit of large representative training sets (AHW)
 	\item[\textsc{ZotBin/ZotNet}] two related neural network methods with a preprocesing normalizing flow (DK)
@@
 \end{description*}
-
-These methods are described in full in \autoref{app:methods}
 
@@
 The DETF results for each method are shown in Table \ref{tab:cosmodc2}.  Complete results for all metrics
-are availanble online.  Missing entries occured in two ways.
+for the nine-bin scenario only are shown in appendix tables \ref{tab:full_dc2} and \ref{tab:full_buzz}.
+
+In all these tables, missing entries are shown for two reasons.
 The first, marked with an asterisk, is when a method did not run for the given scenario,
@@
 we do not include a galaxy bias model in our FOM, and so the constraining power of the clustering
-is artifically high.
-However, since the clustering metric then largely becomes a proxy for how well separated
-galaxy bins are, this is not a major limitation of the metrics.
+is artifically high; this limitation is one reason we explore a wide variety of metrics.
 
@@
 information.  Notably, this is the case even for the UTOPIA entry, which uses the nearest 
-training set neighbour to assign bins and thus was explicitly designed to work only
+training set neighbour to assign bins and thus is explicitly optimised
 for this idealised scenario; it should give an upper limit when using the same fixed bins.
@@
 
-The `Useless Technique that OutPerforms Intelligent Alternatives' is a method
+The `Unreliable Technique that OutPerforms Intelligent Alternatives' is a method
 submitted by AHW that is designed,  primarily, as a demonstration of the care
@@
 increasingly distant regions of the multi-dimensional magnitude-space, resulting
-in poorer tomographic binning.  Therefore, UTOPIA acts as a warning against
+in poorer (i.e. unreliable) tomographic binning.  Therefore, UTOPIA acts as a warning against
 over-interpretation of the tomographic challenge results, as success in the
@@
 
-%=====================
-% END OF THE MAIN TEXT
-%=====================
+\section{Full 9-bin results table}
+Tables \ref{tab:full_dc2} and \ref{tab:full_buzz} show all calculated metrics when generating nine tomographic 
+bins for CosmoDC2 and Buzzard respectively. The \textit{ww} columns
+show weak-lensing metrics, \textit{gg} whow galaxy clustering, and \textit{3x} the full 3x2pt metric.
+
+As in table \ref{tab:cosmodc2}, some values are missing due to failure or time-out of the method (*) or pathological bin assignments (-).
+
+\movetabledown=65mm
+\begin{rotatetable*}
+\begin{deluxetable*}{ |l | c c c | c c c | c c c | c c c | c c c | c c c |}
+\tablecaption{The 9-bin results on the CosmoDC2 data set for all the metrics and all methods.
+ \label{tab:full_dc2}}
+\tablehead{
+\colhead{Method} & \multicolumn{9}{|c|}{riz} & \multicolumn{9}{c|}{griz}
+}
+\startdata
+&  \multicolumn{3}{c|}{SNR}   &  \multicolumn{3}{c|}{$\sigma_8-\Omega_m$ FOM} & \multicolumn{3}{c|}{$w_0-w_a$ FOM}
+&  \multicolumn{3}{c|}{SNR}   &  \multicolumn{3}{c|}{$\sigma_8-\Omega_m$ FOM} & \multicolumn{3}{c|}{$w_0-w_a$ FOM} \\
+\hline
+& ww & gg & 3x & ww & gg & 3x & ww & gg & 3x & ww & gg & 3x & ww & gg & 3x & ww & gg & 3x \\
+\hline
+\input{results/9bin_dc2.tex}
+    \enddata
+\end{deluxetable*}
+\end{rotatetable*}
+
+\movetabledown=65mm
+\begin{rotatetable*}
+\begin{deluxetable*}{ |l | c c c | c c c | c c c | c c c | c c c | c c c |}
+\tablecaption{The 9-bin results for all the metrics calculated for all methods. \textbf{Incomplete.} \label{tab:full_buzz}}
+\tablehead{
+\colhead{Method} & \multicolumn{9}{c|}{riz} & \multicolumn{9}{c|}{griz}
+}
+\startdata
+&  \multicolumn{3}{c|}{SNR}   &  \multicolumn{3}{c|}{$\sigma_8-\Omega_m$ FOM} & \multicolumn{3}{c|}{$w_0-w_a$ FOM}
+&  \multicolumn{3}{c|}{SNR}   &  \multicolumn{3}{c|}{$\sigma_8-\Omega_m$ FOM} & \multicolumn{3}{c|}{$w_0-w_a$ FOM} \\
+\hline
+& ww & gg & 3x & ww & gg & 3x & ww & gg & 3x & ww & gg & 3x & ww & gg & 3x & ww & gg & 3x \\
+\hline
+\input{results/9bin_buzzard.tex}
+    \enddata
+\end{deluxetable*}
+\end{rotatetable*}
+
+% \begin{table}[htbp]
+% \centering
+% \end{table}
 

### 2021-03-12 paper/results/make_figures.py

@@ -413,6 +413,35 @@ def make_tex_tables(dc2, buzzard, dirname):
             print(fn)
             make_table(data, metric, fn)
 
+def make_9bin_table(results, filename):
+    N = np.array([3, 5, 7, 9])
+    data = {}
+    for row in results:
+        name  = row['method']
+        bands = row['bands']
+        bins = row['bins']
+        for metric in metrics:
+            data[name, bands, bins, metric] = row[metric]
+
+    n = 9
+    f = open(filename, 'w')
+    for name in method_names:
+        disp_name = rf"{{\sc {name} }}".replace("_", r"\_").replace("myCombinedClassifiers", "Stacked Generalization")
+        row = [disp_name]
+        for i, bands in enumerate(['riz', 'griz']):
+            for j, metric in enumerate(metrics):
+                val  = data.get((name, bands, n, metric), np.nan)
+                if np.isnan(val):
+                    row.append(f"--")
+                elif val == 0:
+                    row.append(f"*")
+                else:
+                    row.append(f"{val:.1f}")
+
+        f.write(" & ".join(row))
+        f.write("\\\\ \n")
+    f.close()
+
 
 if __name__ == '__main__':
     matplotlib.use('agg')
@@ -428,3 +457,5 @@ if __name__ == '__main__':
     plot_funbins_nz('funbins_nz.pdf')
     make_colour_colour("colour_colour.pdf")
     make_initial_nz("initial_data.pdf")
+    make_9bin_table(buzzard, "9bin_buzzard.tex")
+    make_9bin_table(dc2, "9bin_dc2.tex")

### 2021-03-12 paper/paper.tex
more text

@@
 
-Figure \ref{fig:funbin_nz} shows the number density obtained by the highest scoring method in
-the fiducial metric, {\sc funbins}, for both the riz and griz scenarios.
+The overall winning method in the 3x2pt case in the fiducial griz 3x2pt CosmoDC2 metric was \textsc{funbins},
+which used a random forest to assign bins, but chose nominal bin edges by splitting the range spanned
+by the sample into equal co-moving distance bins. As described in \autoref{sec:funbins},  this was designed
+to minimze the shot noise for the angular power in each bin, and this has proven successful. 
+Figure \ref{fig:funbin_nz} shows the number density obtained by this method, for both the riz and griz 
+scenarios, and an additional post-challenge run using the full Rubin ugrizy bands.
 It is clear from this that fears about restricted photometry catastrophically limiting
@@
 and does not translate into large increases in the figures of merit: the DETF score for CosmoDC2
-is 169.7, barely higher than the score for griz.  While this is only a general indication and not
-a thorough exploration, it is encouraging.
+is 169.7, barely higher than the score for griz, and this is also true for other metrics.   While this is only a general indication and not a thorough exploration, it is encouraging.
+
+\textsc{funbins} seems to be a good and stable general method, though several other methods perform as well
+or slightly better in other scenarios.  Notably, many methods reach the same plateau in lensing-only metrics,
+suggesting this is somewhat easier to optimize; this is consistent with the general reason that photometric
+redshifts suffice for lensing: the broad lensing kernels are less sensitive to small changes.
+
+Several of the 9-bin neural network methods perform well up to seven tomographic bins but fail when run with 
+nine.  This is presumably not an intrinsic feature of these methods, but depends on implementation details.
+Further development solving these challenges should be valuable, since they are the highest scoring seven-bin
+methods.
 
@@
 
-Notably, though, the method that received the highest overall score in Table \ref{tab:cosmodc2}, 
-{\sc funbins} did so with a simpler approach - it used a random forest to assign galaxies to bins,
-but chose nominal bin edges using an equal width bins in \emph{comoving distance} $\chi$, converting
-this to redshift using a fiducial cosmology. The success of this simple solution suggests this
-as an easy heuristic for well-optimized nominal edges.
+Notably, though, the winning \textsc{funbins} method used the simpler approach desribed above.
+The success of this simple and fast solution suggests this as an easy heuristic for well-optimized nominal edges, at least
+for cases where clustering-like metrics are important.
 
@@
 
-\subsection{ {\sc funbins} }
+\subsection{ {\sc funbins} } \label{sec:funbins}
 This method uses the random forest algorithm described in section \ref{sec:randomforest} to assign galaxies to tomographic bins, but modifies the target bin edge selection method.
@@
 bins for CosmoDC2 and Buzzard respectively. The \textit{ww} columns
-show weak-lensing metrics, \textit{gg} whow galaxy clustering, and \textit{3x} the full 3x2pt metric.
+show weak-lensing metrics, \textit{gg} show galaxy clustering, and \textit{3x} the full 3x2pt metric.
 

### 2021-03-12 paper/paper.tex
add hline in app tables

@@
 	\item[\textsc{PCACluster}] a Principal Component Analysis of fluxes then optimized clustering (DG)
-	\item[\textsc{Utopia}] a nearest-neighbour method, optimised in the limit of large representative training sets (AHW)
+	\item[\textsc{UTOPIA}] a nearest-neighbour method, optimised in the limit of large representative training sets (AHW)
 	\item[\textsc{ZotBin/ZotNet}] two related neural network methods with a preprocesing normalizing flow (DK)
@@
 effectively classify by bin, at least in this case of representative, extensive training
-information.  Notably, this is the case even for the UTOPIA entry, which uses the nearest 
+information.  Notably, this is the case even for the \textsc{UTOPIA} entry, which uses the nearest 
 training set neighbour to assign bins and thus is explicitly optimised
@@
 
-In this way UTOPIA is the simplest possible algorithm (utilising all available
+In this way \textsc{UTOPIA} is the simplest possible algorithm (utilising all available
 photometric bands) that one can  implement for tomography. It is also a method
@@
 As the requirements of perfect  training-sample representivity and large size
-are violated, UTOPIA ought to assign redshifts to reference  galaxies from
+are violated, \textsc{UTOPIA} ought to assign redshifts to reference  galaxies from
 increasingly distant regions of the multi-dimensional magnitude-space, resulting
-in poorer (i.e. unreliable) tomographic binning.  Therefore, UTOPIA acts as a warning against
+in poorer (i.e. unreliable) tomographic binning.  Therefore, \textsc{UTOPIA} acts as a warning against
 over-interpretation of the tomographic challenge results, as success in the
@@
 show weak-lensing metrics, \textit{gg} show galaxy clustering, and \textit{3x} the full 3x2pt metric.
-
-As in table \ref{tab:cosmodc2}, some values are missing due to failure or time-out of the method (*) or pathological bin assignments (-).
+Methods above the horizontal line trained bin edges; methods below used fixed fiducial ones, though in some
+cases did some hand-tuning of them before submission.
+
+As in table \ref{tab:cosmodc2}, some values are missing due to failure or time-out of the method (*) or 
+pathological bin assignments (-).  Failure rates were higher for the 9-bin runs than for the lower bin counts.
 

### 2021-03-12 paper/results/make_figures.py

@@ -55,6 +55,9 @@ methods_with_trained_edges = [
     "ZotNet",
 ]
 
+
+methods_reordered = [m for m in method_names if m in methods_with_trained_edges] + [m for m in method_names if m not in methods_with_trained_edges]
+
 metrics = ["SNR_ww", "SNR_gg", "SNR_3x2", "FOM_ww", "FOM_gg", "FOM_3x2", "FOM_DETF_ww", "FOM_DETF_gg", "FOM_DETF_3x2"]
 
 
@@ -425,7 +428,7 @@ def make_9bin_table(results, filename):
 
     n = 9
     f = open(filename, 'w')
-    for name in method_names:
+    for name in methods_reordered:
         disp_name = rf"{{\sc {name} }}".replace("_", r"\_").replace("myCombinedClassifiers", "Stacked Generalization")
         row = [disp_name]
         for i, bands in enumerate(['riz', 'griz']):
@@ -440,6 +443,8 @@ def make_9bin_table(results, filename):
 
         f.write(" & ".join(row))
         f.write("\\\\ \n")
+        if name == methods_with_trained_edges[-1]:
+            f.write("\\hline\n")
     f.close()
 
 

### 2021-03-15 paper/results/make_figures.py
update table and emboldend

@@ -6,7 +6,7 @@ from astropy.table import Table
 import matplotlib
 import subprocess
 import h5py
-
+import collections
 # set default styles
 
 
@@ -416,6 +416,17 @@ def make_tex_tables(dc2, buzzard, dirname):
             print(fn)
             make_table(data, metric, fn)
 
+
+def find_best(data, nbin):
+
+    best = collections.defaultdict(lambda: (-np.inf, ""))
+    for (name, bands, bins, metric), score in data.items():
+        if bins != nbin:
+            continue
+        # print(metric, bands, best[metric, bands, bins], (score, name), max(best[metric, bands, bins], (score, name)))
+        best[metric, bands] = max(best[metric, bands], (score, name))
+    return best
+
 def make_9bin_table(results, filename):
     N = np.array([3, 5, 7, 9])
     data = {}
@@ -427,7 +438,9 @@ def make_9bin_table(results, filename):
             data[name, bands, bins, metric] = row[metric]
 
     n = 9
+    best = find_best(data, n)
     f = open(filename, 'w')
+    print(best)
     for name in methods_reordered:
         disp_name = rf"{{\sc {name} }}".replace("_", r"\_").replace("myCombinedClassifiers", "Stacked Generalization")
         row = [disp_name]
@@ -439,7 +452,11 @@ def make_9bin_table(results, filename):
                 elif val == 0:
                     row.append(f"*")
                 else:
-                    row.append(f"{val:.1f}")
+                    if best[metric, bands][1] == name:
+                        row.append(f"\\textbf{{{val:.1f}}}")
+                        print("best ", metric, bands, name)
+                    else:
+                        row.append(f"{val:.1f}")
 
         f.write(" & ".join(row))
         f.write("\\\\ \n")

### 2021-03-15 paper/paper.tex
affils
fix buzz results and highlight

@@
 ESC acknowledges support from the FAPESP (\#2019/19687-2) and CNPq  (\#308539/2018-4).
+AHW is supported by a European Research Council Consolidator Grant (No. 770935).
 
@@
 Methods above the horizontal line trained bin edges; methods below used fixed fiducial ones, though in some
-cases did some hand-tuning of them before submission.
+cases did some hand-tuning of them before submission. In each section the highest scoring method is highlighted.
 
@@
 \begin{deluxetable*}{ |l | c c c | c c c | c c c | c c c | c c c | c c c |}
-\tablecaption{The 9-bin results on the CosmoDC2 data set for all the metrics and all methods.
+\tablecaption{The 9-bin results on the CosmoDC2 data set for all the metrics and all methods on the CosmoDC2 sample.
  \label{tab:full_dc2}}
@@
 \begin{deluxetable*}{ |l | c c c | c c c | c c c | c c c | c c c | c c c |}
-\tablecaption{The 9-bin results for all the metrics calculated for all methods. \textbf{Incomplete.} \label{tab:full_buzz}}
+\tablecaption{The 9-bin results for all the metrics calculated for all methods on the Buzzard sample. \label{tab:full_buzz}}
 \tablehead{

### 2021-03-15 paper/results/make_figures.py

@@ -55,8 +55,10 @@ methods_with_trained_edges = [
     "ZotNet",
 ]
 
+methods_with_fixed_edges = [m for m in method_names if m not in methods_with_trained_edges]
 
-methods_reordered = [m for m in method_names if m in methods_with_trained_edges] + [m for m in method_names if m not in methods_with_trained_edges]
+
+methods_reordered = methods_with_trained_edges + methods_with_fixed_edges
 
 metrics = ["SNR_ww", "SNR_gg", "SNR_3x2", "FOM_ww", "FOM_gg", "FOM_3x2", "FOM_DETF_ww", "FOM_DETF_gg", "FOM_DETF_3x2"]
 
@@ -417,10 +419,12 @@ def make_tex_tables(dc2, buzzard, dirname):
             make_table(data, metric, fn)
 
 
-def find_best(data, nbin):
+def find_best(data, nbin, methods):
 
     best = collections.defaultdict(lambda: (-np.inf, ""))
     for (name, bands, bins, metric), score in data.items():
+        if name not in methods:
+            continue
         if bins != nbin:
             continue
         # print(metric, bands, best[metric, bands, bins], (score, name), max(best[metric, bands, bins], (score, name)))
@@ -438,9 +442,10 @@ def make_9bin_table(results, filename):
             data[name, bands, bins, metric] = row[metric]
 
     n = 9
-    best = find_best(data, n)
+    best1 = find_best(data, n, methods_with_trained_edges)
+    best2 = find_best(data, n, methods_with_fixed_edges)
     f = open(filename, 'w')
-    print(best)
+
     for name in methods_reordered:
         disp_name = rf"{{\sc {name} }}".replace("_", r"\_").replace("myCombinedClassifiers", "Stacked Generalization")
         row = [disp_name]
@@ -452,7 +457,7 @@ def make_9bin_table(results, filename):
                 elif val == 0:
                     row.append(f"*")
                 else:
-                    if best[metric, bands][1] == name:
+                    if best1[metric, bands][1] == name or best2[metric, bands][1] == name:
                         row.append(f"\\textbf{{{val:.1f}}}")
                         print("best ", metric, bands, name)
                     else:

### 2021-03-15 paper/paper.tex
more text

@@
 
+This is not true of the overall winning method for large bin counts - as shown in the tables
+in \autoref{app:tables}, significantly more methods acheive a top-scoring spot in the Buzzard
+table \ref{tab:full_buzz} than in the CosmoDC2 table \ref{tab:full_dc2}. This reinforces the
+conclusions made below in \autoref{sec:train-vs-fix}, though of course the relatively noisy
+results make this difficult to interpret.
+
 The relatively broader spread in the lensing-only metric illustrates how our metric is dominated
@@
 
-Notably, though, the winning \textsc{funbins} method used the simpler approach desribed above.
-The success of this simple and fast solution suggests this as an easy heuristic for well-optimized nominal edges, at least
-for cases where clustering-like metrics are important.
+Notably, though, the winning CosmoDC2 \textsc{funbins} method used the simpler approach desribed above. The 
+success of this simple and fast solution suggests this as an easy heuristic for well-optimized nominal edges, 
+at least for cases where clustering-like metrics are important.  Its relatively lower
+scores on the Buzzard metrics, as noted in \authoref{sec:metric-results}, does however further highlight
+the importance of science- and sample-specific training.
 
@@
 
-\section{Full 9-bin results table}
+\section{Full 9-bin results table} \label{app:tables}
 Tables \ref{tab:full_dc2} and \ref{tab:full_buzz} show all calculated metrics when generating nine tomographic 

### 2021-03-15 paper/paper.tex
bugs and more concl

@@
  when comparing them on methods including and excluding
-the g-band, in terms of the ratio of the $w_0-w-a$ figure of merit.  The left-hand panel shows
+the g-band, in terms of the ratio of the $w_0-w_a$ figure of merit.  The left-hand panel shows
 results for the CosmoDC2 scenarios, and the right-hand panel for the Buzzard scenarios.  Colours
@@
 at least for cases where clustering-like metrics are important.  Its relatively lower
-scores on the Buzzard metrics, as noted in \authoref{sec:metric-results}, does however further highlight
+scores on the Buzzard metrics, as noted in \autoref{sec:metric-results}, does however further highlight
 the importance of science- and sample-specific training.
@@
 not catastrophic amount.
+
+The wide range in scores for nominally similar methods (for example, the several methods using
+neural network approaches) reminds us that, generically, the implementation details of machine
+learning methods can drastically affect their performance.
 

### 2021-03-16 paper/paper.tex
FFNN fixes

@@
 dense layers of ReLu neurons, were optimized by ADAM on the sparse categorical cross-entropy of the classifications.
-A MinMaxScaler was used to pre-process the data, and the training stopped when there was no improvement in the  
+A StandardScaler was used to pre-process the data, and the training stopped when there was no improvement in the  
 validation loss for three consecutive epochs, to avoid overfitting.
+
+The network is trained using the galaxies' magnitudes, colours and band-triplets as well as their errors to assign them to 
+the redshift bins.
+Prior to processing the training and validation data, the non-detection placeholder magnitudes (30.0) were replaced with 
+an approximation of the $1\sigma$ detection limit as an estimate for the sky noise threshold, where Signal to Noise ratio, 
+S/N, equals 1. We first compute the error equivalent of $dm = 2.5 \log(1 + N/S)$ where $dm \sim 0.7526$ 
+magnitudes for $N/S = 1$, and then fit a logarithmic model to the magnitude as a function of its error 
+for different bands in the training data. This gives us the magnitude corresponding to this limit 
+and in order to be consistent we use this value along with its error to replace the non-detections 
+everywhere during the process.
 

### 2021-03-16 paper/results/make_figures.py

@@ -84,7 +84,7 @@ def make_table(results, metric, filename):
 
     f = open(filename, 'w')
     for name in method_names:
-        disp_name = rf"{{\sc {name} }}".replace("_", r"\_").replace("myCombinedClassifiers", "Stacked Generalization")
+        disp_name = rf"{{\sc {name} }}".replace('TensorFlow_FFNN', 'FFNN').replace("_", r"\_").replace("myCombinedClassifiers", "Stacked Generalization")
         row = [disp_name]
         for i, bands in enumerate(['riz', 'griz']):
             for n in N:
@@ -447,7 +447,7 @@ def make_9bin_table(results, filename):
     f = open(filename, 'w')
 
     for name in methods_reordered:
-        disp_name = rf"{{\sc {name} }}".replace("_", r"\_").replace("myCombinedClassifiers", "Stacked Generalization")
+        disp_name = rf"{{\sc {name} }}".replace('TensorFlow_FFNN', 'FFNN').replace("_", r"\_").replace("myCombinedClassifiers", "Stacked Generalization")
         row = [disp_name]
         for i, bands in enumerate(['riz', 'griz']):
             for j, metric in enumerate(metrics):

### 2021-03-16 paper/paper.tex
add not on stackgen

@@
 Several of the methods with optimized target bin edges break this
-barrier and acheive very high scores.  This included ZotNet, JaxResNet, and JaxCNN.  The
+barrier and acheive very high scores.  This included ZotNet, ZotBin,  and JaxResNet.  The
 Stacked Generalization method similarly included manually tuned bin edges to maximize
-score for this scenario. The typical 15--20\% improvement in FoM scores when optimizing
+score for this scenario (and indeed its ensemble combination method could be applied to
+combine the other successful methods). The typical 15--20\% improvement in FoM scores when optimizing
 nominal bin edges makes clear that this easy win should be compelling.

### 2021-03-18 paper/authors.tex
affils
authors
FFNN fixes
add orcid
more authors and notes

@@
 \affiliation{AIM, CEA, CNRS, Universit\'e Paris-Saclay, Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, F-91191 Gif-sur-Yvette, France}
+
+\author[0000-0001-7363-7932]{Angus Wright}
+\affiliation{Ruhr University Bochum, Faculty of Physics and Astronomy, Astronomical Institute (AIRUB), German Centre for Cosmological Lensing, 44780 Bochum, Germany}
 
@@
 
-\author{Eduardo S. Cypriano} $^3$
+\author{Eduardo S. Cypriano}
 \affiliation{Universidade de S\~ao Paulo, IAG, Rua do Mat\~ao 1225, S\~ao Paulo, SP, Brazil}
@@
 
-\author[0000-0002-0226-9893]{Elizabeth J. Gonzalez} $^{4,5}$.
+\author[0000-0002-0226-9893]{Elizabeth J. Gonzalez}
 \affiliation{Instituto de Astronom\'{\i}a Te\'orica y Experimental (IATE-CONICET), Laprida 854, X5000BGR, C\'ordoba, Argentina.}
@@
 
-\author{Peter Hatfield}
-\noaffiliation
+\author[0000-0002-3065-457X]{Peter Hatfield}
+\affiliation{Astrophysics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford, OX1 3RH, UK}
 
-\author{Kartheik Iyer}
-\noaffiliation
+\author[0000-0001-9298-3523]{Kartheik Iyer}
+\affiliation{Dunlap Institute for Astronomy and Astrophysics, University of Toronto, 50 St George St, Toronto, ON M5S 3H4, Canada}
 
@@
 \author{Alex Malz}
-\noaffiliation
+\affiliation{Ruhr University Bochum, Faculty of Physics and Astronomy, Astronomical Institute (AIRUB), German Centre for Cosmological Lensing, 44780 Bochum, Germany}
 
@@
 
-\author{Erfan Nourbakhsh}
-\noaffiliation
+\author[0000-0003-3827-4691]{Erfan Nourbakhsh}
+\affiliation{Department of Physics and Astronomy, University of California, Davis, Davis, CA 95616, USA}
 
@@
 
-\author{Angus Wright}
-\noaffiliation
-
 % DESC Builder authors

### 2021-03-24 paper/authors.tex
add initial
adding my affiliation and orcid

@@
 
-\author[0000-0002-3065-457X]{Peter Hatfield}
-\affiliation{Astrophysics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford, OX1 3RH, UK}
+\author{Peter Hatfield}
+\noaffiliation
 
-\author[0000-0001-9298-3523]{Kartheik Iyer}
-\affiliation{Dunlap Institute for Astronomy and Astrophysics, University of Toronto, 50 St George St, Toronto, ON M5S 3H4, Canada}
+\author{Kartheik Iyer}
+\noaffiliation
 
@@
 
-\author{Alex Malz}
-\affiliation{Ruhr University Bochum, Faculty of Physics and Astronomy, Astronomical Institute (AIRUB), German Centre for Cosmological Lensing, 44780 Bochum, Germany}
+\author[0000-0002-8676-1622]{Alex I. Malz}
+\affiliation{German Centre of Cosmological Lensing, Ruhr-Universit\"{a}t, Universit\"{a}tsstra{\ss}e 150, 44801 Bochum, Germany}
 

### 2021-03-24 paper/paper.tex

@@
 and computing a complete photometric probability density function (PDF) for each galaxy and assigning
-a bin using the mean or peak of that PDF.
+a bin using the mean or peak of that PDF\footnote{This only prevents us using other photometry for \emph{selection}, not for \emph{characterization} after objects have been selected, such as computing the overall
+number density $n(z)$ for a tomographic bin.}.
 
@@
 
-Once additional issue, though, was that many methods required GPUs to train in a reasonable time,
+One additional issue, though, was that many methods required GPUs to train in a reasonable time,
 and specifying an environment on these systems can be more complicated.

### 2021-03-24 paper/paper.tex
edited abstract and left comments through sec 5.2

@@
 \begin{document}
-\title{The LSST DESC 3x2pt Tomography Optimization Challenge}
+\title{The LSST-DESC 3x2pt Tomography Optimization Challenge}
 \input{authors}
@@
 \begin{abstract}
-In this paper we present the results of the DESC 3x2pt tomography challenge,
-which aimed to compare strategies for optimizing the photometric binning required
-for the main LSST 3x2pt analysis. This task is made particularly delicate in the
-context of a metacalibrated lensing catalog, as only the photometry from the
-bands  included in the metacalibration process (riz and potentially g) can be
-used to define this tomographic binning. Using a set of realistic simulations
-generating galaxy redshifts and simulated photometry, the goal of the
-challenge was to propose bin assignment strategies that maximized various
-figures of merit for a 3x2pt analysis. This challenge was highly idealized: it ignored
-spectroscopic completeness and limited size for the training sample, but was complex enough
-to address the main question of binning optimization. Based on the challenge
-dataset, we review and compare the different methods that have been proposed by
-participants. We find that even from this limited photometry information,
-various methods are able to yield binning schemes that achieve high figure-of-merit scores.
-We further find that the g band adds around 15\% to typical figures of merit, and
-that there is significant value in optimizing different bin assignments for specific
-science cases.
+This paper presents the results of the LSST-DESC 3x2pt tomography challenge, which served as a first step toward optimizing the tomographic binning strategy for the main LSST 3x2pt cosmology analysis.
+The task of choosing an optimal tomographic binning scheme for a photometric survey is made particularly delicate in the context of a metacalibrated lensing catalog, as only the photometry from the bands included in the metacalibration process (traditionally, {\em riz} and potentially {\em g}) can be used in sample definition.
+The goal of the challenge was to collect and compare bin assignment strategies under various metrics of a standard 3x2pt cosmology analysis in a highly idealized setting to establish a baseline for realistically complex follow-up studies;
+in this preliminary study, we used two sets of cosmological simulations of galaxy redshifts and photometry under a simple noise model neglecting outliers and variation in observing conditions, and contributed algorithms were provided with a representative and complete training set.
+We review and evaluate the entries to the challenge, finding that even from this limited photometry information, multiple algorithms yield binning schemes with high figure-of-merit scores.
+We further find that the {\em g} band improves metric performance by $\sim15\%$ and confirm that the optimal bin assignment strategy depends on the specific science case.
 \end{abstract}
@@
 %===========================
+
+% AIM: Suggestion: Use a macro for DESC because it shows up with and without expansion all over
 
@@
 
+% AIM: The next two paragraphs should present tomographic binning as an experimental design choice up front.
+
 Most lensing and 3x2pt analyses have been performed \emph{tomographically}, 
 binning galaxies by redshift.
+% AIM: I get what this paragraph is saying, but it's worded ambiguously and can read as almost its exact opposite. I suggest rewording to specify more of the instances of "this," "that," and "it."
 This approach captures almost all the available information in lensing data, since it measures
@@
 
+% AIM: typo in the next sentence {potentially the ease}
 We are free to assign galaxies to different tomographic bins in any way we wish; changing the choice
@@
 
+% AIM: Move the following sentence two paragraphs up to where tomography is introduced
 Tomographic 3x2pt measurements will be a key science goal in the upcoming \emph{Stage IV} surveys,
@@
 match our bin selections to specific tomographic bin edges.
-
+% AIM: Above sentence should cite other strategies as well.
+
+% AIM: insert comma {paper we}
+% AIM: messaging -- establish scope with the basic info now in the abstract
 In this paper we discuss and evaluate the related challenge
@@
 
+% AIM: 2020, not 2019!
 The Dark Energy Science Collaboration (DESC) Tomographic Challenge was launched in May 2019 and 
@@
 different bins.
+% AIM: We must explicitly state the assumptions that all methods were expected to be ML/training-based and that this is framed as a classification problem, and explain how that fits with the paradigm of how people think about tomography, which is traditionally framed as a very different redshift regression problem.
 
@@
 
+% AIM: Was it actually advertised in Rubin beyond DESC at all?
 The challenge was open to any entrants, not just those already involved in DESC, but it was
 advertised primarily through Rubin, and so is perhaps best described as semi-public.
+% AIM: "tested and ran" --> "developed and tested" 
+% AIM: remove comma {locally, but}
+% AIM: Final tests used to generate results on secret test set were conducted on NERSC, no?
 Participants tested and ran their methods locally, but submitted code to be run as part of a central
 combined analysis on the NERSC supercomputing facility.  No prize was offered apart from recognition.
+% AIM: Don't specify NERSC up here before it's defined in Sec. 3.4.
 
@@
 
+\begin{figure}[htbp]
+	% \centering
+	\includegraphics[width=0.9\linewidth]{results/initial_data.pdf}
+	% AIM: use alpha \neq 1
+	\caption{The underlying number density $n(z)$ for the two catalogs used in the challenge,
+		after the initial cuts described in the text.}
+	% AIM: specify /redshift/ number density
+	\label{fig:initial_nz}
+\end{figure}
+
+\begin{figure}[htbp]
+	% \centering
+	\includegraphics[width=0.9\linewidth]{results/colour_colour.pdf}
+	% AIM: use alpha \neq 1
+	\caption{A colour-colour diagram for the two catalogs used in the challenge, showing the bands
+		available to participants.}
+	\label{fig:colour_colour}
+\end{figure}
+
+% AIM: Rest of paper uses past tense so "run" --> "ran"
+% AIM: Train/Validate/Test set splitting should be described in post-processing rather than preamble, and assumption of ML/training-based approach should be stated explicitly in preamble to entire section.
 We run challenge entries on two data sets, CosmoDC2 and Buzzard, each of which is split into training, 
@@
 
+% AIM: We don't have to justify the second dataset unless we provide a reason why the first was insufficient. . . the presentation here needs some reworking.
 We used the second Buzzard catalog for two reasons.  First, to check the significance of
@@
 
+% AIM: typo {Each of these data set}
 Each of these data set come from cosmological simulations which provide truth values of galaxy magnitudes
@@
 
+% AIM: Say it was also used for DESC DC1 studies, cite e.g. PZ DC1 paper
 The Buzzard catalog was developed to support Dark Energy Survey Year 1 (DES-Y1)
@@
 \subsubsection{Post-processing} \label{sec:data-proc}
+% AIM: Put the info from CosmoDC2/Buzzard pertaining to status after post-processing steps here instead.
 
@@
 
+% AIM: specify /redshift/ number density and provide n(z) symbol because it is used without introduction later
 The overall number density of the two data sets is shown in Figure \ref{fig:initial_nz},
 and a set of colour-colour diagrams  in Figure \ref{fig:colour_colour}.
-
-\begin{figure}[htbp]
-% \centering
-\includegraphics[width=0.9\linewidth]{results/initial_data.pdf}
-\caption{The underlying number density $n(z)$ for the two catalogs used in the challenge,
-after the initial cuts described in the text.}
-\label{fig:initial_nz}
-\end{figure}
-
-\begin{figure}[htbp]
-% \centering
-\includegraphics[width=0.9\linewidth]{results/colour_colour.pdf}
-\caption{A colour-colour diagram for the two catalogs used in the challenge, showing the bands
-available to participants.}
-\label{fig:colour_colour}
-\end{figure}
+% AIM: If submitting to AAS journal, colour --> color; if submitting to MNRAS, catalog --> catalogue
 
@@
 \label{sec:metrics}
+% AIM: Change subsubsection titles to use the shorthand names for the metrics as they are referenced throughout the rest of the paper (SNR, w_{0}-w_{a} FOM, \Omega_{c}-\sigma_{8} FOM).
+% AIM: This should state that the three metrics are evaluated for clustering, lensing, and 3x2pt, so there are really nine metrics, not two!
 
@@
 would be multiple stages in the wider pipeline.
+% AIM: Sounds dumb, but readers will appreciate a clear statement that higher is better for these metrics.
 
 \subsubsection{Metric 1: Sigal-to-noise}
+% AIM: Suggestion: merge Appendix A with this section because this really doesn't stand by itself
+
 Our first metric is the total signal-to-noise ratio of the power spectrum derived from the assigned
@@
 \subsubsection{Metric 2: Figure-of-merit}
+% AIM: The title of this subsubsection should reflect that there are two FOM metrics.
 
@@
 
+% AIM: We must acknowledge GitHub as a platform that enabled this challenge.
+% AIM: Suggestion to specify "subclass" and "superclass" up front for extra-thorough clarity.
 The challenge was structured as a python library, in which each classification
@@
 
+% AIM: Include the code names for these dummy/control methods up front
 Two control methods were used to test the challenge machinery and metrics.
+% AIM: Explicitly put this into the context of a controlled experiment.
+% AIM: Make a new paragraph for each control method.
 The first method randomly assigned galaxies to each tomographic bin, resulting
@@
 \section{Submitted Methods} \label{sec:methods}
-
+% AIM: This should be a subsection of Challenge Design section above since it's really just a table of the submitted methods directing readers to the appendix for any detail.
+
+% AIM: typo: "Twenty five" --> "Twenty-five"
 Twenty five methods were submitted to the challenge, including the control methods described above.
@@
 
+% AIM: These should be a pagewidth table, and we should use macros so the method names display consistently throughout paper.
 These methods are described in full in \autoref{app:methods}.  Briefly, they were (with the submitting authors listed in brackets):
@@
 \section{Results \& Discussion} \label{sec:results}
+% AIM: Convert the first two subsections to the preamble of this section.
+% AIM: Preamble should include the tl;dr of "winners."
 
@@
 
-The methods described above were run on a data set, previously unseen by entrants but drawn
-from the same underlying population, of size 8.6M objects for CosmoDC2 and 5.4M objects for Buzzard.
-A complete suite of the metrics described in Section \ref{sec:metrics} was run on each of the two
-for each method.  In total, each method was run on 16 scenarios, each combination of: griz vs riz,
-Buzzard vs CosmoDC2, and 3,5,7, and 9 bins requested.  Nine metrics were calculated for each
-scenario: clustering, lensing, and 3x2pt for each of the SNR metric (Equation \ref{eq:snr}), the $w_0-w_a$ 
-DETF FOM, and the $\Omega_c - \sigma_8$ FOM (Equation \ref{eq:fom}).
-
-\subsection{Overview}
-
-The DETF results for each method are shown in Table \ref{tab:cosmodc2}.  Complete results for all metrics
-for the nine-bin scenario only are shown in appendix tables \ref{tab:full_dc2} and \ref{tab:full_buzz}.
-
-In all these tables, missing entries are shown for two reasons.
-The first, marked with an asterisk, is when a method did not run for the given scenario,
-either due to memory overflow or taking an extremely long time.
-The second, marked with a dash, is when a method ran, but generated at least one bin with an extremely 
-small number count, such that the spectrum could not be calculated.
-
-The challenge did not require that entrants clasified all galaxies; ambiguous
-galaxies could be discarded at the cost of number density.  Several methods
-took advantage of this.  The ComplexSom method (see \autoref{sec:csom}) found
-it did not improve scores, but as noted in \autoref{sec:pqnld} this may be an artifact
-of the high density of the training data, and so could be explored in future challenges.
-
+% AIM: Move this table to the appendix.
 \begin{table*}[]
@@
 
-The overall winning method in the 3x2pt case in the fiducial griz 3x2pt CosmoDC2 metric was \textsc{funbins},
-which used a random forest to assign bins, but chose nominal bin edges by splitting the range spanned
-by the sample into equal co-moving distance bins. As described in \autoref{sec:funbins},  this was designed
-to minimze the shot noise for the angular power in each bin, and this has proven successful. 
-Figure \ref{fig:funbin_nz} shows the number density obtained by this method, for both the riz and griz 
-scenarios, and an additional post-challenge run using the full Rubin ugrizy bands.
-It is clear from this that fears about restricted photometry catastrophically limiting
-the information needed for tomography are unfounded, at least in our idealised scenario.
-In both the CosmoDC2 and Buzzard analyses,
-even the $riz$ bands alone are sufficient to split objects into nine reasonably separated
-tomographic bins.  Once bins become this narrow the primary analysis concern will become the calibration 
-of overall bin $n(z)$ values, rather than initial tomography, so we consider this to be reassuring.
-
-There is a clear widening of the $n(z)$ when losing the g-band.  This widening leads to a loss in 
-constraining power, which is discussed more in Section \ref{sec:gband}.
-
-For comparison, we also show results using all six LSST bands, ugrizy; these could not be used
-with the metacalibration method. The improvement in the sharpness of the bins is noticeable but not large,
-and does not translate into large increases in the figures of merit: the DETF score for CosmoDC2
-is 169.7, barely higher than the score for griz, and this is also true for other metrics.   While this is only a general indication and not a thorough exploration, it is encouraging.
-
-\textsc{funbins} seems to be a good and stable general method, though several other methods perform as well
-or slightly better in other scenarios.  Notably, many methods reach the same plateau in lensing-only metrics,
-suggesting this is somewhat easier to optimize; this is consistent with the general reason that photometric
-redshifts suffice for lensing: the broad lensing kernels are less sensitive to small changes.
-
-Several of the 9-bin neural network methods perform well up to seven tomographic bins but fail when run with 
-nine.  This is presumably not an intrinsic feature of these methods, but depends on implementation details.
-Further development solving these challenges should be valuable, since they are the highest scoring seven-bin
-methods.
+The methods described above were run on a data set, previously unseen by entrants but drawn
+from the same underlying population, of size 8.6M objects for CosmoDC2 and 5.4M objects for Buzzard.
+A complete suite of the metrics described in Section \ref{sec:metrics} was run on each of the two
+for each method.  In total, each method was run on 16 scenarios, each combination of: $griz$ vs. $riz$,
+Buzzard vs. CosmoDC2, and 3, 5, 7, and 9 bins requested.  Nine metrics were calculated for each
+scenario: clustering, lensing, and 3x2pt for each of the SNR metric (Equation \ref{eq:snr}), the $w_0-w_a$ 
+DETF FOM, and the $\Omega_c - \sigma_8$ FOM (Equation \ref{eq:fom}).
+
+\subsection{Overview}
+% AIM: Suggestion: Convert this subsection to a bulleted list to give it some structure harking to the detailed subsections to come.
+The short versions are long enough to come off as unsupported statements rather than previews of what will be demonstrated in subsequent subsections.
+
+The DETF results for each method are shown in Table \ref{tab:cosmodc2}.  Complete results for all metrics
+for the nine-bin scenario only are shown in appendix tables \ref{tab:full_dc2} and \ref{tab:full_buzz}.  
+% AIM: Use autoref to appendix.
+In all these tables, missing entries are shown for two reasons.
+The first, marked with an asterisk, is when a method did not run for the given scenario,
+either due to memory overflow or taking an extremely long time.
+The second, marked with a dash, is when a method ran, but generated at least one bin with an extremely 
+small number count, such that the spectrum could not be calculated.
+
+% AIM: This is the first mention of this as a classification problem.
+We should mention this up in experimental design along with the assumption of ML/training-based methods, but also change "clasified" [SIC] to "assigned" for overall consistency (and maybe "that entrants clasified" --> "entrants to classify," and maybe reword post-semicolon part as "aside from the impact of reduced final sample size on the metrics, no penalty was applied for discarding galaxies entirely").
+% AIM: Also, the discussion of non-assignment should be its own subsection.
+The challenge did not require that entrants clasified all galaxies; ambiguous
+galaxies could be discarded at the cost of number density.  Several methods
+took advantage of this.  The ComplexSom method (see \autoref{sec:csom}) found
+it did not improve scores, but as noted in \autoref{sec:pqnld} this may be an artifact
+of the high density of the training data, and so could be explored in future challenges.
 
@@
 		generating nine bins with the CosmoDC2 data.
-The upper panel shows results for the method applied to riz bands, and the middle on griz. The
+		The upper panel shows results for the method applied to $riz$ bands, and the middle on $griz$. The
 		increased spread and overlap of the bins in the middle panel decreases the overall constraining
 		power and signal-to-noise.  The lower panel is for comparison and is not part of the challenge,
-and illutrates results using all the ugrizy bands.}
+		and illutrates results using all the $ugrizy$ bands.}
 	\label{fig:funbin_nz}
 \end{figure}
+
+The overall winning method in the 3x2pt case in the fiducial $griz$ 3x2pt CosmoDC2 metric was \textsc{funbins},
+which used a random forest to assign bins, but chose nominal bin edges by splitting the range spanned
+by the sample into equal co-moving distance bins. As described in \autoref{sec:funbins},  this was designed
+to minimze the shot noise for the angular power in each bin, and this has proven successful. 
+Figure \ref{fig:funbin_nz} shows the number density obtained by this method, for both the riz and griz 
+scenarios, and an additional post-challenge run using the full Rubin $ugrizy$ bands.
+% AIM: Split paragraph here to distinguish topics of overall challenge winner and conclusions about band information limitation.
+% AIM: Put the caveat at the beginning of the following sentence rather than the end. Also, this claim is not well-supported even by Fig. 3. . . 
+It is clear from this that fears about restricted photometry catastrophically limiting
+the information needed for tomography are unfounded, at least in our idealised scenario.
+In both the CosmoDC2 and Buzzard analyses,
+% AIM: analyses --> data sets?
+even the $riz$ bands alone are sufficient to split objects into nine reasonably separated
+tomographic bins.  Once bins become this narrow the primary analysis concern will become the calibration 
+of overall bin $n(z)$ values, rather than initial tomography, so we consider this to be reassuring.
+
+There is a clear widening of the $n(z)$ when losing the $g$-band.  This widening leads to a loss in 
+constraining power, which is discussed more in Section \ref{sec:gband}.
+
+For comparison, we also show results using all six LSST bands, $ugrizy$; these could not be used
+with the metacalibration method. The improvement in the sharpness of the bins is noticeable but not large,
+and does not translate into large increases in the figures of merit: the DETF score for CosmoDC2
+is 169.7, barely higher than the score for $griz$, and this is also true for other metrics.   While this is only a general indication and not a thorough exploration, it is encouraging.
+
+\textsc{funbins} seems to be a good and stable general method, though several other methods perform as well
+or slightly better in other scenarios.  Notably, many methods reach the same plateau in lensing-only metrics,
+suggesting this is somewhat easier to optimize; this is consistent with the general reason that photometric
+redshifts suffice for lensing: the broad lensing kernels are less sensitive to small changes.
+
+Several of the 9-bin neural network methods perform well up to seven tomographic bins but fail when run with 
+nine.  This is presumably not an intrinsic feature of these methods, but depends on implementation details.
+Further development solving these challenges should be valuable, since they are the highest scoring seven-bin
+methods.
 
@@
 \section{Mistakes} \label{sec:mistakes}
+% AIM: This should not be an appendix and should instead be part of the Results & Discussion section.
+They should not be presented as "mistakes" -- they are part of the experimental design and constitute caveats, which are simply facts, not subjective judgements.
+
 The challenge organizers made a number of mistakes when building and running the challenge.

### 2021-03-25 paper/paper.tex
First half of comments from AIM looked at

@@
 \begin{abstract}
-This paper presents the results of the LSST-DESC 3x2pt tomography challenge, which served as a first step toward optimizing the tomographic binning strategy for the main LSST 3x2pt cosmology analysis.
-The task of choosing an optimal tomographic binning scheme for a photometric survey is made particularly delicate in the context of a metacalibrated lensing catalog, as only the photometry from the bands included in the metacalibration process (traditionally, {\em riz} and potentially {\em g}) can be used in sample definition.
+This paper presents the results of the Rubin Observatory Dark Energy Science Collaboration (DESC) 3x2pt tomography challenge, which served as a first step toward optimizing the tomographic binning strategy for the main DESC analysis.
+The task of choosing an optimal tomographic binning scheme for a photometric survey is made particularly delicate in the context of a metacalibrated lensing catalog, as only the photometry from the bands included in the metacalibration process (usually {\em riz} and potentially {\em g}) can be used in sample definition.
 The goal of the challenge was to collect and compare bin assignment strategies under various metrics of a standard 3x2pt cosmology analysis in a highly idealized setting to establish a baseline for realistically complex follow-up studies;
@@
 %===========================
-
-% AIM: Suggestion: Use a macro for DESC because it shows up with and without expansion all over
 
@@
 measured in either Fourier or configuration space, and can be predicted from a combination of 
-perturbation theory and simulation results.  The method has been used in the Dark Energy Survey, DES
+perturbation theory and simulation results.  The method has been used in the Dark Energy Survey, DES, 
 \citep{des-3x2pt}, and to combine Baryon Oscillation Spectroscopic Survey, BOSS, with the Kilo-Degree 
@@
 
-% AIM: The next two paragraphs should present tomographic binning as an experimental design choice up front.
-
-Most lensing and 3x2pt analyses have been performed \emph{tomographically}, 
+Most lensing and 3x2pt analyses have chosen to analyze data \emph{tomographically}, 
 binning galaxies by redshift.
-% AIM: I get what this paragraph is saying, but it's worded ambiguously and can read as almost its exact opposite. I suggest rewording to specify more of the instances of "this," "that," and "it."
-This approach captures almost all the available information in lensing data, since it measures
+This approach captures almost all the available information in lensing data, since lensing measures
 an integrated effect and so galaxies nearby in redshift probe very similar fields.  For photometric
-foreground samples it is similarly near-lossless, since redshift estimates of such galaxies have
+foreground samples, tomography also loses little information when reasonably narrow bins are used,
+since redshift estimates of such galaxies have
 large uncertainties\footnote{Spectroscopic foreground samples may be more likely to see significant 
@@
 \citep{heavens,kitching}, the tomographic approach remains the standard within the field.
-
-% AIM: typo in the next sentence {potentially the ease}
+Tomographic 3x2pt measurements will be a key science goal in the upcoming \emph{Stage IV} surveys,
+including the Rubin Observatory \citep{rubin} and the Euclid and Roman space telescopes
+\citep{euclid,roman}.
+
 We are free to assign galaxies to different tomographic bins in any way we wish; changing the choice
-can potentially the ease with which we can calibrate
-the bins, but any choice can lead to correct results. The bins need not even be purely tomographic: we can happily correlate bins with peaks at different redshifts if that is useful.
+can potentially affect the ease with which we can calibrate
+the bins, but any choice can lead to correct results. The bins need not even be purely tomographic: we can happily correlate bins with number density peaks at different redshifts if that is useful.
 We should choose, then, an assignment algorithm that maximises the constraining power for a science
@@
 
-% AIM: Move the following sentence two paragraphs up to where tomography is introduced
-Tomographic 3x2pt measurements will be a key science goal in the upcoming \emph{Stage IV} surveys,
-including the Rubin Observatory \citep{rubin} and the Euclid and Roman space telescopes
-\citep{euclid,roman}. The general question of optimizing tomography was recently discussed in detail
+Various recent work have explored general tomographic binning strategies. The general question of optimization
+was recently discussed in detail
 in \citet{rainbow} using a self-organizing map approach to target up to five tomographic bins.  For this
@@
 for an optimized method, but note, importantly, that we should not be constrained to trying to directly
-match our bin selections to specific tomographic bin edges.
-% AIM: Above sentence should cite other strategies as well.
-
-% AIM: insert comma {paper we}
+match our bin selections to specific tomographic bin edges. \citet{taylor18} considered using
+fine tomographic bins as an alternative to fully 3D lensing, and found that the strategy of equal numbers
+per bin is less effective at high redshift, a result we will echo in this paper. The fine-binning strategy
+also allows a set of nulling methods to be applied, offering various advantages in avoiding poorly-understood
+regimes \citep{taylor18b,bnt,xcut}.
+
 % AIM: messaging -- establish scope with the basic info now in the abstract
-In this paper we discuss and evaluate the related challenge
+This paper is part of preparations for the analysis of Legacy Survey of Space and Time (LSST) data at the Rubin Observatory, being run by the Dark Energy Science Collaboration (DESC).
+In it, we discuss and evaluate the related challenge
 of using a limited set of colour bands to make those assignments, motivated by requirements of using the
@@
 In \autoref{sec:data} we describe the simulated data used in the challenge.
-\autoref{sec:methods} briefly summarizes the submitted methods; they are described in much more detail in \autoref{app:methods}.  In \autoref{sec:results} we discuss the
-results of the challenge, and then conclude in \autoref{sec:conclusion}.
+Section \ref{sec:results} briefly describes the entrants to the challenge, and discusses their performance (full
+method details are given in appendix \ref{app:methods}).
+ We conclude in \autoref{sec:conclusion}.
 
@@
 when PSFs are well-measured and oversampled by the pixel grid.  The method has been successfully
-applied in the Dark Energy Survey Year 1 and Year 3 analyses \citep{des-y1-cat, des-y3-cat},
+applied in the DES Year 1 and Year 3 analyses \citep{des-y1-cat, des-y3-cat},
 and application to early Rubin data is planned.
@@
 
-% AIM: 2020, not 2019!
-The Dark Energy Science Collaboration (DESC) Tomographic Challenge was launched in May 2019 and 
+The DESC Tomographic Challenge was launched in May 2020 and 
 advertised among the LSST cosmology community, though it was open to outside entrants.
@@
 different bins.
-% AIM: We must explicitly state the assumptions that all methods were expected to be ML/training-based and that this is framed as a classification problem, and explain how that fits with the paradigm of how people think about tomography, which is traditionally framed as a very different redshift regression problem.
+
+This differs from many redshift estimation questions because we are not at all concerned with estimates
+of individual galaxy redshifts; instead, this challenge was desgined as a classification and bin-optimization
+problem, amenable to more direct machine-learning methods.  As such we used the training / validation / testing
+approach, as standard in machine-learning analyses.
+The determination of the $n(z)$ of the recovered
+bins (whether with per-galaxy estimates or otherwse) was not part of the challenge.
 
@@
 
-% AIM: Was it actually advertised in Rubin beyond DESC at all?
 The challenge was open to any entrants, not just those already involved in DESC, but it was
 advertised primarily through Rubin, and so is perhaps best described as semi-public.
-% AIM: "tested and ran" --> "developed and tested" 
-% AIM: remove comma {locally, but}
-% AIM: Final tests used to generate results on secret test set were conducted on NERSC, no?
-Participants tested and ran their methods locally, but submitted code to be run as part of a central
-combined analysis on the NERSC supercomputing facility.  No prize was offered apart from recognition.
-% AIM: Don't specify NERSC up here before it's defined in Sec. 3.4.
+Participants developed and tested their methods locally but submitted code to be run as part of a central
+combined analysis at the National Energy Research
+Supercomputing Center (NERSC), where the final tests and metric evaluation were conducted.  No prize was offered apart from recognition.
 
@@
 	\includegraphics[width=0.9\linewidth]{results/initial_data.pdf}
-	% AIM: use alpha \neq 1
-	\caption{The underlying number density $n(z)$ for the two catalogs used in the challenge,
+	\caption{The underlying redshift number density $n(z)$ for the two catalogs used in the challenge,
 		after the initial cuts described in the text.}
-	% AIM: specify /redshift/ number density
 	\label{fig:initial_nz}
@@
 	\includegraphics[width=0.9\linewidth]{results/colour_colour.pdf}
-	% AIM: use alpha \neq 1
 	\caption{A colour-colour diagram for the two catalogs used in the challenge, showing the bands
@@
 
-% AIM: Rest of paper uses past tense so "run" --> "ran"
-% AIM: Train/Validate/Test set splitting should be described in post-processing rather than preamble, and assumption of ML/training-based approach should be stated explicitly in preamble to entire section.
-We run challenge entries on two data sets, CosmoDC2 and Buzzard, each of which is split into training, 
+We ran challenge entries on two data sets, CosmoDC2 and Buzzard, each of which is split into training, 
 validation, and testing components.  In each case participants were supplied with the training and 
@@
 
-% AIM: We don't have to justify the second dataset unless we provide a reason why the first was insufficient. . . the presentation here needs some reworking.
-We used the second Buzzard catalog for two reasons.  First, to check the significance of
-a limitation of galaxy colour assignments in CosmoDC2, described below.  And second, to check
+Using two catalogs allows us to check
 for over-fitting in the design of models themselves, and thus to determine whether two
@@
 
-% AIM: typo {Each of these data set}
-Each of these data set come from cosmological simulations which provide truth values of galaxy magnitudes
+Each of these data sets come from cosmological simulations which provide truth values of galaxy magnitudes
 and sizes.  We simulate and add noise to the objects as described below.
@@
 
-The CosmoDC2 simulation was designed to support the Dark Energy Science
-Collaboration and is described in detail in \citet{cosmodc2}.  It covers 440
+The CosmoDC2 simulation was designed to support the DESC and 
+is described in detail in \citet{cosmodc2}.  It covers 440
 deg${}^2$ of sky, and used the dark matter particles from the Outer Rim
@@
 
-After the post-processing and selections described below, and a cut to a specific geographic region, 
-we used around 36M objects from the CosmoDC2 simulations.
-
 One limitation of the CosmoDC2 catalogs, found as part of the challenge,
@@
 It was unknown whether this limitation would have any practical impact on the challenge,
-and was one of the reasons for adpoting the additional Buzzard catalog.
+and was another reason for adpoting the additional Buzzard catalog.
 
@@
 
-% AIM: Say it was also used for DESC DC1 studies, cite e.g. PZ DC1 paper
-The Buzzard catalog was developed to support Dark Energy Survey Year 1 (DES-Y1)
-data analysis, and is described in detail in \citet{buzzard}.  It used dark
+The Buzzard catalog was developed to support DES Year 1 (DES-Y1)
+data analysis, and is described in detail in \citet{buzzard}.
+It has previously been used in DESC analyses, for example in \citet{dc1_pz}.
+The catalog used dark
 matter simulations from L-GADGET2 \cite{gadget2} and then added galaxies using
@@
 
-After the post-processing and selections described below, and a cut to a
-specific geographic region,  we used around 20M objects from the Buzzard
-simulations.
-
 \subsubsection{Post-processing} \label{sec:data-proc}
-% AIM: Put the info from CosmoDC2/Buzzard pertaining to status after post-processing steps here instead.
 
@@
 
-% AIM: specify /redshift/ number density and provide n(z) symbol because it is used without introduction later
-The overall number density of the two data sets is shown in Figure \ref{fig:initial_nz},
+After this seection, and cuts to contiguous geographic regions, we used around 20M objects from the Buzzard
+simulations and around 36M objects the CosmoDC2 simulations.
+
+The overall number density $n(z)$ of the two data sets is shown in Figure \ref{fig:initial_nz},
 and a set of colour-colour diagrams  in Figure \ref{fig:colour_colour}.
-% AIM: If submitting to AAS journal, colour --> color; if submitting to MNRAS, catalog --> catalogue
 
@@
 \label{sec:metrics}
-% AIM: Change subsubsection titles to use the shorthand names for the metrics as they are referenced throughout the rest of the paper (SNR, w_{0}-w_{a} FOM, \Omega_{c}-\sigma_{8} FOM).
-% AIM: This should state that the three metrics are evaluated for clustering, lensing, and 3x2pt, so there are really nine metrics, not two!
 
@@
 and because clustering-like statistics of lensing samples are important for studies of intrinsic
-alignments. We use two different sets of metrics, one based on the over signal to noise of the spectra 
+alignments. We use three different sets of metrics, one based on the over signal to noise of the spectra 
 derived from the samples, and one based on a figure-of-merit for the constraints that would be obtained 
 using them. The relationships between these metrics in our results is discussed in \autoref{sec:metric-results}.
+We compute each metric on three different sets of summary statistics: lensing-alone, clustering-alone,
+and the full 3x2pt. We therefore compute a total of nine values for each bin selection.
 
@@
 Metric 1 would be approximately measurable on real data without performing a full cosmology analysis,
-whereas metric 2 would be the final output of such an analysis.  They therefore test performance at what
-would be multiple stages in the wider pipeline.
-% AIM: Sounds dumb, but readers will appreciate a clear statement that higher is better for these metrics.
-
-\subsubsection{Metric 1: Sigal-to-noise}
-% AIM: Suggestion: merge Appendix A with this section because this really doesn't stand by itself
-
-Our first metric is the total signal-to-noise ratio of the power spectrum derived from the assigned
+whereas metrics 2 and 3 would be the final output of such an analysis.  They therefore test performance at what
+would be multiple stages in the wider pipeline. In each case the metrics are constructed so that a larger
+value is a better score.
+
+\subsubsection{Metric 1: SNR}
+
+Our first metric is the total signal-to-noise ratio (SNR) of the power spectrum derived from the assigned
 $n_i(z)$ values:
@@
 
-\subsubsection{Metric 2: Figure-of-merit}
-% AIM: The title of this subsubsection should reflect that there are two FOM metrics.
+\subsubsection{Metrics 2 \& 3: $w_{0}-w_{a}$ and $\Omega_c-\sigma_8$ FOMs}
 
@@
 
-% AIM: We must acknowledge GitHub as a platform that enabled this challenge.
-% AIM: Suggestion to specify "subclass" and "superclass" up front for extra-thorough clarity.
 The challenge was structured as a python library, in which each classification
-method was a class.  The abstract base parent class, {\sc Tomographer}, 
-and other challenge machinery, performed initialization,
+method was a subclass of an abstract base parent superclass, {\sc Tomographer}.
+The superclass, and other challenge machinery, performed initialization,
 configuration option handling, data loading, and computing derived quantities
@@
 Methods were submitted to the challenge in the form of a pull request to the challenge
-repostitory\footnote{\url{http://github.com/LSSTDESC/tomo_challenge}}.
+repostitory on GitHub\footnote{\url{http://github.com/LSSTDESC/tomo_challenge}}.
 
@@
 and hyper-parameter tuning. Once the challenge period was complete, the algorithms
-were collected from the pull requests and run at the National Energy Research
-Supercomputing Center (NERSC). If a method failed due to a difference in the
+were collected from the pull requests and run at NERSC. If a method failed due to a difference in the
 runtime environment or the hardware at NERSC, the participants and organizers tried
@@
 
-% AIM: Include the code names for these dummy/control methods up front
-Two control methods were used to test the challenge machinery and metrics.
-% AIM: Explicitly put this into the context of a controlled experiment.
-% AIM: Make a new paragraph for each control method.
-The first method randomly assigned galaxies to each tomographic bin, resulting
+Two ``control'' methods were used to test the challenge machinery and ensure
+that metrics behaved appropriately in specific limits.
+
+The \textsc{Random} algorithm randomly assigned galaxies to each tomographic bin, resulting
 in bins which each had the same $n(z)$, on average (the random noise fluctuations
@@
 
-Second, we used a method which used only the $i$-band to select tomographic bins,
+The \textsc{IBandOnly} method used only the $i$-band to select tomographic bins,
 splitting galaxies in equal-sized bins in this magnitude. We expect only a small
@@
 
-\section{Submitted Methods} \label{sec:methods}
-% AIM: This should be a subsection of Challenge Design section above since it's really just a table of the submitted methods directing readers to the appendix for any detail.
-
-% AIM: typo: "Twenty five" --> "Twenty-five"
-Twenty five methods were submitted to the challenge, including the control methods described above.
-Most used one machine learning method of one type or another to perform the primary classification,
+\section{Methods \& Results} \label{sec:results}
+
+Twenty-five methods were submitted to the challenge, including the control methods described above.
+Most used machine learning methods of various types to perform the primary classification,
 rather than trying to perform a full photometric redshift analysis.  
-
-% AIM: These should be a pagewidth table, and we should use macros so the method names display consistently throughout paper.
-These methods are described in full in \autoref{app:methods}.  Briefly, they were (with the submitting authors listed in brackets):
-% \setlength{\itemsep}{0pt}%
-\begin{description*}
-	\item[\textsc{Random}] the random assignment control method (organizers)
-	\item[\textsc{IBandOnly}] the i-band only control method (organizers)
-	\item[\textsc{RandomForest}] a collection of individual decision trees (organizers)
-	\item[\textsc{LSTM}] a Long Short-Term Memory neural network (CRB, BMOF, GT, ESC, EJG)
-	\item[\textsc{AutokerasLSTM}] an automaticaly configured LSTM variant\footnote{A technical problem, involving conflicting versions of libraries and incompatible hardware, meant that AutoKerasLSTM could not be run in the final challenge.} (CRB, BMOF, GT, ESC, EJG)
-	\item[\textsc{LGBM}] a Light Gradient Boosted Machine, a histogram-based tree algorthm (CRB, BMOF, GT, ESC, EJG)
-	\item[\textsc{TCN}] a Temporal Convolution Network (CRB, BMOF, GT, ESC, EJG)
-	\item[\textsc{CNN}] a Convolutional Neural Network (CRB, BMOF, GT, ESC, EJG)
-	\item[\textsc{Ensemble}] a combination of three other network methods (CRB, BMOF, GT, ESC, EJG)
-	\item[\textsc{SimpleSOM}] a Self-Organizing Map (AHW)
-	\item[\textsc{PQNLD}] an extension of {\sc SimpleSom} combining with template-based redshift estimation (AHW)
-	\item[\textsc{ComplexSOM}] an extension of {\sc SimpleSom} with an additional assignment optimization (DA, AHW, AN, EG, AB)
-	\item[\textsc{GPzBinning}] a Gaussian Process redshift estimator and binner (PH)
-	\item[\textsc{JaxCNN/JaxResnet}] two related CNN-based bin edge optimizers (AP)
-	\item[\textsc{NeuralNetwork1/2}] a dense network optimizing bin assignment for two metrics (AP)
-	\item[\textsc{PCACluster}] a Principal Component Analysis of fluxes then optimized clustering (DG)
-	\item[\textsc{UTOPIA}] a nearest-neighbour method, optimised in the limit of large representative training sets (AHW)
-	\item[\textsc{ZotBin/ZotNet}] two related neural network methods with a preprocesing normalizing flow (DK)
-	\item[\textsc{FFNN}] a feed-forward neural network (EN)
-	\item[\textsc{FunBins}] a random forest with various nominal bin edge selection (AB)
-	\item[\textsc{MLPQNA}] a multi-layer perceptron (SC and MB)
-	\item[\textsc{Stacked Generalization}] an combination of gradient tree boosting and logistic regression (JEC)
-\end{description*}
-
-\section{Results \& Discussion} \label{sec:results}
-% AIM: Convert the first two subsections to the preamble of this section.
-% AIM: Preamble should include the tl;dr of "winners."
+These methods are listed in \autoref{tab:entrants} and described in full in \autoref{app:methods}.
+
+\begin{table*}[]
+	\begin{tabular}{|l|p{8cm}|c|}
+	\hline
+	\textbf{Name} & \textbf{Description} & \textbf{Submitters} \\
+	\hline
+	\textsc{Random} & Random assignment (control method)  & Organizers \\
+	\textsc{IBandOnly} & i-band only (control method)  & Organizers \\
+	\textsc{RandomForest} & A collection of individual decision trees & Organizers \\
+	\textsc{LSTM} & Long Short-Term Memory neural network & CRB, BF, GT, ESC, EJG \\
+	\textsc{AutokerasLSTM}\tablenotemark{a} & Automaticaly configured LSTM variant. & CRB, BF, GT, ESC, EJG \\
+	\textsc{LGBM} & Light Gradient Boosted Machine & CRB, BF, GT, ESC, EJG \\
+	\textsc{TCN} & Temporal Convolution Network & CRB, BF, GT, ESC, EJG \\
+	\textsc{CNN} & Convolutional Neural Network & CRB, BF, GT, ESC, EJG \\
+	\textsc{Ensemble} & Combination of three other network methods & CRB, BF, GT, ESC, EJG \\
+	\textsc{SimpleSOM} & Self-Organizing Map & AHW \\
+	\textsc{PQNLD} & Extension of {\sc SimpleSom} combining with template-based redshift estimation & AHW \\
+	\textsc{UTOPIA} & Nearest-neighbours, optimised in the limit of large representative training sets & AHW \\
+	\textsc{ComplexSOM} & Extension of {\sc SimpleSom} with an additional assignment optimization & DA, AHW, AN, EG, AB \\
+	\textsc{GPzBinning} & Gaussian Process redshift estimator and binner & PH \\
+	\textsc{JaxCNN/JaxResnet} & Two related CNN-based bin edge optimizers & AP \\
+	\textsc{NeuralNetwork1/2} & Dense network optimizing bin assignment for two metrics & AP \\
+	\textsc{PCACluster} & Principal Component Analysis of fluxes then optimized clustering & DG \\
+	\textsc{ZotBin/ZotNet} & Two related neural network methods with a preprocesing normalizing flow & DK \\
+	\textsc{FFNN} & Deed-forward neural network & EN \\
+	\textsc{FunBins} & Random forest with various nominal bin edge selection & AB \\
+	\textsc{MLPQNA} & Multi-layer perceptron & SC, MB \\
+	\textsc{Stacked Generalization} & Combination of gradient tree boosting and logistic regression & JEC\\
+	\hline
+	\end{tabular}
+	\caption{Methods entered into the challenge. The algorithms are described more fully in \autoref{app:methods}.}
+	\label{tab:entrants}
+	\tablenotetext{a}{A technical problem, involving conflicting versions of libraries and incompatible hardware, meant that AutoKerasLSTM could not be run in the final challenge.}
+\end{table*}
 
@@
 
-% AIM: Move this table to the appendix.
 \begin{table*}[]
@@
 
-\subsection{Overview}
+\subsection{Results Overview}
 % AIM: Suggestion: Convert this subsection to a bulleted list to give it some structure harking to the detailed subsections to come.
@@
 Figure \ref{fig:funbin_nz} shows the number density obtained by this method, for both the riz and griz 
-scenarios, and an additional post-challenge run using the full Rubin $ugrizy$ bands.
-% AIM: Split paragraph here to distinguish topics of overall challenge winner and conclusions about band information limitation.
+scenarios, and an additional post-challenge run using the full Rubin $ugrizy$ bands, the metric results for
+which were close to those of the $griz$ analysis.
+
 % AIM: Put the caveat at the beginning of the following sentence rather than the end. Also, this claim is not well-supported even by Fig. 3. . . 
@@
 the information needed for tomography are unfounded, at least in our idealised scenario.
-In both the CosmoDC2 and Buzzard analyses,
-% AIM: analyses --> data sets?
+In both the CosmoDC2 and Buzzard data sets,
 even the $riz$ bands alone are sufficient to split objects into nine reasonably separated
@@
 Methods \ref{ClecioLSTM} - \ref{ClecioEnsemble} were
-submitted by a team of authors CRB, BMOF, GT, ESC, and EJG.  All were built using TensorFlow and/or Keras, and
+submitted by a team of authors CRB, BF, GT, ESC, and EJG.  All were built using TensorFlow and/or Keras, and
 chose bin edges such that an equal number of galaxies is assigned to each bin, and trained the

### 2021-03-25 paper/results/make_figures.py

@@ -7,6 +7,7 @@ import matplotlib
 import subprocess
 import h5py
 import collections
+import matplotlib.pyplot as plt
 # set default styles
 
 
@@ -353,7 +354,7 @@ def make_initial_nz(filename):
     ax.hist(buzz_z, bins=50, histtype='step', label='Buzzard', linewidth=3, weights=w1)
     del buzz_z
     dc2_z = dc2['redshift_true'][:]
-    ax.hist(dc2_z, bins=50, histtype='step', label='CosmoDC2', linewidth=3, weights=w2)
+    ax.hist(dc2_z, bins=50, histtype='step', label='CosmoDC2', linewidth=3, weights=w2, alpha=0.7)
     ax.set_xlabel("z")
     ax.set_ylabel("Counts / $10^5$")
     ax.legend(frameon=True)
@@ -362,7 +363,7 @@ def make_initial_nz(filename):
     plt.close(fig)
 
 
-def colour_colour_cat(ax, cat, colour, thin, label):
+def colour_colour_cat(ax, cat, colour, thin, label, alpha):
     if 'g_mag' in cat.keys():
         fmt = '{}_mag'
     else:
@@ -377,8 +378,8 @@ def colour_colour_cat(ax, cat, colour, thin, label):
     gz = g - z
     iz = i - z
 
-    ax[0].scatter(gr, ri, c=colour, s=0.5, label=label)
-    ax[1].scatter(gz, iz, c=colour, s=0.5, label=label)
+    ax[0].scatter(gr, ri, c=colour, s=0.5, label=label, alpha=alpha)
+    ax[1].scatter(gz, iz, c=colour, s=0.5, label=label, alpha=alpha)
 
     ax[0].set_xlabel("g - r")
     ax[0].set_ylabel("r - i")
@@ -401,8 +402,8 @@ def make_colour_colour(filename):
     buzz = h5py.File(buzzard_file, 'r')
     dc2 = h5py.File(dc2_file, 'r')
 
-    colour_colour_cat(ax, buzz, blue_color, 4000, "Buzzard")
-    colour_colour_cat(ax, dc2, orange_color, 8000, "CosmoDC2")
+    colour_colour_cat(ax, buzz, blue_color, 4000, "Buzzard", alpha=1)
+    colour_colour_cat(ax, dc2, orange_color, 8000, "CosmoDC2", alpha=0.5)
 
     ax[0].legend(frameon=True)
     
[…]

### 2021-03-25 paper/paper.tex
more malzing

@@
 
-% AIM: messaging -- establish scope with the basic info now in the abstract
 This paper is part of preparations for the analysis of Legacy Survey of Space and Time (LSST) data at the Rubin Observatory, being run by the Dark Energy Science Collaboration (DESC).
@@
 redshift as much as possible.  Particpants were free to optimize or simply select nominal (target) 
-redshift bin edges from their training sample, but in either case had to classify galaxies into the 
+redshift bin edges from their training sample, but in either case had to assign galaxies into the 
 different bins.
@@
 
-The challenge was structured as a python library, in which each classification
+The challenge was structured as a python library, in which each assignment
 method was a subclass of an abstract base parent superclass, {\sc Tomographer}.
@@
 \subsection{Results Overview}
-% AIM: Suggestion: Convert this subsection to a bulleted list to give it some structure harking to the detailed subsections to come.
-The short versions are long enough to come off as unsupported statements rather than previews of what will be demonstrated in subsequent subsections.
-
 The DETF results for each method are shown in Table \ref{tab:cosmodc2}.  Complete results for all metrics
-for the nine-bin scenario only are shown in appendix tables \ref{tab:full_dc2} and \ref{tab:full_buzz}.  
-% AIM: Use autoref to appendix.
+for the nine-bin scenario only are shown in appendix tables \ref{tab:full_dc2} and \autoref{tab:full_buzz}.
+
 In all these tables, missing entries are shown for two reasons.
@@
 small number count, such that the spectrum could not be calculated.
-
-% AIM: This is the first mention of this as a classification problem.
-We should mention this up in experimental design along with the assumption of ML/training-based methods, but also change "clasified" [SIC] to "assigned" for overall consistency (and maybe "that entrants clasified" --> "entrants to classify," and maybe reword post-semicolon part as "aside from the impact of reduced final sample size on the metrics, no penalty was applied for discarding galaxies entirely").
-% AIM: Also, the discussion of non-assignment should be its own subsection.
-The challenge did not require that entrants clasified all galaxies; ambiguous
-galaxies could be discarded at the cost of number density.  Several methods
-took advantage of this.  The ComplexSom method (see \autoref{sec:csom}) found
-it did not improve scores, but as noted in \autoref{sec:pqnld} this may be an artifact
-of the high density of the training data, and so could be explored in future challenges.
 
@@
 Figure \ref{fig:funbin_nz} shows the number density obtained by this method, for both the riz and griz 
-scenarios, and an additional post-challenge run using the full Rubin $ugrizy$ bands, the metric results for
-which were close to those of the $griz$ analysis.
-
-% AIM: Put the caveat at the beginning of the following sentence rather than the end. Also, this claim is not well-supported even by Fig. 3. . . 
-It is clear from this that fears about restricted photometry catastrophically limiting
-the information needed for tomography are unfounded, at least in our idealised scenario.
+scenarios, and an additional post-challenge run using the full Rubin $ugrizy$ bands.  The metrics
+obtained with \textsc{funbins} on $ugrizy$ are barely higher than those with $griz$ -- the largest
+increase is 8\% for the $\Omega_c-\sigma_8$ metric, with the remainder all 5\% or less.
+
+It is clear from this that, at least in our idealised scenario,  fears about restricted photometry 
+catastrophically limiting
+the information needed for tomography are unfounded.
 In both the CosmoDC2 and Buzzard data sets,
@@
 methods.
+
+\subsection{Non-assignment}
+The challenge did not require entrants to assign all galaxies to bins; aside from the impact of reduced final sample size on the metrics, no penalty was applied for discarding galaxies entirely.
+Several methods
+took advantage of this.  The ComplexSom method (see \autoref{sec:csom}) found
+it did not improve scores, but as noted in \autoref{sec:pqnld} this may be an artifact
+of the high density of the training data, and so could be explored in future challenges.
 
@@
 \section{Mistakes} \label{sec:mistakes}
-% AIM: This should not be an appendix and should instead be part of the Results & Discussion section.
-They should not be presented as "mistakes" -- they are part of the experimental design and constitute caveats, which are simply facts, not subjective judgements.
-
 The challenge organizers made a number of mistakes when building and running the challenge.
@@
 
-\section{Full 9-bin results table} \label{app:tables}
+\section{Full 9-bin results tables} \label{app:tables}
 Tables \ref{tab:full_dc2} and \ref{tab:full_buzz} show all calculated metrics when generating nine tomographic 

### 2021-03-29 paper/paper.tex
UK spelling of catalogue

@@
 This paper presents the results of the Rubin Observatory Dark Energy Science Collaboration (DESC) 3x2pt tomography challenge, which served as a first step toward optimizing the tomographic binning strategy for the main DESC analysis.
-The task of choosing an optimal tomographic binning scheme for a photometric survey is made particularly delicate in the context of a metacalibrated lensing catalog, as only the photometry from the bands included in the metacalibration process (usually {\em riz} and potentially {\em g}) can be used in sample definition.
+The task of choosing an optimal tomographic binning scheme for a photometric survey is made particularly delicate in the context of a metacalibrated lensing catalogue, as only the photometry from the bands included in the metacalibration process (usually {\em riz} and potentially {\em g}) can be used in sample definition.
 The goal of the challenge was to collect and compare bin assignment strategies under various metrics of a standard 3x2pt cosmology analysis in a highly idealized setting to establish a baseline for realistically complex follow-up studies;
@@
 very accurately, by performing the cut on both
-sheared variants of the catalog, and measuring the difference between the post-cut shear in the two
+sheared variants of the catalogue, and measuring the difference between the post-cut shear in the two
 variants.  In DES the corrected biases were found up to around 4\%, far larger than
@@
 spectra are time-consuming to obtain.  These training sets will also be highly incomplete compared
-to full catalogs, especially at faint magnitudes where spectroscopic coverage is sparse.  In this
+to full catalogues, especially at faint magnitudes where spectroscopic coverage is sparse.  In this
 challenge we avoided both these issues -- the training samples we provided (see \autoref{sec:data}) were comparable
@@
 	\includegraphics[width=0.9\linewidth]{results/initial_data.pdf}
-	\caption{The underlying redshift number density $n(z)$ for the two catalogs used in the challenge,
+	\caption{The underlying redshift number density $n(z)$ for the two catalogues used in the challenge,
 		after the initial cuts described in the text.}
@@
 	\includegraphics[width=0.9\linewidth]{results/colour_colour.pdf}
-	\caption{A colour-colour diagram for the two catalogs used in the challenge, showing the bands
+	\caption{A colour-colour diagram for the two catalogues used in the challenge, showing the bands
 		available to participants.}
@@
 
-Using two catalogs allows us to check
+Using two catalogues allows us to check
 for over-fitting in the design of models themselves, and thus to determine whether two
@@
 
-One limitation of the CosmoDC2 catalogs, found as part of the challenge,
+One limitation of the CosmoDC2 catalogues, found as part of the challenge,
 was that the matching to Galacticus led to only a limited number of SEDs being
@@
 It was unknown whether this limitation would have any practical impact on the challenge,
-and was another reason for adpoting the additional Buzzard catalog.
+and was another reason for adpoting the additional Buzzard catalogue.
 
@@
 
-The Buzzard catalog was developed to support DES Year 1 (DES-Y1)
+The Buzzard catalogue was developed to support DES Year 1 (DES-Y1)
 data analysis, and is described in detail in \citet{buzzard}.
 It has previously been used in DESC analyses, for example in \citet{dc1_pz}.
-The catalog used dark
+The catalogue used dark
 matter simulations from L-GADGET2 \cite{gadget2} and then added galaxies using
@@
 and were shown to be a good
-match (after appropriate selections) to DES-Y1 observable catalogs.  Magnitudes
-using the LSST band passes were provided in the catalog.
+match (after appropriate selections) to DES-Y1 observable catalogues.  Magnitudes
+using the LSST band passes were provided in the catalogue.
 
@@
 
-We add noise to the truth values in the extra-galactic catalogs to simulate real observations.
+We add noise to the truth values in the extra-galactic catalogues to simulate real observations.
 In each case we simulate first year Rubin observations
@@
 In both simulations we add two cuts to approximately simulate the selection used in a real
-lensing catalog.  In both cases we apply a cut on the combined $riz$ signal to noise of
+lensing catalogue.  In both cases we apply a cut on the combined $riz$ signal to noise of
 $S/N > 10$, and a size cut $T / T_\mathrm{psf} > 0.5$, where

### 2021-03-29 paper/paper.tex
discussa mistake earlier

@@
 
+This meant there were two ways to obtain higher scores: either to improve the choice of nominal bin
+edges, or to find better ways to assign objects to those edges.  This was a sub-optimal design decision:
+separating the two issues would have enabled us to more easily delineate the advantages of different methods. 
+This can be explored retrospectively, since we know what approach methods took, but would have been
+easier in advance.  This and other mistakes in designing the challenge are described in \autoref{sec:mistakes}.
+
 This differs from many redshift estimation questions because we are not at all concerned with estimates
@@
 The challenge was open to any entrants, not just those already involved in DESC, but it was
-advertised primarily through Rubin, and so is perhaps best described as semi-public.
+advertised through Rubin, and so is perhaps best described as semi-public.
 Participants developed and tested their methods locally but submitted code to be run as part of a central
@@
 \subsection{Splitting nominal bin choice from assignment}
-There were two ways to obtain good scores in the challenge: either one could choose the best
+As noted in the main test, there were two ways to obtain good scores in the challenge: 
+either a team could choose the best
 nominal bin edges, or find the best method to put galaxies into those bins once they are chosen.

### 2021-04-12 paper/paper.tex
add new figs

@@
 	% \centering
-	\includegraphics[width=1\linewidth]{results/funbins_nz.pdf}
+	\includegraphics[width=1\linewidth]{results/nzbinned_dc2_funbins.pdf}
 	\caption{The tomographic $n(z)$ values generated by the winning method in the challenge, {\sc funbins},
@@
 For comparison, we also show results using all six LSST bands, $ugrizy$; these could not be used
-with the metacalibration method. The improvement in the sharpness of the bins is noticeable but not large,
-and does not translate into large increases in the figures of merit: the DETF score for CosmoDC2
-is 169.7, barely higher than the score for $griz$, and this is also true for other metrics.   While this is only a general indication and not a thorough exploration, it is encouraging.
+with the metacalibration method. The improvement in the sharpness of the bins is noticeable but not large.
+We quantify this in Figure \ref{fig:overlap}, which shows the fraction of objects assigned to each
+bin that are actually in another redshift range, together with the total fraction of mis-assigned galaxies 
+over all the bins, for each set of bands. This illustrates how adding the $g$ band improves the assigment
+for almost every pair of bins, and reduce the total mis-assigned fraction by a factor of three to the 
+point where a significant majority of galaxies are in the right bin.  Further adding the remaining $u$ and $y$
+bands improves the assignment in a way that is smaller in both relative and absolute terms.
+This latter addition also does not translate into large increases in the figures of merit: the 
+DETF score for CosmoDC2 is 169.7, barely higher than the score for $griz$, and this is also true for other 
+metrics.   While this is only a general indication and not a thorough exploration, it is encouraging.
+
+\begin{figure*}[htbp]
+% \centering
+\includegraphics[width=0.9\linewidth]{results/nzoverlap_dc2_funbins.pdf}
+\caption{The fraction of total objects mis-assigned from each bin to each other bin in the winning 
+\textsc{funbins} method for different choices of bands.  The $x$- and $y$-axes show tomographic bin indices, 
+and the color represents the fraction of objects (compared to the total number in the challenge) that are 
+mis-assigned from each bin to each other.  The total-overlap figure gives the total fraction of objects
+that are mis-assigned over all bins, and is the sum of all the values.}
+\label{fig:overlap}
+\end{figure*}
 

### 2021-04-14 paper/paper.tex
more text on overlap

--- 2021-03-29 paper/paper.tex
@@
 For comparison, we also show results using all six LSST bands, $ugrizy$; these could not be used
-with the metacalibration method. The improvement in the sharpness of the bins is noticeable but not large,
-and does not translate into large increases in the figures of merit: the DETF score for CosmoDC2
-is 169.7, barely higher than the score for $griz$, and this is also true for other metrics.   While this is only a general indication and not a thorough exploration, it is encouraging.
+with the metacalibration method. The improvement in the sharpness of the bins is noticeable but not large.
+We quantify this in Figure \ref{fig:overlap}, which shows the fraction of objects that are in ambiguous
+redshift slices, defined as objects in redshift ranges where there are more objects assigned to another 
+tomographic bin.  The total fraction of such objects is also shown. This illustrates how adding the $g$ band 
+improves the assigment
+for almost every pair of bins, and reduce the ambiguous fraction by a factor of three to the 
+point where a significant majority of galaxies are clearly assigned.  Further adding the remaining $u$ and $y$
+bands improves the assignment in a way that is smaller in both relative and absolute terms.
+This latter addition also does not translate into large increases in the figures of merit: the 
+DETF score for CosmoDC2 is 169.7, barely higher than the score for $griz$, and this is also true for other 
+metrics.   While this is only a general indication and not a thorough exploration, it is encouraging.
+
+\begin{figure*}[htbp]
+% \centering
+\includegraphics[width=0.9\linewidth]{results/nzoverlap_dc2_funbins.pdf}
+\caption{The fraction of total objects assigned to each tomographic bin that are ambiguously assigned, for 
+the \textsc{funbins} tomograph method and three choices of bands.  The $x$- and $y$-axes 
+show tomographic bin indices, 
+and the color represents the fraction of objects (compared to the total number in the challenge) that are 
+at redshifts dominated by different tomographic to the ones they are assigned.
+The total-overlap figure gives the total fraction of objects
+that are ambiguous over all bins, and is the sum of all the values.}
+\label{fig:overlap}
+\end{figure*}
 

### 2021-04-14 paper/paper.tex
more text

@@
 and the color represents the fraction of objects (compared to the total number in the challenge) that are 
-at redshifts dominated by different tomographic to the ones they are assigned.
+at redshifts dominated by different tomographic to the ones they are assigned.  This metric corresponds to
+the fraction of the area that is shaded in two colours in Figure \ref{fig:funbin_nz}.
 The total-overlap figure gives the total fraction of objects
-that are ambiguous over all bins, and is the sum of all the values.}
+that are ambiguous over all bins, and is the sum of all the grid values.}
 \label{fig:overlap}

### 2021-04-14 paper/paper.tex
caption update

@@
 \includegraphics[width=0.9\linewidth]{results/nzoverlap_dc2_funbins.pdf}
-\caption{The fraction of total objects assigned to each tomographic bin that are ambiguously assigned, for 
-the \textsc{funbins} tomograph method and three choices of bands.  The $x$- and $y$-axes 
+\caption{The pairwise fractions of total objects in redshift regions where multiple tomographic bins contain galaxies, for 
+the \textsc{funbins} tomography method and three choices of bands.  
+The $x$- and $y$-axes 
 show tomographic bin indices, 
 and the color represents the fraction of objects (compared to the total number in the challenge) that are 
-at redshifts dominated by different tomographic to the ones they are assigned.  This metric corresponds to
+at redshifts where both tomographic bins have members, in bins of width $\delta z=10^{-3}$.  
+This metric corresponds to
 the fraction of the area that is shaded in two colours in Figure \ref{fig:funbin_nz}.
 The total-overlap figure gives the total fraction of objects
-that are ambiguous over all bins, and is the sum of all the grid values.}
+that are ambiguous over all bins, and is the sum of all the pairwise values.}
 \label{fig:overlap}

### 2021-04-15 paper/paper.tex
more on stackgen and afils

@@
 individual estimator by using their outputs as input of a final estimator.  This entry stacked
-a \emph{Gradient Tree Boosting} (GTB) classifier and then used a Logistic Regression as the final step.
-
-GTB is a machine learning method which combines decision trees (like the Random Forest, RF)
+\emph{Gradient Boosted Trees} (GBT) classifiers with standard Random Forests (RFs), and then used a Logistic 
+Regression as the final step.
+
+A GBT is a machine learning method which combines decision trees (like the standard RF)
 by performing a weighted mean average of individual trees \citep{Friedman:2002we,RefWorks:1634}. 
@@
 The two methods have their own systematics; they notably differ in the feature importances. Stacking 
-the two is therefore especially effective.  The submitted method used {\sc SKLearn}, and used 50 models 
-of each type combined using its {\sc StackingClassifier}.
+the two is therefore especially effective.  The submitted method combined 50 GBTs with
+50 RFs using the {\sc SKLearn} {\sc StackingClassifier}.
 
@@
 This procedure was done by hand for this challenge, but could be automated and applied to any kind of classifier.
+
+Although this entry stacked GBTs and RFs, the same approach can be used more generally, such as using
+one classifier optimized for high redshift and another for low, or one classifier for an SNR metric
+and a second for an FOM metric.  This makes stacking approaches particularly flexible.
 

### 2021-05-04 paper/paper.tex
Update paper.tex
Minor typo
typo

@@
 
-After this seection, and cuts to contiguous geographic regions, we used around 20M objects from the Buzzard
+After this section, and cuts to contiguous geographic regions, we used around 20M objects from the Buzzard
 simulations and around 36M objects the CosmoDC2 simulations.
@@
 	\textsc{JaxCNN/JaxResnet} & Two related CNN-based bin edge optimizers & AP \\
-	\textsc{NeuralNetwork1/2} & Dense network optimizing bin assignment for two metrics & AP \\
+	\textsc{NeuralNetwork1/2} & Dense network optimizing bin assignment for two metrics & FL \\
 	\textsc{PCACluster} & Principal Component Analysis of fluxes then optimized clustering & DG \\

### 2021-05-14 paper/paper.tex
address markus and shahab notes

@@
 perturbation theory and simulation results.  The method has been used in the Dark Energy Survey, DES, 
-\citep{des-3x2pt}, and to combine Baryon Oscillation Spectroscopic Survey, BOSS, with the Kilo-Degree 
-Survey, KiDS \citep{kids-3x2pt}.
+\citep{des-3x2pt}, and to combine the Kilo-Degree 
+Survey, KiDS with spectroscopic surveys \citep{kids-3x2pt, kids_gama, kids_2df}.
 
@@
 was recently discussed in detail
-in \citet{rainbow} using a self-organizing map approach to target up to five tomographic bins.  For this
+in \citet{rainbow} using a self-organizing map (SOM) approach to target up to five tomographic bins.  For this
 configuration they find that equally spaced redshift bins are a good proxy
 for an optimized method, but note, importantly, that we should not be constrained to trying to directly
-match our bin selections to specific tomographic bin edges. \citet{taylor18} considered using
+match our bin selections to specific tomographic bin edges. 
+They also highlight the utility of rejecting outlier or hard-to-classify objects from samples, and the ability of SOMs
+to do so cleanly.
+\citet{taylor18} considered using
 fine tomographic bins as an alternative to fully 3D lensing, and found that the strategy of equal numbers
@@
 metacalibration method for galaxy shape measurement to limit biases.  
-A similar question, with different motivations, was explored in \citet{jain}, who find that 3-band 
-gri tomography could be effective for tomographic selection, a result we will echo here in a new context: in this paper we describe the results of a 
+A similar question, with different motivations, was explored in \citet{jain}, who found that 3-band 
+$gri$ tomography could be effective for tomographic selection, a result we will echo here in a new context: in this paper we describe the results of a 
 challenge using simulated Rubin-like data with a range of methods submitted by different teams.
@@
 In \autoref{sec:motivation} we explain the need for methods that work on limited sets of bands in the context of upcoming lensing surveys.
-In \autoref{sec:data} we describe the simulated data used in the challenge.
+In \autoref{sec:design} we describe the design of the challenge and the simulated data used in it.
 Section \ref{sec:results} briefly describes the entrants to the challenge, and discusses their performance (full
@@
 magnitude.  It follows that a cut (or re-weighting) on galaxy magnitudes, or on any quantity derived
-from  them, will induce a shear bias since galaxies will preferentially fall either side of the cut
+from  them, will induce a shear bias since galaxies will preferentially fall on either side of the cut
 depending on their shear.  Within metacal, we can compute and thus account for this bias
@@
 therefore study how well we can perform tomography using only this limited photometric information.  In
-particular, this limitation prevent us from using the most obvious method for tomography,
+particular, this limitation prevents us from using the most obvious method for tomography,
 and computing a complete photometric probability density function (PDF) for each galaxy and assigning
@@
 real data to high accuracy.  If we determine that limiting the bands we can use for tomography results
-in a significant decrease in our overall signal-to-noise, outweighing the gain from the improved shape
+in a significant decrease in our overall 3x2pt signal-to-noise, outweighing the gain from the improved shape
 calibration then this might suggest moving to rely more on such methods\footnote{One can account for
@@
 factor.  Widening the prior on this factor decreases the overall constraining power of the analysis.}.
-
-\section{Challenge Design}
-
-The DESC Tomographic Challenge was launched in May 2020 and 
-advertised among the LSST cosmology community, though it was open to outside entrants.
+Constructing simulations with the required fidelity is challenging at Rubin-level accuracy and requires
+careful comparison to deep field data.
+
+\section{Challenge Design} \label{sec:design}
+
+The DESC Tomographic Challenge opened in May 2020 and accepted entries until September 2020.
 
@@
 using only their (g)riz bands and associated errors, and the galay size.  Their goal was to maximize the cosmological
-information in the final split data set; this is generally acheived by separating different bins by 
+information in the final split data set; this is generally achieved by separating different bins by 
 redshift as much as possible.  Particpants were free to optimize or simply select nominal (target) 
@@
 The determination of the $n(z)$ of the recovered
-bins (whether with per-galaxy estimates or otherwse) was not part of the challenge.
+bins (whether with per-galaxy estimates or otherwise) was not part of the challenge.
 
@@
 
-The CosmoDC2 simulation was designed to support the DESC and 
+The CosmoDC2 simulation was designed to support DESC and 
 is described in detail in \citet{cosmodc2}.  It covers 440
@@
 
-The Buzzard catalogue was developed to support DES Year 1 (DES-Y1)
+The Buzzard catalogue was developed to support the DES Year 1 (DES-Y1)
 data analysis, and is described in detail in \citet{buzzard}.
@@
 After this section, and cuts to contiguous geographic regions, we used around 20M objects from the Buzzard
-simulations and around 36M objects the CosmoDC2 simulations.
+simulations and around 36M objects from the CosmoDC2 simulations.
 
@@
 and because clustering-like statistics of lensing samples are important for studies of intrinsic
-alignments. We use three different sets of metrics, one based on the over signal to noise of the spectra 
-derived from the samples, and one based on a figure-of-merit for the constraints that would be obtained 
+alignments. We use three different sets of metrics, one based on the overall signal to noise of the spectra 
+derived from the samples, and two based on a figure-of-merit for the constraints that would be obtained 
 using them. The relationships between these metrics in our results is discussed in \autoref{sec:metric-results}.
@@
 where $\mu$ is a stack of the theoretical predictions for the $C_\ell$ spectra for each tomographic 
-bin, in turn, and $C$ a Gaussian estimate of the covariance between them, as described in
+bin, in turn, and $C$ a estimate of the covariance between them, making the approximation that the underlying density field has a Gaussian distribution as described in
 \citet{takada_jain} and summarized in \autoref{app:theory}.   We compute this metric for lensing alone, clustering alone, and the full 3x2pt combination.
@@
 evolution, and a version in which $p_1 = \Omega_c$ and $p_2 = \sigma_8$, representing constraining
-power on overall cosmic structure amplitude and its evolution.
+power on overall cosmic structure amplitude and its evolution. In each case the complete parameter space consists
+of the eight $w_0w_aCDM$ cosmological parameters. This simplified measure does not include any nuisance
+parameters modelling systematic errors in the analysis (most importantly the galaxy bias parameters in the clustering 
+spectra), and like all Fisher analyses approximates posterior constraints as Gaussian.
 
@@
 The methods described above were run on a data set, previously unseen by entrants but drawn
-from the same underlying population, of size 8.6M objects for CosmoDC2 and 5.4M objects for Buzzard.
+from the same underlying population, of size 8.6M objects for CosmoDC2 (our fiducial simulation) and 5.4M objects for Buzzard.
 A complete suite of the metrics described in Section \ref{sec:metrics} was run on each of the two
@@
 
-The overall winning method in the 3x2pt case in the fiducial $griz$ 3x2pt CosmoDC2 metric was \textsc{funbins},
+Before the challenge began we somewhat arbitrarily chose 
+the 9-bin 3x2pt $griz$ CosmoDC2 $w_0-w_a$ metric as the fiducial scenario on which
+an overall winning method would be selected.
+By this criterion the best method was, by a small margin, \textsc{funbins},
 which used a random forest to assign bins, but chose nominal bin edges by splitting the range spanned
-by the sample into equal co-moving distance bins. As described in \autoref{sec:funbins},  this was designed
+by the sample into equal co-moving distance bins. 
+Other methods performed better in other branches of the challenge; in particular we note \textsc{ZotNet}, \textsc{Stacked Generalization}, and \textsc{NeuralNetwork} (see sections \ref{sec:zot}, \ref{sec:stackgen}, and \ref{sec:nn} respectively), which each won more than three 9-bin scenarios.
+
+As described in \autoref{sec:funbins},  \textsc{funbins} was designed
 to minimze the shot noise for the angular power in each bin, and this has proven successful. 
@@
 point where a significant majority of galaxies are clearly assigned.  Further adding the remaining $u$ and $y$
-bands improves the assignment in a way that is smaller in both relative and absolute terms.
+bands improves the assignment in a way that is smaller, though not insignificant, in both relative and absolute terms.
 This latter addition also does not translate into large increases in the figures of merit: the 
-DETF score for CosmoDC2 is 169.7, barely higher than the score for $griz$, and this is also true for other 
+DETF score for CosmoDC2 is 169.7, barely higher than its score of 167.2 for $griz$, and this is also true for other 
 metrics.   While this is only a general indication and not a thorough exploration, it is encouraging.
@@
 \includegraphics[width=0.9\linewidth]{results/nzoverlap_dc2_funbins.pdf}
-\caption{The pairwise fractions of total objects in redshift regions where multiple tomographic bins contain galaxies, for 
+\caption{The pairwise fractions $f_\mathrm{mis}$ of total objects in redshift regions where multiple tomographic bins contain galaxies, for 
 the \textsc{funbins} tomography method and three choices of bands.  
@@
 \subsection{Non-assignment}
-The challenge did not require entrants to assign all galaxies to bins; aside from the impact of reduced final sample size on the metrics, no penalty was applied for discarding galaxies entirely.
+The challenge did not require entrants to assign all galaxies to bins; aside from the impact of reduced final sample size on the metrics, no additional penalty was applied for discarding galaxies entirely.
 Several methods
@@
 it did not improve scores, but as noted in \autoref{sec:pqnld} this may be an artifact
-of the high density of the training data, and so could be explored in future challenges.
+of the high density of the training data, and so could be explored in future challenges\footnote{In real data or more realistic simulations discarding galaxies has a more complicated effects than simply decreasing sample size as it does here - removing objects with broad $p(z)$ can narrow the final $n(z)$, not just lower it.
+This was not a factor in this challenge because we use true redshifts to calculate our metrics.}.
 
@@
 This is not true of the overall winning method for large bin counts - as shown in the tables
-in \autoref{app:tables}, significantly more methods acheive a top-scoring spot in the Buzzard
+in \autoref{app:tables}, significantly more methods achieve a top-scoring spot in the Buzzard
 table \ref{tab:full_buzz} than in the CosmoDC2 table \ref{tab:full_dc2}. This reinforces the
@@
 \includegraphics[width=0.9\linewidth]{results/metric_comparisons.pdf}
-\caption{A comparison of the metrics used in our challenge. The x-axis is shared, and shows
+\caption{A comparison of the metrics used in our challenge for the different methods that were entered. The x-axis is shared, and shows
 the 3x2pt $w_0 - w_a$ (DETF) figure of merit on the CosmoDC2 data set for all methods, 
@@
 results for the CosmoDC2 scenarios, and the right-hand panel for the Buzzard scenarios.  Colours
-indicate the number of bins used in the analysis.}
+indicate the number of bins used in the analysis. Some methods perform worse when adding the $g$-band 
+and have ratios greater than unity.}
 \label{fig:loss}
@@
 Some of the methods in the challenge accepted target bin edges as an input to their algorithms,
-and then used classifiers to assign objects to those bins.  Other methods optimized the values
-of bin edges themselves, by optimizing the metric.  For the runs shown here, these methods mostly
+and then used classifiers to assign objects to those bins.
+Entrants could select these edges; many followed the example random forest entry and split into equal counts per bin.
+Other methods optimized the values
+of bin edges themselves, by maximizing a metric.  For the runs shown here, these methods mostly
 targeted one of the 3x2pt metrics.
@@
 for this idealised scenario; it should give an upper limit when using the same fixed bins.
-That other methods with the same nominal bin choice acheive scores close to it once again
+That other methods with the same nominal bin choice achieve scores close to it once again
 confirms that many entries are close to the best possible score for this approach.
@@
 Several of the methods with optimized target bin edges break this
-barrier and acheive very high scores.  This included ZotNet, ZotBin,  and JaxResNet.  The
+barrier and achieve very high scores.  This included ZotNet, ZotBin,  and JaxResNet.  The
 Stacked Generalization method similarly included manually tuned bin edges to maximize
@@
 well-specified science cases, and DESC has subsequently run other challenges along the same lines
-such is its N5K beyond-limber challege\footnote{\url{https://github.com/LSSTDESC/N5K}}.
+such as the N5K beyond-limber challenge\footnote{\url{https://github.com/LSSTDESC/N5K}}.
 We describe in Appendix \ref{sec:mistakes} some lessons we have learned from this challenge
@@
 is sufficient for specifying tomography even up to a large number of bins, if
-the training data is complete enough: degneracies between colour do not make this impossible.
+the training data is complete enough: degeneracies between colour do not make this impossible.
 Excluding the g-band from the data reduces the constraining power by a noticeable but
@@
 and of generating as many potential methods for the task as possible.  With
-both these acheived, future directions for simulating the problem are clear.
+both these achieved, future directions for simulating the problem are clear.
 The most obvious is limited spectroscopic completeness and size.  Realistic
@@
 affected by the latter problem.  Future challenges must explore these important limitations
-using simulated incompleteness.
+using simulated incompleteness, and use realistic photo-z methods to compute metrics.
 
@@
 
+Method implementations used in the challenge may be found in the \textsc{grand\_merge} branch
+of the challenge repository \url{http://github.com/LSSTDESC/tomo_challenge}.
+
 \subsection{ {\sc RandomForest} } \label{sec:randomforest}
@@
 
-We combine the Convolutional layers are with dimensionality reduction (pooling layers) and different
+We combine the convolutional layers with dimensionality reduction (pooling layers) and different
 activation functions. Here, we used Autokeras to optimize a general convolutional architecture,
@@
 ][]{Wright/etal:2020a}, where in this case we combine cells based on the
-similarity of the cells' $n(z) $moments (mean, median, and normalised median absolute
+similarity of the cells' $n(z)$ moments (mean, median, and normalised median absolute
 deviation from median) using complete-linkage clustering. The result of this
@@
 We chose the learning algorithm for both methods to minimize the reciprocal of
-the figure of merit $F$, where $F$ is defined in eqref{E:FOM}. We implement
+the figure of merit $F$, where $F$ is defined in \ref{eq:fom}. We implement
 these models using the JAX-based Flax neural network library \cite{jax}.
 
-\subsection{ {\sc NeuralNetwork 1 \& 2} }
+\subsection{ {\sc NeuralNetwork 1 \& 2} } \label{sec:nn}
 This method was submitted by the author FL.
@@
 	\item Using a simple neural network to parameterize a bin assignment function $f_\theta$ taking available photometry as the input and returning a bin number. 
-	\item Optimizing the parameters of $f_\theta$ as to maximize a target score, either the total SNR \autoref{eq:snr}, or the DETF FoM \autoref{eq:fom}.
+	\item Optimizing the parameters of $f_\theta$ as to maximize a target score, either the total SNR (\autoref{eq:snr}), or the DETF FoM (\autoref{eq:fom}).
 \end{itemize}
@@
 
-\subsection{ {\sc ZotBin and ZotNet} }
+\subsection{ {\sc ZotBin and ZotNet} } \label{sec:zot}
 
@@
 \subsection{ {\sc funbins} } \label{sec:funbins}
+This entry was submitted by author AB.
+
 This method uses the random forest algorithm described in section \ref{sec:randomforest} to assign galaxies to tomographic bins, but modifies the target bin edge selection method.
@@
 
-\subsection{ {\sc Stacked Generalization} }
+\subsection{ {\sc Stacked Generalization} } \label{sec:stackgen}
 This method was submitted by author JEC.
@@
 \begin{deluxetable*}{ |l | c c c | c c c | c c c | c c c | c c c | c c c |}
-\tablecaption{The 9-bin results on the CosmoDC2 data set for all the metrics and all methods on the CosmoDC2 sample.
+\tablecaption{The 9-bin results for all metrics calculated for all methods on the CosmoDC2 sample.
  \label{tab:full_dc2}}

### 2021-05-14 paper/paper.tex
more caveats

@@
 
+No one method dominates the metrics across the different scenarios considered.
 Before the challenge began we somewhat arbitrarily chose 
@@
 
-It is clear from this that, at least in our idealised scenario,  fears about restricted photometry 
-catastrophically limiting
-the information needed for tomography are unfounded.
+These results suggest that, at least in our idealised scenario of perfect training sets and given the limitations of our simulations, restricted photometry does does not necessarily catastrophically limit our ability to use tomography.
 In both the CosmoDC2 and Buzzard data sets,

### 2021-05-17 paper/paper.tex
add horiz lone

@@
 	\caption{Values of the 3x2pt DETF ($w_0,w_a$) figure-of-merit achieved by entrant methods on the 
-		CosmoDC2 part of the challenge.}
+		CosmoDC2 part of the challenge. The horizontal line separates two general approaches to the problem, as discussed in \autoref{sec:train-vs-fix}.}
 	\label{tab:cosmodc2}
@@
 it did not improve scores, but as noted in \autoref{sec:pqnld} this may be an artifact
-of the high density of the training data, and so could be explored in future challenges\footnote{In real data or more realistic simulations discarding galaxies has a more complicated effects than simply decreasing sample size as it does here - removing objects with broad $p(z)$ can narrow the final $n(z)$, not just lower it.
+of the high density of the training data, and so could be explored in future challenges\footnote{In real data or more realistic simulations discarding galaxies has more complicated effects than simply decreasing sample size as it does here - removing objects with broad $p(z)$ can narrow the final $n(z)$, not just lower it.
 This was not a factor in this challenge because we use true redshifts to calculate our metrics.}.

### 2021-05-17 paper/results/make_figures.py

@@ -60,7 +60,7 @@ methods_with_fixed_edges = [m for m in method_names if m not in methods_with_tra
 
 
 methods_reordered = methods_with_trained_edges + methods_with_fixed_edges
-
+method_names = methods_reordered
 metrics = ["SNR_ww", "SNR_gg", "SNR_3x2", "FOM_ww", "FOM_gg", "FOM_3x2", "FOM_DETF_ww", "FOM_DETF_gg", "FOM_DETF_3x2"]
 
 
@@ -99,6 +99,8 @@ def make_table(results, metric, filename):
 
         f.write(row_format.format(*row))
         f.write("\n")
+        if name == methods_with_trained_edges[-1]:
+            f.write("\\hline\n")
     f.close()
 
 

### 2021-05-17 paper/paper.tex
more in 4.6

@@
 should make it straightforward to calibrate multiple configurations and pass them to
-subsequent science anlysis.
+subsequent science anlysis; this will probably be true even when comparing models with
+different systematic parametrizations.  
+This feature is highlighted by the comparison between the two variants
+of the \textsc{NeuralNetwork} entry, which differed only in which metric they were targeting;
+optimizing one metric does not optimize the other.
 

### 2021-05-19 paper/paper.tex
two more fixes

@@
 should make it straightforward to calibrate multiple configurations and pass them to
-subsequent science anlysis; this will probably be true even when comparing models with
-different systematic parametrizations.  
+subsequent science anlysis; this will probably be true even when comparing models
+differing only in their systematic parametrizations.  
 This feature is highlighted by the comparison between the two variants
@@
 We chose the learning algorithm for both methods to minimize the reciprocal of
-the figure of merit $F$, where $F$ is defined in \ref{eq:fom}. We implement
+the figure of merit $F$, where $F$ is defined in \autoref{eq:fom}. We implement
 these models using the JAX-based Flax neural network library \cite{jax}.

### 2021-05-21 bin/batch_metrics.py
change path

@@ -62,7 +62,7 @@ def setup(queue):
 def task(name):
     print(name)
     sys.stdout.flush()
-    results_dir = "/global/cscratch1/sd/zuntz/tomo_challenge_secret_testing_results"
+    results_dir = "/global/cscratch1/sd/zuntz/tomo_challenge_dc2_results"
     result_file = f'{results_dir}/{name}.npy'
     img_file = f'{results_dir}/plots/{name}.png'
     metric_file = f'{results_dir}/metrics/{name}.yml'

### 2021-06-02 paper/results/make_figures.py
add grid plotting code

@@ -138,6 +138,63 @@ def color_plot(data, x, y, ax):
         ax.plot(x[s], y[s], '.', markersize=8, color=cmap.colors[i], label=f'$n_b={nbin}$')
 
 
+def plot_metric_grid(data, data_set, bands, filename):
+    fig, axes = plt.subplots(3, 3, figsize=(10,9), sharex=True, sharey=True)
+    min_sel = (data['bands'] == bands) & (data['method'] == 'Random')
+    ymin_data = data[min_sel]
+
+    # pick out equal
+    if data_set == "dc2":
+        ymax_data = Table.read(f"output_dc2_FIEGE_validation_riz.txt", format='ascii')
+    else:
+        ymax_data = Table.read(f"output_buzzard_FIEGE_validation_griz.txt", format='ascii')
+
+    cmap = plt.get_cmap("tab10")
+    x = np.array([3, 5, 7, 9])
+    for i, metric in enumerate(metrics):
+        ymin = ymin_data[metric]
+        ymax = ymax_data[metric][:4]
+        ymax_equalz = ymax_data[metric][4:]
+        ax = axes[i//3, i%3]
+        method_sets = [methods_with_fixed_edges, methods_with_trained_edges]
+        for j, method_set in enumerate(method_sets):
+            c = cmap(j)
+            for method in method_set:
+                if method == "Random":
+                    continue
+                sel = (data['bands'] == bands) & (data['method'] == method)
+                y = data[sel][metric]
+                b = data[sel]['bins']
+                s = y != 0
+                ax.plot(x[s], ((y - ymin) / (ymax - ymin))[s], color=c, lw=1)
+                ax.set_xticks([3, 5, 7, 9])
+                ax.tick_params(axis='x', which='minor', bottom=False)
+                ax.text(3.5, 1.7, metric, bbox={'facecolor': 'white'})
+                if i//3 == 0 :
+                    ax.set_yticks([0.0, 0.5, 1.0, 1.5])
+                if i == 3:
+                    ax.set_ylabel("Normalized metric")
+        ax.plot(x, ((ymax_equalz - ymin) / (ymax - ymin)), color=cmap(2), lw=3)
+        ax.plot(x, [0, 0, 0, 0], color='k', lw=3)
[…]

### 2021-06-03 paper/paper.tex

@@
 are performed using only bands in which the PSF is measured well enough for deconvolution to be
-performed.  For Rubin, this means using only the $r$, $i$, $z$, and perhaps $g$ bands.  In this work we
+performed\footnote{Alternative or extended methods to the metacalibration process described here, modelling the full likelihood space of deconvolved galaxy properties in a unified way might one day be feasible, but are likely to be computationally challenging.}.
+For Rubin, this means using only the $r$, $i$, $z$, and perhaps $g$ bands.
+In this work we
 therefore study how well we can perform tomography using only this limited photometric information.  In

### 2021-06-24 paper/paper.tex
add new plots

@@
 
+An illustration of the rank acheived by each method within each challenge configuration and for each metric
+is shown in \autoref{fig:rank_grid}.
+
+\begin{figure*}[htbp]
+	% \centering
+	\includegraphics[width=\linewidth]{results/rank_grid.pdf}
+	\caption{The rank of each method (1 = highest scoring, 25 = lowest scoring) within each configuration of the challenge, for each metric separately.
+A method that was good everywhere would appear as yellow strip across an entire horizontal range.}
+	\label{fig:rank_grid}
+\end{figure*}
+
 As described in \autoref{sec:funbins},  \textsc{funbins} was designed
@@
 A comparison of trained-edge and fixed-edge methods on the CosmoDC2 griz scenarios
-is shown in Figure \ref{fig:edges}.  One important point to note is that while trained-edge
+is shown in Figure \ref{fig:edges}, and the same data normalized against the score acheived
+by objects binned in true redshift (i.e. a perfect assigment method) are shown in Figure \ref{fig:metric_grid_dc2_griz}.
+
+One important point to note is that while trained-edge
 methods dominate the highest scores for the 3x2pt DETF metric, on which they trained, they
@@
 \end{figure}
+
+\begin{figure*}
+\includegraphics[width=1\linewidth]{results/metric_grid_dc2_griz.pdf}
+\caption{Normalized scores for each method, normalized between the score for random assignment (zero) and truth assignment to bins with equal number counts (one).
+It can be seen that both equal-number and equal-redshift spacings can be beaten for certain metrics.
+In each case, classification is done using griz bands on the CosmoDC2 data set; equivalent plots for Buzzard and riz are shown in \autoref{app:metric_grids}.}
+\label{fig:metric_grid_dc2_griz}
+\end{figure*}
 
@@
 Again, a requirement enforced by continuous integration could have checked this.
+
+\section{Normalized metric scores} \label{app:metric_grids}
+Figures \ref{fig:metric_grid_dc2_riz}, \ref{fig:metric_grid_buzzard_griz}, and \ref{fig:metric_grid_buzzard_riz} show metric scores normalized bewteen random assignment and perfect assignemtn to equal number density bins for CosmoDC2/riz, Buzzard/griz and Buzzard/riz respectively
+
+\begin{figure*}
+\includegraphics[width=1\linewidth]{results/metric_grid_dc2_riz.pdf}
+\caption{Normalized scores for each method, following the scheme of Figure \ref{fig:metric_grid_dc2_griz}
+but for CosmoDC2 with riz bands}
+\label{fig:metric_grid_dc2_riz}
+\end{figure*}
+
+\begin{figure*}
+\includegraphics[width=1\linewidth]{results/metric_grid_buzzard_griz.pdf}
+\caption{Normalized scores for each method, following the scheme of Figure \ref{fig:metric_grid_dc2_griz}
+but for Buzzard with griz bands}
+\label{fig:metric_grid_buzzard_griz}
+\end{figure*}
+
+\begin{figure*}
+\includegraphics[width=1\linewidth]{results/metric_grid_buzzard_riz.pdf}
+\caption{Normalized scores for each method, following the scheme of Figure \ref{fig:metric_grid_dc2_griz}
+but for Buzzard with riz bands}
+\label{fig:metric_grid_buzzard_riz}
+\end{figure*}
 

### 2021-06-24 paper/results/make_figures.py

@@ -8,6 +8,7 @@ import subprocess
 import h5py
 import collections
 import matplotlib.pyplot as plt
+import scipy.stats
 # set default styles
 
 
@@ -530,6 +531,81 @@ def make_9bin_table(results, filename):
     f.close()
 
 
+def make_metric_grid(data_set, bands):
+    
+    # cut doen to the correct bands
+    data_set = data_set[data_set['bands'] == bands]
+    metrics = ['SNR_ww','SNR_gg','SNR_3x2','FOM_ww','FOM_gg','FOM_3x2','FOM_DETF_ww','FOM_DETF_gg','FOM_DETF_3x2']
+    bins = [3, 5, 7, 9]
+    
+    nmetric = len(metrics)
+    nbin = len(bins)
+
+    nrow = nbin * nmetric
+    row_meanings = [(b, metric) for b in bins for metric in metrics]
+    row_labels = [f'n={b} {metric}' for b in bins for metric in metrics]
+    
+    methods = np.unique(data_set['method'])
+    ncol = len(methods)
+
+    grid = np.zeros((nrow, ncol), dtype=int)
+    for i, (row, row_label) in enumerate(zip(row_meanings, row_labels)):
+        b, metric = row
+        d = data_set[data_set['bins']==b]
+        # check that the ordering is the same.  should be
+        assert (d['method'] == method_names).all()
+        ranks = len(method_names) - scipy.stats.rankdata(d[metric]) + 1
+        grid[i, :] = ranks
+    
+    return row_labels, grid
+        
+def plot_rank_grid(dcz, buzz, filename):
+    fig, axes = plt.subplots(2, 2, figsize=(14.5, 13))
+    # make_metric_grid(dc2, 'riz')
+    for i, (data_set_name, data_set) in enumerate([("CosmoDC2", dc2), ("Buzzard", buzz)]):
+        for j, bands in enumerate(["riz", "griz"]):
+            row_labels, grid = make_metric_grid(data_set, bands)
+            im = axes[i, j].imshow(grid.T, cmap='summer_r')
+            ax = axes[i, j]
+            
+            if i == 0:
+                ax.title.set_text(f"{bands}")
+            
+            xlabels = row_labels if i == 1 else []
+            
+            ylabels = method_names if j == 0 else []
+            ax.set_yticks(np.arange(len(method_names)))
[…]

### 2021-06-24 paper/paper.tex
typo

@@
 \section{Normalized metric scores} \label{app:metric_grids}
-Figures \ref{fig:metric_grid_dc2_riz}, \ref{fig:metric_grid_buzzard_griz}, and \ref{fig:metric_grid_buzzard_riz} show metric scores normalized bewteen random assignment and perfect assignemtn to equal number density bins for CosmoDC2/riz, Buzzard/griz and Buzzard/riz respectively
+Figures \ref{fig:metric_grid_dc2_riz}, \ref{fig:metric_grid_buzzard_griz}, and \ref{fig:metric_grid_buzzard_riz} show metric scores normalized bewteen random assignment and perfect assigment to equal number density bins for CosmoDC2/riz, Buzzard/griz and Buzzard/riz respectively
 

### 2021-06-28 paper/paper.tex
author tweak and avoid repetition
Deed-forward --> Feed-forward
Typo
update table 1 description

@@
 
-The \emph{3x2pt} method has become a standard method for performing this combination.
+The \emph{3x2pt} method has become a standard tool for performing this combination.
 In this method, two-point correlations are computed among and between two samples, the shapes of 
@@
 	\textsc{MLPQNA} & Multi-layer perceptron & SC, MB \\
-	\textsc{Stacked Generalization} & Combination of gradient tree boosting and logistic regression & JEC\\
+	\textsc{Stacked Generalization} & Combination of  Gradient Boosted Trees classifiers with standard Random Forests (50/50) & JEC\\
 	\hline

### 2021-08-02 paper/paper.tex
typos

@@
 	\textsc{ZotBin/ZotNet} & Two related neural network methods with a preprocesing normalizing flow & DK \\
-	\textsc{FFNN} & Deed-forward neural network & EN \\
+	\textsc{FFNN} & Feed-forward neural network & EN \\
 	\textsc{FunBins} & Random forest with various nominal bin edge selection & AB \\
@@
 
-An illustration of the rank acheived by each method within each challenge configuration and for each metric
+An illustration of the rank achieved by each method within each challenge configuration and for each metric
 is shown in \autoref{fig:rank_grid}.
@@
 A comparison of trained-edge and fixed-edge methods on the CosmoDC2 griz scenarios
-is shown in Figure \ref{fig:edges}, and the same data normalized against the score acheived
+is shown in Figure \ref{fig:edges}, and the same data normalized against the score achieved
 by objects binned in true redshift (i.e. a perfect assigment method) are shown in Figure \ref{fig:metric_grid_dc2_griz}.
@@
 ][]{Wright/etal:2020a}, where in this case we combine cells based on the
-similarity of the cells' $n(z)$ moments (mean, median, and normalised median absolute
+similarity of the cells' $n(z)$ moments (mean, median, and normalized median absolute
 deviation from median) using complete-linkage clustering. The result of this
@@
 \section{Normalized metric scores} \label{app:metric_grids}
-Figures \ref{fig:metric_grid_dc2_riz}, \ref{fig:metric_grid_buzzard_griz}, and \ref{fig:metric_grid_buzzard_riz} show metric scores normalized bewteen random assignment and perfect assigment to equal number density bins for CosmoDC2/riz, Buzzard/griz and Buzzard/riz respectively
+Figures \ref{fig:metric_grid_dc2_riz}, \ref{fig:metric_grid_buzzard_griz}, and \ref{fig:metric_grid_buzzard_riz} show metric scores normalized between random assignment and perfect assigment to equal number density bins for CosmoDC2/riz, Buzzard/griz and Buzzard/riz respectively
 

### 2021-08-02 paper/results/make_figures.py
fix x axis in figs 9-12

@@ -182,7 +182,7 @@ def plot_metric_grid(data, data_set, bands, filename):
     ax.set_ylim(-0.01, 1.98)
     ax.set_xlim(3, 9)
     fig.tight_layout()
-    fig.subplots_adjust(hspace=0.025, wspace=0.0)
+    fig.subplots_adjust(hspace=0.025, wspace=0.06)
 
     legend_elements = [
         plt.Line2D([0], [0], color='k', lw=3),

### 2021-08-02 paper/paper.tex
one more sentence and some spelling

@@
 
-\subsubsection{CosmoDC2}
+\subsubsection{CosmoDC2} \label{sec:cosmodc2}
 
@@
 show tomographic bin indices, 
-and the color represents the fraction of objects (compared to the total number in the challenge) that are 
+and the colour represents the fraction of objects (compared to the total number in the challenge) that are 
 at redshifts where both tomographic bins have members, in bins of width $\delta z=10^{-3}$.  
@@
 of the \textsc{NeuralNetwork} entry, which differed only in which metric they were targeting;
-optimizing one metric does not optimize the other.
+optimizing one metric does not optimize the other (though the changes between results on the two simulations
+could also indicate that this is related to the high-redshift behaviour in the CosmoDC2 simulation
+described in \ref{sec:cosmodc2}).
 

### 2021-08-03 paper/paper.tex
author initial fix
ref fix

@@
 	\textsc{UTOPIA} & Nearest-neighbours, optimised in the limit of large representative training sets & AHW \\
-	\textsc{ComplexSOM} & Extension of {\sc SimpleSom} with an additional assignment optimization & DA, AHW, AN, EG, AB \\
+	\textsc{ComplexSOM} & Extension of {\sc SimpleSom} with an additional assignment optimization & DA, AHW, AN, EG, BA \\
 	\textsc{GPzBinning} & Gaussian Process redshift estimator and binner & PH \\
@@
 
-The first option, \textit{log}, calculates bin edges such that the number of galaxies in each bin grows logarithmically instead of being constant as in section 4.1.
+The first option, \textit{log}, calculates bin edges such that the number of galaxies in each bin grows logarithmically instead of being constant as in section \ref{sec:randomforest}.
 This option uses increasingly large bins for more distant galaxies and tests the relative importance of nearby galaxies.

### 2021-08-05 paper/paper.tex
add author statement and standard acks

@@
 \section{Acknowledgements}
+
+This paper has undergone internal review in the LSST Dark Energy Science Collaboration.
+The internal reviewers were Markus Michael Rau, Christopher Morrison, and Shahab Joudaki.
+
 ESC acknowledges support from the FAPESP (\#2019/19687-2) and CNPq  (\#308539/2018-4).
 AHW is supported by a European Research Council Consolidator Grant (No. 770935).
+
+The DESC acknowledges ongoing support from the Institut National de
+Physique Nucl\'eaire et de Physique des Particules in France; the
+Science \& Technology Facilities Council in the United Kingdom; and the
+Department of Energy, the National Science Foundation, and the LSST
+Corporation in the United States.  DESC uses resources of the IN2P3
+Computing Center (CC-IN2P3--Lyon/Villeurbanne - France) funded by the
+Centre National de la Recherche Scientifique; the National Energy
+Research Scientific Computing Center, a DOE Office of Science User
+Facility supported by the Office of Science of the U.S.\ Department of
+Energy under Contract No.\ DE-AC02-05CH11231; STFC DiRAC HPC Facilities,
+funded by UK BIS National E-infrastructure capital grants; and the UK
+particle physics grid, supported by the GridPP Collaboration.  This
+work was performed in part under DOE Contract DE-AC02-76SF00515
+
+\subsection{Author Contributions}
+Zuntz led the project, ran the challenge, wrote most paper text, and made plots.
+Lanusse submitted methods, made plots, and contributed to metrics, infrastructure, analysis, and conclusions.
+Malz and Wright submitted methods, made plots, and contributed to analysis and conclusions.
+Slosar submitted methods, wrote initial infrastructure, and assisted with challenge design.
+Authors Abolfathi to Teixeira submitted methods to the challenge.
+Subsequent authors generated the data used in the challenge and contributed to infrastructure used in the project.
+
+Authors AP, CRB, ESC, BMOF, EJG, and GT are collaborators from outside the
+Dark Energy Science Collaboration included here under a DESC publication policy exemption.
 

### 2021-08-06 paper/paper.tex
correction

@@
 
-Authors AP, CRB, ESC, BMOF, EJG, and GT are collaborators from outside the
+Authors CRB, ESC, BMOF, EJG, and GT are collaborators from outside the
 Dark Energy Science Collaboration included here under a DESC publication policy exemption.

### 2021-08-17 paper/paper.tex
update malz contrib
changes from Danielle

@@
 
-The \emph{3x2pt} method has become a standard tool for performing this combination.
+The \emph{3$\times$2pt} method has become a standard tool for performing this combination.
 In this method, two-point correlations are computed among and between two samples, the shapes of 
-background (source) galaxies and the locations of foreground galaxies, which trace foreground
+background (source) galaxies and the locations of foreground (lens) galaxies, which trace foreground
 dark matter haloes.  The three combinations (source-source, source-lens, and lens-lens) are
@@
 can potentially affect the ease with which we can calibrate
-the bins, but any choice can lead to correct results. The bins need not even be purely tomographic: we can happily correlate bins with number density peaks at different redshifts if that is useful.
+the bins, but any choice can lead to correct cosmology results.
+The bins need not even be assigned contiguous redshift ranges: we can happily correlate bins with multiple redshifts if that is useful, or by some other galaxy property than redshift entirely.
 We should choose, then, an assignment algorithm that maximises the constraining power for a science
@@
 
-This paper is part of preparations for the analysis of Legacy Survey of Space and Time (LSST) data at the Rubin Observatory, being run by the Dark Energy Science Collaboration (DESC).
+This paper is part of preparations for the analysis to be run by the Dark Energy Science Collaboration (DESC) of Legacy Survey of Space and Time (LSST) data from the Rubin Observatory.
 In it, we discuss and evaluate the related challenge
@@
 In the challenge, participants assigned simulated galaxies to tomographic bins,
-using only their (g)riz bands and associated errors, and the galay size.  Their goal was to maximize the cosmological
+using only their (g)riz bands and associated errors, and the galaxy size.  Their goal was to maximize the cosmological
 information in the final split data set; this is generally achieved by separating different bins by 
@@
 The challenge was open to any entrants, not just those already involved in DESC, but it was
-advertised through Rubin, and so is perhaps best described as semi-public.
+advertised through Rubin channels, and so is perhaps best described as semi-public.
 Participants developed and tested their methods locally but submitted code to be run as part of a central
@@
 the full data each and the validation 50\%. In practice many entrants trained on a subset of
-this, so to make the final evaluation phase tractable we trained on 1M objects for both data sets.
+this, so to make the final evaluation of methods tractable we used 1M objects for both data sets when
+computing the metrics presented in this paper.
 
@@
 power on overall cosmic structure amplitude and its evolution. In each case the complete parameter space consists
-of the eight $w_0w_aCDM$ cosmological parameters. This simplified measure does not include any nuisance
+of the seven $w_0w_aCDM$ cosmological parameters; we used the parameter combination used natively by the DESC core cosmology library \citep{ccl} ($\Omega_c, \Omega_b, \mathrm{H}_0, \sigma_8, n_s, w_0, w_a$).
+This simplified measure does not include any nuisance
 parameters modelling systematic errors in the analysis (most importantly the galaxy bias parameters in the clustering 
@@
 The DETF results for each method are shown in Table \ref{tab:cosmodc2}.  Complete results for all metrics
-for the nine-bin scenario only are shown in appendix tables \ref{tab:full_dc2} and \autoref{tab:full_buzz}.
+for the nine-bin scenario only are shown in appendix Tables \ref{tab:full_dc2} and \autoref{tab:full_buzz}.
 
@@
 		generating nine bins with the CosmoDC2 data.
-		The upper panel shows results for the method applied to $riz$ bands, and the middle on $griz$. The
-		increased spread and overlap of the bins in the middle panel decreases the overall constraining
-		power and signal-to-noise.  The lower panel is for comparison and is not part of the challenge,
-		and illutrates results using all the $ugrizy$ bands.}
+		The upper panel shows results for the method applied to $riz$ bands, and the middle on $griz$.
+		The lower panel is for comparison and is not part of the challenge,
+		and illustrates results using all the $ugrizy$ bands.
+		The increased spread and overlap of the bins as bands are removed decreases the overall constraining
+		power and signal-to-noise.
+	}
 	\label{fig:funbin_nz}
@@
 
-These results suggest that, at least in our idealised scenario of perfect training sets and given the limitations of our simulations, restricted photometry does does not necessarily catastrophically limit our ability to use tomography.
+These results suggest that, at least in our idealised scenario of perfect training sets and given the limitations of our simulations, restricted photometry does not necessarily catastrophically limit our ability to use tomography.
 In both the CosmoDC2 and Buzzard data sets,
@@
 combine the other successful methods). The typical 15--20\% improvement in FoM scores when optimizing
-nominal bin edges makes clear that this easy win should be compelling.
+nominal bin edges makes clear that re-selecting edges for each science case is worth the time and effort required.
 
@@
 Lanusse submitted methods, made plots, and contributed to metrics, infrastructure, analysis, and conclusions.
-Malz and Wright submitted methods, made plots, and contributed to analysis and conclusions.
+Malz and Wright made plots, and contributed to analysis and conclusions; Wright also submitted methods.
 Slosar submitted methods, wrote initial infrastructure, and assisted with challenge design.

### 2021-08-23 paper/authors.tex
more affils
add another orcid
First half of comments from AIM looked at
add new figs
more text on overlap
author tweak and avoid repetition
add author statement and standard acks
CWR changes

--- 2021-03-18 paper/authors.tex
@@
-\collaboration{1000}{The LSST Dark Energy Science Collaboration (LSST DESC)}
-\noaffiliation
-
 % paper leads
@@
 
-\author[0000-0001-7363-7932]{Angus Wright}
+\author[0000-0002-8676-1622]{Alex I. Malz}
+\affiliation{German Centre of Cosmological Lensing, Ruhr-Universit\"{a}t, Universit\"{a}tsstra{\ss}e 150, 44801 Bochum, Germany}
+
+\author[0000-0001-7363-7932]{Angus H. Wright}
 \affiliation{Ruhr University Bochum, Faculty of Physics and Astronomy, Astronomical Institute (AIRUB), German Centre for Cosmological Lensing, 44780 Bochum, Germany}
+
+\author[0000-0002-8713-3695]{An\v{z}e Slosar}
+\affiliation{Brookhaven National Laboratory, Upton NY 11973, USA}
 
@@
 
-\author{Adam Broussard}
-\noaffiliation
+\author[0000-0002-7767-5044]{Adam Broussard}
+\affiliation{Department of Physics and Astronomy, Rutgers, the State University, Piscataway, NJ 08854, USA}
 
@@
 
-\author{Eduardo S. Cypriano}
+\author[0000-0001-9196-7038]{Eduardo S. Cypriano}
 \affiliation{Universidade de S\~ao Paulo, IAG, Rua do Mat\~ao 1225, S\~ao Paulo, SP, Brazil}
 
-\author{Bernardo M. O. Fraga}
+\author[0000-0002-6395-3410 ]{Bernardo M. O. Fraga}
 \affiliation{Centro Brasileiro de Pesquisas F\'isicas, Rua Dr. Xavier Sigaud 150, 22290-180 Rio de Janeiro, RJ, Brazil}
@@
 \author[0000-0002-0226-9893]{Elizabeth J. Gonzalez}
+\affiliation{Centro Brasileiro de Pesquisas F\'isicas, Rua Dr. Xavier Sigaud 150, 22290-180 Rio de Janeiro, RJ, Brazil}
 \affiliation{Instituto de Astronom\'{\i}a Te\'orica y Experimental (IATE-CONICET), Laprida 854, X5000BGR, C\'ordoba, Argentina.}
-
 \affiliation{Observatorio Astron\'omico de C\'ordoba, Universidad Nacional de C\'ordoba, Laprida 854, X5000BGR, C\'ordoba, Argentina.}
@@
 
-\author{Alex Malz}
-\affiliation{Ruhr University Bochum, Faculty of Physics and Astronomy, Astronomical Institute (AIRUB), German Centre for Cosmological Lensing, 44780 Bochum, Germany}
-
 \author{Andrina Nicola}
-\noaffiliation
+\affiliation{Department of Astrophysical Sciences, Princeton University, Peyton Hall, Princeton, NJ 08544, USA}
 
@@
 
-\author{Andy Park}
+\author[0000-0002-8318-8226]{Andy Park}
 \affiliation{Department of Physics and Astronomy, University of California, Irvine, CA 92697, USA}
 
-\author{Anze Slosar}
-\affiliation{Brookhaven National Laboratory, Upton NY 11973, USA}
-
-\author{Gabriel Teixeira}
+\author[0000-0002-1594-208X]{Gabriel Teixeira}
 \affiliation{Centro Brasileiro de Pesquisas F\'isicas, Rua Dr. Xavier Sigaud 150, 22290-180 Rio de Janeiro, RJ, Brazil}
 
+\author[0000-0003-1468-8232]{Katrin Heitmann}
+\affiliation{Argonne National Laboratory, 9700 S Cass Ave, Lemont, IL 60439, USA}
+
+\author[0000-0002-2545-1989]{Eve Kovacs}
+\affiliation{Argonne National Laboratory, 9700 S Cass Ave, Lemont, IL 60439, USA}
+
 % DESC Builder authors
-\author{Yao-Yuan Mao}
+\author[0000-0002-1200-0820]{Yao-Yuan Mao}
+\affiliation{Department of Physics and Astronomy, Rutgers University, Piscataway, NJ 08854, USA}
+
+\collaboration{1000}{The LSST Dark Energy Science Collaboration (LSST DESC)}
 \noaffiliation

### 2021-08-23 paper/paper.tex
CWR changes

@@
 \usepackage{lineno}
-\linenumbers
+% \linenumbers
 %\modulolinenumbers[2]
@@
 The goal of the challenge was to collect and compare bin assignment strategies under various metrics of a standard 3x2pt cosmology analysis in a highly idealized setting to establish a baseline for realistically complex follow-up studies;
-in this preliminary study, we used two sets of cosmological simulations of galaxy redshifts and photometry under a simple noise model neglecting outliers and variation in observing conditions, and contributed algorithms were provided with a representative and complete training set.
-We review and evaluate the entries to the challenge, finding that even from this limited photometry information, multiple algorithms yield binning schemes with high figure-of-merit scores.
-We further find that the {\em g} band improves metric performance by $\sim15\%$ and confirm that the optimal bin assignment strategy depends on the specific science case.
+in this preliminary study, we used two sets of cosmological simulations of galaxy redshifts and photometry under a simple noise model neglecting photometric outliers and variation in observing conditions, and contributed algorithms were provided with a representative and complete training set.
+We review and evaluate the entries to the challenge, finding that even from this limited photometry information, multiple algorithms can separate tomographic bins reasonably well, reaching figures-of-merit scores close to the attainable maximum.
+We further find that adding the \emph{g} band to \emph{riz} photometry improves metric performance by $\sim15\%$ and that the optimal bin assignment strategy depends strongly on the science case: which figure-of-merit is to be optimized, and which observables (clustering, lensing, or both) are included.
 \end{abstract}
@@
 Tomographic 3x2pt measurements will be a key science goal in the upcoming \emph{Stage IV} surveys,
-including the Rubin Observatory \citep{rubin} and the Euclid and Roman space telescopes
+including the Vera C. Rubin Observatory \citep{rubin} and the Euclid and Roman space telescopes
 \citep{euclid,roman}.
@@
 challenge using simulated Rubin-like data with a range of methods submitted by different teams.
-We explore how many tomographic bins we can effectively generate using only a subset of bins, and compare methods submitted to the challenge.
+We explore how many tomographic bins we can effectively generate using only a subset of bands, and compare methods submitted to the challenge.
 
@@
 particular, this limitation prevents us from using the most obvious method for tomography,
-and computing a complete photometric probability density function (PDF) for each galaxy and assigning
+and computing a complete photometric redshift probability density function (PDF) for each galaxy and assigning
 a bin using the mean or peak of that PDF\footnote{This only prevents us using other photometry for \emph{selection}, not for \emph{characterization} after objects have been selected, such as computing the overall
@@
 This meant there were two ways to obtain higher scores: either to improve the choice of nominal bin
-edges, or to find better ways to assign objects to those edges.  This was a sub-optimal design decision:
+edges, or to find better ways to assign objects to those bins.  This was a sub-optimal design decision:
 separating the two issues would have enabled us to more easily delineate the advantages of different methods. 
@@
 	\includegraphics[width=0.9\linewidth]{results/colour_colour.pdf}
-	\caption{A colour-colour diagram for the two catalogues used in the challenge, showing the bands
+	\caption{Colour-colour diagrams for the two catalogues used in the challenge, showing the bands
 		available to participants.}
@@
 data sets were random sub-selections of the full data sets, with the training and testing sets 25\% of 
-the full data each and the validation 50\%. In practice many entrants trained on a subset of
-this, so to make the final evaluation of methods tractable we used 1M objects for both data sets when
-computing the metrics presented in this paper.
+the full data each and the validation 50\%.
 
@@
 similar characteristics at these redshifts, such as colour-colour relations.
-It was unknown whether this limitation would have any practical impact on the challenge,
-and was another reason for adpoting the additional Buzzard catalogue.
+It was unknown whether this limitation would have any practical impact on the challenge;
+this was a reason for adpoting the additional Buzzard catalogue.
 
@@
 $S/N > 10$, and a size cut $T / T_\mathrm{psf} > 0.5$, where
-$T$ is the trace $I_{xx} + I_{yy}$ of the moments matrix and measures the squared radius of the
-galaxy compared to that of a fixed Rubin PSF of size $0.75$ arc-seconds.
+$T$ is the trace $I_{xx} + I_{yy}$ of the moments matrix and measures the squared (deconvolved) 
+radius of the
+galaxy, and $T_\mathrm{psf}$ is a fixed Rubin PSF size of $0.75$ arc-seconds.
 
@@
 simulations and around 36M objects from the CosmoDC2 simulations.
+The nominal 25\% training data sets were therefore 5M and 9M objects respectively, but many entrants trained their methods on a reduced subset of the data.
+To make training tractable but consistent for the full suite we therefore trained all methods with 1M objects for each data set.
 
@@
 and because clustering-like statistics of lensing samples are important for studies of intrinsic
-alignments. We use three different sets of metrics, one based on the overall signal to noise of the spectra 
+alignments. We use three different sets of metrics, one based on the overall signal to noise of the angular power spectra 
 derived from the samples, and two based on a figure-of-merit for the constraints that would be obtained 
-using them. The relationships between these metrics in our results is discussed in \autoref{sec:metric-results}.
+using them. The relationships between these metrics in our results are discussed in \autoref{sec:metric-results}.
 We compute each metric on three different sets of summary statistics: lensing-alone, clustering-alone,
@@
     F &= \left( \frac{\partial \mu}{\partial \theta} \right)^T C^{-1} \left( \frac{\partial \mu}{\partial \theta} \right), \\
-    S &= \frac{1}{2 \pi \sqrt{\det{([F^{-1}]_{p_1, p_2})}}}
+    \mathrm{FOM} &= \frac{1}{2 \pi \sqrt{\det{([F^{-1}]_{p_1, p_2})}}}
 \label{eq:fom}
@@
 evolution, and a version in which $p_1 = \Omega_c$ and $p_2 = \sigma_8$, representing constraining
-power on overall cosmic structure amplitude and its evolution. In each case the complete parameter space consists
+power on overall cosmic structure amplitude ($\sigma_8$), and its evolution (a function of $\Omega_c$). In each case the complete parameter space consists
 of the seven $w_0w_aCDM$ cosmological parameters; we used the parameter combination used natively by the DESC core cosmology library \citep{ccl} ($\Omega_c, \Omega_b, \mathrm{H}_0, \sigma_8, n_s, w_0, w_a$).
@@
 	\textsc{UTOPIA} & Nearest-neighbours, optimised in the limit of large representative training sets & AHW \\
-	\textsc{ComplexSOM} & Extension of {\sc SimpleSom} with an additional assignment optimization & DA, AHW, AN, EG, BA \\
+	\textsc{ComplexSOM} & Extension of {\sc SimpleSom} with an additional assignment optimization & DA, AHW, AN, EG, BA, ABr \\
 	\textsc{GPzBinning} & Gaussian Process redshift estimator and binner & PH \\
@@
 	\textsc{FFNN} & Feed-forward neural network & EN \\
-	\textsc{FunBins} & Random forest with various nominal bin edge selection & AB \\
+	\textsc{FunBins} & Random forest with various nominal bin edge selection & ABa \\
 	\textsc{MLPQNA} & Multi-layer perceptron & SC, MB \\
@@
 	\caption{Values of the 3x2pt DETF ($w_0,w_a$) figure-of-merit achieved by entrant methods on the 
-		CosmoDC2 part of the challenge. The horizontal line separates two general approaches to the problem, as discussed in \autoref{sec:train-vs-fix}.}
+		CosmoDC2 part of the challenge. The horizontal line separates two general approaches to the problem, as discussed in \autoref{sec:train-vs-fix}: those below the line used fixed
+		target edges for tomographic bins and optimized the assignment of objects to those bins,
+		whereas those above also optimized the target edges themselves.}
 	\label{tab:cosmodc2}
@@
 		The upper panel shows results for the method applied to $riz$ bands, and the middle on $griz$.
-		The lower panel is for comparison and is not part of the challenge,
-		and illustrates results using all the $ugrizy$ bands.
+		The lower panel is for comparison and is not part of the challenge;
+		it illustrates {\sc funbins} results using all the $ugrizy$ bands.
 		The increased spread and overlap of the bins as bands are removed decreases the overall constraining
@@
 By this criterion the best method was, by a small margin, \textsc{funbins},
-which used a random forest to assign bins, but chose nominal bin edges by splitting the range spanned
+which used a random forest to assign galaxies to after choosing nominal bin edges that split 
+the range spanned
 by the sample into equal co-moving distance bins. 
+As described in \autoref{sec:funbins},  this splitting was designed
+to minimze the shot noise for the angular power in each bin, and this has proven successful. 
+Figure \ref{fig:funbin_nz} shows the number density obtained by this method, for both the riz and griz 
+scenarios, and an additional post-challenge run using the full Rubin $ugrizy$ bands.
+The metrics \textsc{funbins}
+obtained on $ugrizy$ are barely higher than those with $griz$ -- the largest
+increase is 8\% for the $\Omega_c-\sigma_8$ metric, with the remainder all 5\% or less.
+
 Other methods performed better in other branches of the challenge; in particular we note \textsc{ZotNet}, \textsc{Stacked Generalization}, and \textsc{NeuralNetwork} (see sections \ref{sec:zot}, \ref{sec:stackgen}, and \ref{sec:nn} respectively), which each won more than three 9-bin scenarios.
-
 An illustration of the rank achieved by each method within each challenge configuration and for each metric
@@
 \end{figure*}
-
-As described in \autoref{sec:funbins},  \textsc{funbins} was designed
-to minimze the shot noise for the angular power in each bin, and this has proven successful. 
-Figure \ref{fig:funbin_nz} shows the number density obtained by this method, for both the riz and griz 
-scenarios, and an additional post-challenge run using the full Rubin $ugrizy$ bands.  The metrics
-obtained with \textsc{funbins} on $ugrizy$ are barely higher than those with $griz$ -- the largest
-increase is 8\% for the $\Omega_c-\sigma_8$ metric, with the remainder all 5\% or less.
 
@@
 
-Several of the 9-bin neural network methods perform well up to seven tomographic bins but fail when run with 
+Several of the neural network methods perform well up to seven tomographic bins but fail when run with 
 nine.  This is presumably not an intrinsic feature of these methods, but depends on implementation details.
-Further development solving these challenges should be valuable, since they are the highest scoring seven-bin
+Further development solving these challenges should be valuable, since they are among the highest scoring seven-bin
 methods.
@@
 table \ref{tab:full_buzz} than in the CosmoDC2 table \ref{tab:full_dc2}. This reinforces the
-conclusions made below in \autoref{sec:train-vs-fix}, though of course the relatively noisy
+conclusion discusssed below in \autoref{sec:train-vs-fix} that target bin edges should be 
+re-optimized for each new scenario, though the relatively noisy
 results make this difficult to interpret.
@@
 
-The results of the challenge have shown first that three or four-band photometry
+The results of the challenge have shown that three or four-band photometry
 is sufficient for specifying tomography even up to a large number of bins, if
@@
 Excluding the g-band from the data reduces the constraining power by a noticeable but
-not catastrophic amount.
+not catastrophic amount ($\sim 15$\%).
 
@@
 Despite this, evenly spaced bins in fiducial co-moving distance have proven
-a good general choice in the winning method in the challenge, {\sc funbins}.
+a good general choice, as shown in the winning method in the challenge, {\sc funbins}.
 
@@
 training data sets will be far smaller than the millions of objects we used here,
-and the difficulty of measuring spectra for faint galaxies will make the them
+and the difficulty of measuring spectra for faint galaxies will make them
 incomplete and highly unrepresentative. Machine learning methods will be particularly
@@
 ESC acknowledges support from the FAPESP (\#2019/19687-2) and CNPq  (\#308539/2018-4).
-AHW is supported by a European Research Council Consolidator Grant (No. 770935).
+AHW is supported by a European Research Council Consolidator Grant (No.
+770935).
+The work of EG and ABr was supported by the U.S.
+Department of Energy, Office of Science, Office of High Energy Physics Cosmic Frontier Research program under Award Number DE-SC0010008.
 
@@
 \begin{align}
-    q^{\mathrm{cluster}}_i(\chi) &= n_i(\chi)\\
-    q^{\mathrm{lensing}}_i(\chi) &= \frac{3}{2}\Omega_m \left(\frac{\mathrm{H}_0}{c}\right)^2 \frac{\chi}{a(\chi)} \int_\chi^{\infty} \frac{\chi' - \chi}{\chi'} n_i(\chi')\,\,\mathrm{d}\chi'
+    q^{\mathrm{clust}}_i(\chi) &= n_i(\chi)\\
+    q^{\mathrm{lens}}_i(\chi) &= \frac{3}{2}\Omega_m \left(\frac{\mathrm{H}_0}{c}\right)^2 \frac{\chi}{a(\chi)} \int_\chi^{\infty} \frac{\chi' - \chi}{\chi'} n_i(\chi')\,\,\mathrm{d}\chi'
 \end{align}
@@
 \begin{equation}
-    \mathrm{Cov}(C^{ij}_\ell, C^{mn}_\ell) = \frac{1}{(2 \ell + 1)\Delta\ell f_\mathrm{sky}}(D^{im}_\ell + D^{jn}_\ell)(D^{in}_\ell + D^{jm}_\ell)
+    \mathrm{Cov}(C^{ij}_\ell, C^{mn}_\ell) = \frac{1}{(2 \ell + 1)\Delta\ell f_\mathrm{sky}}(D^{im}_\ell  D^{jn}_\ell + D^{in}_\ell D^{jm}_\ell)
 \end{equation}
-where $D^{ij}_\ell = C^{ij}_\ell + N^{ij}_\ell$ and we assume an $f_\mathrm{sky}=0.25$.  The noise spectra are:
+where $D^{ij}_\ell = C^{ij}_\ell + N^{ij}_\ell$, $\Delta_\ell$ is the size in $\ell$ of the binned spectra, and we assume an $f_\mathrm{sky}=0.25$.  The noise spectra are:
 \begin{align}
-N^{\mathrm{cluster},ij}_\ell = \delta_{ij} / n_i \\
-N^{\mathrm{lensing},ij}_\ell = \delta_{ij} \sigma_e^2 / n_i
+N^{\mathrm{clust},ij}_\ell = \delta_{ij} / n_i \\
+N^{\mathrm{lens},ij}_\ell = \delta_{ij} \sigma_e^2 / n_i
 \end{align}
-where $n_i$ is the number density of galaxies in bin $i$ (we asssume equally weighted galaxies, and a 
-total number density over all bins of 20 galaxies per square arcminute) and $\sigma_e=0.26$.
+where $n_i$ is the angular number density of galaxies in bin $i$ (we asssume equally weighted galaxies, and a 
+total number density over all bins of 20 galaxies per square arcminute) and $\sigma_e=0.26$, roughly matching that found in previous surveys \citep{des,kid,hsc}.
 
@@
 template set of \cite{Capak:2004} in combination with 
-the Bayesian redshift prior from \cite{Raichoor/etal:2014}.
+the Bayesian redshift prior from \cite{Raichoor/etal:2014}, using only the (g)riz bands supplied in the challenge.
 
@@
 \subsection{ {\sc ComplexSOM} } \label{sec:csom}
-This method was submitted by authors DA, AHW, AN, EG, and AB.
+This method was submitted by authors DA, AHW, AN, EG, BA, and ABr.
 
@@
 \subsection{ {\sc funbins} } \label{sec:funbins}
-This entry was submitted by author AB.
+This entry was submitted by author ABa.
 
@@
 \section{Normalized metric scores} \label{app:metric_grids}
-Figures \ref{fig:metric_grid_dc2_riz}, \ref{fig:metric_grid_buzzard_griz}, and \ref{fig:metric_grid_buzzard_riz} show metric scores normalized between random assignment and perfect assigment to equal number density bins for CosmoDC2/riz, Buzzard/griz and Buzzard/riz respectively
+Figures \ref{fig:metric_grid_dc2_riz}, \ref{fig:metric_grid_buzzard_griz}, and \ref{fig:metric_grid_buzzard_riz} show metric scores normalized between random assignment and perfect assigment to equal number density bins for CosmoDC2/riz, Buzzard/griz and Buzzard/riz respectively:
+\begin{equation}
+S_\mathrm{norm} = \frac{S - S_\mathrm{random}} {S_\mathrm{perfect,eq} - S_\mathrm{random}}
+\end{equation}
 

### 2021-08-27 paper/paper.tex
cite and acks
acks

@@
 
+AHW is supported by a European Research Council Consolidator Grant (No. 770935).
+The work of EG and ABr was supported by the U.S. Department of Energy, Office of Science, Office of High Energy Physics Cosmic Frontier Research program under Award Number DE-SC0010008.
+CRB, ESC, BMOF, EJG, and GT made use of multi-GPU Sci-Mind Brunelleschi machines developed and tested for Artificial Intelligence and would like to thank CBPF Multi GPU development team of LITCOMP/COTEC, for supporting this infrastructure.
 ESC acknowledges support from the FAPESP (\#2019/19687-2) and CNPq  (\#308539/2018-4).
-AHW is supported by a European Research Council Consolidator Grant (No.
-770935).
-The work of EG and ABr was supported by the U.S.
-Department of Energy, Office of Science, Office of High Energy Physics Cosmic Frontier Research program under Award Number DE-SC0010008.
+MB acknowledges financial contributions from the agreement ASI/INAF 2018-23-HH.0, Euclid ESA mission - Phase D.
+PH acknowledges generous support from the Hintze Family Charitable Foundation through the Oxford Hintze Centre for Astrophysical Surveys.
 
@@
 training and reference samples; a problem which does not exist (by construction) in this challenge.
+
+For a further discussion of the merits of band triplets in machine learning photo-z see \citet{broussard_gawiser}
 

### 2021-08-27 paper/paper.tex
fix cite and change short title

@@
 % header settings
-\shorttitle{Tomographic binning optimization}
+\shorttitle{DESC Tomography Challenge}
 \shortauthors{Zuntz et al.\ (LSST~DESC)}
@@
 where $n_i$ is the angular number density of galaxies in bin $i$ (we asssume equally weighted galaxies, and a 
-total number density over all bins of 20 galaxies per square arcminute) and $\sigma_e=0.26$, roughly matching that found in previous surveys \citep{des,kid,hsc}.
+total number density over all bins of 20 galaxies per square arcminute) and $\sigma_e=0.26$, roughly matching that found in previous surveys \citep{des,kids,hsc}.
 

### 2021-09-01 paper/paper.tex
remove old text
citation updates

@@
 \keywords{methods: statistical -- dark energy  -- large-scale structure of the universe}
-
-%\accepted{}
-%\submitjournal{the Astrophysical Journal Supplement}
 
@@
 where $n_i$ is the angular number density of galaxies in bin $i$ (we asssume equally weighted galaxies, and a 
-total number density over all bins of 20 galaxies per square arcminute) and $\sigma_e=0.26$, roughly matching that found in previous surveys \citep{des,kids,hsc}.
+total number density over all bins of 20 galaxies per square arcminute) and $\sigma_e=0.26$, roughly matching that found in previous surveys \citep{des,desy3a,desy3b,kids,hsc}.
 
@@
 
-%During the training stage of the algorithm GPz is inferring the locations and
-%spreads that describe the basis functions; during the validation stage it is
-%inferring the appropriate complexity of the model, essentially how many basis
-%functions to use.
-
 GPzBin is a simple extension to GPz for the problem of tomographic binning. For
@@
 
-%GPz requires very little fine-tuning. The most important parameter is $m$, the number of basis functions. A higher $m$ corresponds to higher model complexity and longer training times.
-
 \subsection{{\sc JaxCNN \& JaxResNet} }
@@
 \end{itemize}
-
-%\subsection{ {\sc NeuralNetwork 2} }
 

### 2021-09-24 paper/paper.tex
address referee comments up to end of section 3

@@
 Weak gravitational lensing (WL) has emerged over the last decade as a powerful
-cosmological probe \citep{cfhtlens,rcslens,des,kids,hsc}.  WL
+cosmological probe \citep{cfhtlens,rcslens,desy3a,desy3b,kids,hsc}.  WL
 uses measurements of coherent shear distortion to the observed shapes of galaxies
@@
 perturbation theory and simulation results.  The method has been used in the Dark Energy Survey, DES, 
-\citep{des-3x2pt}, and to combine the Kilo-Degree 
+\citep{des-3x2pt, desy3-3x2pt}, and to combine the Kilo-Degree 
 Survey, KiDS with spectroscopic surveys \citep{kids-3x2pt, kids_gama, kids_2df}.
@@
 regimes \citep{taylor18b,bnt,xcut}.
+Most recently, \citet{euclid-nz} studied this question with respect to the Euclid mission, using the Flagship
+simulation \citep{euclid-flagship} and applying realistic photometric redshift estimation methods to the
+sample (we defer this question in this work). They highlight the power of using large numbers of bins,
+differences in ideal binnings when
+changing the choice of data (the inclusion galaxy galaxy-lensing), and
+note the value in discarding galaxies with high redshift uncertainty. We will echo many of these issues below.
 
@@
 information in the final split data set; this is generally achieved by separating different bins by 
-redshift as much as possible.  Particpants were free to optimize or simply select nominal (target) 
+redshift as much as possible.  Participants were free to optimize or simply select nominal (target) 
 redshift bin edges from their training sample, but in either case had to assign galaxies into the 
@@
 cases would probably be impossible.
+These simplifications will increase scores overall, since one source of uncertainty in the calculation
+is removed. We expect that they will disproportionately increase scores for large numbers of
+tomographic bins, since the relative widening from individual galaxy uncertainties will be larger. Faint
+galaxy samples, including those at higher redshift, will also be more affected.
 
@@
 In each case we simulate first year Rubin observations
-using the DESC {\sc TXPipe} framework\footnote{\url{https://github.com/LSSTDESC/TXPipe}}, following the methodology of \citet{ivezic_jones_lupton}
-and assuming the numerical values for telescope and sky characteristics therein.
+using the DESC {\sc TXPipe} framework\footnote{\url{https://github.com/LSSTDESC/TXPipe}}.
+This follows the methodology of \citet{ivezic_jones_lupton}
+and assuming the numerical values for telescope and sky properties therein:
+for each galaxy it generates a poisson sample of the number of photons per band based on
+its true magnitude, the sky brightness, and instrumental characteristics.
+No noise is added to the redshifts of the galaxy; we defer discussion of photometric
+redshift uncertainty to future work.
 
@@
 
+Entrants to the challenge were permitted to use the metrics here on the training data, and several did
+so to optimize their target bin edges (see \autoref{sec:train-vs-fix}).
+
 \subsubsection{Metric 1: SNR}
@@
 where $\mu$ is a stack of the theoretical predictions for the $C_\ell$ spectra for each tomographic 
-bin, in turn, and $C$ a estimate of the covariance between them, making the approximation that the underlying density field has a Gaussian distribution as described in
+bin pair, in turn (including auto- and cross-correlations for both clustering and lensing), and $C$ a estimate of the covariance between them, making the approximation that the underlying density field has a Gaussian distribution as described in
 \citet{takada_jain} and summarized in \autoref{app:theory}.   We compute this metric for lensing alone, clustering alone, and the full 3x2pt combination.
@@
 This simplified measure does not include any nuisance
-parameters modelling systematic errors in the analysis (most importantly the galaxy bias parameters in the clustering 
-spectra), and like all Fisher analyses approximates posterior constraints as Gaussian.
+parameters modelling systematic errors in the analysis, most importantly the galaxy bias parameters in the clustering 
+spectra; this artifically increases the constraining power of the galaxy clustering-only metric, as we will see below, but the metric still acts as a useful proxy for science cases dependent on finding well separated bins.
+Like all Fisher analyses, this metric approximates posterior constraints as Gaussian.
+
+The JAX library provides automatic differentiation of the calculation of the $C_\ell$ values, and so avoids the painful numerical sensitivity which usually affects Fisher matrix calculations.
+We verified this by comparing the code code to tuned finite difference estimates on CCL spectrum calculations.
 
@@
 
-The \textsc{IBandOnly} method used only the $i$-band to select tomographic bins,
-splitting galaxies in equal-sized bins in this magnitude. We expect only a small
+The \textsc{IBandOnly} method used only the $i$-band to select bin assignments.
+Edge values of bins in $i$ were chosen such that each bin had the same number of training galaxies in,
+and then test sample galaxies assigned to a bin based only on their $i$ band value. The band correlates
+only weakly with redshift, so the scores of the method are poor.
+We expect only a small
 increase in metric score with the number of bins, and also that the addition of the $g$-band
-should make no difference to scores.  Again, we find this to be true.
+should make no difference to scores, because that column of data is not used in the method.  Again, we find this to be true.
 
@@
 This was not a factor in this challenge because we use true redshifts to calculate our metrics.}.
+See \citet{euclid-flagship} for a discussion of this topic in the Euclid context.
 

### 2021-09-27 paper/paper.tex
changes for referee report

@@
 regimes \citep{taylor18b,bnt,xcut}.
-Most recently, \citet{euclid-nz} studied this question with respect to the Euclid mission, using the Flagship
-simulation \citep{euclid-flagship} and applying realistic photometric redshift estimation methods to the
+Most recently, \citet{euclid-nz} studied this question with respect to the Euclid mission, using the \textit{Flagship}
+simulation and applying realistic photometric redshift estimation methods to the
 sample (we defer this question in this work). They highlight the power of using large numbers of bins,
@@
 	\textsc{LSTM} & Long Short-Term Memory neural network & CRB, BF, GT, ESC, EJG \\
-	\textsc{AutokerasLSTM}\tablenotemark{a} & Automaticaly configured LSTM variant. & CRB, BF, GT, ESC, EJG \\
+	\textsc{AutokerasLSTM} & Automaticaly configured LSTM variant. & CRB, BF, GT, ESC, EJG \\
 	\textsc{LGBM} & Light Gradient Boosted Machine & CRB, BF, GT, ESC, EJG \\
@@
 	\label{tab:entrants}
-	\tablenotetext{a}{A technical problem, involving conflicting versions of libraries and incompatible hardware, meant that AutoKerasLSTM could not be run in the final challenge.}
 \end{table*}
@@
 		target edges for tomographic bins and optimized the assignment of objects to those bins,
-		whereas those above also optimized the target edges themselves.}
+		whereas those above also optimized the target edges themselves.
+        Asterisks indicate runs which failed during classification, and dashes indicate runs with
+        number counts too small for spectra to be generated.
+        }
 	\label{tab:cosmodc2}
@@
 	\includegraphics[width=1\linewidth]{results/nzbinned_dc2_funbins.pdf}
-	\caption{The tomographic $n(z)$ values generated by the winning method in the challenge, {\sc funbins},
+	\caption{The tomographic $n(z)$ values generated by the winning method in the challenge, {\sc FunBins},
 		generating nine bins with the CosmoDC2 data.
@@
 		The lower panel is for comparison and is not part of the challenge;
-		it illustrates {\sc funbins} results using all the $ugrizy$ bands.
+		it illustrates {\sc FunBins} results using all the $ugrizy$ bands.
 		The increased spread and overlap of the bins as bands are removed decreases the overall constraining
@@
 an overall winning method would be selected.
-By this criterion the best method was, by a small margin, \textsc{funbins},
+By this criterion the best method was, by a small margin, \textsc{FunBins},
 which used a random forest to assign galaxies to after choosing nominal bin edges that split 
@@
 As described in \autoref{sec:funbins},  this splitting was designed
-to minimze the shot noise for the angular power in each bin, and this has proven successful. 
+to minimize the shot noise for the angular power in each bin, and this has proven successful. 
 Figure \ref{fig:funbin_nz} shows the number density obtained by this method, for both the riz and griz 
 scenarios, and an additional post-challenge run using the full Rubin $ugrizy$ bands.
-The metrics \textsc{funbins}
+The metrics \textsc{FunBins}
 obtained on $ugrizy$ are barely higher than those with $griz$ -- the largest
@@
 \caption{The pairwise fractions $f_\mathrm{mis}$ of total objects in redshift regions where multiple tomographic bins contain galaxies, for 
-the \textsc{funbins} tomography method and three choices of bands.  
+the \textsc{FunBins} tomography method and three choices of bands.  
 The $x$- and $y$-axes 
@@
 
-\textsc{funbins} seems to be a good and stable general method, though several other methods perform as well
+\textsc{FunBins} seems to be a good and stable general method, though several other methods perform as well
 or slightly better in other scenarios.  Notably, many methods reach the same plateau in lensing-only metrics,
@@
 Several methods
-took advantage of this.  The ComplexSom method (see \autoref{sec:csom}) found
+took advantage of this.  The \textsc{ComplexSOM} method (see \autoref{sec:csom}) found
 it did not improve scores, but as noted in \autoref{sec:pqnld} this may be an artifact
@@
 This was not a factor in this challenge because we use true redshifts to calculate our metrics.}.
-See \citet{euclid-flagship} for a discussion of this topic in the Euclid context.
+See \citet{euclid-nz} for a discussion of this topic in the Euclid context.
 
@@
 distribution, and hence on the survey properties.  The largest outlier comes from the 
-{\sc Neural Network} entry.
+{\sc NeuralNetwork} entry.
 
@@
 
-The value of the extra FoM that could be gained should be weighed against 
+The value of the extra FoM that could be gained by adding the band should be weighed against 
 the challenge of high-accuracy
@@
 but until the end of the survey this is unlikely to be the limiting factor in overall precision.
+
+Some methods perform worse when adding the $g$-band, which we ascribe to reduced performance
+in their machine learning algorithms when increasing the dimensionality of their inputs; further tuning
+of their hyperparameters could perhaps alleviate the issue.
 
@@
 and then used classifiers to assign objects to those bins.
-Entrants could select these edges; many followed the example random forest entry and split into equal counts per bin.
-Other methods optimized the values
+Entrants could select these edges; many followed the example random forest entry and split into equal counts per bin; some like \textsc{FunBins} used alternative
+criteria. Other methods optimized the values
 of bin edges themselves, by maximizing a metric.  For the runs shown here, these methods mostly
@@
 should make it straightforward to calibrate multiple configurations and pass them to
-subsequent science anlysis; this will probably be true even when comparing models
+subsequent science analysis; this will probably be true even when comparing models
 differing only in their systematic parametrizations.  
@@
 It can be seen that both equal-number and equal-redshift spacings can be beaten for certain metrics.
-In each case, classification is done using griz bands on the CosmoDC2 data set; equivalent plots for Buzzard and riz are shown in \autoref{app:metric_grids}.}
+In each case, classification is done using griz bands on the CosmoDC2 data set; equivalent plots for Buzzard and riz are shown in \autoref{app:metric_grids}.
+\emph{ww} indicates lensing-only metrics, \emph{gg} indicates clustering only, and \emph{3x2pt} their combination.
+The three metrics are defined in \autoref{sec:metrics}.
+}
 \label{fig:metric_grid_dc2_griz}
@@
 Several of the methods with optimized target bin edges break this
-barrier and achieve very high scores.  This included ZotNet, ZotBin,  and JaxResNet.  The
-Stacked Generalization method similarly included manually tuned bin edges to maximize
+barrier and achieve very high scores.  This included \textsc{ZotNet}, \textsc{ZotBin}, and \textsc{JaxResNet}.  The
+\textsc{Stacked Generalization} method similarly included manually tuned bin edges to maximize
 score for this scenario (and indeed its ensemble combination method could be applied to
@@
 
-Notably, though, the winning CosmoDC2 \textsc{funbins} method used the simpler approach desribed above. The 
+Notably, though, the winning CosmoDC2 \textsc{FunBins} method used the simpler approach desribed above. The 
 success of this simple and fast solution suggests this as an easy heuristic for well-optimized nominal edges, 
@@
 Despite this, evenly spaced bins in fiducial co-moving distance have proven
-a good general choice, as shown in the winning method in the challenge, {\sc funbins}.
+a good general choice, as shown in the winning method in the challenge, {\sc FunBins}.
 
@@
 \end{equation}
-where $D^{ij}_\ell = C^{ij}_\ell + N^{ij}_\ell$, $\Delta_\ell$ is the size in $\ell$ of the binned spectra, and we assume an $f_\mathrm{sky}=0.25$.  The noise spectra are:
+where $D^{ij}_\ell = C^{ij}_\ell + N^{ij}_\ell$, $\Delta\ell$ is the size in $\ell$ of the binned spectra, and we assume an $f_\mathrm{sky}=0.25$.  The noise spectra are:
 \begin{align}
 N^{\mathrm{clust},ij}_\ell = \delta_{ij} / n_i \\
-N^{\mathrm{lens},ij}_\ell = \delta_{ij} \sigma_e^2 / n_i
+N^{\mathrm{lens},ij}_\ell = \delta_{ij} \sigma_{e,\mathrm{tot}}^2 / n_i
 \end{align}
 where $n_i$ is the angular number density of galaxies in bin $i$ (we asssume equally weighted galaxies, and a 
-total number density over all bins of 20 galaxies per square arcminute) and $\sigma_e=0.26$, roughly matching that found in previous surveys \citep{des,desy3a,desy3b,kids,hsc}.
+total number density over all bins of 20 galaxies per square arcminute) and $\sigma_{ei}=0.26$ per shear component, giving $\sigma_{e,\mathrm{tot}} = 0.37$, roughly matching that found in previous surveys \citep{des,desy3a,desy3b,kids,hsc}.
+
+This model does not include a number of theoretical effects that will be important in real data, including
+intrinsic alignments, non-Gaussian covariance, and a range of effects on spectra at small scales, like non-linear 
+bias and beyond-Limber effects. Collectively these effects mean that our tests are exaggerating the constraining
+power of small-scale signals, and increasing all our metrics (especially as our $\ell$ range is fixed with $z$).  
+Although this unrealistic aspect mostly affects algorithms equally, the relation between redshift, angle, and physical scale means that the exaggeration is especially strong for low redshift, and so it
+could plausibly favour methods that are better in that range.
 
@@
 This appendix details the different methods submitted to the challenge.
+Methods described as using optimized bin edges are those that appear as orange solid
+lines in \autoref{fig:edges}; those without this note appear as dashed blue lines there.
 
@@
 
-In this implementation we chose nominal bin edges by splittng ensuring equal numbers of training set 
+In this implementation we chose nominal bin edges by splitting ensuring equal numbers of training set 
 objects were in each bin, and then trained the forest to map magnitudes, magnitude errors, and
@@
 
-\subsection{ {\sc Autokeras LSTM}}
+\subsection{ {\sc AutokerasLSTM}}
 Autokeras \citep{autokeras} is an AutoML system based on Keras. Given a general neural network
@@
 
-A technical problem, involving conflicting versions of libraries and incompatible hardware,
-meant that this method could not be run in the final challenge.
-
 \subsection{ {\sc LGBM} }
@@
 \subsection{ {\sc SimpleSOM} }
-The SimpleSOM algorithm submitted by author AHW utilises a self-organising map
+The \textsc{SimpleSOM} algorithm submitted by author AHW utilises a self-organising map
 \citep[SOM,][]{Kohonen:1982} to perform  a discretisation of the
@@
 
-We utilise a branched version of the {\tt
+We utilise a branched version of the {\sc
 kohonen}\footnote{\url{https://github.com/ AngusWright/kohonen.git}} package
 \citep{Wright/etal:2020b, Wehrens/Kruisselbrink:2018, Wehrens/Lutgarde:2007}
-within R \citep{R}, which we integrate with  the challenge's {\tt python}
-classes using the {\tt rpy2} interface\footnote{\url{https://rpy2.github.io}}. 
+within R \citep{R}, which we integrate with  the challenge's python
+classes using the {\sc rpy2} interface\footnote{\url{https://rpy2.github.io}}. 
 We train a $101 \times 101$ hexagonal-cell SOM with toroidal topology on the
@@
 
-In addition to this base functionality, the SimpleSOM method includes a
+In addition to this base functionality, the \textsc{SimpleSOM} method includes a
 mechanism for distilling $N_{\rm cell}$ SOM cells into $N_{\rm group}<N_{\rm
@@
 
-The `por que no los dos' algorithm, submitted by author AHW, is a development of the SimpleSOM algorithm, adding the 
+The `por que no los dos' algorithm, submitted by author AHW, is a development of the \textsc{SimpleSOM} algorithm, adding the 
 additional complexity of also computing template-based photometric redshift point-estimates using BPZ 
@@
 This method was submitted by authors DA, AHW, AN, EG, BA, and ABr. 
-
-The {\tt ComplexSOM} method implements an additional optimization layer on the
-methodology used by {\tt SimpleSOM}.
+It used optimized bin edges.
+
+The {\sc ComplexSOM} method implements an additional optimization layer on the
+methodology used by {\sc SimpleSOM}.
 
@@
 
-The rationale behind the {\tt ComplexSOM} method is based on the observation
+The rationale behind the {\sc ComplexSOM} method is based on the observation
 that, while ${\sf C}^{\rm group}_\ell$ is a slow quantity to calculate for a
@@
 
-Once the SOM has been generated, the {\tt ComplexSOM} method proceeds as
+Once the SOM has been generated, the {\sc ComplexSOM} method proceeds as
 follows. First, the full SOM is distilled into a smaller set of ${\cal O}(100)$
@@
 
-This method was submitted by the author AP.
-
-JaxCNN and JaxResNet are convolutional neural network based algorithms that
+These method were submitted by the author AP.
+They used optimized bin edges.
+
+\textsc{JaxCNN} and \textsc{JaxResNet} are convolutional neural network based algorithms that
 map the relationship between the colour-magnitude data of galaxies to their bin
@@
 
-JaxCNN uses 2-layer convolutional neural networks and JaxResNet uses a stack of
+\textsc{JaxCNN} uses 2-layer convolutional neural networks and \textsc{JaxResNet} uses a stack of
 3 layers consisting of $1 \times 1, 3 \times 3$, and $1 \times 1$ convolutions
@@
 \subsection{ {\sc NeuralNetwork 1 \& 2} } \label{sec:nn}
-This method was submitted by the author FL.
+These methods were submitted by the author FL.
+They used optimized bin edges
 
@@
 Training is performed using batches of 5000 galaxies, over 2000 iterations under the ADAM optimizer \citep{adam} with a base learning rate of 0.001.
-The loss functions $\mathcal{L}$ used for training directly derive from the challenge metrics, and constitute the only difference between the NeuralNetwork 1 \& NeuralNetwork 2 entries:
+The loss functions $\mathcal{L}$ used for training directly derive from the challenge metrics, and constitute the only difference between the \textsc{NeuralNetwork1} \& \textsc{neurNetwork2} entries:
 
 \begin{itemize}
-	\item NeuralNetwork 1: $\mathcal{L}_1 = 1 / S_{FoM}$
-	\item NeuralNetwork 2: $\mathcal{L}_2 = - S_{SNR} $
+	\item \textsc{NeuralNetwork1}: $\mathcal{L}_1 = 1 / S_{FoM}$
+	\item \textsc{NeuralNetwork2}: $\mathcal{L}_2 = - S_{SNR} $
 \end{itemize}
@@
 This entry was submitted by author DG.
+It used optimized bin edges.
 
@@
 \url{https://github.com/dkirkby/zotbin}.
+It used optimized bin edges.
 
@@
 
-\subsection{ {\sc funbins} } \label{sec:funbins}
+\subsection{ {\sc FunBins} } \label{sec:funbins}
 This entry was submitted by author ABa.
@@
 
-Once the bin edges are determined by either of the methods above, the random forest is trained as described in section \ref{sec:randomforest}.
+Once the bin edges are fixed using either of the methods above, the random forest is trained as described in section \ref{sec:randomforest}.
 
@@
 
-\subsection{ {\sc mlpqna} }
+\subsection{ {\sc MLPQNA} }
 This entry was submitted by authors SC and MB.
@@
 cached values were then incorrectly picked up by subsequent runs on new data sets, leading
-to (at best) crashes and at worst silently poor scores.  We should have provided an explicit
+to (at best) crashes and at worst silently poor scores\footnote{We checked for this problem in 
+our final results by running each method in turn and checking for new files in the working
+directory.}.  
+We should have provided an explicit
 caching mechanism for entrants to avoid such issues.
@@
 \tablecaption{The 9-bin results for all metrics calculated for all methods on the CosmoDC2 sample.
+\textit{ww} are weak-lensing metrics, \textit{gg} clustering, and \textit{3x} the full 3x2pt.
  \label{tab:full_dc2}}
@@
 \begin{deluxetable*}{ |l | c c c | c c c | c c c | c c c | c c c | c c c |}
-\tablecaption{The 9-bin results for all the metrics calculated for all methods on the Buzzard sample. \label{tab:full_buzz}}
+\tablecaption{The 9-bin results for all the metrics calculated for all methods on the Buzzard sample. 
+\textit{ww} are weak-lensing metrics, \textit{gg} clustering, and \textit{3x} the full 3x2pt.
+\label{tab:full_buzz}
+}
 \tablehead{

### 2021-09-27 paper/results/make_figures.py

@@ -73,6 +73,32 @@ class EverythingIsNan:
     def __getitem__(self, x):
         return np.nan
 
+def make_disp_name(name):
+    repls = [
+        ('TensorFlow_FFNN', 'FFNN'),
+        ("myCombinedClassifiers", "Stacked Generalization"),
+        ('Autokeras_LSTM', 'AutokerasLSTM'),
+        ('mlpqna', 'MLPQNA'),
+        ('funbins', 'FunBins'),
+        ('ENSEMBLE1', 'ENSEMBLE'),
+    ]
+    for a, b in repls:
+        name = name.replace(a, b)
+    return rf"{{\sc {name} }}"
+
+def make_disp_name_plot(name):
+    repls = [
+        ('TensorFlow_FFNN', 'FFNN'),
+        ("myCombinedClassifiers", "Stacked Generalization"),
+        ('Autokeras_LSTM', 'AutokerasLSTM'),
+        ('mlpqna', 'MLPQNA'),
+        ('funbins', 'FunBins'),
+        ('ENSEMBLE1', 'ENSEMBLE'),
+    ]
+    for a, b in repls:
+        name = name.replace(a, b)
+    return name
+
 
 def make_table(results, metric, filename):
     N = np.array([3, 5, 7, 9])
@@ -86,7 +112,7 @@ def make_table(results, metric, filename):
 
     f = open(filename, 'w')
     for name in method_names:
-        disp_name = rf"{{\sc {name} }}".replace('TensorFlow_FFNN', 'FFNN').replace("_", r"\_").replace("myCombinedClassifiers", "Stacked Generalization")
+        disp_name = make_disp_name(name)
         row = [disp_name]
         for i, bands in enumerate(['riz', 'griz']):
             for n in N:
@@ -508,7 +534,7 @@ def make_9bin_table(results, filename):
     f = open(filename, 'w')
 
     for name in methods_reordered:
-        disp_name = rf"{{\sc {name} }}".replace('TensorFlow_FFNN', 'FFNN').replace("_", r"\_").replace("myCombinedClassifiers", "Stacked Generalization")
+        disp_name = make_disp_name(name)
         row = [disp_name]
         for i, bands in enumerate(['riz', 'griz']):
             for j, metric in enumerate(metrics):
@@ -573,9 +599,9 @@ def plot_rank_grid(dcz, buzz, filename):
             
             xlabels = row_labels if i == 1 else []
             
-            ylabels = method_names if j == 0 else []
[…]

### 2021-09-27 paper/paper.tex
more tweaks

@@
 simulation and applying realistic photometric redshift estimation methods to the
-sample (we defer this question in this work). They highlight the power of using large numbers of bins,
+sample (we defer the latter in this work). They highlight the power of using large numbers of bins,
 differences in ideal binnings when
@@
 In each case, classification is done using griz bands on the CosmoDC2 data set; equivalent plots for Buzzard and riz are shown in \autoref{app:metric_grids}.
-\emph{ww} indicates lensing-only metrics, \emph{gg} indicates clustering only, and \emph{3x2pt} their combination.
+\emph{ww} indicates lensing-only metrics, \emph{gg} indicates clustering only, and \emph{3x2} their combination.
 The three metrics are defined in \autoref{sec:metrics}.
@@
 Although this unrealistic aspect mostly affects algorithms equally, the relation between redshift, angle, and physical scale means that the exaggeration is especially strong for low redshift, and so it
-could plausibly favour methods that are better in that range.
+could plausibly favour methods that create narrower bins in that range.
 
@@
 
-These method were submitted by the author AP.
+These methods were submitted by the author AP.
 They used optimized bin edges.

### 2021-09-29 paper/paper.tex
fix typos

@@
 and assuming the numerical values for telescope and sky properties therein:
-for each galaxy it generates a poisson sample of the number of photons per band based on
+for each galaxy it generates a Poisson sample of the number of photons per band based on
 its true magnitude, the sky brightness, and instrumental characteristics.
@@
 The JAX library provides automatic differentiation of the calculation of the $C_\ell$ values, and so avoids the painful numerical sensitivity which usually affects Fisher matrix calculations.
-We verified this by comparing the code code to tuned finite difference estimates on CCL spectrum calculations.
+We verified this by comparing the code to tuned finite difference estimates on CCL spectrum calculations.
 

### 2021-10-14 paper/authors2.tex
putting things into OJA format

\author{Joe Zuntz\altaffilmark{1}, Fran\c{c}ois Lanusse\altaffilmark{2}, Alex I. Malz\altaffilmark{3}, Angus H. Wright\altaffilmark{3}, An\v{z}e Slosar\altaffilmark{4}, Bela Abolfathi\altaffilmark{5}, David Alonso\altaffilmark{6}, Abby Bault\altaffilmark{5}, Cl\'{e}cio R. Bom\altaffilmark{7}\altaffilmark{8}, Massimo Brescia\altaffilmark{9}, Adam Broussard\altaffilmark{10}, Jean-Eric Campagne\altaffilmark{11}, Stefano Cavuoti\altaffilmark{9}\altaffilmark{12}\altaffilmark{13}, Eduardo S. Cypriano\altaffilmark{14}, Bernardo M. O. Fraga\altaffilmark{7}, Eric Gawiser\altaffilmark{10}, Elizabeth J. Gonzalez\altaffilmark{7}\altaffilmark{15}\altaffilmark{16}, Dylan Green\altaffilmark{5}, Peter Hatfield\altaffilmark{6}, Kartheik Iyer\altaffilmark{17}, David Kirkby\altaffilmark{5}, Andrina Nicola\altaffilmark{18}, Erfan Nourbakhsh\altaffilmark{19}, Andy Park\altaffilmark{5}, Gabriel Teixeira\altaffilmark{7}, Katrin Heitmann\altaffilmark{20}, Eve Kovacs\altaffilmark{20}, Yao-Yuan Mao\altaffilmark{21},}
\collaboration{The LSST Dark Energy Science Collaboration, LSST DESC}

\altaffiltext{1}{Institute for Astronomy, University of Edinburgh, Edinburgh EH9 3HJ, United Kingdom}
\altaffiltext{2}{AIM, CEA, CNRS, Universit\'e Paris-Saclay, Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, F-91191 Gif-sur-Yvette, France}
\altaffiltext{3}{German Centre of Cosmological Lensing, Ruhr-Universit\"{a}t, Universit\"{a}tsstra{\ss}e 150, 44801 Bochum, Germany}
\altaffiltext{4}{Brookhaven National Laboratory, Upton NY 11973, USA}
\altaffiltext{5}{Department of Physics and Astronomy, University of California, Irvine, CA 92697, USA}
\altaffiltext{6}{Astrophysics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford, OX1 3RH, UK}
\altaffiltext{7}{Centro Brasileiro de Pesquisas F\'isicas, Rua Dr. Xavier Sigaud 150, 22290-180 Rio de Janeiro, RJ, Brazil}
\altaffiltext{8}{Centro Federal de Educa\c{c}\~{a}o Tecnol\'{o}gica Celso Suckow da Fonseca, Rodovia M\'{a}rcio Covas, lote J2, quadra J - Itagua\'{i} (Brazil)}
\altaffiltext{9}{INAF - Osservatorio Astronomico di Capodimonte, via Moiariello 16, I-80131, Napoli Italy}
\altaffiltext{10}{Department of Physics and Astronomy, Rutgers, the State University, Piscataway, NJ 08854, USA}
\altaffiltext{11}{Universit\'e Paris-Saclay, CNRS/IN2P3, IJCLab, 91405 Orsay, France}
\altaffiltext{12}{INFN - Sezione di Napoli, via Cinthia 21, I-80126 Napoli, Italy}
\altaffiltext{13}{Department of Physics ``E. Pancini'', University of Naples Federico II, via Cintia, 21, I-80126 Napoli, Italy}
\altaffiltext{14}{Universidade de S\~ao Paulo, IAG, Rua do Mat\~ao 1225, S\~ao Paulo, SP, Brazil}
\altaffiltext{15}{Instituto de Astronom\'{\i}a Te\'orica y Experimental (IATE-CONICET), Laprida 854, X5000BGR, C\'ordoba, Argentina.}
\altaffiltext{16}{Observatorio Astron\'omico de C\'ordoba, Universidad Nacional de C\'ordoba, Laprida 854, X5000BGR, C\'ordoba, Argentina.}
\altaffiltext{17}{Dunlap Institute for Astronomy and Astrophysics, University of Toronto, 50 St George St, Toronto, ON M5S 3H4, Canada}
\altaffiltext{18}{Department of Astrophysical Sciences, Princeton University, Peyton Hall, Princeton, NJ 08544, USA}
\altaffiltext{19}{Department of Physics and Astronomy, University of California, Davis, Davis, CA 95616, USA}
\altaffiltext{20}{Argonne National Laboratory, 9700 S Cass Ave, Lemont, IL 60439, USA}
\altaffiltext{21}{Department of Physics and Astronomy, Rutgers University, Piscataway, NJ 08854, USA}

### 2021-10-14 paper/paper.tex
Correction spotted by Irene
putting things into OJA format

@@
+\RequirePackage{rotating}
 \PassOptionsToPackage{usenames,dvipsnames}{xcolour}
-\documentclass[twocolumn,twocolappendix]{aastex63}
+\documentclass[twocolumn,twocolappendix]{openjournal}
 
@@
 \usepackage[utf8]{inputenc}
+\usepackage{hyperref}
 \usepackage[figure,figure*]{hypcap}
@@
 \usepackage{rotating}
-% \usepackage{pdflscape}
+\usepackage{multirow}
 \urlstyle{same}
@@
 \title{The LSST-DESC 3x2pt Tomography Optimization Challenge}
-\input{authors}
+\input{authors2}
 
@@
 \begin{equation}
-    S = \mu^{T} C^{-1} \mu
+    S = \sqrt{\mu^{T} C^{-1} \mu}
 \label{eq:snr}
@@
 \begin{table*}[]
+\centering
 	\begin{tabular}{|l|p{8cm}|c|}
@@
 \begin{table*}[]
+\centering
 	\begin{tabular}{|l|llll|llll|}
@@
 to (at best) crashes and at worst silently poor scores\footnote{We checked for this problem in 
-our final results by running each method in turn and checking for new files in the working
+our final results by running each method and checking for new files in the working
 directory.}.  
@@
 pathological bin assignments (-).  Failure rates were higher for the 9-bin runs than for the lower bin counts.
-
-\movetabledown=65mm
-\begin{rotatetable*}
-\begin{deluxetable*}{ |l | c c c | c c c | c c c | c c c | c c c | c c c |}
-\tablecaption{The 9-bin results for all metrics calculated for all methods on the CosmoDC2 sample.
-\textit{ww} are weak-lensing metrics, \textit{gg} clustering, and \textit{3x} the full 3x2pt.
- \label{tab:full_dc2}}
-\tablehead{
-\colhead{Method} & \multicolumn{9}{|c|}{riz} & \multicolumn{9}{c|}{griz}
-}
-\startdata
+\vspace{0.5cm}
+
+\renewcommand{\arraystretch}{1.8}
+\begin{sidewaystable}
+\vspace{10cm}
+\begin{tabular}{ |l | c c c | c c c | c c c | c c c | c c c | c c c |}
+\hline
+\multirow{3}{*}{Method} & \multicolumn{9}{|c|}{riz} & \multicolumn{9}{c|}{griz} \\
+\cline{2-19}
 &  \multicolumn{3}{c|}{SNR}   &  \multicolumn{3}{c|}{$\sigma_8-\Omega_m$ FOM} & \multicolumn{3}{c|}{$w_0-w_a$ FOM}
 &  \multicolumn{3}{c|}{SNR}   &  \multicolumn{3}{c|}{$\sigma_8-\Omega_m$ FOM} & \multicolumn{3}{c|}{$w_0-w_a$ FOM} \\
-\hline
+\cline{2-19}
 & ww & gg & 3x & ww & gg & 3x & ww & gg & 3x & ww & gg & 3x & ww & gg & 3x & ww & gg & 3x \\
@@
 \input{results/9bin_dc2.tex}
-    \enddata
-\end{deluxetable*}
-\end{rotatetable*}
-
-\movetabledown=65mm
-\begin{rotatetable*}
-\begin{deluxetable*}{ |l | c c c | c c c | c c c | c c c | c c c | c c c |}
-\tablecaption{The 9-bin results for all the metrics calculated for all methods on the Buzzard sample. 
+\hline
+\end{tabular}
+\caption{The 9-bin results for all metrics calculated for all methods on the CosmoDC2 sample.
+\textit{ww} are weak-lensing metrics, \textit{gg} clustering, and \textit{3x} the full 3x2pt.
+\label{tab:full_dc2}
+}
+\end{sidewaystable}
+
+\begin{sidewaystable}
+\vspace{10cm}
+\begin{tabular}{ |l | c c c | c c c | c c c | c c c | c c c | c c c |}
+\hline
+\multirow{3}{*}{Method} & \multicolumn{9}{|c|}{riz} & \multicolumn{9}{c|}{griz} \\
+\cline{2-19}
+&  \multicolumn{3}{c|}{SNR}   &  \multicolumn{3}{c|}{$\sigma_8-\Omega_m$ FOM} & \multicolumn{3}{c|}{$w_0-w_a$ FOM}
+&  \multicolumn{3}{c|}{SNR}   &  \multicolumn{3}{c|}{$\sigma_8-\Omega_m$ FOM} & \multicolumn{3}{c|}{$w_0-w_a$ FOM} \\
+\cline{2-19}
+& ww & gg & 3x & ww & gg & 3x & ww & gg & 3x & ww & gg & 3x & ww & gg & 3x & ww & gg & 3x \\
+\hline
+\input{results/9bin_buzzard.tex}
+\hline
+\end{tabular}
+\caption{The 9-bin results for all metrics calculated for all methods on the Buzzard sample.
 \textit{ww} are weak-lensing metrics, \textit{gg} clustering, and \textit{3x} the full 3x2pt.
@@
 }
-\tablehead{
-\colhead{Method} & \multicolumn{9}{c|}{riz} & \multicolumn{9}{c|}{griz}
-}
-\startdata
-&  \multicolumn{3}{c|}{SNR}   &  \multicolumn{3}{c|}{$\sigma_8-\Omega_m$ FOM} & \multicolumn{3}{c|}{$w_0-w_a$ FOM}
-&  \multicolumn{3}{c|}{SNR}   &  \multicolumn{3}{c|}{$\sigma_8-\Omega_m$ FOM} & \multicolumn{3}{c|}{$w_0-w_a$ FOM} \\
-\hline
-& ww & gg & 3x & ww & gg & 3x & ww & gg & 3x & ww & gg & 3x & ww & gg & 3x & ww & gg & 3x \\
-\hline
-\input{results/9bin_buzzard.tex}
-    \enddata
-\end{deluxetable*}
-\end{rotatetable*}
-
-% \begin{table}[htbp]
-% \centering
-% \end{table}
+\end{sidewaystable}
 

### 2021-10-15 paper/authors.tex
acks
pretend DESC is an author

@@
 \author[0000-0001-7363-7932]{Angus H. Wright}
-\affiliation{Ruhr University Bochum, Faculty of Physics and Astronomy, Astronomical Institute (AIRUB), German Centre for Cosmological Lensing, 44780 Bochum, Germany}
+\affiliation{German Centre of Cosmological Lensing, Ruhr-Universit\"{a}t, Universit\"{a}tsstra{\ss}e 150, 44801 Bochum, Germany}
 
@@
 \author[0000-0002-4598-9719]{David Alonso}
-\affiliation{Astrophysics, University of Oxford, DWB, Keble Road, Oxford OX1 3RH, UK}
+\affiliation{Astrophysics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford, OX1 3RH, UK}
 
@@
 
-\author{Andrina Nicola}
+\author[0000-0003-2792-6252]{Andrina Nicola}
 \affiliation{Department of Astrophysical Sciences, Princeton University, Peyton Hall, Princeton, NJ 08544, USA}
@@
 
-\collaboration{1000}{The LSST Dark Energy Science Collaboration (LSST DESC)}
+\author{The LSST Dark Energy Science Collaboration (LSST DESC)}
 \noaffiliation

### 2021-10-15 paper/authors2.tex

@@
 \collaboration{The LSST Dark Energy Science Collaboration, LSST DESC}
-
-\altaffiltext{1}{Institute for Astronomy, University of Edinburgh, Edinburgh EH9 3HJ, United Kingdom}
-\altaffiltext{2}{AIM, CEA, CNRS, Universit\'e Paris-Saclay, Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, F-91191 Gif-sur-Yvette, France}
-\altaffiltext{3}{German Centre of Cosmological Lensing, Ruhr-Universit\"{a}t, Universit\"{a}tsstra{\ss}e 150, 44801 Bochum, Germany}
-\altaffiltext{4}{Brookhaven National Laboratory, Upton NY 11973, USA}
-\altaffiltext{5}{Department of Physics and Astronomy, University of California, Irvine, CA 92697, USA}
-\altaffiltext{6}{Astrophysics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford, OX1 3RH, UK}
-\altaffiltext{7}{Centro Brasileiro de Pesquisas F\'isicas, Rua Dr. Xavier Sigaud 150, 22290-180 Rio de Janeiro, RJ, Brazil}
-\altaffiltext{8}{Centro Federal de Educa\c{c}\~{a}o Tecnol\'{o}gica Celso Suckow da Fonseca, Rodovia M\'{a}rcio Covas, lote J2, quadra J - Itagua\'{i} (Brazil)}
-\altaffiltext{9}{INAF - Osservatorio Astronomico di Capodimonte, via Moiariello 16, I-80131, Napoli Italy}
-\altaffiltext{10}{Department of Physics and Astronomy, Rutgers, the State University, Piscataway, NJ 08854, USA}
-\altaffiltext{11}{Universit\'e Paris-Saclay, CNRS/IN2P3, IJCLab, 91405 Orsay, France}
-\altaffiltext{12}{INFN - Sezione di Napoli, via Cinthia 21, I-80126 Napoli, Italy}
-\altaffiltext{13}{Department of Physics ``E. Pancini'', University of Naples Federico II, via Cintia, 21, I-80126 Napoli, Italy}
-\altaffiltext{14}{Universidade de S\~ao Paulo, IAG, Rua do Mat\~ao 1225, S\~ao Paulo, SP, Brazil}
-\altaffiltext{15}{Instituto de Astronom\'{\i}a Te\'orica y Experimental (IATE-CONICET), Laprida 854, X5000BGR, C\'ordoba, Argentina.}
-\altaffiltext{16}{Observatorio Astron\'omico de C\'ordoba, Universidad Nacional de C\'ordoba, Laprida 854, X5000BGR, C\'ordoba, Argentina.}
-\altaffiltext{17}{Dunlap Institute for Astronomy and Astrophysics, University of Toronto, 50 St George St, Toronto, ON M5S 3H4, Canada}
-\altaffiltext{18}{Department of Astrophysical Sciences, Princeton University, Peyton Hall, Princeton, NJ 08544, USA}
-\altaffiltext{19}{Department of Physics and Astronomy, University of California, Davis, Davis, CA 95616, USA}
-\altaffiltext{20}{Argonne National Laboratory, 9700 S Cass Ave, Lemont, IL 60439, USA}
-\altaffiltext{21}{Department of Physics and Astronomy, Rutgers University, Piscataway, NJ 08854, USA}

### 2021-10-15 paper/paper.tex

@@
 \input{authors2}
+\footnote{Author affiliations may be found before the references.}
 
@@
 Dark Energy Science Collaboration included here under a DESC publication policy exemption.
+\vspace{1cm}
+
+\subsection{Author Affiliations}
+
+$^{1}$Institute for Astronomy, University of Edinburgh, Edinburgh EH9 3HJ, United Kingdom\\
+$^{2}$AIM, CEA, CNRS, Universit\'e Paris-Saclay, Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, F-91191 Gif-sur-Yvette, France\\
+$^{3}$German Centre of Cosmological Lensing, Ruhr-Universit\"{a}t, Universit\"{a}tsstra{\ss}e 150, 44801 Bochum, Germany\\
+$^{4}$Brookhaven National Laboratory, Upton NY 11973, USA\\
+$^{5}$Department of Physics and Astronomy, University of California, Irvine, CA 92697, USA\\
+$^{6}$Astrophysics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford, OX1 3RH, UK\\
+$^{7}$Centro Brasileiro de Pesquisas F\'isicas, Rua Dr. Xavier Sigaud 150, 22290-180 Rio de Janeiro, RJ, Brazil\\
+$^{8}$Centro Federal de Educa\c{c}\~{a}o Tecnol\'{o}gica Celso Suckow da Fonseca, Rodovia M\'{a}rcio Covas, lote J2, quadra J - Itagua\'{i} (Brazil)\\
+$^{9}$INAF - Osservatorio Astronomico di Capodimonte, via Moiariello 16, I-80131, Napoli Italy\\
+$^{10}$Department of Physics and Astronomy, Rutgers, the State University, Piscataway, NJ 08854, USA\\
+$^{11}$Universit\'e Paris-Saclay, CNRS/IN2P3, IJCLab, 91405 Orsay, France\\
+$^{12}$INFN - Sezione di Napoli, via Cinthia 21, I-80126 Napoli, Italy\\
+$^{13}$Department of Physics ``E. Pancini'', University of Naples Federico II, via Cintia, 21, I-80126 Napoli, Italy\\
+$^{14}$Universidade de S\~ao Paulo, IAG, Rua do Mat\~ao 1225, S\~ao Paulo, SP, Brazil\\
+$^{15}$Instituto de Astronom\'{\i}a Te\'orica y Experimental (IATE-CONICET), Laprida 854, X5000BGR, C\'ordoba, Argentina.\\
+$^{16}$Observatorio Astron\'omico de C\'ordoba, Universidad Nacional de C\'ordoba, Laprida 854, X5000BGR, C\'ordoba, Argentina.\\
+$^{17}$Dunlap Institute for Astronomy and Astrophysics, University of Toronto, 50 St George St, Toronto, ON M5S 3H4, Canada\\
+$^{18}$Department of Astrophysical Sciences, Princeton University, Peyton Hall, Princeton, NJ 08544, USA\\
+$^{19}$Department of Physics and Astronomy, University of California, Davis, Davis, CA 95616, USA\\
+$^{20}$Argonne National Laboratory, 9700 S Cass Ave, Lemont, IL 60439, USA\\
+$^{21}$Department of Physics and Astronomy, Rutgers University, Piscataway, NJ 08854, USA
 
